### 2018-02-28 ms.Rmd
draft of ms outline with some intro stuff

---
title: Phylogenetic adaptions to urban living in Australian birds
author: "Corey T. Callaghan, Will Cornwell, John Wilshire, XXXX, XXXX, Richard T. Kingsford"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  bookdown::tufte_html2:
    number_sections: no
    toc: no
  bookdown::word_document2: null
bibliography: references.bib
fontsize: 12pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/global-ecology-and-biogeography.csl
---

# ABSTRACT

https://www.youtube.com/watch?v=9Gc4QTqslN4

# INTRODUCTION
Urbanization fragments and isolates natural habitat [@marzluff2001restoration; @melles2003urban; @loram2007urban], thus globally threatening biodiversity [@mckinney2006urbanization; @mckinney2008effects]. Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization], some species have successfully adapted to, colonized and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @lowry2013behavioural; @mcdonnell2015adaptation].

Within urban ecosystems there are generally urban avoiders, urban adapters, and urban exploiters [@mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation]. Urban avoiders are sensitive to urban environments [@mcdonnell2015adaptation] and tend to apportion more energy towards reproduction, relative to adapters and exploiters [@croci2008does]. Urban adapters can utilize urban environmental resources but may not reside in them [@mcdonnell2015adaptation] and are generally sedentary, social, high fecundity, with wide distribution [@kark2007living, @croci2008does]. Urban exploiters thrive in urban environments [@mcdonnell2015adaptation] and are generally more social and sedentary than adapters [@kark2007living]. Species’ traits such as social status, behavioral flexibility, migratory status, nesting, diet preferences, habitat preferences, and overall species range have all been used to predict presence in urban environments [@mcclure1989characterizes; @mills1989effects; @jokimaki1998distribution; @lim2004responses; @clergeau2006avifauna; @chace2006urban; @lowry2013behavioural]. Understanding which favored traits are associated with urban adapters and avoiders [@croci2008does] and their habitats can inform management decisions for conservation of particular species and improvements of habitats within urban environments.

Rather than a specific trait that predicts a species’ ability to be an urban avoider, adapter or exploiter, it is more likely a combination of traits. For instance, urban adapters’ presence in Jerusalem was predicted by a combination of traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living]. While a regional analysis throughout France and Switzerland demonstrated urban adapters prefer forested habitats, are sedentary, omnivorous, high-nesters, and were widely distributed [@croci2008does].

Despite the number of studies which have investigated the relationships between species' traits and urban adaptedness, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or two month survey period [@kark2007living]]. Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].

Fortunately, citizen science [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen] can provide broad-scale empirical datasets, allowing us to test hypotheses at spatial and temporal scales previously difficult to attain. This data provides a cost-effective method to research the relationship between species' traits and urban observations. Various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].

Here, we use data from eBird [@sullivan2009ebird; -@sullivan2014ebird], a successful citizen science project with over 500 million observations, submitted by more than 250,000 participants [@sullivan2017using], in order to elucidate adaptations in Australian birds to urban living. We combine eBird data with multiple input rasters, that represent a continuous levels of urbanness, over the whole of Australia. This is opposed to assigning species to a priori groups (adapters, avoiders, or exploiters) and investigating differences in traits among groups [e.g., @kark2007living]. We then apply phylogenetically controlled models in order to assess species traits that predict pre-adaptation to urban living. We hypothesize that a combination of traits allows species to thrive in urban environments.

# METHODS
## Study area

## Bird observation data
We downloaded the eBird basic dataset

## Assignment of landsat variables
We used Google Earth Engine [@gorelick2017google] to assign various landsat variables to each sampling checklist.

# RESULTS

A total of ```nrow(all_eBird_data)``` observations were used in the analysis.

# DISCUSSION

# REFERENCES

### 2018-03-21 scripts/creating_df_for_analysis.R
change negative avg_rad values to 0.00001

@@ -32,7 +32,7 @@ for (fileName in fileNames) {
 }
 
 ### combine to one file
-all_eBird_data <- bind_rows(df1, df2, df3, df4, df5, df6, df7, df8, df9)
+all_eBird_data <- bind_rows(df1, df2, df3, df4, df5, df6, df7, df8)
 
 ### clean up column names
 colnames(all_eBird_data) <- gsub(" ", "_", colnames(all_eBird_data))
@@ -45,6 +45,8 @@ rm(list=setdiff(ls(), c("checklists", "all_eBird_data")))
 ### Merge all eBird data with response variables file
 ### subset the sampling data to what is necessary and join each species with its own urbanness value
 species_urban <- checklists %>%
+  ## truncate any negative avg_rad values to zero
+  mutate(avg_rad = ifelse(.$avg_rad < 0, 0.00001, .$avg_rad)) %>%
   replace_na(list(EFFORT_DISTANCE_KM=0, EFFORT_AREA_HA=0)) %>%
   ## filter out any checklists > 5 km in distance travelled
   filter(EFFORT_DISTANCE_KM <= 5) %>%
@@ -141,6 +143,9 @@ species_urban <- species_urban %>%
 
 ## remove everything but the final subsetted eBird dataframe
 rm(list=setdiff(ls(), "species_urban"))
+
+## set wd back
+setwd("../../../")
 
 ## save as RData file
 save.image("Data/eBird data/Final eBird data for analysis/species_urban.RData")

### 2018-03-21 R/calculating_response_variables.R
calculating response variables and plotting the distribution

@@ -1,0 +1,47 @@
+### Calculating response variables for each species
+### This script is used to create the response variables for all species included in the analysis
+
+
+### load eBird data
+load_ebird_data <- function() {
+  load('Data/eBird data/Final eBird data for analysis/species_urban.RData')
+  species_urban
+}
+
+### calculate response variable df
+response_variables <- function() {
+  
+  response_variables <- species_urban %>%
+  group_by(COMMON_NAME) %>%
+  summarise(adapter= quantile(log(avg_rad), .90),
+            exploiter = quantile(log(avg_rad), .50),
+            avoider = quantile(log(avg_rad), .10),
+            N = n(),
+            unique_localities = length(unique(LOCALITY_ID)))
+  
+  response_variables
+  
+}
+
+
+### Plot the histogram of the response variables
+distribution_response_variables <- function(response_variables) {
+  
+  pdf("figures/distribution_response_variables.pdf")
+  
+  print(response_variables %>%
+    gather(key = "response_level", value = "value", adapter:avoider) %>%
+    ggplot(., aes(value))+
+    geom_histogram(bins = 50, 
+                   col="black", 
+                   fill="green", 
+                   alpha = .2)+
+    theme_bw()+
+    xlab("Value")+
+    ylab("Count")+
+    facet_wrap(~response_level, scales="free"))
+  
+  dev.off()
+}
+
+

### 2018-03-21 scripts/calculating_summary_statistics_per_species.R
remakifying R scripts

@@ -1,132 +1,0 @@
-### Calculating summary statistics for each species
-
-### load data
-load("Data/eBird data/Final eBird data for analysis/species_urban.RData")
-
-### load checklist data
-source("R/urbanness_functions.R")
-checklists <- get_urbanness_table()
-
-### packages
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(purrr)
-
-
-### select one species as an example
-AUPI <- species_urban %>%
-  filter(COMMON_NAME=="Australasian Pipit")
-
-### density plot of avg_radiance
-ggplot(AUPI, aes(avg_rad))+
-  geom_density()
-
-### 90th percentile
-quantile(AUPI$avg_rad, .90)
-
-
-### Attempt to account for sampling biases
-n <- nrow(AUPI)
-
-
-resample_lists <- function(data, reps, variable, species) {
-  
-  checklist_df <- data %>%
-    select(variable)
-  
-  response_vector <- species_urban %>%
-    filter(COMMON_NAME==species) %>%
-    select(variable)
-  
-  n <- nrow(response_vector)
-  
-  list_of_resamps <- replicate(n = reps, expr = {sample_n(checklist_df, n, replace = T)}, simplify = F)
-  
-  resamps_df <- data.frame(matrix(unlist(list_of_resamps), nrow=n, byrow=T), stringsAsFactors=FALSE)
-  
-  regression <- resamps_df %>%
-    map(~fitted.values(lm(.~response_vector[[1]]))) %>%
-    as.data.frame()
-  
-  quantile_results <- regression %>%
-    map(quantile, .90) %>%
-    as.data.frame() %>%
-    gather()
-  
-  out_df <- data.frame(raw_data=quantile(response_vector[[1]], .90),
-                       mean_sampled=mean(quantile_results$value),
-                       sd=sd(quantile_results$value),
-                       COMMON_NAME=species)
-  return(out_df)
-}
-
-AUPI_gam <- gam(avg_rad ~ s(LATITUDE, LONGITUDE), data=AUPI, family=gaussian)
-
-AUPI_gam_values <- predict(AUPI_gam)
-
-anova(AUPI_gam)
-
-## COMY
-COMY <- species_urban %>%
[…]

### 2018-03-21 R/calculating_response_variables.R
adding scientific name

@@ -12,7 +12,7 @@ load_ebird_data <- function() {
 response_variables <- function() {
   
   response_variables <- species_urban %>%
-  group_by(COMMON_NAME) %>%
+  group_by(COMMON_NAME, SCIENTIFIC_NAME) %>%
   summarise(adapter= quantile(log(avg_rad), .90),
             exploiter = quantile(log(avg_rad), .50),
             avoider = quantile(log(avg_rad), .10),

### 2018-03-21 R/calculating_response_variables.R
fixing remake and changing tree to plot urbanness

@@ -9,13 +9,13 @@ load_ebird_data <- function() {
 }
 
 ### calculate response variable df
-response_variables <- function() {
+make_response_variables <- function(species_urban) {
   
   response_variables <- species_urban %>%
   group_by(COMMON_NAME, SCIENTIFIC_NAME) %>%
-  summarise(adapter= quantile(log(avg_rad), .90),
-            exploiter = quantile(log(avg_rad), .50),
-            avoider = quantile(log(avg_rad), .10),
+  summarise(adapter= quantile(log(avg_rad), .90,na.rm=T),
+            exploiter = quantile(log(avg_rad), .50,na.rm=T),
+            avoider = quantile(log(avg_rad), .10,na.rm=T),
             N = n(),
             unique_localities = length(unique(LOCALITY_ID)))
   

### 2018-03-21 R/phylo_fxns.R

@@ -22,18 +22,31 @@ subset_tree<-function(trees,traits){
   }
 
 
-plot_bird_tree_traits<-function(aus_bird_tree,traits){
+plot_bird_tree_traits<-function(aus_bird_tree,traits,response_variables){
   trait<-as.array(traits$mean_body_size)
   row.names(trait)<-row.names(traits)
   trait<-subset(trait,trait!="NaN")
   trait<-subset(trait,names(trait)%in%aus_bird_tree$tip.label)
   tree_plotting<-drop.tip(aus_bird_tree,aus_bird_tree$tip.label[!aus_bird_tree$tip.label%in%row.names(trait)])
   
-  obj <- contMap(tree_plotting, log10(trait), fsize = c(0.6, 1), outline = FALSE)
+  response_variables$SCIENTIFIC_NAME<-gsub(" ","_",response_variables$SCIENTIFIC_NAME)
+  rv<-filter(response_variables,SCIENTIFIC_NAME%in%tree_plotting$tip.label)
+  exploiter<-as.array(rv$exploiter)
+  row.names(exploiter)<-rv$SCIENTIFIC_NAME
+  exploiter2<-exploiter-mean(exploiter)
   
-  pdf("figures/bird_plot.pdf")
-  plotTree.wBars(obj$tree, log10(trait), method = "plotSimmap", colors = obj$cols,
-                 type = "fan", scale = 5)
+  
+  tree_plotting_2<-drop.tip(tree_plotting,tree_plotting$tip.label[!tree_plotting$tip.label%in%row.names(exploiter2)])
+  trait<-subset(trait,names(trait)%in%tree_plotting_2$tip.label)
+  
+  obj <- contMap(tree_plotting_2, exploiter2, fsize = c(0.6, 1), outline = FALSE,plot=FALSE,type="fan")
+
+  pdf("figures/bird_urbanness_phylo.pdf")
+  plotTree.wBars(obj$tree, exploiter2, method = "plotSimmap", colors = obj$cols,
+                 type = "fan", scale = 5,tip.labels=FALSE)
   dev.off()
-}
+  pdf("figures/ref_tree.pdf")
+  plot(tree_plotting_2,type="f",cex=0.2)
+  dev.off()
+  }
 

### 2018-03-22 reports/phylo_report.rmd
allowing for multiple traits in phylo analysis

---
title: "phylo_analyses"
author: "Everyone"
date: "3/22/2018"
output: pdf_document
---

```{r}
  library(phylolm)

  trait<-data.frame(mean_body_size=traits$mean_body_size,clutch_size=traits$clutch_size)
  row.names(trait)<-row.names(traits)
  trait<-subset(trait,trait$mean_body_size!="NaN"&trait$clutch_size!="NaN")
  trait<-subset(trait,row.names(trait)%in%aus_bird_tree$tip.label)
  tree_plotting<-drop.tip(aus_bird_tree,aus_bird_tree$tip.label[!aus_bird_tree$tip.label%in%row.names(trait)])

  response_variables$SCIENTIFIC_NAME<-gsub(" ","_",response_variables$SCIENTIFIC_NAME)
  rv<-filter(response_variables,SCIENTIFIC_NAME%in%tree_plotting$tip.label)
  exploiter<-as.array(rv$exploiter)
  row.names(exploiter)<-rv$SCIENTIFIC_NAME
  exploiter2<-exploiter-mean(exploiter)

  tree_plotting_2<-drop.tip(tree_plotting,tree_plotting$tip.label[!tree_plotting$tip.label%in%row.names(exploiter2)])
  trait<-subset(trait,row.names(trait)%in%tree_plotting_2$tip.label)
  dd<-data.frame(urb=exploiter2)
  dd$body_size<-trait$mean_body_size[match(row.names(dd),row.names(trait))]
  dd$clutch_size<-trait$clutch_size[match(row.names(dd),row.names(trait))]
  summary(phylolm(urb~log10(body_size)+clutch_size,data=dd,phy=tree_plotting_2))
```

```{r}
  library(phylosignal)
  library(phylobase)
  dd$body_size<-log10(dd$body_size)
  out<-phylo4d(tree_plotting_2,tip.data=dd)
  phyloSignal(out)
```

### 2018-03-22 R/phylo_fxns.R

@@ -48,5 +48,8 @@ plot_bird_tree_traits<-function(aus_bird_tree,traits,response_variables){
   pdf("figures/ref_tree.pdf")
   plot(tree_plotting_2,type="f",cex=0.2)
   dev.off()
-  }
+}
+
+
+
 

### 2018-03-22 R/trait_fxns.R

@@ -10,8 +10,9 @@ read_process_trait_data<-function(){
   traits<-read_trait_data_in()
   traits %>%
     mutate(body_mass=ifelse(`99_Body_mass_average_8`=="NAV",NA,as.numeric(`99_Body_mass_average_8`))) %>%
+    mutate(clutch_size=ifelse(`178_Clutch_size_average_12`=="NAV",NA,as.numeric(`178_Clutch_size_average_12`))) %>%
     group_by(binom) %>%
-    summarise(mean_body_size=mean(body_mass,na.rm=T)) -> ms
+    summarise(mean_body_size=mean(body_mass,na.rm=T),clutch_size=mean(clutch_size,na.rm=T)) -> ms
               
     ms<-data.frame(ms)
    row.names(ms)<-ms$binom

### 2018-03-22 R/trait_fxns.R
added a few more traits

@@ -11,6 +11,9 @@ read_process_trait_data<-function(){
   traits %>%
     mutate(body_mass=ifelse(`99_Body_mass_average_8`=="NAV",NA,as.numeric(`99_Body_mass_average_8`))) %>%
     mutate(clutch_size=ifelse(`178_Clutch_size_average_12`=="NAV",NA,as.numeric(`178_Clutch_size_average_12`))) %>%
+    mutate(category=`29_Population_description_4`) %>%
+    mutate(status=`49_Australian_status_July_2015_5`) %>%
+    mutate(brain_size=ifelse(`113_Brain_mass_8`=="NAV", NA, as.numeric(`113_Brain_mass_8`))) %>%
     group_by(binom) %>%
     summarise(mean_body_size=mean(body_mass,na.rm=T),clutch_size=mean(clutch_size,na.rm=T)) -> ms
               

### 2018-03-22 R/trait_fxns.R
adding more traits

@@ -5,17 +5,32 @@ traits<-read_csv("Data/Raw trait data/Australian_Bird_Data_Version_1.csv")
          return(traits)
 }
 
+Mode <- function(x) {
+  ux <- unique(x)
+  ux[which.max(tabulate(match(x, ux)))]
+}
+
 
 read_process_trait_data<-function(){
   traits<-read_trait_data_in()
-  traits %>%
+    traits %>%
     mutate(body_mass=ifelse(`99_Body_mass_average_8`=="NAV",NA,as.numeric(`99_Body_mass_average_8`))) %>%
     mutate(clutch_size=ifelse(`178_Clutch_size_average_12`=="NAV",NA,as.numeric(`178_Clutch_size_average_12`))) %>%
     mutate(category=`29_Population_description_4`) %>%
     mutate(status=`49_Australian_status_July_2015_5`) %>%
     mutate(brain_size=ifelse(`113_Brain_mass_8`=="NAV", NA, as.numeric(`113_Brain_mass_8`))) %>%
+    mutate(feeding_habitat_generalism=rowSums(.[115:145], na.rm=TRUE)) %>%
+    mutate(breeding_generalism=rowSums(.[146:162], na.rm=TRUE)) %>%
+    mutate(diet_generalism=rowSums(.[163:173], na.rm=TRUE))
     group_by(binom) %>%
-    summarise(mean_body_size=mean(body_mass,na.rm=T),clutch_size=mean(clutch_size,na.rm=T)) -> ms
+    summarise(mean_body_size=mean(body_mass,na.rm=T), 
+              clutch_size=mean(clutch_size,na.rm=T),
+              category=Mode(category),
+              status=Mode(status),
+              brain_size=mean(brain_size, na.rm=T),
+              feeding_habitat_generalism=mean(feeding_habitat_generalism),
+              breeding_habitat_generalism=mean(breeding_habitat_generalism),
+              diet_generalism=mean(diet_generalism)) -> ms
               
     ms<-data.frame(ms)
    row.names(ms)<-ms$binom

### 2018-03-22 R/trait_fxns.R
fixed errors

@@ -20,8 +20,8 @@ read_process_trait_data<-function(){
     mutate(status=`49_Australian_status_July_2015_5`) %>%
     mutate(brain_size=ifelse(`113_Brain_mass_8`=="NAV", NA, as.numeric(`113_Brain_mass_8`))) %>%
     mutate(feeding_habitat_generalism=rowSums(.[115:145], na.rm=TRUE)) %>%
-    mutate(breeding_generalism=rowSums(.[146:162], na.rm=TRUE)) %>%
-    mutate(diet_generalism=rowSums(.[163:173], na.rm=TRUE))
+    mutate(breeding_habitat_generalism=rowSums(.[146:162], na.rm=TRUE)) %>%
+    mutate(diet_generalism=rowSums(.[163:173], na.rm=TRUE)) %>%
     group_by(binom) %>%
     summarise(mean_body_size=mean(body_mass,na.rm=T), 
               clutch_size=mean(clutch_size,na.rm=T),

### 2018-03-22 R/density_plot_per_species.R
added plots of histograms in addition to the density plots

@@ -26,6 +26,25 @@ avg_rad_density_for_each_species <- function (species_urban) {
   
   dev.off()
   
+  pdf("figures/histogram_plot_per_species.pdf")
+  
+  for (i in 1:n_pages) {
+    
+    print(ggplot(species_urban, aes(avg_rad)) +
+            geom_histogram(bins = 50, 
+                           col="black", 
+                           fill="green", 
+                           alpha = .2) + 
+            scale_x_log10() + 
+            theme_classic() +
+            xlab("Average radiance") +
+            ylab("Count") +
+            facet_wrap_paginate(~COMMON_NAME, ncol = 4, nrow = 5, scales="free_y", page = i))
+    
+  }
+  
+  dev.off()
+  
   
 }
 

### 2018-03-22 scripts/playing_with_response_variable.R
scratchings...

@@ -1,0 +1,134 @@
+## packages
+library(ggplot2)
+library(dplyr)
+library(fitdistrplus)
+
+
+## These would be the theoretical groups of birds
+## based on the beta distribution
+avoider <- data.frame(values=rbeta(10000, 0.5, 2),
+                      type="avoider", stringsAsFactors = FALSE)
+adapter <- data.frame(values=rbeta(10000, 2, 0.5),
+                      type="adapter", stringsAsFactors = FALSE)
+exploiter <- data.frame(values=rbeta(10000, 2, 2),
+                      type="exploiter", stringsAsFactors = FALSE)
+stork <- bind_rows(avoider, adapter, exploiter)
+
+## plot them
+ggplot(stork, aes(x=values, fill=type))+
+  geom_histogram()+
+  facet_wrap(~type)
+
+
+## The theoretical distributions are made by alpha and beta
+## So, if a species is a typical exploiter it should have a value of 1
+## avoider should have 4 and adapter should have 0.25, etc.
+# avoider <- 0.5/2
+# adapter <- 2/0.5
+# exploiter <- 2/2
+
+## Showing this with simulated data
+## avoider
+beta_fit <- qmedist(avoider$values, "beta", probs=c(1/3, 2/3))
+beta_fit_estimate <- as.data.frame(beta_fit$estimate)
+beta_fit_estimate[1,1]/beta_fit_estimate[2,1]
+
+## adapter
+beta_fit <- qmedist(adapter$values, "beta", probs=c(1/3, 2/3))
+beta_fit_estimate <- as.data.frame(beta_fit$estimate)
+beta_fit_estimate[1,1]/beta_fit_estimate[2,1]
+
+## exploiter
+beta_fit <- qmedist(exploiter$values, "beta", probs=c(1/3, 2/3))
+beta_fit_estimate <- as.data.frame(beta_fit$estimate)
+beta_fit_estimate[1,1]/beta_fit_estimate[2,1]
+
+
+
+## pick a couple species and test this theory!
+## load data
+load('Data/eBird data/Final eBird data for analysis/species_urban.RData')
+
+## Dusky Moorhen
+DUMO <- species_urban %>%
+  filter(COMMON_NAME == "Dusky Moorhen") %>%
+  dplyr::select(avg_rad) %>%
+  mutate(avg_rad_log = log(avg_rad))
+
[…]

### 2018-03-22 scripts/playing_with_response_variable.R
adding a bootstrap component

@@ -49,6 +49,8 @@ beta_fit_estimate[1,1]/beta_fit_estimate[2,1]
 ## load data
 load('Data/eBird data/Final eBird data for analysis/species_urban.RData')
 
+species_urban$avg_rad_log <- log(species_urban$avg_rad)
+
 ## Dusky Moorhen
 DUMO <- species_urban %>%
   filter(COMMON_NAME == "Dusky Moorhen") %>%
@@ -63,7 +65,7 @@ ggplot(DUMO, aes(x=avg_rad_log))+
 
 # Need to put values between 0-1
 # function for this
-range01 <- function(x){(x-min(x))/(max(x)-min(x))}
+range01 <- function(x){(x-min(species_urban$avg_rad_log))/(max(species_urban$avg_rad_log)-min(species_urban$avg_rad_log))}
 
 # transform data
 DUMO$avg_rad_log_transform <- range01(DUMO$avg_rad_log)
@@ -100,7 +102,6 @@ beta_fit <- qmedist(DUGR$avg_rad_log_transform, "beta", probs=c(1/3, 2/3))
 beta_fit_estimate <- as.data.frame(beta_fit$estimate)
 beta_fit_estimate[1,1]/beta_fit_estimate[2,1]
 
-
 ## Still not convinced. Need to test it
 
 ## write function to calculate "value" to test it a little bit
@@ -132,3 +133,64 @@ rv("Spotted Dove")
 rv("Spotted Bowerbird")
 rv("Varied Lorikeet")
 rv("Western Rosella")
+
+
+## Looks like it might work
+## Want to see about using bootstrap to 
+## account for sample size of the distibrutions it is estimating
+
+# Want to see about using bootstrap
+# need to use fitdist function instead of qmedist
+# does the same thing though, but returns different object
+beta_fit <- fitdist(DUGR$avg_rad_log_transform, "beta", method="qme", probs=c(1/3, 2/3))
+beta_fit_estimate <- as.data.frame(beta_fit$estimate)
+beta_fit_estimate[1,1]/beta_fit_estimate[2,1]
+
+# function to run bootstraps
+# can automatically parallelize it!
+boot_results <- bootdist(beta_fit, niter=1000, bootmethod="nonparam", parallel="snow", ncpus=10)
+
+boot_results_df <- boot_results[['estim']]
+
+estimate <- boot_results_df %>%
+  mutate(rv=shape1/shape2) %>%
[…]

### 2018-03-22 scripts/playing_with_response_variable.R
more testing. not sure it works well

@@ -180,6 +180,54 @@ estimate <- boot_results_df %>%
 
 estimate
 
+# Or another way would be to use the summary and CI they provide?
+test_df <- as.data.frame(t(boot_results[["CI"]]))
+
+test_df %>%
+  mutate(rv=shape1/shape2)
+
+# that doesn't seem to work
+
+## rewrite function to calculate the bootstrapped estimates
+rv <- function(species) {
+  
+  data <- species_urban %>%
+    filter(COMMON_NAME == species) %>%
+    dplyr::select(avg_rad) %>%
+    mutate(avg_rad_log = log(avg_rad))
+  
+  # transform data
+  data$avg_rad_log_transform <- range01(data$avg_rad_log)
+  
+  # fit beta distribution
+  beta_fit <- fitdist(data$avg_rad_log_transform, "beta", method="qme", probs=c(1/3, 2/3))
+  beta_fit_estimate <- as.data.frame(beta_fit$estimate)
+  beta_fit_estimate[1,1]/beta_fit_estimate[2,1]
+  
+  boot_results <- bootdist(beta_fit, niter=1000, bootmethod="nonparam", parallel="snow", ncpus=10)
+  
+  boot_results_df <- boot_results[['estim']]
+  
+  estimate <- boot_results_df %>%
+    mutate(rv=shape1/shape2) %>%
+    summarise(mean_rv = mean(rv),
+              sd_rv = sd(rv))
+  
+  return(estimate)
+  
+}
+
+
+rv("Common Myna")
+rv("Australian Ibis")
+rv("Southern Emuwren")
+rv("Noisy Miner")
+rv("Dusky Grasswren")
+rv("Spinifex Pigeon")
+rv("Spotted Dove")
+rv("Spotted Bowerbird")
+rv("Varied Lorikeet")
+rv("Western Rosella")
 
 
 

### 2018-03-27 scripts/mixtools_means.R
mixtools examples

@@ -1,0 +1,41 @@
+library(dplyr)
+library(mixtools)
+# https://www.r-bloggers.com/fitting-mixture-distributions-with-the-r-package-mixtools/
+load('Data/eBird data/Final eBird data for analysis/species_urban.RData')
+
+test_species <- c("Common Myna", "Australian Ibis", "Southern Emuwren", "Noisy Miner", 
+                  "Dusky Grasswren", "Spinifex Pigeon", "Spotted Dove", "Spotted Bowerbird",
+                  "Varied Lorikeet", "Western Rosella")
+
+
+
+
+species_urban %>% group_by(COMMON_NAME) %>% 
+  summarise(
+    nobs = n(),
+    log_mean = log(mean(avg_rad))
+  ) %>% arrange(desc(log_mean)) -> sp_summary
+
+sp_summary %>% 
+  ggplot(aes(log_mean)) + geom_histogram()
+
+sp_summary$mean %>% log %>% density(bw = 0.2) %>% plot
+mixtools::normalmixEM(sp_summary$mean %>% log, k = 3) -> mxdist3
+mixtools::normalmixEM(sp_summary$mean %>% log, k = 2) -> mxdist2
+mixtools::normalmixEM(sp_summary$mean %>% log, k = 4) -> mxdist4
+
+mxdist2 %>% summary
+mxdist3 %>% summary
+mxdist4 %>% summary
+# ok
+plot(mxdist2, which = 2)
+# good
+plot(mxdist3, which = 2)
+# bad
+plot(mxdist4, which = 2)
+
+plot(mxdist3, which = 2)
+favs <- sp_summary %>% filter(COMMON_NAME %in% test_species)
+abline(v = favs$log_mean)
+
+

### 2018-03-27 scripts/mixtools_means.R
no log

@@ -13,11 +13,11 @@ test_species <- c("Common Myna", "Australian Ibis", "Southern Emuwren", "Noisy M
 species_urban %>% group_by(COMMON_NAME) %>% 
   summarise(
     nobs = n(),
-    log_mean = log(mean(avg_rad))
-  ) %>% arrange(desc(log_mean)) -> sp_summary
+    mean = mean(avg_rad)
+  ) %>% arrange(desc(mean)) -> sp_summary
 
 sp_summary %>% 
-  ggplot(aes(log_mean)) + geom_histogram()
+  ggplot(aes(mean)) + geom_histogram()
 
 sp_summary$mean %>% log %>% density(bw = 0.2) %>% plot
 mixtools::normalmixEM(sp_summary$mean %>% log, k = 3) -> mxdist3

### 2018-03-27 R/trait_fxns.R
added traits and cleaned the previous work

@@ -11,28 +11,85 @@ Mode <- function(x) {
 }
 
 
-read_process_trait_data<-function(){
-  traits<-read_trait_data_in()
+read_process_trait_data <- function(){
+  
+  traits <- read_trait_data_in()
     traits %>%
     mutate(body_mass=ifelse(`99_Body_mass_average_8`=="NAV",NA,as.numeric(`99_Body_mass_average_8`))) %>%
     mutate(clutch_size=ifelse(`178_Clutch_size_average_12`=="NAV",NA,as.numeric(`178_Clutch_size_average_12`))) %>%
     mutate(category=`29_Population_description_4`) %>%
     mutate(status=`49_Australian_status_July_2015_5`) %>%
     mutate(brain_size=ifelse(`113_Brain_mass_8`=="NAV", NA, as.numeric(`113_Brain_mass_8`))) %>%
-    mutate(feeding_habitat_generalism=rowSums(.[115:145], na.rm=TRUE)) %>%
-    mutate(breeding_habitat_generalism=rowSums(.[146:162], na.rm=TRUE)) %>%
     mutate(diet_generalism=rowSums(.[163:173], na.rm=TRUE)) %>%
     group_by(binom) %>%
     summarise(mean_body_size=mean(body_mass,na.rm=T), 
               clutch_size=mean(clutch_size,na.rm=T),
               category=Mode(category),
               status=Mode(status),
-              brain_size=mean(brain_size, na.rm=T),
-              feeding_habitat_generalism=mean(feeding_habitat_generalism),
-              breeding_habitat_generalism=mean(breeding_habitat_generalism),
-              diet_generalism=mean(diet_generalism)) -> ms
+              brain_size=mean(brain_size, na.rm=T)) -> ms
               
-    ms<-data.frame(ms)
-   row.names(ms)<-ms$binom
+    
+## calculate df for feeding habitat generalism
+    feeding_habitat_generalism <- traits %>%
+      dplyr::select(binom, 115:144) %>%
+      replace(is.na(.), 0) %>%
+      group_by(binom) %>%
+      summarise_all(sum) %>%
+      mutate(feeding_habitat_generalism = rowSums(.[2:31])) %>%
+      dplyr::select(binom, feeding_habitat_generalism)
+    
+## calculate df for breeding habitat generalism
+    breeding_habitat_generalism <- traits %>%
+      dplyr::select(binom, 146:161) %>%
+      replace(is.na(.), 0) %>%
+      group_by(binom) %>%
+      summarise_all(sum) %>%
[…]

### 2018-03-28 R/trait_fxns.R
prelim traits

@@ -56,6 +56,32 @@ read_process_trait_data <- function(){
       mutate(diet_generalism = rowSums(.[2:12])) %>%
       dplyr::select(binom, diet_generalism)
     
+## Nesting generalism
+    nest_generalism <- traits %>%
+      dplyr::select(binom, 183:188) %>%
+      replace(is.na(.), 0) %>%
+      group_by(binom) %>%
+      summarise_all(sum) %>%
+      mutate(`183_Nest_location_Ground_level_12` = as.integer(as.character(gsub(2, 1, .$`183_Nest_location_Ground_level_12`)))) %>%
+      mutate(`184_Nest_location_Supported_12` = as.integer(as.character(gsub(2, 1, .$`184_Nest_location_Supported_12`)))) %>%
+      mutate(nest_generalism = rowSums(.[2:7])) %>%
+      dplyr::select(binom, nest_generalism)
+    
+## nesting categories of specific interest
+    nest_types <- traits %>%
+      dplyr::select(binom, 183, 187) %>%
+      replace(is.na(.), 0) %>%
+      group_by(binom) %>%
+      summarise_all(sum) %>%
+      rename(ground_nesting = `183_Nest_location_Ground_level_12`) %>%
+      rename(hollow_nesting = `187_Nest_location_Hollow_12`) %>%
+      mutate(ground_nesting = gsub(2, 1, .$ground_nesting)) %>%
+      mutate(ground_nesting = ifelse(ground_nesting == "1", "Yes", "No")) %>%
+      mutate(hollow_nesting = ifelse(hollow_nesting == 1, "Yes", "No"))
+    
+## join nesting
+    nesting <- inner_join(nest_types, nest_generalism, by="binom")
+    
 ## movement
     movement <- traits %>%
       dplyr::select(binom, 193:195) %>%
@@ -80,11 +106,24 @@ read_process_trait_data <- function(){
       mutate(movement_class = gsub(" and partial_migrant and ", "partial_migrant", .$movement_class)) %>%
       mutate(movement_class = gsub(" and  and ", "none", .$movement_class))
     
+## gregariousness
+    gregariousness <- traits %>%
+      dplyr::select(binom, 192) %>%
+      replace(is.na(.), 0) %>%
+      mutate_all(funs(str_replace(., "NAV", "0"))) %>%
+      rename(gregariousness = `192_Breeding_system_Cooperative_12`) %>%
+      mutate(gregariousness = as.integer(as.character(.$gregariousness))) %>%
[…]

### 2018-03-28 R/calculating_response_variables.R
updated response variable to include mean and median

@@ -13,9 +13,8 @@ make_response_variables <- function(species_urban) {
   
   response_variables <- species_urban %>%
   group_by(COMMON_NAME, SCIENTIFIC_NAME) %>%
-  summarise(adapter= quantile(log(avg_rad), .90,na.rm=T),
-            exploiter = quantile(log(avg_rad), .50,na.rm=T),
-            avoider = quantile(log(avg_rad), .10,na.rm=T),
+  summarise(urban_mean = mean(avg_rad),
+            urban_median = median(avg_rad),
             N = n(),
             unique_localities = length(unique(LOCALITY_ID)))
   
@@ -30,7 +29,7 @@ distribution_response_variables <- function(response_variables) {
   pdf("figures/distribution_response_variables.pdf")
   
   print(response_variables %>%
-    gather(key = "response_level", value = "value", adapter:avoider) %>%
+    gather(key = "response_level", value = "value", urban_mean:urban_median) %>%
     ggplot(., aes(value))+
     geom_histogram(bins = 50, 
                    col="black", 

### 2018-03-28 reports/phylo_report.rmd
bigger phylo model

@@
 
+match and subset to tree.  this is annoying because the phylo packages all use row.names and the tidyverse hates rownames.
+
+first create a new data.frame (not tibble) with just the traits of interest
 ```{r}
   library(phylolm)
+  trait<-data.frame(mean_body_size=traits$mean_body_size,clutch_size=traits$clutch_size,gregariousness=traits$gregariousness)
+  row.names(trait)<-row.names(traits)
+```
 
-  trait<-data.frame(mean_body_size=traits$mean_body_size,clutch_size=traits$clutch_size)
-  row.names(trait)<-row.names(traits)
-  trait<-subset(trait,trait$mean_body_size!="NaN"&trait$clutch_size!="NaN")
+subset that dataframe to those names that are both in the trait data base AND the tree
+```{r}  
+  trait<-subset(trait,trait$mean_body_size!="NaN"&trait$clutch_size!="NaN"&trait$gregariousness!="NaN")
   trait<-subset(trait,row.names(trait)%in%aus_bird_tree$tip.label)
+```
+
+subset the tree
+```{r}
   tree_plotting<-drop.tip(aus_bird_tree,aus_bird_tree$tip.label[!aus_bird_tree$tip.label%in%row.names(trait)])
-
   response_variables$SCIENTIFIC_NAME<-gsub(" ","_",response_variables$SCIENTIFIC_NAME)
   rv<-filter(response_variables,SCIENTIFIC_NAME%in%tree_plotting$tip.label)
-  exploiter<-as.array(rv$exploiter)
+  exploiter<-as.array(rv$exploiter) #the median
   row.names(exploiter)<-rv$SCIENTIFIC_NAME
@@
   dd<-data.frame(urb=exploiter2)
+```
+
+match traits into the right dataframe  
+```{r}    
   dd$body_size<-trait$mean_body_size[match(row.names(dd),row.names(trait))]
   dd$clutch_size<-trait$clutch_size[match(row.names(dd),row.names(trait))]
-  summary(phylolm(urb~log10(body_size)+clutch_size,data=dd,phy=tree_plotting_2))
+  dd$gregariousness<-trait$gregariousness[match(row.names(dd),row.names(trait))]
+
 ```
+
+run the model
+
+```{r}
+  mod1<-phylolm(urb~log10(body_size)+clutch_size+gregariousness,data=dd,phy=tree_plotting_2)
+summary(mod1)
+```
+
+maybe we eventually want to do model selection
+
+```{r}
+library(MuMIn)
+dredge(mod1)
+```
+
+this is sort of cool.  phylogenetic signal in all the traits
 
@@
   dd$body_size<-log10(dd$body_size)
-  out<-phylo4d(tree_plotting_2,tip.data=dd)
-  phyloSignal(out)
+  dd2<-subset(dd,select=-gregariousness)
+  out<-phylo4d(tree_plotting_2,tip.data=dd2)
+  ps<-phyloSignal(out)
+  knitr::kable(ps$stat)
 ```

### 2018-03-28 R/phylo_fxns.R
matching new response variables - using median - dropping terms

@@ -17,7 +17,7 @@ plot_bird_tree<-function(aus_bird_tree){
 
 subset_tree<-function(trees,traits){
   non_aus_sp<-trees$tip.label[!trees$tip.label %in% traits$binom]
-  aus_bird_tree<-drop.tip(trees,non_aus_sp)
+  aus_bird_tree<-drop.tip(bird_trees,non_aus_sp)
   return(aus_bird_tree)
   }
 
@@ -31,18 +31,18 @@ plot_bird_tree_traits<-function(aus_bird_tree,traits,response_variables){
   
   response_variables$SCIENTIFIC_NAME<-gsub(" ","_",response_variables$SCIENTIFIC_NAME)
   rv<-filter(response_variables,SCIENTIFIC_NAME%in%tree_plotting$tip.label)
-  exploiter<-as.array(rv$exploiter)
-  row.names(exploiter)<-rv$SCIENTIFIC_NAME
-  exploiter2<-exploiter-mean(exploiter)
+  median_rv<-as.array(rv$urban_median)
+  row.names(median_rv)<-rv$SCIENTIFIC_NAME
+  median_rv2<-median_rv-mean(median_rv)
   
   
-  tree_plotting_2<-drop.tip(tree_plotting,tree_plotting$tip.label[!tree_plotting$tip.label%in%row.names(exploiter2)])
+  tree_plotting_2<-drop.tip(tree_plotting,tree_plotting$tip.label[!tree_plotting$tip.label%in%row.names(median_rv2)])
   trait<-subset(trait,names(trait)%in%tree_plotting_2$tip.label)
   
-  obj <- contMap(tree_plotting_2, exploiter2, fsize = c(0.6, 1), outline = FALSE,plot=FALSE,type="fan")
+  obj <- contMap(tree_plotting_2, median_rv2, fsize = c(0.6, 1), outline = FALSE,plot=FALSE,type="fan")
 
   pdf("figures/bird_urbanness_phylo.pdf")
-  plotTree.wBars(obj$tree, exploiter2, method = "plotSimmap", colors = obj$cols,
+  plotTree.wBars(obj$tree, median_rv2, method = "plotSimmap", colors = obj$cols,
                  type = "fan", scale = 5,tip.labels=FALSE)
   dev.off()
   pdf("figures/ref_tree.pdf")

### 2018-03-28 reports/phylo_report.rmd
getting rid of "exploiter" terminology and just sticking with median.

@@
   rv<-filter(response_variables,SCIENTIFIC_NAME%in%tree_plotting$tip.label)
-  exploiter<-as.array(rv$exploiter) #the median
-  row.names(exploiter)<-rv$SCIENTIFIC_NAME
-  exploiter2<-exploiter-mean(exploiter)
+  median_rv<-as.array(rv$urban_median)
+  row.names(median_rv)<-rv$SCIENTIFIC_NAME
+  median_rv2<-median_rv-mean(median_rv)
 
-  tree_plotting_2<-drop.tip(tree_plotting,tree_plotting$tip.label[!tree_plotting$tip.label%in%row.names(exploiter2)])
+  tree_plotting_2<-drop.tip(tree_plotting,tree_plotting$tip.label[!tree_plotting$tip.label%in%row.names(median_rv2)])
   trait<-subset(trait,row.names(trait)%in%tree_plotting_2$tip.label)
-  dd<-data.frame(urb=exploiter2)
+
+  dd <- data.frame(urb=median_rv2)
+
 ```

### 2018-03-28 R/phylo_fxns.R
solving #18

@@ -17,7 +17,7 @@ plot_bird_tree<-function(aus_bird_tree){
 
 subset_tree<-function(trees,traits){
   non_aus_sp<-trees$tip.label[!trees$tip.label %in% traits$binom]
-  aus_bird_tree<-drop.tip(bird_trees,non_aus_sp)
+  aus_bird_tree<-drop.tip(trees,non_aus_sp)
   return(aus_bird_tree)
   }
 

### 2018-03-30 scripts/phylo_lme4.R
ways to use lme4 for phylogenies

@@ -1,0 +1,184 @@
+library(ape)
+library(lme4)
+library(Matrix)
+## additional possibilities for PGLMMs ...
+library(MCMCglmm)
+library(MASS) ## glmmPQL + ape::corBrownian()
+library(pez)
+## n.b. at present need 'modular' devel branch:
+## devtools::install_github("glmmTMB/glmmTMB/glmmTMB",ref="modular")
+library(glmmTMB)
+## utils
+library(dplyr)
+library(coda)
+library(lattice)
+library(broom) ## need install_github("bbolker/broom")
+library(dotwhisker)
+library(nlme)
+
+
+
+phylo.to.Z <- function(r) {
+  ntip <- length(r$tip.label)
+  Zid <- Matrix(0.0,ncol=length(r$edge.length),nrow=ntip)
+  nodes <- (ntip+1):max(r$edge)
+  root <- nodes[!(nodes %in% r$edge[,2])]
+  for (i in 1:ntip) {
+    cn <- i  ## current node
+    while (cn != root) {
+      ce <- which(r$edge[,2]==cn)   ## find current edge
+      Zid[i,ce] <- 1   ## set Zid to 1
+      cn <- r$edge[ce,1]            ## find previous node
+    }
+  }
+  Z <- t(r$edge.length * t(Zid))
+  return(Z)
+}
+
+#' split a square (block) matrix into component blocks 
+#' @param M square matrix
+#' @param ind indices (0,n1,n2,...) giving the endpoint of each block
+split_blkMat <- function(M,ind) {
+  res <- list()
+  if (length(ind)==1) return(list(M))
+  for (i in 1:(length(ind)-1)) {
+    v <- (ind[i]+1):ind[i+1]
+    res[[i]] <- M[v,v]
+  }
+  return(res)
+}
+
+#' modify reTrms object
+#' @param rt a reTrms object
+#' @param phylo a phylo object (phylogenetic tree)
+#' @param phylonm name of phylogenetic term in model
+#' @param phyloZ Z matrix built on branch length
+modify_phylo_retrms <- function(rt,phylo,phylonm,
+                                phyloZ=phylo.to.Z(phylo),sp) {
+  rep_phylo <- rt$Zt@Dim[2]/length(unique(sp)) ## number of column (aka same as number of obs)
[…]

### 2018-03-30 R/phylo_fxns.R
more functions

@@ -1,34 +1,122 @@
-read_bird_tree<-function(){
+
+
+
+read_bird_tree <- function() {
   require(ape)
-  if(!exists("Data/phylo/")) dir.create("Data/phylo/", showWarnings = FALSE)
-  download.file("https://www.dropbox.com/s/gfackoogv6n0bv0/AllBirdsEricson1.tre?raw=1","Data/phylo/phy.tre")
-  bird_trees<-ape::read.tree(file="Data/phylo/phy.tre")[[1]]
+  if (!exists("Data/phylo/"))
+    dir.create("Data/phylo/", showWarnings = FALSE)
+  download.file(
+    "https://www.dropbox.com/s/gfackoogv6n0bv0/AllBirdsEricson1.tre?raw=1",
+    "Data/phylo/phy.tre"
+  )
+  bird_trees <- ape::read.tree(file = "Data/phylo/phy.tre")[[1]]
   #drop tips to get down to Aus species here
   return(bird_trees)
 }
 
 
-plot_bird_tree<-function(aus_bird_tree){
+plot_bird_tree <- function(aus_bird_tree) {
   pdf("figures/bird_plot.pdf")
-  plot(aus_bird_tree,type="r",show.tip.label = TRUE,cex=0.4,show.node.label=TRUE)
+  plot(
+    aus_bird_tree,
+    type = "r",
+    show.tip.label = TRUE,
+    cex = 0.4,
+    show.node.label = TRUE
+  )
   dev.off()
 }
 
 
-subset_tree<-function(trees,traits){
-  non_aus_sp<-trees$tip.label[!trees$tip.label %in% traits$binom]
-  aus_bird_tree<-drop.tip(trees,non_aus_sp)
+subset_tree <- function(trees, traits) {
+  non_aus_sp <- trees$tip.label[!trees$tip.label %in% traits$binom]
+  aus_bird_tree <- drop.tip(trees, non_aus_sp)
   return(aus_bird_tree)
+}
+
+
+plot_bird_tree_traits <-
+  function(aus_bird_tree,
+           traits,
+           response_variables) {
+    trait <- as.array(traits$mean_body_size)
+    row.names(trait) <- row.names(traits)
+    trait <- subset(trait, trait != "NaN")
+    trait <- subset(trait, names(trait) %in% aus_bird_tree$tip.label)
+    tree_plotting <-
+      drop.tip(aus_bird_tree, aus_bird_tree$tip.label[!aus_bird_tree$tip.label %in%
+                                                        row.names(trait)])
+    
[…]

### 2018-03-30 reports/phylo_report.rmd
solving #18
not much to report at the moment

--- 2018-03-22 reports/phylo_report.rmd
@@
 
+run phylo model
+
 ```{r}
-  library(phylolm)
 
-  trait<-data.frame(mean_body_size=traits$mean_body_size,clutch_size=traits$clutch_size)
-  row.names(trait)<-row.names(traits)
-  trait<-subset(trait,trait$mean_body_size!="NaN"&trait$clutch_size!="NaN")
-  trait<-subset(trait,row.names(trait)%in%aus_bird_tree$tip.label)
-  tree_plotting<-drop.tip(aus_bird_tree,aus_bird_tree$tip.label[!aus_bird_tree$tip.label%in%row.names(trait)])
+mod2<-run_phylo_lm(traits,
+                         response_variables,
+                         aus_bird_tree)
 
-  response_variables$SCIENTIFIC_NAME<-gsub(" ","_",response_variables$SCIENTIFIC_NAME)
-  rv<-filter(response_variables,SCIENTIFIC_NAME%in%tree_plotting$tip.label)
-  exploiter<-as.array(rv$exploiter)
-  row.names(exploiter)<-rv$SCIENTIFIC_NAME
-  exploiter2<-exploiter-mean(exploiter)
+summary(mod2)
+```
+maybe we eventually want to do model selection
 
-  tree_plotting_2<-drop.tip(tree_plotting,tree_plotting$tip.label[!tree_plotting$tip.label%in%row.names(exploiter2)])
-  trait<-subset(trait,row.names(trait)%in%tree_plotting_2$tip.label)
-  dd<-data.frame(urb=exploiter2)
-  dd$body_size<-trait$mean_body_size[match(row.names(dd),row.names(trait))]
-  dd$clutch_size<-trait$clutch_size[match(row.names(dd),row.names(trait))]
-  summary(phylolm(urb~log10(body_size)+clutch_size,data=dd,phy=tree_plotting_2))
+```{r}
+# library(MuMIn)
+# dredge(mod2)
 ```
 
+this is sort of cool.  phylogenetic signal in all the traits
+
 ```{r}
-  library(phylosignal)
-  library(phylobase)
-  dd$body_size<-log10(dd$body_size)
-  out<-phylo4d(tree_plotting_2,tip.data=dd)
-  phyloSignal(out)
+  # library(phylosignal)
+  # library(phylobase)
+  # dd$body_size<-log10(dd$body_size)
+  # dd2<-subset(dd,select=-gregariousness)
+  # out<-phylo4d(tree_plotting_2,tip.data=dd2)
+  # ps<-phyloSignal(out)
+  # knitr::kable(ps$stat)
 ```

### 2018-03-31 R/phylo_fxns.R
gaussian process regression is bugged, but normal mixed model is working

@@ -185,20 +185,23 @@ run_phylo_lme4 <- function(traits,
   dd$gregariousness<-trait$gregariousness[match(row.names(dd),row.names(trait))]
   # run the model
   phylo<-tree_plotting_2
-  birdZ <- phylo.to.Z(tree_plotting_2)
+#  birdZ <- phylo.to.Z(tree_plotting_2)
   dd$phylo <- row.names(dd)
   dd$obs <- factor(seq(nrow(dd)))
   dd$urb_cat<-as.integer(dd$urb>0)
   dd$obs <- factor(seq(nrow(dd)))
 #  +(1|obs)
   
-  phylo_lme4_fit <- phylo_lmm(urb~log10(body_size)+(1|phylo)+(1|obs),sp=dd$phylo,
-                              data=dd,phylo=phylo,phylonm="phylo",
-                              control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"),
-                              phyloZ=birdZ)
+#  phylo_lme4_fit <- phylo_lmm(urb~log10(body_size)+(1|phylo),sp=dd$phylo,
+#                              data=dd,phylo=phylo,phylonm="phylo",
+#                              control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"),
+#                              phyloZ=birdZ)
   
-  
-  return(phylo_lme4_fit)
+  basic_gaus_fit <- lmer(urb~clutch_size+log10(body_size)+(1|phylo),
+                         data=dd,
+                         control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"))
+
+  return(basic_gaus_fit)
 }
 
 
@@ -206,7 +209,7 @@ run_phylo_lme4 <- function(traits,
 
 phylolm_r2 <- function(phylo_lme4_fit) {
   #stuff
-  Fixed<-fixef(phylo_lme4_fit)[2]*phylo_lme4_fit@frame$`log10(body_size)`
+  Fixed<-fixef(phylo_lme4_fit)[2]*phylo_lme4_fit@frame$clutch_size+fixef(phylo_lme4_fit)[3]*phylo_lme4_fit@frame$`log10(body_size)`
     varF<-var(Fixed)
     r2<-varF/(varF + VarCorr(phylo_lme4_fit)$phylo[1]  + attr(VarCorr(phylo_lme4_fit), "sc")^2)
     return(r2)  

### 2018-03-31 scripts/phylo_lme4.R

@@ -172,13 +172,16 @@ basic_gaus_fit <- lmer(phen~cofactor+(1|phylo),
 basic_gaus_res <- rbind(coef(summary(basic_gaus_fit))[,1:2],
                         cbind(matrix(unlist(VarCorr(basic_gaus_fit))),NA))
 
-phylo_gaus_fit <- phylo_lmm(phen~cofactor+(1|phylo)+(1|obs),sp=dat_gaus$phylo,
+phylo_gaus_fit <- phylo_lmm(phen~cofactor+(1|phylo),sp=dat_gaus$phylo,
                             data=dat_gaus,phylo=phylo,phylonm="phylo",
                             control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"),
                             phyloZ=phyloZ)
 
 lme4_gaus_res <- rbind(coef(summary(phylo_gaus_fit))[,1:2],
                        cbind(matrix(unlist(VarCorr(phylo_gaus_fit))),NA))
+
+tt_lmer_nocor_gaus <- tidy(basic_gaus_fit)
+tt_lmer_cor_gaus <- tidy(phylo_gaus_fit)
 
 print(basic_gaus_res)
 print(lme4_gaus_res)

### 2018-04-05 scripts/creating_lookup_table_of_species.R
write out file as separate dataset

@@ -264,7 +264,8 @@ rm(list=setdiff(ls(), "final_lookup"))
 
 
 
-
+## Write out lookup table
+write_csv(final_lookup, "Data/Taxonomy/taxonomy_key.csv")
 
 
 

### 2018-04-05 R/trait_fxns.R
subsetting to the species in our data in this R script using the created taxonomy key

@@ -1,9 +1,29 @@
-read_trait_data_in<-function(){
-  require(readr)
-traits<-read_csv("Data/Raw trait data/Australian_Bird_Data_Version_1.csv")
-         traits$binom<- paste(traits$`4_Genus_name_2`, traits$`5_Species_name_2`,sep="_")
-         return(traits)
+read_lookup_table_in <- function(){
+  taxonomy_key <- read_csv("Data/Taxonomy/taxonomy_key.csv")
 }
+
+
+## this will read in trait data
+## cut the trait data down to all those that
+## are in eBird and then assign the "binom" to
+## the scientific names of the tree so we get the best matches
+read_trait_data_in <- function(){
+  
+  require(readr)
+  
+traits <- read_csv("Data/Raw trait data/Australian_Bird_Data_Version_1.csv")
+
+traits$SCIENTIFIC_NAME_traits <- paste(traits$`4_Genus_name_2`, traits$`5_Species_name_2`,sep="_")
+
+traits <- traits %>%
+  inner_join(., taxonomy_key, by="SCIENTIFIC_NAME_traits") %>%
+  rename(binom = SCIENTIFIC_NAME_tree) %>%
+  mutate(binom = gsub(" ", "_", .$binom))
+
+  return(traits)
+}
+
+
 
 Mode <- function(x) {
   ux <- unique(x)
@@ -11,6 +31,8 @@ Mode <- function(x) {
 }
 
 
+## This uses the trait data to collect all
+## The necessary trait characteristics
 read_process_trait_data <- function(){
   
   traits <- read_trait_data_in()

### 2018-04-05 R/trait_fxns.R
needed to call dplyr::

@@ -88,7 +88,7 @@ read_process_trait_data <- function(){
       rename(plants2 = `167_Food_Corms_or_tubers`) %>%
       rename(carrion_eater = `170_Food_Carrion_10`) %>%
       mutate(plant_eater = plants1+plants2) %>%
-      select(-plants1, -plants2) %>%
+      dplyr::select(-plants1, -plants2) %>%
       mutate(granivore = ifelse(granivore == 1, "Yes", "no")) %>%
       mutate(insectivore = ifelse(insectivore == 1, "Yes", "no")) %>%
       mutate(plant_eater = ifelse(plant_eater == 2, "Yes", "no")) %>%

### 2018-04-05 R/calculating_response_variables.R
join taxonomy right to the response variables file as it makes it easier for later on!

@@ -8,15 +8,28 @@ load_ebird_data <- function() {
   species_urban
 }
 
+
+### read in taxonomy key
+read_lookup_table_in <- function(){
+  taxonomy_key <- read_csv("Data/Taxonomy/taxonomy_key.csv")
+  taxonomy_key
+}
+
+
 ### calculate response variable df
 make_response_variables <- function(species_urban) {
   
   response_variables <- species_urban %>%
-  group_by(COMMON_NAME, SCIENTIFIC_NAME) %>%
-  summarise(urban_mean = mean(avg_rad),
+    group_by(COMMON_NAME, SCIENTIFIC_NAME) %>%
+    summarise(urban_mean = mean(avg_rad),
             urban_median = median(avg_rad),
             N = n(),
-            unique_localities = length(unique(LOCALITY_ID)))
+            unique_localities = length(unique(LOCALITY_ID))) %>%
+    ungroup() %>%
+    rename(SCIENTIFIC_NAME_ebird = SCIENTIFIC_NAME) %>%
+    rename(COMMON_NAME_ebird = COMMON_NAME) %>%
+    inner_join(., taxonomy_key, by=c("COMMON_NAME_ebird", "SCIENTIFIC_NAME_ebird")) %>%
+    mutate(SCIENTIFIC_NAME_tree = gsub(" ", "_", .$SCIENTIFIC_NAME_tree))
   
   response_variables
   

### 2018-04-05 R/phylo_fxns.R
lots of changes! @wcornwell should probably check these. turn "traits" into "ms" which is created in the traits funx script.

@@ -28,32 +28,30 @@ plot_bird_tree <- function(aus_bird_tree) {
 }
 
 
-subset_tree <- function(trees, traits) {
-  non_aus_sp <- trees$tip.label[!trees$tip.label %in% traits$binom]
-  aus_bird_tree <- drop.tip(trees, non_aus_sp)
+subset_tree <- function(bird_trees, traits) {
+  non_aus_sp <- bird_trees$tip.label[!bird_trees$tip.label %in% traits$binom]
+  aus_bird_tree <- drop.tip(bird_trees, non_aus_sp)
   return(aus_bird_tree)
 }
 
 
 plot_bird_tree_traits <-
   function(aus_bird_tree,
-           traits,
+           ms,
            response_variables) {
-    trait <- as.array(traits$mean_body_size)
-    row.names(trait) <- row.names(traits)
+    trait <- as.array(ms$mean_body_size)
+    row.names(trait) <- row.names(ms)
     trait <- subset(trait, trait != "NaN")
     trait <- subset(trait, names(trait) %in% aus_bird_tree$tip.label)
     tree_plotting <-
       drop.tip(aus_bird_tree, aus_bird_tree$tip.label[!aus_bird_tree$tip.label %in%
                                                         row.names(trait)])
     
-    response_variables$SCIENTIFIC_NAME <-
-      gsub(" ", "_", response_variables$SCIENTIFIC_NAME)
     rv <-
       filter(response_variables,
-             SCIENTIFIC_NAME %in% tree_plotting$tip.label)
+             SCIENTIFIC_NAME_tree %in% tree_plotting$tip.label)
     median_rv <- as.array(rv$urban_median)
-    row.names(median_rv) <- rv$SCIENTIFIC_NAME
+    row.names(median_rv) <- rv$SCIENTIFIC_NAME_tree
     median_rv2 <- median_rv - mean(median_rv)
     
     
@@ -139,38 +137,38 @@ run_phylo_lm <- function(traits,
 }
 
 
-run_phylo_lme4 <- function(traits,
+run_phylo_lme4 <- function(ms,
                          response_variables,
                          aus_bird_tree) {
   
   #
   # match and subset to tree.
   # this is annoying because the phylo packages all use row.names and the tidyverse hates rownames.
+  ## CTC - the tibble package is helpful here with rownames to column and column to rownames within a pipe
[…]

### 2018-04-06 R/calculating_response_variables.R
reading in taxonomy look up

@@ -18,7 +18,7 @@ read_lookup_table_in <- function(){
 
 ### calculate response variable df
 make_response_variables <- function(species_urban) {
-  
+  taxonomy_key<-read_csv("Data/Taxonomy/taxonomy_key.csv")
   response_variables <- species_urban %>%
     group_by(COMMON_NAME, SCIENTIFIC_NAME) %>%
     summarise(urban_mean = mean(avg_rad),

### 2018-04-06 R/trait_fxns.R

@@ -10,7 +10,8 @@ read_lookup_table_in <- function(){
 read_trait_data_in <- function(){
   
   require(readr)
-  
+
+  taxonomy_key<-read_csv("Data/Taxonomy/taxonomy_key.csv")
 traits <- read_csv("Data/Raw trait data/Australian_Bird_Data_Version_1.csv")
 
 traits$SCIENTIFIC_NAME_traits <- paste(traits$`4_Genus_name_2`, traits$`5_Species_name_2`,sep="_")

### 2018-04-06 R/trait_fxns.R
getting rid of taxonomy key function that wasn't necessary

@@ -1,8 +1,3 @@
-read_lookup_table_in <- function(){
-  taxonomy_key <- read_csv("Data/Taxonomy/taxonomy_key.csv")
-}
-
-
 ## this will read in trait data
 ## cut the trait data down to all those that
 ## are in eBird and then assign the "binom" to
@@ -11,7 +6,7 @@ read_trait_data_in <- function(){
   
   require(readr)
 
-  taxonomy_key<-read_csv("Data/Taxonomy/taxonomy_key.csv")
+taxonomy_key <- read_csv("Data/Taxonomy/taxonomy_key.csv")
 traits <- read_csv("Data/Raw trait data/Australian_Bird_Data_Version_1.csv")
 
 traits$SCIENTIFIC_NAME_traits <- paste(traits$`4_Genus_name_2`, traits$`5_Species_name_2`,sep="_")

### 2018-04-06 R/trait_fxns.R
filtering out for final set of species to include in analyses

@@ -234,7 +234,12 @@ read_process_trait_data <- function(){
     
     ms <- data.frame(ms)
     
+    
+    ms <- ms %>%
+      filter(complete.cases(.))
+    
    row.names(ms) <- ms$binom
    
+  
   return(ms)  
 }

### 2018-04-06 R/hacking_phytools.R
better tree figure, with legend

@@ -1,0 +1,99 @@
+contMap<-function (tree, x, res = 100, fsize = NULL, ftype = NULL, lwd = 4, 
+          legend = NULL, lims = NULL, outline = TRUE, sig = 3, type = "phylogram", 
+          direction = "rightwards", plot = TRUE, ...) 
+{
+  if (!inherits(tree, "phylo")) 
+    stop("tree should be an object of class \"phylo\".")
+  if (hasArg(mar)) 
+    mar <- list(...)$mar
+  else mar <- rep(0.3, 4)
+  if (hasArg(offset)) 
+    offset <- list(...)$offset
+  else offset <- NULL
+  if (hasArg(method)) 
+    method <- list(...)$method
+  else method <- "fastAnc"
+  if (hasArg(hold)) 
+    hold <- list(...)$hold
+  else hold <- TRUE
+  if (hasArg(leg.txt)) 
+    leg.txt <- list(...)$leg.txt
+  else leg.txt <- "trait value"
+  h <- max(nodeHeights(tree))
+  steps <- 0:res/res * max(h)
+  H <- nodeHeights(tree)
+  if (method == "fastAnc") 
+    a <- fastAnc(tree, x)
+  else if (method == "anc.ML") {
+    fit <- anc.ML(tree, x)
+    a <- fit$ace
+    if (!is.null(fit$missing.x)) 
+      x <- c(x, fit$missing.x)
+  }
+  else if (method == "user") {
+    if (hasArg(anc.states)) 
+      anc.states <- list(...)$anc.states
+    else {
+      cat("No ancestral states have been provided. Using states estimated with fastAnc.\n\n")
+      a <- fastAnc(tree, x)
+    }
+    if (length(anc.states) < tree$Nnode) {
+      nodes <- as.numeric(names(anc.states))
+      tt <- tree
+      for (i in 1:length(nodes)) {
+        M <- matchNodes(tt, tree, method = "distances")
+        ii <- M[which(M[, 2] == nodes[i]), 1]
+        tt <- bind.tip(tt, nodes[i], edge.length = 0, 
+                       where = ii)
+      }
+      a <- fastAnc(tt, c(x, anc.states))
+      M <- matchNodes(tree, tt, method = "distances")
+      a <- a[as.character(M[, 2])]
+      names(a) <- M[, 1]
+    }
+    else {
+      if (is.null(names(anc.states))) 
[…]

### 2018-04-06 R/phylo_fxns.R

@@ -52,7 +52,7 @@ plot_bird_tree_traits <-
              SCIENTIFIC_NAME_tree %in% tree_plotting$tip.label)
     median_rv <- as.array(rv$urban_median)
     row.names(median_rv) <- rv$SCIENTIFIC_NAME_tree
-    median_rv2 <- median_rv - mean(median_rv)
+    median_rv2 <- median_rv #- mean(median_rv)
     
     
     tree_plotting_2 <-
@@ -70,7 +70,7 @@ plot_bird_tree_traits <-
         type = "fan"
       )
     
-    pdf("figures/bird_urbanness_phylo.pdf")
+    pdf("figures/bird_urbanness_phylo.pdf",width=8.5,height=8.5)
     plotTree.wBars(
       obj$tree,
       median_rv2,
@@ -80,8 +80,9 @@ plot_bird_tree_traits <-
       scale = 5,
       tip.labels = FALSE
     )
+    add.color.bar(100, obj$cols, title = "trait value", lims = obj$lims, prompt = FALSE,x = 0.9 * par()$usr[1], y = 0.9 * par()$usr[3])
     dev.off()
-    pdf("figures/ref_tree.pdf")
+    pdf("figures/ref_tree.pdf",width=8.5,height=8.5)
     plot(tree_plotting_2, type = "f", cex = 0.2)
     dev.off()
   }

### 2018-04-06 reports/non_phylo_model.Rmd
model selection (no phylo)

---
title: "non_phylo_analyses"
author: "Everyone"
date: "3/22/2018"
output: pdf_document
---

matching

```{r}
response_variables$binom<-response_variables$SCIENTIFIC_NAME_tree
df<-inner_join(response_variables,traits)
```

look at the continuous ones
```{r}
library(corrplot)
select_if(df, is.numeric) %>%
  cor(use="pairwise.complete.obs") %>%
  corrplot
```

completely naive model, plant eater taken out cause it only had one level.

Logging response cause it is right-skewed and produced crazy residual plots.

```{r}
names(df)
df %>%
filter_all(all_vars(!is.na(.)))->df2

naive_mod<-lm(log(urban_median)~log(mean_body_size)+clutch_size+iucn_status+log(brain_size)+feeding_habitat_generalism+
     breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+
     movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail")

plot(naive_mod)

```

looking at the model in more depth

```{r}
anova(naive_mod)
```
and some not-so-quick model selection.  dredge takes a little while because it considers all combinations of predictor variables.

```{r}
#library(MuMIn)
#dredge(naive_mod,trace = TRUE)
modselect<-step(naive_mod)
```

```{r}
library(visreg)
visreg(modselect)
```

### 2018-04-06 ms.Rmd
adding a bit of text to the methods

@@
 title: Phylogenetic adaptions to urban living in Australian birds
-author: "Corey T. Callaghan, Will Cornwell, John Wilshire, XXXX, XXXX, Richard T. Kingsford"
+author: "Corey T. Callaghan, John Wilshire, XXXX, XXXX, Richard T. Kingsford, William K. Cornwell"
 date: '`r format(Sys.time(), "%d %B, %Y")`'
@@
 ## Study area
+We looked at bird observations throughout the whole of continental Australia. Bird observations from outlying islands were eliminated from potential analysis.
 
 ## Bird observation data
-We downloaded the eBird basic dataset
+Bird observations originated from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset collected by citizen scientists.
+Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird].
+We downloaded the eBird basic dataset (version ebd_relFeb-2018), and used all observations between January 1, 2010 - February 28, 2018.
+This temporal scale was used because eBird data are richest.
+This time frame was also used to minimize undue leverage of mismatch between changes in the response variable measurement (see response variable below) over the given time frame.
 
-## Assignment of landsat variables
-We used Google Earth Engine [@gorelick2017google] to assign various landsat variables to each sampling checklist.
+We filtered the bird observations (sensu [@callaghan2017assessing; @larole2014role]) by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barret2003new].
+Further, we filtered the checklists to those with a distance travelled of < 5 km or an area covered which was < 500 Ha.
+Any checklists which were shared among multiple individuals were subsampled to include only one checklist in order to avoid duplicate results.
+
+Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we do not expect any correlation with seabirds using urban areas under normal circumstances.
+Only species which included > 100 observations in the eBird database over the specified time-frame, and that met our criteria were considered for further analyses.
+
+## Response variable - an urbanization index
+We used VIIRS night-time lights [@eldvidge2017viirs] as a proxy of a bird's association with urban habitat.
+The VIIRS night-time lights measures the electric lighting present on Earth's surface which is generally associated with human settlement.
+There are also a significant number of steps (e.g., filtering out background noise, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@eldvidge2017virrs].
+For all checklists which met our criteria above (N=length(unique(urbanness$SAMPLING_EVENT_IDENTIFIER))), we used the associated spatial coordinates and Google Earth Engine [@gorelick2017google] to assign the average radiance value calculated from the VIIRS layer.
+For each location, we calculated the average radiance within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias.
+
+Each checklist's average radiance value was then passed to every bird on that checklist, which provided every species with a distribution of average radiance values (Fig.
+1).
+For all species, we then used the median of this distribution as a continuous urbanization index.
+
+## Explanatory variables
+Explanatory variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological].
+Potential traits were chosen on the basis of their existing support and hypotheses in the current literature (e.g., [@mcclure1989characterizes; @kark2007living; @evans2011makes]).
+Traits were parsed in different manners, and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat).
+Supplementary table 1 provides an overview of all candidate traits, associated references for each trait, and more detailed information on how the traits were calculated.
+
+Only species which had complete data for the list of candidate traits were considered in the analyses.
+
+## Phylogenetic analysis
+For the phylogenetic portion of our analyses, we used the phylogenetic tree provided by Jetz et al.
+[@jetz2012global].
+This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global].
+
+In order to assess whether there was a phylogenetic signal of the urbanization index used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
+
+### Taxonomic considerations
+Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed.
+The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/].
+There were minor differences which existed in recognized species among the three datasets.
+A taxonomic key was made which combined the three different taxonomies (Supplementary Table 2).
+Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset.
+If it did not match both, then it was not considered for analysis (e.g., species currently only recognized by eBird/clements such as Western Whistler, Silver-backed Butcherbird, or Paperbark Flycatcher).
+Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia.
+
+## Models
 
 # RESULTS
-
 A total of ```nrow(all_eBird_data)``` observations were used in the analysis.

### 2018-04-06 reports/non_phylo_model.Rmd
model selection is a nightmare

@@
 
-naive_mod<-lm(log(urban_median)~log(mean_body_size)+clutch_size+iucn_status+log(brain_size)+feeding_habitat_generalism+
+naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+log(brain_size)+feeding_habitat_generalism+
      breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+
      movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail")
+summary(naive_mod)
+dim(df2)
+```
 
+```{r}
 plot(naive_mod)

### 2018-04-06 reports/phylo_report.rmd
gaussian process regression is bugged, but normal mixed model is working
model selection is a nightmare

@@
 ```{r}
+traits %>%
+    rownames_to_column('names') %>%
+    filter(complete.cases(.)) %>%
+    column_to_rownames('names') -> trait
 
-mod2<-run_phylo_lm(traits,
-                         response_variables,
-                         aus_bird_tree)
+  # subset the tree
+  matched_tree <- drop.tip(aus_bird_tree,aus_bird_tree$tip.label[!aus_bird_tree$tip.label%in%row.names(trait)])
 
-summary(mod2)
+  response_variables$binom<-response_variables$SCIENTIFIC_NAME_tree
+  inner_join(response_variables,trait) %>%
+    filter(binom!="not_treated_as_species") %>%
+    data.frame(.)-> trait
+
+  dd<-trait
+  dd$phylo<-dd$binom
+  phylo<-matched_tree # this is hilarious: tree has to be called "phylo" for lmer to work
 ```
-maybe we eventually want to do model selection
 
 ```{r}
-# library(MuMIn)
-# dredge(mod2)
+
+dd$lbody<-log10(dd$mean_body_size)
+dd$lbrain<-log10(dd$brain_size)
+
+ basic_gaus_fit <- lmer(urban_median~lbody+clutch_size+iucn_status+lbrain+feeding_habitat_generalism+
+     breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+
+     movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation+(1|phylo),
+                         data=dd,
+                         control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"),na.action = "na.fail")
+
+ AIC(basic_gaus_fit)
 ```
 
-this is sort of cool.  phylogenetic signal in all the traits
+Ben Bolker hates model selection (and signficance testing) so no easy way to do it on lmer objects.
+
+for now just test the model that the non-phylo found
 
 ```{r}
-  # library(phylosignal)
-  # library(phylobase)
-  # dd$body_size<-log10(dd$body_size)
-  # dd2<-subset(dd,select=-gregariousness)
-  # out<-phylo4d(tree_plotting_2,tip.data=dd2)
-  # ps<-phyloSignal(out)
-  # knitr::kable(ps$stat)
+dd$urban_median_rescale<-(dd$urban_median-mean(dd$urban_median))/sd(dd$urban_median)
+dd$clutch_size_rescale<-(dd$clutch_size-mean(dd$clutch_size))/sd(dd$clutch_size)
+dd$lclutch<-log(dd$clutch_size)
+
+ basic_gaus_fit_2 <- lmer(urban_median~lclutch+lbody+
+                          (1|phylo),
+                         data=dd,
+                         control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"),
+                        na.action = "na.fail")
+AIC(basic_gaus_fit_2)
+summary(basic_gaus_fit_2)
 ```
+
+trying with some other software--i think lmer is bugged for phylogenies.
+
+```{r}
+ basic_gaus_fit_phylolm <- phylolm(urban_median~lclutch+lbody,phy=phylo,
+                         data=dd)
+
+ summary(basic_gaus_fit_phylolm)
+ AIC(basic_gaus_fit_phylolm)
+```
+
+trying to find another factor that improves the model but it seems tough
+```{r}
+  basic_gaus_fit_phylolm <- phylolm(urban_median~lclutch+
+                                    lbody+feeding_habitat_generalism+
+                                    lbrain+insectivore+diet_generalism+
+                                    hollow_nesting,
+                                    phy=phylo,data=dd)
+
+logLik(basic_gaus_fit_phylolm)
+```
+
+lmer is totally bugged for phylo applications.  continue with phylolm next time.
+
+```{r}
+ basic_gaus_fit_phylolm_full <- phylolm(urban_median~iucn_status+lbrain+feeding_habitat_generalism+
+     carrion_eater+lclutch+lbody+
+     movement_class+ground_nesting+nest_generalism+nest_aggregation,
+                                    phy=phylo,data=dd)
+
+dredge(basic_gaus_fit_phylolm_full)
+
+```
+
+```{r}
+
+visreg(basic_gaus_fit_phylolm_full)
+```

### 2018-04-07 reports/non_phylo_model.Rmd
removed species which aren't species based on Jetz's tree. Makes sense to keep the same species set for both phylo and non-phylo models, I think.

@@
 response_variables$binom<-response_variables$SCIENTIFIC_NAME_tree
-df<-inner_join(response_variables,traits)
+df<-left_join(traits,response_variables, by="binom") %>%
+  filter(SCIENTIFIC_NAME_tree != "not_treated_as_species")
 ```

### 2018-04-07 ms.Rmd
Need help with including figures

@@
 ---
-title: Phylogenetic adaptions to urban living in Australian birds
-author: "Corey T. Callaghan, John Wilshire, XXXX, XXXX, Richard T. Kingsford, William K. Cornwell"
-date: '`r format(Sys.time(), "%d %B, %Y")`'
+title: "Phylogenetic adaptions to urban living in Australian birds"
+author: Corey T. Callaghan, John Wilshire, XXXX, XXXX, Richard T. Kingsford, William K. Cornwell
+date: ""
 output:
+  html_document:
+    df_print: paged
+    toc: no
+  bookdown::word_document2: null
+  pdf_document: null
   bookdown::tufte_html2:
@@
     toc: no
-  bookdown::word_document2: null
-bibliography: references.bib
 fontsize: 12pt
@@
 csl: https://raw.githubusercontent.com/citation-style-language/styles/master/global-ecology-and-biogeography.csl
+bibliography: references.bib
 ---
@@
 ## Study area
-We looked at bird observations throughout the whole of continental Australia. Bird observations from outlying islands were eliminated from potential analysis.
+We looked at bird observations throughout the whole of continental Australia. Bird observations from outlying islands were eliminated from potential analyses.
 
@@
 Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird].
-We downloaded the eBird basic dataset (version ebd_relFeb-2018), and used all observations between January 1, 2010 - February 28, 2018.
-This temporal scale was used because eBird data are richest.
-This time frame was also used to minimize undue leverage of mismatch between changes in the response variable measurement (see response variable below) over the given time frame.
+We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018.
+This temporal scale was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see response variable below) over the given time frame in which bird observations were collected.
 
-We filtered the bird observations (sensu [@callaghan2017assessing; @larole2014role]) by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barret2003new].
+We filtered the bird observations (sensu [@callaghan2017assessing; @la2014role]) by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barrett2003new].
+For further information on protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types].
 Further, we filtered the checklists to those with a distance travelled of < 5 km or an area covered which was < 500 Ha.
 Any checklists which were shared among multiple individuals were subsampled to include only one checklist in order to avoid duplicate results.
+After this sampling, we were left with 637,482 eligible checklists throughout Australia.
 
-Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we do not expect any correlation with seabirds using urban areas under normal circumstances.
+Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we did not expect any correlation with seabirds using urban areas under normal circumstances.
 Only species which included > 100 observations in the eBird database over the specified time-frame, and that met our criteria were considered for further analyses.
+This left us with 580 potential species from the eBird dataset which had > 100 observations on eligible checklists in the specified time frame.
 
 ## Response variable - an urbanization index
-We used VIIRS night-time lights [@eldvidge2017viirs] as a proxy of a bird's association with urban habitat.
+We used VIIRS night-time lights [@elvidge2017viirs] as a proxy of a bird's association with urban habitat.
 The VIIRS night-time lights measures the electric lighting present on Earth's surface which is generally associated with human settlement.
-There are also a significant number of steps (e.g., filtering out background noise, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@eldvidge2017virrs].
-For all checklists which met our criteria above (N=length(unique(urbanness$SAMPLING_EVENT_IDENTIFIER))), we used the associated spatial coordinates and Google Earth Engine [@gorelick2017google] to assign the average radiance value calculated from the VIIRS layer.
-For each location, we calculated the average radiance within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias.
+There are also a significant number of steps (e.g., filtering out background noise, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@elvidge2017viirs].
+For the 637,482 checklists which met our criteria above, we used the checklists' associated spatial coordinates and Google Earth Engine [@gorelick2017google] to assign the average radiance value calculated from the VIIRS layer.
+For each location, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
 
@@
 1).
-For all species, we then used the median of this distribution as a continuous urbanization index.
+Our response variable, which served as an urbanization index, was the median for each species' given distribution of average radiance values.
 
@@
 Traits were parsed in different manners, and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat).
-Supplementary table 1 provides an overview of all candidate traits, associated references for each trait, and more detailed information on how the traits were calculated.
+Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated.
 
@@
 The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/].
-There were minor differences which existed in recognized species among the three datasets.
-A taxonomic key was made which combined the three different taxonomies (Supplementary Table 2).
+There were minor differences which existed in common names, English names, and recognized species among the three datasets.
+A taxonomic key was made which combined the three different taxonomies (Appendix 2).
 Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset.
 If it did not match both, then it was not considered for analysis (e.g., species currently only recognized by eBird/clements such as Western Whistler, Silver-backed Butcherbird, or Paperbark Flycatcher).
-Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia.
+Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia and thus not included in the analysis.
 
@@
 # RESULTS
-A total of ```nrow(all_eBird_data)``` observations were used in the analysis.
+A total of 637,482 checklists provided a total of 5,944,819 observations of 580 different species which had > 100 observations.
+The least recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
+The mean +- SD number of observations for a species was 10250 +- 19269.19.
+
+We ended with 450 candidate species (Appendix 3) from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data.
+
+## Urbanization index
+
+### Phylogenetic signal
+There was a distinct phylogenetic signal in our urbanization index measure (Fig. 2; XXXXX SIGNIFICANCE TEST).
+
+## Model results
+### Non phylogenetic models
+### Phylogenetically constrained models
 
@@
 # REFERENCES
+
+# Figures
+
+```{r Figure 2, echo=FALSE, fig.cap="Figure 2. A phylogenetic tree of the birds yata yata yata...."}
+knitr::include_graphics(path = "figures/bird_urbanness_phylo.pdf")
+```

### 2018-04-07 ms.Rmd
differences between knit to html and knit to tufte_html_2. Can't figure out how to knit to pdf

@@
 output:
-  html_document:
-    df_print: paged
+  bookdown::tufte_html2:
+    number_sections: no
     toc: no
@@
   pdf_document: null
-  bookdown::tufte_html2:
-    number_sections: no
+  html_document:
+    df_print: paged
     toc: no

### 2018-04-07 R/trait_fxns.R
fixed "plant_eater" error.

@@ -87,9 +87,9 @@ read_process_trait_data <- function(){
       dplyr::select(-plants1, -plants2) %>%
       mutate(granivore = ifelse(granivore == 1, "Yes", "no")) %>%
       mutate(insectivore = ifelse(insectivore == 1, "Yes", "no")) %>%
-      mutate(plant_eater = ifelse(plant_eater == 2, "Yes", "no")) %>%
+      mutate(plant_eater = as.integer(as.character(gsub("2", "1", .$plant_eater)))) %>%
       mutate(plant_eater = ifelse(plant_eater == 1, "Yes", "no")) %>%
-      mutate(carrion_eater = ifelse(carrion_eater == 1, "Yes", "no")) 
+      mutate(carrion_eater = ifelse(carrion_eater == 1, "Yes", "no"))
     
 ## join diet
     diet <- inner_join(diet_types, diet_generalism, by="binom")

### 2018-04-07 reports/non_phylo_model.Rmd
changing the location in which we choose the traits to use, and switching that from the trait funxs script to the model report(s).

--- 2018-04-06 reports/non_phylo_model.Rmd
@@
 response_variables$binom<-response_variables$SCIENTIFIC_NAME_tree
-df<-inner_join(response_variables,traits)
+df<-left_join(traits,response_variables, by="binom") %>%
+  filter(SCIENTIFIC_NAME_tree != "not_treated_as_species") %>%
+  # drop columns that we aren't inerested in
+  select(-brain_size)
 ```

### 2018-04-07 R/trait_fxns.R

@@ -235,9 +235,6 @@ read_process_trait_data <- function(){
     ms <- data.frame(ms)
     
     
-    ms <- ms %>%
-      filter(complete.cases(.))
-    
    row.names(ms) <- ms$binom
    
   

### 2018-04-07 scripts/results_script.R
a script to assist in writing the results section of the paper and to create all final figures for the paper in one place

@@ -1,0 +1,102 @@
+### A script for summary of the results and to create final figures using the final set of species
+
+## first dump the remake environment
+remake::dump_environment()
+
+## get the final dataframe which was used for analysis
+## code copies from the non_phylo_model.Rmd
+response_variables$binom <- response_variables$SCIENTIFIC_NAME_tree
+
+df <- left_join(traits,response_variables, by="binom") %>%
+  filter(SCIENTIFIC_NAME_tree != "not_treated_as_species") %>%
+  # drop columns that we aren't inerested in
+  dplyr::select(-brain_size)
+
+df2 <- df %>%
+  filter_all(all_vars(!is.na(.)))
+
+## total number of species included in models
+length(unique(df2))
+
+## summary of urbanization index
+summary(df2$urban_median)
+
+sd(df2$urban_median)
+
+min(df2$urban_median)
+
+max(df2$urban_median)
+
+
+## Figure 1
+## A figure showing the distribution of 5 different species included in the analysis
+## Need to load the species_urban .RData file first
+load('Data/eBird data/Final eBird data for analysis/species_urban.RData')
+
+species_to_plot <- c("Common Myna", "Scaly-breasted Lorikeet", "Welcome Swallow", 
+                     "Banded Honeyeater", "White-winged Fairywren")
+
+Figure1 <- species_urban %>%
+  filter(COMMON_NAME %in% species_to_plot) %>%
+  ggplot(., aes(avg_rad, fill=COMMON_NAME)) +
+  geom_density(alpha = 0.7) + 
+  scale_x_log10() + 
+  theme_classic() +
+  xlab("Average radiance") +
+  ylab("Density")+
+  guides(fill=guide_legend(title="  Common Name"))
+
+pdf("final figures for paper/Figure1.pdf", width=5.5, height=4.5)
+print(Figure1)
+dev.off()
+
+
+## Figure 2
+## Figure 2 is the distribution of response variable for those species included in the analyses
+## Use the df2 dataframe from above
+## I want to add a line for each of the example species on the distribution
+
+species_to_plot <- c("Common Myna", "Scaly-breasted Lorikeet", "Welcome Swallow", 
+                     "Banded Honeyeater", "White-winged Fairywren")
+
+species_values <- df2 %>%
+  filter(COMMON_NAME_ebird %in% species_to_plot) %>%
[…]

### 2018-04-07 reports/non_phylo_model.Rmd
adding a weights vector to the model fitting. I think this is an important step and it seems to increase model fit (by just looking at the summary). But, happy to be told this is a crazy idea.

--- 2018-04-06 reports/non_phylo_model.Rmd
@@
 response_variables$binom<-response_variables$SCIENTIFIC_NAME_tree
-df<-inner_join(response_variables,traits)
+df<-left_join(traits,response_variables, by="binom") %>%
+  filter(SCIENTIFIC_NAME_tree != "not_treated_as_species") %>%
+  # drop columns that we aren't inerested in
+  dplyr::select(-brain_size)
 ```
@@
 
-naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+log(brain_size)+feeding_habitat_generalism+
-     breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+
-     movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail")
+naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+feeding_habitat_generalism+
+     breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+plant_eater+
+     movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail", weights=(df2$N/df2$unique_localities))
 summary(naive_mod)

### 2018-04-07 ms.Rmd
methods getting pretty close... and results started a little bit.

@@
 For each location, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
+Any radiance values which were negative were forced to 0.00001 (sensu CITATION!!!).
 
@@
 ## Models
+Before modelling, the explanatory variables (Appendix 1) were investigated for collinearity, and brain size and body size were highly correlated.
+Due to the affect that collinearity can have on modelling [@cade2015model], we were inclined to drop one of the two correlated explanatory variables.
+Given that others using a continuous response measure of urbanization have found brain size to have no evidence of species' response to urban environments [@evans2011makes], we eliminated this trait from consideration.
+In so doing, it also increased the number of species to be included in our models.
+
+We had two linear modelling approaches; one with and one without phylogenetic constraints.
+Our urbanization index was log-transformed in order to meet model assumptions.
+The explanatory variables used in the models were those listed in Appendix 1, with the exception of brain size.
+
+For both approaches, we employed a model-averaging technique [@grueber2011multimodel] in which we 'dredged' all possible subsets of the model [@barton2016mumin].
+The global model was weighted by a measure of error associated with the response variable.
+This was assessed as the proportion of the number of total observations of a species divided by the number of unique localities a species was observed from.
+This helps to give more weight in the model-fitting procedure to those species that had more observations in more unique locations (i.e., an assessment of reliability of the observations).
+The global model fit was assessed by averaging the R2 values [@nakagawa2013general; @nakagawa2017coefficient] for each model in the top-selected model-set.
+Additionally, we calculated the relative contribution of each explanatory variable [sensu @evans2011make] by **XXXXXXXXXXXXXXXX**.
+The phylogenetically constrained model was fit using the 'phylolm' function from the phylolm package [@ho2014phylolm].
+In order to assess the model fit and relative effect sizes of explanatory variables in the phylogenetically constrained models we **XXXXXXXXXXXX**.
+
+All models and analyses were conducted in the R computing environment [@r2017coreteam] and relied heavily on the tidyverse environment [@wickham2017tidyverse].
 
@@
 ## Urbanization index
+For the 450 species included in the models the mean +- SD urbanization index was 1.00 +- 1.81, with a range from 0.00135 (Black Noddy and Black-naped Tern) to 11.851 (Powerful Owl; Fig. 2).
 
 ### Phylogenetic signal
-There was a distinct phylogenetic signal in our urbanization index measure (Fig. 2; XXXXX SIGNIFICANCE TEST).
+There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; XXXXX SIGNIFICANCE TEST).
+
+**MAYBE DO SOME MEAN index by FAMILY PRESENTATIONS HERE**.
 

### 2018-04-07 ms.Rmd
flushing out intro

@@
 # INTRODUCTION
-Urbanization fragments and isolates natural habitat [@marzluff2001restoration; @melles2003urban; @loram2007urban], thus globally threatening biodiversity [@mckinney2006urbanization; @mckinney2008effects].
-Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization], some species have successfully adapted to, colonized and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @lowry2013behavioural; @mcdonnell2015adaptation].
+Urbanization fragments and isolates natural habitat [@marzluff2001restoration; @melles2003urban; @loram2007urban], drastically altering species' composition and ecosystem functions.
+Thus, urbanization globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects] and is a major threat to conservation [@czech2000economic; @dearborn2010motivations].
+Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @lowry2013behavioural; @mcdonnell2015adaptation].
+For instance, some species which rarely occur within urban areas can be relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
 
-Within urban ecosystems there are generally urban avoiders, urban adapters, and urban exploiters [@mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation].
+In order to best predict species' responses to future urbanization changes, it is essential to understand what biological traits are associated with urban birds [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
+To date, a wide array of traits have been used to assess presence of birds in urban environments.
+For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate (CITATION), diet (CITATIONS), distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@kark2007living; @evans2011make], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have all been hypothesized to affect a birds' presence in urban environments.
+However, a wide range of results have been reported in the literature, often contrasting one another.
+Brain size was found to be important by [@moller2015brain & @maklakov2011brains] but [@evans2011make & @kark2007living] found that brain size had no effect on species' response to urbanization.
+Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011make with @moller2009successful & @croci2008does] and generalism or niche breadth [cf., @kark2007living with @evans2011make].
+
+The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even birds which do or do not occur within urban environments [e.g., @bonier2007urban; @moller2009successful].
+This severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011make].
+Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
+In a notable exception, [@evans2011make] advocates for a continuous measure of urbanization in which a species' is assessed on their entire distribution in response to urban environments.
+
+Despite the number of studies which have investigated the relationships between species' traits and urban adaptedness, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two month survey period [@kark2007living]].
+Most studies also neglect any phylogenetic constraints in their study design [but see @bonier2007urban; @evans2011make], or their analyses are conducted within a single closely-related group of birds [@maklakov2011brains].
+We note that the generally small sample sizes (i.e., number of species) might not require the use such phylogenetically-controlled analyses.
+
+Fortunately, citizen science [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen] can provide broad-scale empirical datasets, allowing us to test hypotheses at spatial and temporal scales previously difficult to attain.
+This data provides a cost-effective method to research the relationship between species' traits and urban observations.
+Various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
+
 Urban avoiders are sensitive to urban environments [@mcdonnell2015adaptation] and tend to apportion more energy towards reproduction, relative to adapters and exploiters [@croci2008does].
@@
 While a regional analysis throughout France and Switzerland demonstrated urban adapters prefer forested habitats, are sedentary, omnivorous, high-nesters, and were widely distributed [@croci2008does].
-
-Despite the number of studies which have investigated the relationships between species' traits and urban adaptedness, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or two month survey period [@kark2007living]].
-Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
-
-Fortunately, citizen science [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen] can provide broad-scale empirical datasets, allowing us to test hypotheses at spatial and temporal scales previously difficult to attain.
-This data provides a cost-effective method to research the relationship between species' traits and urban observations.
-Various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
 

### 2018-04-07 ms.Rmd
intro getting pretty flushed out with a reasonable structure, I think.

@@
 To date, a wide array of traits have been used to assess presence of birds in urban environments.
-For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate (CITATION), diet (CITATIONS), distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@kark2007living; @evans2011make], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have all been hypothesized to affect a birds' presence in urban environments.
-However, a wide range of results have been reported in the literature, often contrasting one another.
-Brain size was found to be important by [@moller2015brain & @maklakov2011brains] but [@evans2011make & @kark2007living] found that brain size had no effect on species' response to urbanization.
+For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011make], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@kark2007living; @evans2011make], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have all been hypothesized to affect a birds' presence in urban environments.
+Rather than a specific trait that predicts a species’ ability to be found in urban environments, it is most likely a combination of traits.
+For instance, urban birds' presence in Jerusalem was predicted by a combination of traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-nesters, and widely distributed [@croci2008does].
+Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting one another.
+Indeed, brain size was found to be important by [@moller2015brain & @maklakov2011brains] but [@evans2011make & @kark2007living] found that brain size had no effect on species' response to urbanization.
 Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011make with @moller2009successful & @croci2008does] and generalism or niche breadth [cf., @kark2007living with @evans2011make].
@@
 The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even birds which do or do not occur within urban environments [e.g., @bonier2007urban; @moller2009successful].
-This severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011make].
+This type of discrete response variable severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011make].
 Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
@@
 Despite the number of studies which have investigated the relationships between species' traits and urban adaptedness, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two month survey period [@kark2007living]].
-Most studies also neglect any phylogenetic constraints in their study design [but see @bonier2007urban; @evans2011make], or their analyses are conducted within a single closely-related group of birds [@maklakov2011brains].
-We note that the generally small sample sizes (i.e., number of species) might not require the use such phylogenetically-controlled analyses.
-
 Fortunately, citizen science [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen] can provide broad-scale empirical datasets, allowing us to test hypotheses at spatial and temporal scales previously difficult to attain.
 This data provides a cost-effective method to research the relationship between species' traits and urban observations.
-Various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
+Indeed, various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
+However, we unaware of any projects which use citizen science observations to assess species' traits in relation to urban environments over a large spatial and temporal scale.
 
-Urban avoiders are sensitive to urban environments [@mcdonnell2015adaptation] and tend to apportion more energy towards reproduction, relative to adapters and exploiters [@croci2008does].
-Urban adapters can utilize urban environmental resources but may not reside in them [@mcdonnell2015adaptation] and are generally sedentary, social, high fecundity, with wide distribution [@kark2007living, @croci2008does].
-Urban exploiters thrive in urban environments [@mcdonnell2015adaptation] and are generally more social and sedentary than adapters [@kark2007living].
-Species’ traits such as social status, behavioral flexibility, migratory status, nesting, diet preferences, habitat preferences, and overall species range have all been used to predict presence in urban environments [@mcclure1989characterizes; @mills1989effects; @jokimaki1998distribution; @lim2004responses; @clergeau2006avifauna; @chace2006urban; @lowry2013behavioural].
-Understanding which favored traits are associated with urban adapters and avoiders [@croci2008does] and their habitats can inform management decisions for conservation of particular species and improvements of habitats within urban environments.
-
-Rather than a specific trait that predicts a species’ ability to be an urban avoider, adapter or exploiter, it is more likely a combination of traits.
-For instance, urban adapters’ presence in Jerusalem was predicted by a combination of traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living].
-While a regional analysis throughout France and Switzerland demonstrated urban adapters prefer forested habitats, are sedentary, omnivorous, high-nesters, and were widely distributed [@croci2008does].
-
-Here, we use data from eBird [@sullivan2009ebird; -@sullivan2014ebird], a successful citizen science project with over 500 million observations, submitted by more than 250,000 participants [@sullivan2017using], in order to elucidate adaptations in Australian birds to urban living.
-We combine eBird data with multiple input rasters, that represent a continuous levels of urbanness, over the whole of Australia.
-This is opposed to assigning species to a priori groups (adapters, avoiders, or exploiters) and investigating differences in traits among groups [e.g., @kark2007living].
-We then apply phylogenetically controlled models in order to assess species traits that predict pre-adaptation to urban living.
-We hypothesize that a combination of traits allows species to thrive in urban environments.
+Here, we improve upon current methods of assessing biological traits which predicts urbanization in birds by using a series of publicly available datasets available for Australian birds.
+Importantly, we developed and used a continuous urbanization index as a response variable in the analyses.
+We had two objectives.
+Our first objective was to investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011make].
+Second, we developed models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using continuous and categorical explanatory variables.
+These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth.
 
@@
 
-We filtered the bird observations (sensu [@callaghan2017assessing; @la2014role]) by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barrett2003new].
+We filtered the bird observations [@callaghan2017assessing; @la2014role] by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barrett2003new].
 For further information on protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types].
@@
 For each location, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
-Any radiance values which were negative were forced to 0.00001 (sensu CITATION!!!).
+Any radiance values which were negative were forced to 0.00001 [*sensu* @ou2015evaluation].
 
@@
 The global model fit was assessed by averaging the R2 values [@nakagawa2013general; @nakagawa2017coefficient] for each model in the top-selected model-set.
-Additionally, we calculated the relative contribution of each explanatory variable [sensu @evans2011make] by **XXXXXXXXXXXXXXXX**.
+Additionally, we calculated the relative contribution of each explanatory variable [*sensu* @evans2011make] by **XXXXXXXXXXXXXXXX**.
 The phylogenetically constrained model was fit using the 'phylolm' function from the phylolm package [@ho2014phylolm].

### 2018-04-07 scripts/results_script.R
move legend for better aesthetics

@@ -44,7 +44,8 @@ Figure1 <- species_urban %>%
   theme_classic() +
   xlab("Average radiance") +
   ylab("Density")+
-  guides(fill=guide_legend(title="  Common Name"))
+  guides(fill=guide_legend(title="  Common Name")) +
+  theme(legend.position = c(0.2, 0.8))
 
 pdf("final figures for paper/Figure1.pdf", width=5.5, height=4.5)
 print(Figure1)

### 2018-04-07 ms.Rmd
intro draft and figures now plotting in knitted ms.Rmd

@@
 ---
-title: "Phylogenetic adaptions to urban living in Australian birds"
+title: "A continuous urbanization index to assess phylogenetic adaptations in urban bird living"
 author: Corey T. Callaghan, John Wilshire, XXXX, XXXX, Richard T. Kingsford, William K. Cornwell
@@
 output:
+  pdf_document: null
+  bookdown::word_document2: null
   bookdown::tufte_html2:
@@
     toc: no
-  bookdown::word_document2: null
-  pdf_document: null
   html_document:
@@
 Thus, urbanization globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects] and is a major threat to conservation [@czech2000economic; @dearborn2010motivations].
-Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @lowry2013behavioural; @mcdonnell2015adaptation].
+Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @lowry2013behavioural; @mcdonnell2015adaptation].
 For instance, some species which rarely occur within urban areas can be relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
@@
 In order to best predict species' responses to future urbanization changes, it is essential to understand what biological traits are associated with urban birds [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
-To date, a wide array of traits have been used to assess presence of birds in urban environments.
-For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011make], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@kark2007living; @evans2011make], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have all been hypothesized to affect a birds' presence in urban environments.
-Rather than a specific trait that predicts a species’ ability to be found in urban environments, it is most likely a combination of traits.
+To date, a wide array of traits have been used to predict presence of birds in urban environments.
+For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have all been hypothesized to influence a birds' presence in urban environments.
+Rather than a specific trait that predicts a species' ability to be found in urban environments, it is most likely a combination of traits.
 For instance, urban birds' presence in Jerusalem was predicted by a combination of traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-nesters, and widely distributed [@croci2008does].
 Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting one another.
-Indeed, brain size was found to be important by [@moller2015brain & @maklakov2011brains] but [@evans2011make & @kark2007living] found that brain size had no effect on species' response to urbanization.
-Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011make with @moller2009successful & @croci2008does] and generalism or niche breadth [cf., @kark2007living with @evans2011make].
+Indeed, brain size was found to be important by Møller and Erritzøe [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains] but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on species' response to urbanization.
+Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011makes with @moller2009successful & @croci2008does] and generalism or niche breadth [cf., @kark2007living with @evans2011makes].
 
 The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even birds which do or do not occur within urban environments [e.g., @bonier2007urban; @moller2009successful].
-This type of discrete response variable severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011make].
+This type of discrete response variable severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011makes].
 Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
-In a notable exception, [@evans2011make] advocates for a continuous measure of urbanization in which a species' is assessed on their entire distribution in response to urban environments.
+In a notable exception, [@evans2011makes] advocates for a continuous measure of urbanization in which a species' is assessed on their entire distribution in response to urban environments.
 
@@
 We had two objectives.
-Our first objective was to investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011make].
+Our first objective was to investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
 Second, we developed models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using continuous and categorical explanatory variables.
@@
 The global model fit was assessed by averaging the R2 values [@nakagawa2013general; @nakagawa2017coefficient] for each model in the top-selected model-set.
-Additionally, we calculated the relative contribution of each explanatory variable [*sensu* @evans2011make] by **XXXXXXXXXXXXXXXX**.
+Additionally, we calculated the relative contribution of each explanatory variable [*sensu* @evans2011makes] by **XXXXXXXXXXXXXXXX**.
 The phylogenetically constrained model was fit using the 'phylolm' function from the phylolm package [@ho2014phylolm].
@@
 The least recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
-The mean +- SD number of observations for a species was 10250 +- 19269.19.
+The mean ± SD number of observations for a species was 10,250 ± 19,269.19.
 
-We ended with 450 candidate species (Appendix 3) from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data.
+We ended with 569 candidate species (Appendix 3) from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data.
 
 ## Urbanization index
-For the 450 species included in the models the mean +- SD urbanization index was 1.00 +- 1.81, with a range from 0.00135 (Black Noddy and Black-naped Tern) to 11.851 (Powerful Owl; Fig. 2).
+For the 569 species included in the models the mean ± SD urbanization index was 1.11 ± 2.17, with a range from 0.001 (Black Noddy and Black-naped Tern) to 20.15 (Laughing Dove; Fig. 2).
 
@@
 
+<div id="refs"></div>
+
 # Figures
 
-```{r Figure 2, echo=FALSE, fig.cap="Figure 2. A phylogenetic tree of the birds yata yata yata...."}
-knitr::include_graphics(path = "figures/bird_urbanness_phylo.pdf")
+```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance values"}
+knitr::include_graphics("final figures for paper/Figure1.pdf")
 ```
+
+```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a log-scale, of the response variable for the 569 species included in the analysis, with five example species highlighted."}
+knitr::include_graphics("final figures for paper/Figure2.pdf")
+```
+
+```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree of the birds yata yata yata...."}
+knitr::include_graphics("final figures for paper/Figure3.pdf")
+```

### 2018-04-08 R/trait_fxns.R
added possibly specific habitat type predictors into the candidate traits df

@@ -62,6 +62,45 @@ read_process_trait_data <- function(){
       summarise_all(sum) %>%
       mutate(breeding_habitat_generalism = rowSums(.[2:17])) %>%
       dplyr::select(binom, breeding_habitat_generalism)
+    
+    
+## df for specific habitat types (trees/forests and shrubs/grasslands and agricultural)
+    tree_forest <- traits %>%
+      dplyr::select(binom, 115:144, 146:161) %>%
+      replace(is.na(.), 0) %>%
+      group_by(binom) %>%
+      summarise_all(sum) %>%
+      dplyr::select(binom, 8:12, 38:42) %>%
+      mutate(tree_forest = rowSums(.[2:11])) %>%
+      dplyr::select(binom, tree_forest) %>%
+      mutate(tree_forest = ifelse(.$tree_forest > 1, 1, .$tree_forest))
+    
+    
+    grass_shrubland <- traits %>%
+      dplyr::select(binom, 115:144, 146:161) %>%
+      replace(is.na(.), 0) %>%
+      group_by(binom) %>%
+      summarise_all(sum) %>%
+      dplyr::select(binom, 1:2, 4:5, 32:33, 35:36) %>%
+      mutate(grass_shrubland = rowSums(.[2:8])) %>%
+      dplyr::select(binom, grass_shrubland) %>%
+      mutate(grass_shrubland = ifelse(.$grass_shrubland > 1, 1, .$grass_shrubland))
+    
+    
+    agricultural <- traits %>%
+      dplyr::select(binom, 115:144, 146:161) %>%
+      replace(is.na(.), 0) %>%
+      group_by(binom) %>%
+      summarise_all(sum) %>%
+      dplyr::select(binom, 31, 37) %>%
+      mutate(agricultural = rowSums(.[2:3])) %>%
+      dplyr::select(binom, agricultural) %>%
+      mutate(agricultural = ifelse(.$agricultural > 1, 1, .$agricultural))
+  
+    habitat_types <- agricultural %>%
+      inner_join(., grass_shrubland, by="binom") %>%
+      inner_join(., tree_forest, by="binom")
+      
 
 ## Diet generalism
     diet_generalism <- traits %>%
@@ -229,7 +268,8 @@ read_process_trait_data <- function(){
       inner_join(., diet, by="binom") %>%
       inner_join(., movement, by="binom") %>%
       inner_join(., nesting, by="binom") %>%
-      inner_join(., gregariousness, by="binom")
[…]

### 2018-04-08 reports/non_phylo_model.Rmd
added the habitat predictors

@@
 
-naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+feeding_habitat_generalism+
-     breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+plant_eater+
-     movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail", weights=(df2$N/df2$unique_localities))
+naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+feeding_habitat_generalism+agricultural+grass_shrubland+tree_forest+breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+plant_eater+movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail", weights=(df2$N/df2$unique_localities))
 summary(naive_mod)

### 2018-04-08 reports/non_phylo_model.Rmd
changing habitat traits to yes/no to match rest of traits

--- 2018-04-07 reports/non_phylo_model.Rmd
@@
 df<-left_join(traits,response_variables, by="binom") %>%
-  filter(SCIENTIFIC_NAME_tree != "not_treated_as_species") %>%
-  # drop columns that we aren't inerested in
-  select(-brain_size)
+  filter(SCIENTIFIC_NAME_tree != "not_treated_as_species")
 ```
@@
 names(df)
-df %>%
-filter_all(all_vars(!is.na(.)))->df2
+df2 <- df %>%
+  # drop columns that we aren't inerested in
+  dplyr::select(-brain_size) %>%
+filter_all(all_vars(!is.na(.)))
 
-naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+log(brain_size)+feeding_habitat_generalism+
-     breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+
-     movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail")
+naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+feeding_habitat_generalism+Habitat_agricultural+Habitat_grass_shrubland+Habitat_tree_forest+breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+plant_eater+movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail", weights=(df2$N/df2$unique_localities))
 summary(naive_mod)

### 2018-04-08 R/trait_fxns.R

@@ -73,7 +73,9 @@ read_process_trait_data <- function(){
       dplyr::select(binom, 8:12, 38:42) %>%
       mutate(tree_forest = rowSums(.[2:11])) %>%
       dplyr::select(binom, tree_forest) %>%
-      mutate(tree_forest = ifelse(.$tree_forest > 1, 1, .$tree_forest))
+      mutate(tree_forest = ifelse(.$tree_forest > 1, 1, .$tree_forest)) %>%
+      mutate(Habitat_tree_forest = ifelse(.$tree_forest == 1, "Yes", "no")) %>%
+      dplyr::select(-tree_forest)
     
     
     grass_shrubland <- traits %>%
@@ -84,7 +86,9 @@ read_process_trait_data <- function(){
       dplyr::select(binom, 1:2, 4:5, 32:33, 35:36) %>%
       mutate(grass_shrubland = rowSums(.[2:8])) %>%
       dplyr::select(binom, grass_shrubland) %>%
-      mutate(grass_shrubland = ifelse(.$grass_shrubland > 1, 1, .$grass_shrubland))
+      mutate(grass_shrubland = ifelse(.$grass_shrubland > 1, 1, .$grass_shrubland)) %>%
+      mutate(Habitat_grass_shrubland = ifelse(.$grass_shrubland == 1, "Yes", "no")) %>%
+      dplyr::select(-grass_shrubland)
     
     
     agricultural <- traits %>%
@@ -95,7 +99,9 @@ read_process_trait_data <- function(){
       dplyr::select(binom, 31, 37) %>%
       mutate(agricultural = rowSums(.[2:3])) %>%
       dplyr::select(binom, agricultural) %>%
-      mutate(agricultural = ifelse(.$agricultural > 1, 1, .$agricultural))
+      mutate(agricultural = ifelse(.$agricultural > 1, 1, .$agricultural)) %>%
+      mutate(Habitat_agricultural = ifelse(.$agricultural == 1, "Yes", "no")) %>%
+      dplyr::select(-agricultural)
   
     habitat_types <- agricultural %>%
       inner_join(., grass_shrubland, by="binom") %>%

### 2018-04-09 reports/non_phylo_model.Rmd
looking at VIF and also model selection via best subset selection

@@
 
-completely naive model, plant eater taken out cause it only had one level.
-
 Logging response cause it is right-skewed and produced crazy residual plots.
@@
 
-naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+feeding_habitat_generalism+Habitat_agricultural+Habitat_grass_shrubland+Habitat_tree_forest+breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+plant_eater+movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail", weights=(df2$N/df2$unique_localities))
+naive_mod<-lm(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+feeding_habitat_generalism+brain_residual+Habitat_agricultural+Habitat_grass_shrubland+Habitat_tree_forest+breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+plant_eater+movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail", weights=(df2$N/df2$unique_localities))
 summary(naive_mod)
@@
 ```
+
+checking for multicolinearity
+```{r}
+library(car)
+vif()
+```
+
 and some not-so-quick model selection.  dredge takes a little while because it considers all combinations of predictor variables.
@@
 ```
+
+another type of potential model selection
+
+```{r}
+library(leaps)
+
+model_all_select <- regsubsets(log(urban_median)~log(mean_body_size)+log(clutch_size)+iucn_status+feeding_habitat_generalism+brain_residual+Habitat_agricultural+Habitat_grass_shrubland+Habitat_tree_forest+breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+plant_eater+movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation,data=df2,na.action = "na.fail", weights=(df2$N/df2$unique_localities), method = "exhaustive")
+
+results <- summary(model_all_select)
+
+which.max(results$adjr2)
+
+which.min(results$bic)
+
+which.min(results$cp)
+
+```

### 2018-04-09 R/trait_fxns.R
adding brain residual (which is what I should have been looking at all along, as opposed to brain size), which isn't correlated with body size, but still a measure of behavioural flexibility

@@ -37,12 +37,14 @@ read_process_trait_data <- function(){
     mutate(clutch_size=ifelse(`178_Clutch_size_average_12`=="NAV",NA,as.numeric(`178_Clutch_size_average_12`))) %>%
     mutate(iucn_status=`49_Australian_status_July_2015_5`) %>%
     mutate(brain_size=ifelse(`113_Brain_mass_8`=="NAV", NA, as.numeric(`113_Brain_mass_8`))) %>%
+    mutate(brain_residual=ifelse(`114_Brain_mass_residual_8`=="NAV", NA, as.numeric(`114_Brain_mass_residual_8`))) %>%
     mutate(diet_generalism=rowSums(.[163:173], na.rm=TRUE)) %>%
     group_by(binom) %>%
     summarise(mean_body_size=mean(body_mass,na.rm=T), 
               clutch_size=mean(clutch_size,na.rm=T),
               iucn_status=Mode(iucn_status),
-              brain_size=mean(brain_size, na.rm=T)) -> ms
+              brain_size=mean(brain_size, na.rm=T),
+              brain_residual=mean(brain_residual, na.rm=T)) -> ms
               
     
 ## calculate df for feeding habitat generalism

### 2018-04-09 reports/non_phylo_model.Rmd
adding some dredging and model-averaging potential code

@@
 ```
+
+a model averaging approach
+```{r}
+#### Automated Approach ####
+#### Computes all possible subsets of global model
+df3 <- df2 %>%
+  mutate(response = log(urban_median)) %>%
+  mutate(body_size_logged = log(mean_body_size)) %>%
+  mutate(clutch_size_logged = log(clutch_size))
+
+glob.mod <- lm(response~body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + Habitat_agricultural,
+               data=df3, na.action = "na.fail", weights=(df3$N/df3$unique_localities))
+
+library(arm)
+stdz.model <- standardize(glob.mod)
+
+model.set <- dredge(stdz.model)
+
+model.set
+
+#### selects all models with deltaAic < 4
+top.models <- get.models(model.set, cumsum(weight) <= 0.95)
+
+#### how many top models
+length(top.models)
+
+#### Ranks these models based on AICc
+my.models <- model.sel(top.models, rank="AICc")
+
+ #### Prints the top models (selected above)
+print.data.frame(my.models, digits=2)
+
+Averaged_models <- model.avg(top.models)
+
+summary(Averaged_models)
+confint(Averaged_models, full=TRUE)
+```
+
+Now repeat on the full model (will take a while so let it run overnight)
+```{r}
+df3 <- df2 %>%
+  mutate(response = log(urban_median)) %>%
+  mutate(body_size_logged = log(mean_body_size)) %>%
+  mutate(clutch_size_logged = log(clutch_size))
+
+glob.mod <- lm(response ~ body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + Habitat_agricultural +
+                 breeding_habitat_generalism + granivore + insectivore + carrion_eater + plant_eater + diet_generalism + movement_class +
+                 groun_nesting + hollow_nesting + nest_generalism + breeding + nest_aggregation + feeding_aggregation + Habitat_grass_shrubland +
+                 Habitat_tree_forest, data=df3, na.action = "na.fail", weights=(df3$N/df3$unique_localities))
+
+library(arm)
+stdz.model <- standardize(glob.mod)
+
+model.set <- dredge(stdz.model, extra="R^2")
+
+model.set
+
+#### selects all models with deltaAic < 4
+top.models <- get.models(model.set, cumsum(weight) <= 0.95)
+
+#### how many top models
+length(top.models)
+
+#### Ranks these models based on AICc
+my.models <- model.sel(top.models, rank="AICc")
+
+ #### Prints the top models (selected above)
+print.data.frame(my.models, digits=2)
+
+Averaged_models <- model.avg(top.models)
+
+summary(Averaged_models)
+confint(Averaged_models, full=TRUE)
+
+```

### 2018-04-11 reports/non_phylo_model.Rmd
some modelling work

--- 2018-04-09 reports/non_phylo_model.Rmd
@@
 library(car)
-vif()
+vif(naive_mod)
 ```
@@
 ```
+
+```{r}
+df3 <- df2 %>%
+  mutate(response = log(urban_median)) %>%
+  mutate(body_size_logged = log(mean_body_size)) %>%
+  mutate(clutch_size_logged = log(clutch_size))
+```
+
+a model averaging approach
+small model first
+
+```{r}
+glob.mod <- lm(response~body_size_logged+clutch_size_logged+feeding_habitat_generalism+brain_residual+Habitat_agricultural,data=df3,na.action="na.fail",weights=(df3$N/df3$unique_localities))
+
+library(arm)
+
+stdz.model <- standardize(glob.mod)
+model.set <- dredge(stdz.model)
+
+model.set
+
+#### selects all models with deltaAic < 4
+top.models <- get.models(model.set, cumsum(weight) <= 0.95)
+
+#### how many top models
+length(top.models)
+
+#### Ranks these models based on AICc
+my.models <- model.sel(top.models, rank="AICc")
+
+ #### Prints the top models (selected above)
+print.data.frame(my.models, digits=2)
+
+Averaged_models <- model.avg(top.models)
+
+summary(Averaged_models)
+confint(Averaged_models, full=TRUE)
+```
+
+run a parallel version
+```{r}
+library(snow)
+clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
+clust <- try(makeCluster(getOption("cl.cores", 10), type = clusterType))
+
+clusterExport(clust, "df3")
+
+glob.mod <- lm(response~body_size_logged+clutch_size_logged+feeding_habitat_generalism+brain_residual+Habitat_agricultural,data=df3,na.action="na.fail",weights=(df3$N/df3$unique_localities))
+
+library(arm)
+
+stdz.model <- standardize(glob.mod)
+
+model.set <- pdredge(stdz.model, cluster=clust)
+
+model.set
+
+#### selects all models with deltaAic < 4
+top.models <- get.models(model.set, cumsum(weight) <= 0.95)
+
+#### how many top models
+length(top.models)
+
+#### Ranks these models based on AICc
+my.models <- model.sel(top.models, rank="AICc")
+
+#### Prints the top models (selected above)
+print.data.frame(my.models, digits=2)
+
+Averaged_models <- model.avg(top.models)
+
+summary(Averaged_models)
+confint(Averaged_models, full=TRUE)
+
+```
+
+Now repeat on the full model (will take a while so let it run overnight)
+```{r}
+library(snow)
+clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
+clust <- try(makeCluster(getOption("cl.cores", 10), type = clusterType))
+
+clusterExport(clust, "df3")
+
+glob.mod <- lm(response ~ body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + Habitat_agricultural +
+                 breeding_habitat_generalism + granivore + insectivore + carrion_eater + plant_eater + diet_generalism + movement_class +
+                 ground_nesting + hollow_nesting + nest_generalism + breeding + nest_aggregation + feeding_aggregation + Habitat_grass_shrubland +
+                 Habitat_tree_forest, data=df3, na.action = "na.fail", weights=(df3$N/df3$unique_localities))
+
+library(arm)
+
+stdz.model <- standardize(glob.mod)
+
+model.set <- pdredge(stdz.model, m.lim=c(0, 11), cluster=clust, extra="R^2")
+
+model.set
+
+#### selects all models with deltaAic < 4
+top.models <- get.models(model.set, subset=delta<4)
+
+#### how many top models
+length(top.models)
+
+#### Ranks these models based on AICc
+my.models <- model.sel(top.models, rank="AICc")
+
+ #### Prints the top models (selected above)
+print.data.frame(my.models, digits=2)
+
+Averaged_models <- model.avg(top.models)
+
+summary(Averaged_models)
+confint(Averaged_models, full=TRUE)
+
+```

### 2018-04-18 R/non_phylo_analysis.R
working on the non_phylo_analyses and putting them in the workflow

@@ -1,0 +1,166 @@
+## Script with functions for the results of the
+## Non-phylo models
+
+
+## create df for the model fitting
+matched_df <- function(response_variables, traits){
+  
+  response_variables$binom <- response_variables$SCIENTIFIC_NAME_tree
+  
+  analysis_data <- left_join(traits, response_variables, by="binom") %>%
+    filter(SCIENTIFIC_NAME_tree != "not_treated_as_species")
+  
+  # clean data to final dataset
+  analysis_data <- analysis_data %>%
+    # drop any columns that we aren't interested in
+    dplyr::select(-brain_size, -iucn_status) %>%
+    # filter for only complete variables
+    filter_all(all_vars(!is.na(.))) %>%
+    # rename columns so that they are logged
+    # necessary for the model selection step
+    mutate(response = log(urban_median)) %>%
+    mutate(body_size_logged = log(mean_body_size)) %>%
+    mutate(clutch_size_logged = log(clutch_size))
+  
+  analysis_data
+}
+
+# make coorplot of continuous variables
+corrplot_figure <- function(analysis_data){
+  
+  library(corrplot)
+  
+ continuous_df <- select_if(analysis_data, is.numeric) %>%
+    dplyr::select(-urban_mean, -clutch_size, -mean_body_size, -urban_median, -response) %>%
+    rename(`Brain residual` = brain_residual) %>%
+    rename(`Feeding habitat generalism` = feeding_habitat_generalism) %>%
+    rename(`Breeding habitat generalism` = breeding_habitat_generalism) %>%
+    rename(`Diet generalism` = diet_generalism) %>%
+    rename(`Nest generalism` = nest_generalism) %>%
+    rename(`Number of records` = N) %>%
+    rename(`Unique localitieis observed` = unique_localities) %>%
+    rename(`log(Body size)` = body_size_logged) %>%
+    rename(`log(Clutch size)` = clutch_size_logged) %>%
+    cor(use="pairwise.complete.obs")
+  
+  pdf("figures/corrplot_of_continuous_variables.pdf")
+  
+  print(
+  corrplot::corrplot(continuous_df)
+  )
+  
+  dev.off()
+}
+
+## run one large model and standardize the model
+## so that effect sizes are comparable
+
+global_model <- function(analysis_data) {
+  
+  glob.mod <- lm(response ~ body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + 
+                   Habitat_agricultural + breeding_habitat_generalism + granivore + insectivore + 
[…]

### 2018-04-18 R/non_phylo_analysis.R
trying to get dredged model to work

@@ -160,6 +160,54 @@ plot_params_globmod <- function(global_model) {
 }
 
 
+## do a model averaging approach using dredging
+model_averaging <- function(analysis_data, global_model) {
+  
+  library(snow)
+  clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
+  clust <- try(makeCluster(getOption("cl.cores", 10), type = clusterType))
+  
+  clusterExport(clust, "analysis_data")
+  
+  library(arm)
+  
+  stdz.model <- standardize(glob.mod)
+  
+  model.set <- pdredge(stdz.model, m.lim=c(0, 2), cluster=clust, extra="R^2") 
+  
+  return(model.set)
+  
+}
+
+
+model_averaging_results <- function(model_averaging) {
+  
+  #### selects all models with deltaAic < 4
+  top.models <- get.models(model.set, subset=delta<4) 
+  
+  #### how many top models
+  length(top.models)
+  
+  #### Ranks these models based on AICc
+  my.models <- model.sel(top.models, rank="AICc") 
+  
+  Averaged_models <- model.avg(top.models)
+  
+  summary(Averaged_models)
+  confidence_intervals <- confint(Averaged_models, full=TRUE)
+  
+  library(gridExtra)
+  
+  pdf("tables/Averaged_models.pdf")
+  grid.table(confidence_intervals)
+  dev.off()
+}
+
+
+
+
+
+
 
 
 

### 2018-04-19 R/non_phylo_analysis.R
dredging the model OUTSIDE of the remake workflow

@@ -98,7 +98,6 @@ collinearity_investigation_function <- function(global_model) {
 
 
 ## plot parameter estimates of standardized global model
-
 plot_params_globmod <- function(global_model) {
   
   library(sjPlot)
@@ -161,26 +160,14 @@ plot_params_globmod <- function(global_model) {
 
 
 ## do a model averaging approach using dredging
-model_averaging <- function(analysis_data, global_model) {
+get_dredged_model <- function() {
   
-  library(snow)
-  clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
-  clust <- try(makeCluster(getOption("cl.cores", 10), type = clusterType))
-  
-  clusterExport(clust, "analysis_data")
-  
-  library(arm)
-  
-  stdz.model <- standardize(glob.mod)
-  
-  model.set <- pdredge(stdz.model, m.lim=c(0, 2), cluster=clust, extra="R^2") 
-  
-  return(model.set)
+  load("Data/dredged_model.rds")
   
 }
 
 
-model_averaging_results <- function(model_averaging) {
+model_averaging_results <- function(dredged_model) {
   
   #### selects all models with deltaAic < 4
   top.models <- get.models(model.set, subset=delta<4) 

### 2018-04-19 scripts/build_dredged_model_for_non_phylo_analysis.R

@@ -1,0 +1,25 @@
+## This is an R script to dredge the model for the non
+## phylo models and then save it as an object
+## to save headaches with remake workflow
+## build the dredged model and export it
+## done outside of the remake workflow as 
+## the environment issue was very difficult to navigate!
+
+build_dredged_model <- function() {
+  remake::dump_environment()
+  
+  library(snow)
+  clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
+  clust <- try(makeCluster(getOption("cl.cores", 10), type = clusterType))
+  
+  clusterExport(clust, "analysis_data", envir = environment())
+  
+  stdz.model <- standardize(global_model)
+  
+  model.set <- pdredge(global_model, m.lim=c(0, 11), cluster=clust, extra="R^2") 
+  
+  saveRDS(model.set, file="Data/dredged_model.rds")
+  
+}
+
+build_dredged_model()

### 2018-04-19 reports/phylo_report.rmd
comparing phylo and non-phylo

--- 2018-03-30 reports/phylo_report.rmd
@@
 
-run phylo model
+non phylo model
 
 ```{r}
+setwd("..") #take this out later
+remake::dump_environment() #take this out later
+glob.mod <- lm(response ~ body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + 
+                 Habitat_agricultural + breeding_habitat_generalism + granivore + insectivore + 
+                 carrion_eater + plant_eater + diet_generalism + movement_class +
+                 ground_nesting + hollow_nesting + nest_generalism + breeding + 
+                 nest_aggregation + feeding_aggregation + Habitat_grass_shrubland +
+                 Habitat_tree_forest, data=analysis_data, 
+               na.action = "na.fail", weights=(analysis_data$N/analysis_data$unique_localities))
 
-mod2<-run_phylo_lm(traits,
-                         response_variables,
-                         aus_bird_tree)
+plot(glob.mod)
+```
 
-summary(mod2)
-```
-maybe we eventually want to do model selection
+run phylo model, exactly the same as the non-phylo model at this point
 
 ```{r}
-# library(MuMIn)
-# dredge(mod2)
+non_aus_sp <- aus_bird_tree$tip.label[!aus_bird_tree$tip.label %in% analysis_data$binom]
+aus_bird_tree_ss <- drop.tip(aus_bird_tree, non_aus_sp)
+
+row.names(analysis_data)<-analysis_data$binom
+
+phy_mod<-phylolm(response~body_size_logged + clutch_size_logged + feeding_habitat_generalism +   
+                   brain_residual + Habitat_agricultural + breeding_habitat_generalism + 
+                   granivore + insectivore + 
+                   carrion_eater + plant_eater + diet_generalism + movement_class +
+                   ground_nesting + hollow_nesting + nest_generalism + breeding + 
+                   nest_aggregation + feeding_aggregation + Habitat_grass_shrubland +
+                   Habitat_tree_forest,data=analysis_data,phy=aus_bird_tree_ss,na.action = "na.fail", weights=(analysis_data$N/analysis_data$unique_localities))
+
+plot(phy_mod)
 ```
 
-this is sort of cool.  phylogenetic signal in all the traits
+looks ok from a first glance. now let's look at the model summaries.
 
 ```{r}
-  # library(phylosignal)
-  # library(phylobase)
-  # dd$body_size<-log10(dd$body_size)
-  # dd2<-subset(dd,select=-gregariousness)
-  # out<-phylo4d(tree_plotting_2,tip.data=dd2)
-  # ps<-phyloSignal(out)
-  # knitr::kable(ps$stat)
+summary(glob.mod)
+
 ```
+
+and then the phylo model (this is most often called pgls in the literature)
+
+```{r}
+summary(phy_mod)
+```
+
+look a little bit more at the coefficients
+
+```{r}
+library(ggplot2)
+cc<-data.frame(phy_mod_coefs=coef(phy_mod),non_phy_mod=coef(glob.mod))
+cc$parameter<-substr(row.names(cc), start = 1, stop = 7)
+
+ggplot(cc,aes(x=non_phy_mod,y=phy_mod_coefs))+geom_point()+geom_text(aes(label=parameter),hjust=0, vjust=0)+theme_bw()+geom_abline(slope=1,intercept=0)
+
+```
+
+Wwith the phylogeny in the model, brain residuals association with urbanness becomes stronger.
+The implication is that brain residuals varies across the phylogeny and some clades are big and some are small which means the non-phylo relationship is weak.
+But **within each clade** it's the species with the large brain residuals that are doing fine in urban areas.
+
+IDK, might be something real there????
+
+What do you think?

### 2018-04-19 R/non_phylo_analysis.R
working on dredged portion

@@ -162,7 +162,7 @@ plot_params_globmod <- function(global_model) {
 ## do a model averaging approach using dredging
 get_dredged_model <- function() {
   
-  load("Data/dredged_model.rds")
+  readRDS("Data/dredged_model.rds")
   
 }
 

### 2018-04-19 scripts/build_dredged_model_for_non_phylo_analysis.R
chunking dredged model and exporting rds files

@@ -18,8 +18,16 @@ build_dredged_model <- function() {
   
   model.set <- pdredge(global_model, m.lim=c(0, 11), cluster=clust, extra="R^2") 
   
-  saveRDS(model.set, file="Data/dredged_model.rds")
+  split <- split(model.set, (seq(nrow(model.set))-1) %/% 200000)
   
+  list2env(setNames(split,paste0("df",1:4)),environment())
+  
+  rm(split)
+  
+  saveRDS(df1, file="Data/dredged_models_1.rds")
+  saveRDS(df2, file="Data/dredged_models_2.rds")
+  saveRDS(df3, file="Data/dredged_models_3.rds")
+  saveRDS(df4, file="Data/dredged_models_4.rds")
 }
 
 build_dredged_model()

### 2018-04-19 R/non_phylo_analysis.R
adding the dredged model results into the workflow

@@ -159,11 +159,21 @@ plot_params_globmod <- function(global_model) {
 }
 
 
-## do a model averaging approach using dredging
+## this pulls in the dredged model results which was
+## done out of the workflow
+## the data was chunked in order to get around github file size limits
+## will likely have to alter this once we have a finalized
+## set of predictor variables
 get_dredged_model <- function() {
   
-  readRDS("Data/dredged_model.rds")
+  df1 <- readRDS("Data/dredged_models_1.rds")
+  df2 <- readRDS("Data/dredged_models_2.rds")
+  df3 <- readRDS("Data/dredged_models_3.rds")
+  df4 <- readRDS("Data/dredged_models_4.rds")
   
+  model.set <- bind_rows(df1, df2, df3, df4)
+  
+  return(model.set)
 }
 
 

### 2018-04-19 R/non_phylo_analysis.R
fixing plot

@@ -107,7 +107,7 @@ plot_params_globmod <- function(global_model) {
   
   print(
   get_model_data(global_model, type="std2") %>%
-    arrange(estimate) %>%
+    arrange(desc(estimate)) %>%
     mutate(term2 = c("log(Clutch size)",
                      "Feeding habitat generalism", 
                      "Migratory type \n (partial migrant)",
@@ -143,6 +143,7 @@ plot_params_globmod <- function(global_model) {
                      "Insectivore",
                      "Feeding aggregation \n (solitary)",
                      "Migratory type \n (partial migrant & nomadic/irruptive)")) %>%
+    arrange(estimate) %>%
     mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
     ggplot(., aes(x=fct_inorder(term2), y=estimate, color=trend))+
     geom_point()+

### 2018-04-19 R/non_phylo_analysis.R
adding code for param plots of averaged models (but there is some remake error!)

@@ -194,6 +194,62 @@ model_averaging_results <- function(dredged_model) {
   summary(Averaged_models)
   confidence_intervals <- confint(Averaged_models, full=TRUE)
   
+  a <- as.data.frame(coefficients(Averaged_models, full=TRUE))
+  b <- as.data.frame(confint(Averaged_models, full=TRUE))
+  model_results <- cbind(a, b)
+  names(model_results)[1] <- "estimate"
+  names(model_results)[2] <- "lwr"
+  names(model_results)[3] <- "upr"
+  model_results$variable <- row.names(model_results)
+  row.names(model_results) <- NULL
+  
+  pdf("figures/param_plot_averaged_results.pdf")
+  print(
+  model_results %>%
+    filter(variable != "(Intercept)") %>%
+    droplevels() %>%
+    arrange(desc(estimate)) %>%
+    mutate(variable2 = c("log(Clutch size)",
+                         "Feeding habitat generalism",
+                         "Diet generalism",
+                         "Habitat - agricultural",
+                         "Breeding habitat generalism",
+                         "Feeding aggregation \n (solitary & flocks)",
+                         "log(Body size)",
+                         "Nest generalism",
+                         "Brain residual",
+                         "Ground-nesting",
+                         "Plant eater",
+                         "Hollow-nesting",
+                         "Cooperative breeding",
+                         "Feeding aggregation \n (solitary, pairs, & flocks)",
+                         "Nest aggregation \n (solitary)",
+                         "Carrion eater",
+                         "Nest aggregation \n (none)",
+                         "Nest aggregation \n (colonial)",
+                         "Habitat - tree/forest",
+                         "Feeding aggregation \n (solitary & pairs)",
+                         "Habitat - grass/shrubland",
+                         "Feeding aggregation \n (pairs)", 
+                         "Granivore", 
+                         "Feeding aggregation \n (pairs & flocks)",
+                         "Feeding aggregation \n (solitary)",
+                         "Insectivore")) %>%
+    arrange(estimate) %>%
+    mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
+    ggplot(., aes(x=fct_inorder(variable2), y=estimate, color=trend))+
+    geom_point()+
+    geom_errorbar(aes(ymin=lwr, ymax=upr, color=trend))+
+    ylab("Parameter estimates")+
+    xlab("")+
+    coord_flip()+
[…]

### 2018-04-19 R/non_phylo_analysis.R
changing how i calculate df for plot, and it seems to work in remake again...

@@ -100,60 +100,66 @@ collinearity_investigation_function <- function(global_model) {
 ## plot parameter estimates of standardized global model
 plot_params_globmod <- function(global_model) {
   
-  library(sjPlot)
-  library(forcats)
+    library(forcats)
   
-  pdf("figures/param_plot_global_model.pdf", height=11, width=9)
-  
-  print(
-  get_model_data(global_model, type="std2") %>%
-    arrange(desc(estimate)) %>%
-    mutate(term2 = c("log(Clutch size)",
-                     "Feeding habitat generalism", 
-                     "Migratory type \n (partial migrant)",
-                     "Migratory type \n (dispersal, partial migrant, & nomadic/irruptive",
-                     "Diet generalism", 
-                     "Breeding habitat generalism", 
-                     "Migratory type \n (dispersal & nomadic/irruptive)",
-                     "Migratory type \n (total migrant)", 
-                     "Migratory type \n (none)", 
-                     "Habitat - agricultural", 
-                     "Migratory type \n (nomadic/irruptive)",
-                     "log(Body size)", 
-                     "Brain residual", 
-                     "Migratory type \n (dispersal & partial migrant)",
-                     "Nest generalism", 
-                     "Feeding aggregation \n (solitary & flocks)",
-                     "Feeding aggregation \n (solitary, pairs, & flocks)", 
-                     "Plant eater", 
-                     "Ground-nesting", 
-                     "Migratory type \n (total migrant & nomadic/irruptive)", 
-                     "Cooperative breeding",
-                     "Hollow-nesting", 
-                     "Feeding aggregation \n (solitary & pairs)", 
-                     "Nest aggregation \n (solitary)",
-                     "Habitat - tree/forest", 
-                     "Nest aggregation \n (none)", 
-                     "Nest aggregation \n (colonial)",
-                     "Feeding aggregation \n (pairs)", 
-                     "Granivore", 
-                     "Carrion eater", 
-                     "Feeding aggregation \n (pairs & flocks)", 
-                     "Habitat - grass/shrubland", 
-                     "Insectivore",
-                     "Feeding aggregation \n (solitary)",
-                     "Migratory type \n (partial migrant & nomadic/irruptive)")) %>%
-    arrange(estimate) %>%
-    mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
[…]

### 2018-04-19 R/non_phylo_analysis.R
probably a poor work-around for the remake workflow to make the figure for param estimates. still error with averaged model results though

@@ -101,7 +101,7 @@ collinearity_investigation_function <- function(global_model) {
 plot_params_globmod <- function(global_model) {
   
     library(forcats)
-  
+    remake::dump_environment()
     std.mod <- standardize(global_model) 
     lwr <- data.frame(lwr=confint(std.mod)[,1])
     upr <- data.frame(upr=confint(std.mod)[,2])
@@ -163,6 +163,8 @@ plot_params_globmod <- function(global_model) {
   )
   
   dev.off()
+  
+  rm(list = ls())
 }
 
 
@@ -185,6 +187,8 @@ get_dredged_model <- function() {
 
 
 model_averaging_results <- function(dredged_model) {
+  
+  remake::dump_environment()
   
   #### selects all models with deltaAic < 4
   top.models <- get.models(model.set, subset=delta<4) 
@@ -261,6 +265,8 @@ dev.off()
   pdf("tables/Averaged_models.pdf")
   grid.table(confidence_intervals)
   dev.off()
+  
+  rm(list = ls())
 }
 
 

### 2018-04-19 R/non_phylo_analysis.R
doing all the model averaging on my local

@@ -188,30 +188,7 @@ get_dredged_model <- function() {
 
 model_averaging_results <- function(dredged_model) {
   
-  remake::dump_environment()
-  
-  #### selects all models with deltaAic < 4
-  top.models <- get.models(model.set, subset=delta<4) 
-  
-  #### how many top models
-  length(top.models)
-  
-  #### Ranks these models based on AICc
-  my.models <- model.sel(top.models, rank="AICc") 
-  
-  Averaged_models <- model.avg(top.models)
-  
-  summary(Averaged_models)
-  confidence_intervals <- confint(Averaged_models, full=TRUE)
-  
-  a <- as.data.frame(coefficients(Averaged_models, full=TRUE))
-  b <- as.data.frame(confint(Averaged_models, full=TRUE))
-  model_results <- cbind(a, b)
-  names(model_results)[1] <- "estimate"
-  names(model_results)[2] <- "lwr"
-  names(model_results)[3] <- "upr"
-  model_results$variable <- row.names(model_results)
-  row.names(model_results) <- NULL
+  model_results <- readRDS("Data/dredged_model_averaged_param_est.rds")
   
   pdf("figures/param_plot_averaged_results.pdf")
   print(

### 2018-04-19 scripts/build_dredged_model_for_non_phylo_analysis.R

@@ -16,18 +16,33 @@ build_dredged_model <- function() {
   
   stdz.model <- standardize(global_model)
   
-  model.set <- pdredge(global_model, m.lim=c(0, 11), cluster=clust, extra="R^2") 
+  model.set <- pdredge(global_model, m.lim=c(0, 11), cluster=clust, extra="R^2")
   
-  split <- split(model.set, (seq(nrow(model.set))-1) %/% 200000)
+  #### selects all models with deltaAic < 4
+  top.models <- get.models(model.set, subset=delta<4) 
   
-  list2env(setNames(split,paste0("df",1:4)),environment())
+  #### how many top models
+  length(top.models)
   
-  rm(split)
+  #### Ranks these models based on AICc
+  my.models <- model.sel(top.models, rank="AICc") 
   
-  saveRDS(df1, file="Data/dredged_models_1.rds")
-  saveRDS(df2, file="Data/dredged_models_2.rds")
-  saveRDS(df3, file="Data/dredged_models_3.rds")
-  saveRDS(df4, file="Data/dredged_models_4.rds")
+  Averaged_models <- model.avg(top.models)
+  
+  summary <- summary(Averaged_models)
+  
+  a <- as.data.frame(coefficients(Averaged_models, full=TRUE))
+  b <- as.data.frame(confint(Averaged_models, full=TRUE))
+  model_results <- cbind(a, b)
+  names(model_results)[1] <- "estimate"
+  names(model_results)[2] <- "lwr"
+  names(model_results)[3] <- "upr"
+  model_results$variable <- row.names(model_results)
+  row.names(model_results) <- NULL
+  
+  save(model_results, file="Data/dredged_model_averaged_param_est.rds")
+  save(summary, file="Data/dredged_model_summary_results.rds")
+
 }
 
 build_dredged_model()

### 2018-04-19 scripts/build_dredged_model_for_non_phylo_analysis.R
running entire model averaging script on a local machine and saving results and summary for manipulation in remake workflow

@@ -40,8 +40,8 @@ build_dredged_model <- function() {
   model_results$variable <- row.names(model_results)
   row.names(model_results) <- NULL
   
-  save(model_results, file="Data/dredged_model_averaged_param_est.rds")
-  save(summary, file="Data/dredged_model_summary_results.rds")
+  saveRDS(model_results, file="Data/dredged_model_averaged_param_est.rds")
+  saveRDS(summary, file="Data/dredged_model_summary_results.rds")
 
 }
 

### 2018-04-20 R/non_phylo_analysis.R
seemingly working remake workflow for the non_phylo_analysis. (someone should definitely check!)

@@ -186,7 +186,7 @@ get_dredged_model <- function() {
 }
 
 
-model_averaging_results <- function(dredged_model) {
+model_averaging_results <- function() {
   
   model_results <- readRDS("Data/dredged_model_averaged_param_est.rds")
   
@@ -236,12 +236,7 @@ model_averaging_results <- function(dredged_model) {
   )
 
 dev.off()
-  
-  library(gridExtra)
-  
-  pdf("tables/Averaged_models.pdf")
-  grid.table(confidence_intervals)
-  dev.off()
+
   
   rm(list = ls())
 }

### 2018-04-22 R/phylo_fxns.R
getting set up for many phylos

@@ -1,18 +1,31 @@
 
 
 
-read_bird_tree <- function() {
+download_phylo <- function() {
   require(ape)
   if (!exists("Data/phylo/"))
     dir.create("Data/phylo/", showWarnings = FALSE)
   download.file(
-    "https://www.dropbox.com/s/gfackoogv6n0bv0/AllBirdsEricson1.tre?raw=1",
-    "Data/phylo/phy.tre"
+    "https://data.vertlife.org/birdtree/Stage2/EricsonStage2_0001_1000.zip",
+    "Data/phylo/EricsonStage2_0001_1000.zip"
   )
-  bird_trees <- ape::read.tree(file = "Data/phylo/phy.tre")[[1]]
-  #drop tips to get down to Aus species here
-  return(bird_trees)
+  
 }
+
+read_one_tree<-function(path,x=1){
+  unzip(path,overwrite=TRUE,exdir="Data/phylo/")
+  one_bird_tree <- ape::read.tree(file = "/Users/wcornwell/Documents/bird_urbanness/Data/phylo/mnt/data/projects/birdphylo/Tree_sets/Stage2_full_data/CombinedTrees/AllBirdsEricson1.tre")[[x]]
+  #drop tips to get down to Aus species here
+  return(one_bird_tree)
+}
+
+read_all_trees<-function(path){
+  unzip(path,overwrite=TRUE,exdir="Data/phylo/")
+  ape::read.tree(file = "/Users/wcornwell/Documents/bird_urbanness/Data/phylo/mnt/data/projects/birdphylo/Tree_sets/Stage2_full_data/CombinedTrees/AllBirdsEricson1.tre")
+}
+  
+  
+
 
 
 plot_bird_tree <- function(aus_bird_tree) {

### 2018-04-22 R/phylo_fxns.R
new figure into workflow

@@ -23,8 +23,8 @@ read_all_trees<-function(path){
   unzip(path,overwrite=TRUE,exdir="Data/phylo/")
   ape::read.tree(file = "/Users/wcornwell/Documents/bird_urbanness/Data/phylo/mnt/data/projects/birdphylo/Tree_sets/Stage2_full_data/CombinedTrees/AllBirdsEricson1.tre")
 }
-  
-  
+
+
 
 
 
@@ -101,10 +101,47 @@ plot_bird_tree_traits <-
   }
 
 
+
+
+
+
+run_one_phylo_model<-function(analysis_data,aus_bird_tree){
+  #non_aus_sp <- aus_bird_tree$tip.label[!aus_bird_tree$tip.label %in% analysis_data$binom]
+  #aus_bird_tree_ss <- diversitree:::drop.tip.fixed(aus_bird_tree, non_aus_sp)
+  row.names(analysis_data)<-analysis_data$binom
+  print(class(aus_bird_tree))
+  phy_mod<-phylolm(response~body_size_logged + clutch_size_logged + feeding_habitat_generalism +   
+                     brain_residual + Habitat_agricultural + breeding_habitat_generalism + 
+                     granivore + insectivore + 
+                     carrion_eater + plant_eater + diet_generalism + movement_class +
+                     ground_nesting + hollow_nesting + nest_generalism + breeding + 
+                     nest_aggregation + feeding_aggregation + Habitat_grass_shrubland +
+                     Habitat_tree_forest,data=analysis_data,phy=aus_bird_tree,
+                   na.action = "na.fail", weights=(analysis_data$N/analysis_data$unique_localities))
+  
+  return(phy_mod)
+}
+
+
+phy_v_non_phy<-function(glob.mod,phy_mod){
+  library(ggplot2)
+  cc<-data.frame(phy_mod_coefs=coef(phy_mod),non_phy_mod=coef(glob.mod))
+  cc$parameter<-substr(row.names(cc), start = 1, stop = 7)
+  
+  p <- ggplot(cc,aes(x=non_phy_mod,y=phy_mod_coefs))+geom_point()+geom_text(aes(label=parameter),hjust=0, vjust=0)+theme_bw()+geom_abline(slope=1,intercept=0)
+  pdf("figures/phy_v_non_phy.pdf")
+  print(p)
+  dev.off()
+}
+
+
+############# BELOW HERE IS OBSOLETE 
+
+
[…]

### 2018-04-22 R/phylo_fxns.R
working on #33

@@ -101,15 +101,31 @@ plot_bird_tree_traits <-
   }
 
 
+run_many_phylo_models<-function(analysis_data,list_bird_trees,n=1000){
+  bird_trees_ss<-list_bird_trees[1:n]
+  ss_trees<-lapply(bird_trees_ss,subset_tree,analysis_data)
+  list_o<-lapply(ss_trees,run_one_phylo_model,analysis_data=analysis_data)
+  return(list_o)
+}
+
+extract_brain<-function(mod){
+  nn<-names(mod$coefficients)
+  return(mod$coefficients[nn=="brain_residual"])
+}
+
+plot_dist_parameter<-function(list_of_models){
+  df<-data.frame(coef_brain=sapply(list_of_models,extract_brain))
+  p<-ggplot(df,aes(x=coef_brain))+geom_histogram()
+  pdf("figures/brains.pdf")
+  print(p)
+  dev.off()
+}
 
 
-
-
-run_one_phylo_model<-function(analysis_data,aus_bird_tree){
+run_one_phylo_model<-function(aus_bird_tree,analysis_data){
   #non_aus_sp <- aus_bird_tree$tip.label[!aus_bird_tree$tip.label %in% analysis_data$binom]
   #aus_bird_tree_ss <- diversitree:::drop.tip.fixed(aus_bird_tree, non_aus_sp)
   row.names(analysis_data)<-analysis_data$binom
-  print(class(aus_bird_tree))
   phy_mod<-phylolm(response~body_size_logged + clutch_size_logged + feeding_habitat_generalism +   
                      brain_residual + Habitat_agricultural + breeding_habitat_generalism + 
                      granivore + insectivore + 
@@ -132,138 +148,4 @@ phy_v_non_phy<-function(glob.mod,phy_mod){
   pdf("figures/phy_v_non_phy.pdf")
   print(p)
   dev.off()
-}
-
-
-############# BELOW HERE IS OBSOLETE 
-
-
-run_phylo_lm <- function(traits,
-                         response_variables,
-                         aus_bird_tree) {
-  
-  #
-  # match and subset to tree.
-  # this is annoying because the phylo packages all use row.names and the tidyverse hates rownames.
-  # first create a new data.frame (not tibble) with just the traits of interest
-  #
-  trait <-
-    data.frame(
-      mean_body_size = traits$mean_body_size,
[…]

### 2018-04-23 R/phylo_fxns.R
changing directory for me to work remake

@@ -13,14 +13,14 @@ download_phylo <- function() {
 }
 
 read_one_tree<-function(path,x=1){
-  unzip(path,overwrite=TRUE,exdir="Data/phylo/")
+  unzip(path,overwrite=TRUE,exdir="Data/phylo")
   one_bird_tree <- ape::read.tree(file = "Data/phylo/mnt/data/projects/birdphylo/Tree_sets/Stage2_full_data/CombinedTrees/AllBirdsEricson1.tre")[[x]]
   #drop tips to get down to Aus species here
   return(one_bird_tree)
 }
 
 read_all_trees<-function(path){
-  unzip(path,overwrite=TRUE,exdir="Data/phylo/")
+  unzip(path,overwrite=TRUE,exdir="Data/phylo")
   ape::read.tree(file = "Data/phylo/mnt/data/projects/birdphylo/Tree_sets/Stage2_full_data/CombinedTrees/AllBirdsEricson1.tre")
 }
 

### 2018-04-23 R/non_phylo_analysis.R
adding range size, converting to two movement classes, and remakifying it for non_phylo analysis

@@ -3,6 +3,8 @@
 
 
 ## create df for the model fitting
+## read in range size file here
+## subset data based on complete variables as well
 matched_df <- function(response_variables, traits){
   
   response_variables$binom <- response_variables$SCIENTIFIC_NAME_tree
@@ -10,7 +12,7 @@ matched_df <- function(response_variables, traits){
   analysis_data <- left_join(traits, response_variables, by="binom") %>%
     filter(SCIENTIFIC_NAME_tree != "not_treated_as_species")
   
-  # clean data to final dataset
+  # clean data
   analysis_data <- analysis_data %>%
     # drop any columns that we aren't interested in
     dplyr::select(-brain_size, -iucn_status) %>%
@@ -21,6 +23,14 @@ matched_df <- function(response_variables, traits){
     mutate(response = log(urban_median)) %>%
     mutate(body_size_logged = log(mean_body_size)) %>%
     mutate(clutch_size_logged = log(clutch_size))
+  
+  # filter data based on range size availability
+  range_size <- read_csv("Data/Raw trait data/temp_range_size.csv")
+  
+  # filter one more time to final dataset
+  analysis_data <- analysis_data %>%
+    inner_join(., range_size, by=c("COMMON_NAME_traits", "SCIENTIFIC_NAME_traits")) %>%
+    filter_all(all_vars(!is.na(.)))
   
   analysis_data
 }
@@ -41,6 +51,7 @@ corrplot_figure <- function(analysis_data){
     rename(`Unique localitieis observed` = unique_localities) %>%
     rename(`log(Body size)` = body_size_logged) %>%
     rename(`log(Clutch size)` = clutch_size_logged) %>%
+    rename(`Range size (1000s km2)` = range_size) %>%
     cor(use="pairwise.complete.obs")
   
   pdf("figures/corrplot_of_continuous_variables.pdf")
@@ -59,9 +70,9 @@ get_global_model <- function(analysis_data) {
   
   glob.mod <- lm(response ~ body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + 
                    Habitat_agricultural + breeding_habitat_generalism + granivore + insectivore + 
-                   carrion_eater + plant_eater + diet_generalism + movement_class +
+                   carrion_eater + plant_eater + diet_generalism + migrate + nomadic_irruptive +
                    ground_nesting + hollow_nesting + nest_generalism + breeding + 
[…]

### 2018-04-23 R/trait_fxns.R

@@ -32,7 +32,8 @@ Mode <- function(x) {
 read_process_trait_data <- function(){
   
   traits <- read_trait_data_in()
-    traits %>%
+  
+  ms <- traits %>%
     mutate(body_mass=ifelse(`99_Body_mass_average_8`=="NAV",NA,as.numeric(`99_Body_mass_average_8`))) %>%
     mutate(clutch_size=ifelse(`178_Clutch_size_average_12`=="NAV",NA,as.numeric(`178_Clutch_size_average_12`))) %>%
     mutate(iucn_status=`49_Australian_status_July_2015_5`) %>%
@@ -44,7 +45,7 @@ read_process_trait_data <- function(){
               clutch_size=mean(clutch_size,na.rm=T),
               iucn_status=Mode(iucn_status),
               brain_size=mean(brain_size, na.rm=T),
-              brain_residual=mean(brain_residual, na.rm=T)) -> ms
+              brain_residual=mean(brain_residual, na.rm=T))
               
     
 ## calculate df for feeding habitat generalism
@@ -76,7 +77,7 @@ read_process_trait_data <- function(){
       mutate(tree_forest = rowSums(.[2:11])) %>%
       dplyr::select(binom, tree_forest) %>%
       mutate(tree_forest = ifelse(.$tree_forest > 1, 1, .$tree_forest)) %>%
-      mutate(Habitat_tree_forest = ifelse(.$tree_forest == 1, "Yes", "no")) %>%
+      mutate(Habitat_tree_forest = ifelse(.$tree_forest == 1, "Yes", "No")) %>%
       dplyr::select(-tree_forest)
     
     
@@ -89,7 +90,7 @@ read_process_trait_data <- function(){
       mutate(grass_shrubland = rowSums(.[2:8])) %>%
       dplyr::select(binom, grass_shrubland) %>%
       mutate(grass_shrubland = ifelse(.$grass_shrubland > 1, 1, .$grass_shrubland)) %>%
-      mutate(Habitat_grass_shrubland = ifelse(.$grass_shrubland == 1, "Yes", "no")) %>%
+      mutate(Habitat_grass_shrubland = ifelse(.$grass_shrubland == 1, "Yes", "No")) %>%
       dplyr::select(-grass_shrubland)
     
     
@@ -102,7 +103,7 @@ read_process_trait_data <- function(){
       mutate(agricultural = rowSums(.[2:3])) %>%
[…]

### 2018-04-23 R/phylo_fxns.R
everything remaking up to date for non_phylo and phylo analyses

@@ -126,13 +126,12 @@ run_one_phylo_model<-function(aus_bird_tree,analysis_data){
   #non_aus_sp <- aus_bird_tree$tip.label[!aus_bird_tree$tip.label %in% analysis_data$binom]
   #aus_bird_tree_ss <- diversitree:::drop.tip.fixed(aus_bird_tree, non_aus_sp)
   row.names(analysis_data)<-analysis_data$binom
-  phy_mod<-phylolm(response~body_size_logged + clutch_size_logged + feeding_habitat_generalism +   
-                     brain_residual + Habitat_agricultural + breeding_habitat_generalism + 
-                     granivore + insectivore + 
-                     carrion_eater + plant_eater + diet_generalism + movement_class +
+  phy_mod<-phylolm(response ~ body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + 
+                     Habitat_agricultural + breeding_habitat_generalism + granivore + insectivore + 
+                     carrion_eater + plant_eater + diet_generalism + migrate + nomadic_irruptive +
                      ground_nesting + hollow_nesting + nest_generalism + breeding + 
-                     nest_aggregation + feeding_aggregation + Habitat_grass_shrubland +
-                     Habitat_tree_forest,data=analysis_data,phy=aus_bird_tree,
+                     nest_aggregation + feeding_aggregation + Habitat_grass_shrubland + range_size +
+                     Habitat_tree_forest, data=analysis_data, phy=aus_bird_tree,
                    na.action = "na.fail", weights=(analysis_data$N/analysis_data$unique_localities))
   
   return(phy_mod)

### 2018-04-23 R/phylo_fxns.R
#37 and #38 - standardize phylo model and make param plot

@@ -122,21 +122,117 @@ plot_dist_parameter<-function(list_of_models){
 }
 
 
-run_one_phylo_model<-function(aus_bird_tree,analysis_data){
+run_one_phylo_model<-function(aus_bird_tree, analysis_data){
   #non_aus_sp <- aus_bird_tree$tip.label[!aus_bird_tree$tip.label %in% analysis_data$binom]
   #aus_bird_tree_ss <- diversitree:::drop.tip.fixed(aus_bird_tree, non_aus_sp)
-  row.names(analysis_data)<-analysis_data$binom
-  phy_mod<-phylolm(response ~ body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + 
+  row.names(analysis_data) <- analysis_data$binom
+  
+  phy_mod <- phylolm(response ~ body_size_logged + clutch_size_logged + feeding_habitat_generalism + brain_residual + 
                      Habitat_agricultural + breeding_habitat_generalism + granivore + insectivore + 
                      carrion_eater + plant_eater + diet_generalism + migrate + nomadic_irruptive +
                      ground_nesting + hollow_nesting + nest_generalism + breeding + 
                      nest_aggregation + feeding_aggregation + Habitat_grass_shrubland + range_size +
                      Habitat_tree_forest, data=analysis_data, phy=aus_bird_tree,
-                   na.action = "na.fail", weights=(analysis_data$N/analysis_data$unique_localities))
+                     na.action = "na.fail", weights=(analysis_data$N/analysis_data$unique_localities))
+  
+  phy_mod_rescaled <- phylolm(response ~ rescale(body_size_logged) + rescale(clutch_size_logged) + 
+                                rescale(feeding_habitat_generalism) + rescale(brain_residual) + 
+                                rescale(Habitat_agricultural) + rescale(breeding_habitat_generalism) + 
+                                rescale(granivore) + rescale(insectivore) + 
+                                rescale(carrion_eater) + rescale(plant_eater) + rescale(diet_generalism) + 
+                                rescale(migrate) + rescale(nomadic_irruptive) +
+                                rescale(ground_nesting) + rescale(hollow_nesting) + 
+                                rescale(nest_generalism) + rescale(breeding) + 
+                                nest_aggregation + feeding_aggregation + 
+                                rescale(Habitat_grass_shrubland) + rescale(range_size) +
+                                rescale(Habitat_tree_forest), data=analysis_data, phy=aus_bird_tree,
+                                na.action = "na.fail", 
[…]

### 2018-04-23 R/phylo_fxns.R
sharpening up the coefficient comparison plot between the two models

@@ -236,11 +236,49 @@ plot_params_phymod <- function(phy_mod_rescaled) {
 
 phy_v_non_phy<-function(glob.mod,phy_mod){
   library(ggplot2)
-  cc<-data.frame(phy_mod_coefs=coef(phy_mod),non_phy_mod=coef(glob.mod))
-  cc$parameter<-substr(row.names(cc), start = 1, stop = 7)
+  cc <- data.frame(phy_mod_coefs=coef(phy_mod_rescaled), 
+                   non_phy_mod=coef(global_model)) %>%
+    mutate(parameter = c("Intercept",
+                         "log(Body size)",
+                         "log(Clutch size)",
+                         "Feeding habitat generalism",
+                         "Brain residual", 
+                         "Habitat - agricultural", 
+                         "Breeding habitat generalism",
+                         "Granivore", 
+                         "Insectivore", 
+                         "Carrion eater", 
+                         "Plant eater", 
+                         "Diet generalism",
+                         "Movement - migratory", 
+                         "Movement - nomadic/irruptive", 
+                         "Ground-nesting", 
+                         "Hollow-nesting", 
+                         "Nest generalism", 
+                         "Cooperative breeding", 
+                         "Nest aggregation \n (colonial)",
+                         "Nest aggregation \n (none)",
+                         "Nest aggregation  \n (solitary)",
+                         "Feeding aggregation \n (pairs)",
+                         "Feeding aggregation \n (pairs & flocks)",
+                         "Feeding aggregation \n (solitary)",
+                         "Feeding aggregation \n (solitary & flocks)",
+                         "Feeding aggregation \n (solitary & pairs)",
+                         "Feeding aggregation \n (solitary, pairs, & flocks)",
+                         "Habitat - grass/shrubland", 
+                         "Range size (1000s km2)", 
+                         "Habitat - tree/forest"))
   
-  p <- ggplot(cc,aes(x=non_phy_mod,y=phy_mod_coefs))+geom_point()+geom_text(aes(label=parameter),hjust=0, vjust=0)+theme_bw()+geom_abline(slope=1,intercept=0)
-  pdf("figures/phy_v_non_phy.pdf")
+  p <- ggplot(cc,aes(x=non_phy_mod,y=phy_mod_coefs))+
+    geom_point(color="royalblue4")+
+    theme_bw()+
+    geom_abline(slope=1,intercept=0)+
[…]

### 2018-04-24 R/phylo_fxns.R
a bit of renaming objects

@@ -234,7 +234,7 @@ plot_params_phymod <- function(phy_mod_rescaled) {
   rm(list = ls())
 }
 
-phy_v_non_phy<-function(glob.mod,phy_mod){
+phy_v_non_phy<-function(global_model,phy_mod_rescaled){
   library(ggplot2)
   cc <- data.frame(phy_mod_coefs=coef(phy_mod_rescaled), 
                    non_phy_mod=coef(global_model)) %>%

### 2018-04-24 scripts/results_script.R
minor changes to some figures

@@ -3,29 +3,17 @@
 ## first dump the remake environment
 remake::dump_environment()
 
-## get the final dataframe which was used for analysis
-## code copies from the non_phylo_model.Rmd
-response_variables$binom <- response_variables$SCIENTIFIC_NAME_tree
-
-df <- left_join(traits,response_variables, by="binom") %>%
-  filter(SCIENTIFIC_NAME_tree != "not_treated_as_species") %>%
-  # drop columns that we aren't inerested in
-  dplyr::select(-brain_size)
-
-df2 <- df %>%
-  filter_all(all_vars(!is.na(.)))
-
 ## total number of species included in models
-length(unique(df2))
+length(unique(analysis_data))
 
 ## summary of urbanization index
-summary(df2$urban_median)
+summary(analysis_data$urban_median)
 
-sd(df2$urban_median)
+sd(analysis_data$urban_median)
 
-min(df2$urban_median)
+min(analysis_data$urban_median)
 
-max(df2$urban_median)
+max(analysis_data$urban_median)
 
 
 ## Figure 1
@@ -33,8 +21,11 @@ max(df2$urban_median)
 ## Need to load the species_urban .RData file first
 load('Data/eBird data/Final eBird data for analysis/species_urban.RData')
 
-species_to_plot <- c("Common Myna", "Scaly-breasted Lorikeet", "Welcome Swallow", 
-                     "Banded Honeyeater", "White-winged Fairywren")
+species_to_plot <- c("Dusky Moorhen",
+                     "Yellow-tailed Black-Cockatoo",
+                     "Hooded Robin",
+                     "Noisy Friarbird",
+                     "Squatter Pigeon")
 
 Figure1 <- species_urban %>%
   filter(COMMON_NAME %in% species_to_plot) %>%
@@ -42,10 +33,11 @@ Figure1 <- species_urban %>%
   geom_density(alpha = 0.7) + 
   scale_x_log10() + 
   theme_classic() +
-  xlab("Average radiance") +
+  xlab("log(Average radiance)") +
   ylab("Density")+
   guides(fill=guide_legend(title="  Common Name")) +
-  theme(legend.position = c(0.2, 0.8))
+  theme(legend.position = c(0.26, 0.8))+
+  theme(axis.text.x=element_text(hjust=0.7))
 
 pdf("final figures for paper/Figure1.pdf", width=5.5, height=4.5)
 print(Figure1)
@@ -57,21 +49,25 @@ dev.off()
 ## Use the df2 dataframe from above
[…]

### 2018-04-24 reports/phylo_report.rmd
phylo signal stuff

--- 2018-03-30 reports/phylo_report.rmd
@@
 ---
-title: "phylo_analyses"
+title: "phylo_signal_analyses"
 author: "Everyone"
@@
 
-run phylo model
+```{r}
+require(phylosignal)
+   dplyr::select(analysis_data,body_size_logged,clutch_size_logged,response,range_size, #select appropriate rows
+          nest_generalism,diet_generalism,breeding_habitat_generalism,feeding_habitat_generalism,brain_residual) ->dd
 
+row.names(dd)<-analysis_data$binom #name rows so that it matches the tree
+dd$rand<-rnorm(dim(dd)[2]) #random numbers to test package
+p4d <- phylo4d(aus_bird_tree, dd) #create phylobase object
+
+ps<-phyloSignal(p4d,reps = 9999) #run calculation, p values a bit unstable at 999 reps
+```
+There are two output table the summary statistics and the p-value first the summary stats
 ```{r}
-
-mod2<-run_phylo_lm(traits,
-                         response_variables,
-                         aus_bird_tree)
-
-summary(mod2)
+knitr::kable(ps$stat)
 ```
-maybe we eventually want to do model selection
-
+and then the p-values
 ```{r}
-# library(MuMIn)
-# dredge(mod2)
+knitr::kable(ps$pvalue)
 ```
-
-this is sort of cool.  phylogenetic signal in all the traits
-
-```{r}
-  # library(phylosignal)
-  # library(phylobase)
-  # dd$body_size<-log10(dd$body_size)
-  # dd2<-subset(dd,select=-gregariousness)
-  # out<-phylo4d(tree_plotting_2,tip.data=dd2)
-  # ps<-phyloSignal(out)
-  # knitr::kable(ps$stat)
-```

### 2018-04-24 readme.Rmd
readme things

---
title: "Traits that affect urbanness in birds"
output: github_document
---

This repository contains code needed to reproduce the analysis for the article: Callaghan et al. 2019

## Instructions

### Install relevant software

All analyses were done in `R`. You need to download this repository, and then open an R session with working directory set to the root of the project.

To compile the paper, we use the [remake](https://github.com/richfitz/remake) package for R. You can install remake using the `devtools` package (run `install.packages("devtools")` to install devtools if needed):

```{r,eval=FALSE}
devtools::install_github("richfitz/remake", dependencies=TRUE)
devtools::install_github("liamrevell/phytools", dependencies=TRUE)
```

### Recreating the figures and paper

To generate all figures and analyses simply do:

```{r}
remake::make()
```

### Understanding the workflow of the analysis

Here is a figure that describes the workflow from the cannonical data files to the final figures and tables.  In the following figure, internal data steps (intermediate files) are shown in blue and the final outputs are shown in red.  Each arrow is performed by a function written in R and contained in this repository.

```{r,eval=FALSE}
remake::diagram(remake::make())
```

![](workflow.png)

### 2018-04-25 R/non_phylo_analysis.R
removing old code

@@ -174,23 +174,9 @@ plot_params_globmod <- function(global_model) {
 
 
 ## this pulls in the dredged model results which was
-## done out of the workflow
-## the data was chunked in order to get around github file size limits
+## done out of the workflow (currently takes about 24 hours to run)
 ## will likely have to alter this once we have a finalized
 ## set of predictor variables
-get_dredged_model <- function() {
-  
-  df1 <- readRDS("Data/dredged_models_1.rds")
-  df2 <- readRDS("Data/dredged_models_2.rds")
-  df3 <- readRDS("Data/dredged_models_3.rds")
-  df4 <- readRDS("Data/dredged_models_4.rds")
-  
-  model.set <- bind_rows(df1, df2, df3, df4)
-  
-  return(model.set)
-}
-
-
 model_averaging_results <- function() {
   
   model_results <- readRDS("Data/dredged_model_averaged_param_est.rds")

### 2018-04-25 R/non_phylo_analysis.R
updated dredged non-phylo results (takes about 24 hours to dredge the models on my machine)

@@ -188,31 +188,29 @@ model_averaging_results <- function() {
     droplevels() %>%
     arrange(desc(estimate)) %>%
     mutate(variable2 = c("log(Clutch size)",
-                         "Feeding habitat generalism",
-                         "Diet generalism",
                          "Habitat - agricultural",
-                         "Breeding habitat generalism",
+                         "Diet generalism",
                          "Feeding aggregation \n (solitary & flocks)",
+                         "Breeding habitat generalism",
+                         "Feeding habitat generalism",
+                         "Movement - migratory",
+                         "Movement - nomadic/irruptive",
                          "log(Body size)",
-                         "Nest generalism",
-                         "Brain residual",
                          "Ground-nesting",
-                         "Plant eater",
-                         "Hollow-nesting",
                          "Cooperative breeding",
-                         "Feeding aggregation \n (solitary, pairs, & flocks)",
-                         "Nest aggregation \n (solitary)",
                          "Carrion eater",
-                         "Nest aggregation \n (none)",
                          "Nest aggregation \n (colonial)",
-                         "Habitat - tree/forest",
+                         "Nest aggregation \n (solitary)",
+                         "Nest aggregation \n (none)",
+                         "Feeding aggregation \n (solitary, pairs, & flocks)",
                          "Feeding aggregation \n (solitary & pairs)",
-                         "Habitat - grass/shrubland",
-                         "Feeding aggregation \n (pairs)", 
+                         "Feeding aggregation \n (pairs)",
+                         "Habitat - tree/forest",
                          "Granivore", 
                          "Feeding aggregation \n (pairs & flocks)",
-                         "Feeding aggregation \n (solitary)",
-                         "Insectivore")) %>%
+                         "Habitat - grass/shrubland",
+                         "Insectivore",
+                         "Feeding aggregation \n (solitary)")) %>%
     arrange(estimate) %>%
     mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
     ggplot(., aes(x=fct_inorder(variable2), y=estimate, color=trend))+

### 2018-04-25 R/non_phylo_analysis.R
extracting p-values for each variable from all model results so far

@@ -123,6 +123,7 @@ plot_params_globmod <- function(global_model) {
     tidy(std.mod) %>%
       mutate(lwr=lwr$lwr) %>%
       mutate(upr=upr$upr) %>%
+      mutate(significance = ifelse(p.value <=0.05, "Significant", "Non-significant")) %>%
       filter(term != "(Intercept)") %>%
       arrange(desc(estimate)) %>%
       mutate(term2 = c("Feeding habitat generalism",
@@ -180,10 +181,15 @@ plot_params_globmod <- function(global_model) {
 model_averaging_results <- function() {
   
   model_results <- readRDS("Data/dredged_model_averaged_param_est.rds")
+  summary <- readRDS("Data/dredged_model_summary_results.rds")
+  
+  p_values <- data.frame(p_value=summary$coefmat.full[,5]) %>%
+    rownames_to_column("variable")
   
   pdf("figures/param_plot_averaged_results.pdf")
   print(
   model_results %>%
+    inner_join(., p_values, by="variable") %>%
     filter(variable != "(Intercept)") %>%
     droplevels() %>%
     arrange(desc(estimate)) %>%
@@ -213,6 +219,7 @@ model_averaging_results <- function() {
                          "Feeding aggregation \n (solitary)")) %>%
     arrange(estimate) %>%
     mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
+    mutate(significance=ifelse(.$p_value <= 0.05, "Significant", "Non-significant")) %>%
     ggplot(., aes(x=fct_inorder(variable2), y=estimate, color=trend))+
     geom_point()+
     geom_errorbar(aes(ymin=lwr, ymax=upr, color=trend))+
@@ -226,8 +233,6 @@ model_averaging_results <- function() {
 
 dev.off()
 
-  
-  rm(list = ls())
 }
 
 

### 2018-04-25 R/phylo_fxns.R

@@ -176,9 +176,12 @@ standard_phylo_model <- function(aus_bird_tree, analysis_data) {
 plot_params_phymod <- function(phy_mod_rescaled) {
   
   results <- data.frame(estimate = phy_mod_rescaled$coefficients, 
-                             lwr = confint(phy_mod_rescaled)[,1],
-                             upr = confint(phy_mod_rescaled)[,2],
-                             stringsAsFactors = FALSE)
+                        lwr = confint(phy_mod_rescaled)[,1],
+                        upr = confint(phy_mod_rescaled)[,2],
+                        p_value = summary(phy_mod_rescaled)$coefficients[,4],
+                        stringsAsFactors = FALSE)
+  
+  summary(phy_mod_rescaled)$coefficients[,4]
   
   pdf("figures/param_plot_phylo_model.pdf", height=11, width=9)
   
@@ -218,6 +221,7 @@ plot_params_phymod <- function(phy_mod_rescaled) {
                        "Feeding aggregation \n (pairs)")) %>%
       arrange(estimate) %>%
       mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
+      mutate(significance=ifelse(.$p_value <= 0.05, "Significant", "Non-significant")) %>%
       ggplot(., aes(x=fct_inorder(term2), y=estimate, color=trend))+
       geom_point()+
       geom_errorbar(aes(ymin=lwr, ymax=upr, color=trend))+

### 2018-04-25 R/phylo_fxns.R
getting rid of line I accidentally left in

@@ -185,8 +185,6 @@ plot_params_phymod <- function(phy_mod_rescaled) {
                         p_value = summary(phy_mod_rescaled)$coefficients[,4],
                         stringsAsFactors = FALSE)
   
-  summary(phy_mod_rescaled)$coefficients[,4]
-  
   pdf("figures/param_plot_phylo_model.pdf", height=11, width=9)
   
   print( 

### 2018-04-25 scripts/results_script.R
messing with making final figure of param estimates

@@ -92,8 +92,207 @@ my.file.copy(from = "figures/bird_urbanness_phylo.pdf",
 
 
 
+## All parameter estimates
+## A script which plots results for each of the four 'models'
+## only three for now though
+
+## first make dfs for each separate model approach
+### Non -phylo global model
+std.mod <- standardize(global_model) 
+lwr <- data.frame(lwr=confint(std.mod)[,1])
+upr <- data.frame(upr=confint(std.mod)[,2])
+
+estimates_non_phylo <- tidy(std.mod) %>%
+  mutate(lwr=lwr$lwr) %>%
+  mutate(upr=upr$upr) %>%
+  mutate(significance = ifelse(p.value <=0.05, "Significant", "Non-significant")) %>%
+  filter(term != "(Intercept)") %>%
+  arrange(desc(estimate)) %>%
+  mutate(term2 = c("Feeding habitat generalism",
+                   "Breeding habitat generalism",
+                   "log(Clutch size)",
+                   "Diet generalism",
+                   "Habitat - agricultural",
+                   "Movement - migratory",
+                   "log(Body size)",
+                   "Movement - nomadic/irruptive",
+                   "Brain residual",
+                   "Feeding aggregation \n (solitary & flocks)",
+                   "Ground-nesting",
+                   "Feeding aggregation \n (solitary, pairs, & flocks)",
+                   "Plant eater",
+                   "Nest generalism",
+                   "Feeding aggregation \n (solitary & pairs)",
+                   "Range size (1000s km2)",
+                   "Hollow-nesting",
+                   "Cooperative breeding",
+                   "Nest aggregation \n (colonial)",
+                   "Feeding aggregation \n (pairs)",
+                   "Nest aggregation \n (solitary)",
+                   "Carrion eater",
+                   "Nest aggregation \n (none)",
+                   "Feeding aggregation \n (pairs & flocks)",
+                   "Granivore",
+                   "Habitat - tree/forest",
+                   "Habitat - grass/shrubland",
+                   "Insectivore",
+                   "Feeding aggregation \n (solitary)")) %>%
+  arrange(estimate) %>%
+  mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
+  dplyr::select(term2, estimate, lwr, upr, p.value, significance, trend) %>%
+  rename(variable=term2) %>%
+  rename(p_value=p.value) %>%
+  mutate(model = "Non-phylo global model")
+
[…]

### 2018-05-03 scripts/build_dredged_model_for_non_phylo_analysis.R
averaging across my.models instead of top.models. same thing, but it works now.

@@ -29,7 +29,7 @@ build_dredged_model <- function() {
   #### Ranks these models based on AICc
   my.models <- model.sel(top.models, rank="AICc") 
   
-  Averaged_models <- model.avg(top.models)
+  Averaged_models <- model.avg(my.models)
   
   summary <- summary(Averaged_models)
   

### 2018-05-03 scripts/build_dredged_model_for_phylo_analysis.R

@@ -77,7 +77,7 @@ build_dredged_model <- function() {
   #### Ranks these models based on AICc
   my.models <- model.sel(top.models, rank="AICc") 
   
-  Averaged_models <- model.avg(top.models)
+  Averaged_models <- model.avg(my.models)
   
   summary <- summary(Averaged_models)
   

### 2018-05-03 R/phylo_fxns.R
adding phylo model averaged param plot to remake workflow

@@ -242,6 +242,69 @@ plot_params_phymod <- function(phy_mod_rescaled) {
   rm(list = ls())
 }
 
+## this pulls in the dredged model results which was
+## done out of the workflow (currently takes about 24 hours to run)
+## will likely have to alter this once we have a finalized
+## set of predictor variables
+phylomodel_averaging_results <- function() {
+  
+  model_results <- readRDS("Data/PHYLO_dredged_model_averaged_param_est.rds")
+  summary <- readRDS("Data/PHYLO_dredged_model_summary_results.rds")
+  
+  p_values <- data.frame(p_value=summary$coefmat.full[,4]) %>%
+    rownames_to_column("variable")
+  
+  pdf("figures/phylo_param_plot_averaged_results.pdf")
+  print(
+    model_results %>%
+      inner_join(., p_values, by="variable") %>%
+      filter(variable != "(Intercept)") %>%
+      droplevels() %>%
+      arrange(desc(estimate)) %>%
+      mutate(variable2 = c("Feeding habitat generalism",
+                           "Brain residual",
+                           "log(Body size)",
+                           "log(Clutch size)",
+                           "Breeding habitat generalism",
+                           "Nest aggregation \n (colonial)",
+                           "Diet generalism",
+                           "Plant eater",
+                           "Habitat - agricultural",
+                           "Insectivore",
+                           "Nest aggregation \n (solitary)",
+                           "Movement - migratory",
+                           "Habitat - tree/forest",
+                           "Ground-nesting",
+                           "Movement - nomadic/irruptive",
+                           "Habitat - grass/shrubland",
+                           "Hollow-nesting",
+                           "Granivore",
+                           "Nest generalism",
+                           "Cooperative breeding",
+                           "Carrion eater",
+                           "Range size (1000s km2)",
+                           "Nest aggregation \n (none)")) %>%
+      arrange(estimate) %>%
+      mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
+      mutate(significance=ifelse(.$p_value <= 0.05, "Significant", "Non-significant")) %>%
+      ggplot(., aes(x=fct_inorder(variable2), y=estimate, color=trend))+
+      geom_point()+
+      geom_errorbar(aes(ymin=lwr, ymax=upr, color=trend))+
+      ylab("Parameter estimates")+
[…]

### 2018-05-03 scripts/results_script.R
working on figure presentation

@@ -235,8 +235,52 @@ estimates_phylo <- results %>%
     rename(variable=term2) %>%
     mutate(model = "Phylo global model")
 
+# Phylo model-averaged results
+model_results <- readRDS("Data/PHYLO_dredged_model_averaged_param_est.rds")
+summary <- readRDS("Data/PHYLO_dredged_model_summary_results.rds")
+
+p_values <- data.frame(p_value=summary$coefmat.full[,4]) %>%
+  rownames_to_column("variable")
+
+estimates_phylo_averaged <- model_results %>%
+    inner_join(., p_values, by="variable") %>%
+    filter(variable != "(Intercept)") %>%
+    droplevels() %>%
+    arrange(desc(estimate)) %>%
+    mutate(variable2 = c("Feeding habitat generalism",
+                         "Brain residual",
+                         "log(Body size)",
+                         "log(Clutch size)",
+                         "Breeding habitat generalism",
+                         "Nest aggregation \n (colonial)",
+                         "Diet generalism",
+                         "Plant eater",
+                         "Habitat - agricultural",
+                         "Insectivore",
+                         "Nest aggregation \n (solitary)",
+                         "Movement - migratory",
+                         "Habitat - tree/forest",
+                         "Ground-nesting",
+                         "Movement - nomadic/irruptive",
+                         "Habitat - grass/shrubland",
+                         "Hollow-nesting",
+                         "Granivore",
+                         "Nest generalism",
+                         "Cooperative breeding",
+                         "Carrion eater",
+                         "Range size (1000s km2)",
+                         "Nest aggregation \n (none)")) %>%
+    arrange(estimate) %>%
+    mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
+    mutate(significance=ifelse(.$p_value <= 0.05, "Significant", "Non-significant")) %>%
+    mutate(model = "Phylo model-averaged") %>%
+    dplyr::select(variable2, estimate, lwr, upr, p_value, significance, trend, model) %>%
+    rename(variable = variable2)
+
+
+
 ## Now need to merge the dfs together
-param_estimates <- bind_rows(estimates_non_phylo, estimates_phylo, estimates_non_phylo_averaged)
[…]

### 2018-05-03 R/phylo_fxns.R
minor tweaks to uncertainty figure

@@ -115,11 +115,12 @@ extract_brain<-function(mod,term_to_extract="brain_residual"){
 
 plot_dist_parameter<-function(list_phy_models){
   df<-data.frame(`Feeding habitat generalism`=sapply(list_phy_models,extract_brain,term_to_extract="feeding_habitat_generalism"),
-                 `Clutch size (logged)`=sapply(list_phy_models,extract_brain,term_to_extract="clutch_size_logged"),
-                 `Diet generalism`=sapply(list_phy_models,extract_brain,term_to_extract="diet_generalism"))
+                 `log(Clutch size)`=sapply(list_phy_models,extract_brain,term_to_extract="clutch_size_logged"),
+                 `Diet generalism`=sapply(list_phy_models,extract_brain,term_to_extract="diet_generalism"),
+                 `Breeding habitat generalism`=sapply(list_phy_models,extract_brain,term_to_extract="breeding_habitat_generalism"))
                  #`Granivore`=sapply(list_phy_models,extract_brain,term_to_extract="granivoreYes"))
   df_long<-gather(df,key="Trait",value="Coefficient estimate")               
-  p<-ggplot(df_long,aes(x=`Coefficient estimate`,fill=Trait))+geom_density(alpha=0.5)+theme_bw()
+  p<-ggplot(df_long,aes(x=`Coefficient estimate`,fill=Trait))+geom_density(alpha=0.5)+theme_bw()+ylab("Density")
   pdf("figures/accounting_for_phylo_uncertainty.pdf")
   print(p)
   dev.off()

### 2018-05-03 scripts/results_script.R
tweaking results script and gathering figures for paper

@@ -4,7 +4,7 @@
 remake::dump_environment()
 
 ## total number of species included in models
-length(unique(analysis_data))
+nrow(analysis_data)
 
 ## summary of urbanization index
 summary(analysis_data$urban_median)
@@ -67,7 +67,7 @@ Figure2 <- ggplot(data=analysis_data, aes(log(urban_median))) +
                  fill="green", 
                  alpha = 0.3)+
   theme_classic()+
-  xlab("log(Response variable)")+
+  xlab("log(Urbanization index)")+
   ylab("Count")+
   geom_vline(data=species_values, aes(xintercept=urban_median), color='red', size=1.62, alpha=0.9)
 
@@ -91,10 +91,9 @@ my.file.copy(from = "figures/bird_urbanness_phylo.pdf",
 
 
 
-
+## Figure 4
 ## All parameter estimates
 ## A script which plots results for each of the four 'models'
-## only three for now though
 
 ## first make dfs for each separate model approach
 ### Non -phylo global model
@@ -282,39 +281,16 @@ estimates_phylo_averaged <- model_results %>%
 ## Now need to merge the dfs together
 param_estimates <- bind_rows(estimates_non_phylo, estimates_phylo, estimates_non_phylo_averaged, estimates_phylo_averaged)
 
+
+## Make figure which combines all model results into one
 p <- param_estimates %>%
-  arrange(model, estimate) %>%
-  mutate(order=row_number())
-
-ggplot(p, aes(x=order, y=estimate, color=trend))+
-    geom_point(aes(shape=significance), size=3)+
-    geom_errorbar(aes(ymin=lwr, ymax=upr, color=trend))+
-    facet_wrap(~model, scales="free")+
-    scale_x_continuous(breaks=p$order, labels=p$variable, expand=c(0,0))+
-    coord_flip()+
-    xlab("Parameter estimates")+
-    ylab("")+
-    theme_bw()+
-    geom_hline(yintercept=0, color="black")+
-    theme(panel.grid.major = element_blank(), 
-          panel.grid.minor = element_blank(),
-          axis.line = element_blank())+
-    guides(color=FALSE)+
-    scale_shape_manual(values=c(19, 8))+
-    guides(shape=FALSE)
-    
-ggsave(filename="C:/Users/CTC/Desktop/ex1.png", width=14, height=10, units="in", dpi=300)
-
-
-## OR, potentially this
[…]

### 2018-05-04 ms.Rmd
a few formatting things and some discussion started.

@@
 title: "A continuous urbanization index to assess phylogenetic adaptations in urban bird living"
-author: Corey T. Callaghan, John Wilshire, XXXX, XXXX, Richard T. Kingsford, William K. Cornwell
+
+author: | 
+  | Corey T.
+Callaghan^[Corresponding author: c.callaghan@unsw.edu.au] $^1$, Richard E.
+Major $^1$$^,$$^2$, John H.
+Wilshire $^1$, John M.
+Martin $^1$$^,$$^3$, Richard T.
+Kingsford $^1$, William K.
+Cornwell $^1$$^,$$^4$
+  | $^1$Center for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
 date: ""
@@
 link-citations: no
-csl: https://raw.githubusercontent.com/citation-style-language/styles/master/global-ecology-and-biogeography.csl
+csl: https://raw.githubusercontent.com/citation-style-language/styles/master/ecology-letters.csl
 bibliography: references.bib
+header-includes: 
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}
+  - \linenumbers
 ---
+
+\newpage
 
@@
 
-https://www.youtube.com/watch?v=9Gc4QTqslN4
+TBD
 
@@
 To date, a wide array of traits have been used to predict presence of birds in urban environments.
-For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have all been hypothesized to influence a birds' presence in urban environments.
+For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does; @conole2011functional], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have been hypothesized to influence a birds' presence in urban environments.
 Rather than a specific trait that predicts a species' ability to be found in urban environments, it is most likely a combination of traits.
-For instance, urban birds' presence in Jerusalem was predicted by a combination of traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-nesters, and widely distributed [@croci2008does].
+For instance, urban birds' presence in Jerusalem was predicted by traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-nesters, and widely distributed [@croci2008does].
 Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting one another.
-Indeed, brain size was found to be important by Møller and Erritzøe [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains] but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on species' response to urbanization.
-Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011makes with @moller2009successful & @croci2008does] and generalism or niche breadth [cf., @kark2007living with @evans2011makes].
+Indeed, residual brain size was found to be important by Mï¿½ller and Erritzï¿½e [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains] but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on species' response to urbanization.
+Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011makes with @moller2009successful & @croci2008does] and niche breadth [cf., @kark2007living with @evans2011makes].
 
@@
 Indeed, various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
-However, we unaware of any projects which use citizen science observations to assess species' traits in relation to urban environments over a large spatial and temporal scale.
+However, we are unaware of any projects which use citizen science observations to assess species' traits in relation to urban environments over a large spatial and temporal scale.
 
@@
 ## Bird observation data
-Bird observations originated from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset collected by citizen scientists.
+We collated observations from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists.
 Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird].
 We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018.
-This temporal scale was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see response variable below) over the given time frame in which bird observations were collected.
+This time-frame was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see below) over the given time frame in which bird observations were collected.
 
 We filtered the bird observations [@callaghan2017assessing; @la2014role] by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barrett2003new].
-For further information on protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types].
+For further information on eBird protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types].
 Further, we filtered the checklists to those with a distance travelled of < 5 km or an area covered which was < 500 Ha.
@@
 
-## Response variable - an urbanization index
+## Urbanization index
 We used VIIRS night-time lights [@elvidge2017viirs] as a proxy of a bird's association with urban habitat.
@@
 For the 637,482 checklists which met our criteria above, we used the checklists' associated spatial coordinates and Google Earth Engine [@gorelick2017google] to assign the average radiance value calculated from the VIIRS layer.
-For each location, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
+For each checklist, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
 Any radiance values which were negative were forced to 0.00001 [*sensu* @ou2015evaluation].
@@
 1).
-Our response variable, which served as an urbanization index, was the median for each species' given distribution of average radiance values.
+The median for each species' given distribution of average radiance values served as our response variable in our analyses.
 
-## Explanatory variables
+## Biological traits
 Explanatory variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological].
-Potential traits were chosen on the basis of their existing support and hypotheses in the current literature (e.g., [@mcclure1989characterizes; @kark2007living; @evans2011makes]).
-Traits were parsed in different manners, and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat).
+Potential traits (Appendix 1a) were chosen on the basis of their existing support and hypotheses in the current literature (e.g., [@mcclure1989characterizes; @kark2007living; @evans2011makes]).
+Traits were parsed in different manners (Appendix 1b), and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat).
 Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated.
+Preliminary analyses including IUCN status (exotic/native) showed that exotic birds were strongly associated with urbanization, but given the preferential selection by acclimatization societies for introduction of species with proven synanthropy [@blackburn2009avian], exotic species were omitted from the final analysis to avoid bias.
+Therefore, we excluded IUCN status as a trait given that the majority of our candidate species were of Least Concern.
 
-Only species which had complete data for the list of candidate traits were considered in the analyses.
+Only species which had complete data for the list of candidate traits were considered in the analyses, and we thereby ended with 477 species which met this criteria.
 
@@
 This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global].
+In order to assess whether there was a phylogenetic signal of the urbanization index used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
+This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the explanatory variables.
 
-In order to assess whether there was a phylogenetic signal of the urbanization index used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
+## Statistical analyses
+All models and analyses were conducted in the R computing environment [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse].
 
-### Taxonomic considerations
+Before modelling, the explanatory variables were investigated for multi-collinearity.
+We used the corrplot package [@corrplot2017] to show correlation among continuous variables (Supplementary Figure 1), and used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other (Supplementary Table 1).
+
+We employed two linear modelling approaches; one with and one without phylogenetic constraints.
+Our urbanization index was log-transformed in order to meet model assumptions.
+The explanatory variables used in the models were those listed in Appendix 1, with the exception of IUCN status.
+
+We weighted the linear regression models [@solon2015weighting] using a source of error associated with our urbanization indices.
+This error was the  proportion of the number of total observations of a species divided by the number of unique localities a species was observed from.
+This gave more weight in the model-fitting procedure to those species that had more observations in more unique locations.
+This was done based on the assumption that those species which had the most observations would have the most reliable urbanization indices.
+This approach also addresses biases in sampling of eBird data in which some birds can have a high number of observations, but from known roosting sites [e.g., @callaghanavian].
+
+### Non-phylogenetically constrained models
+Given that we hypothesized any of the potential candidate traits could significantly drive the response in our urbanization index, we first ran one global model including all 22 explanatory variables.
+We then used the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes.
+In order to confirm our linear model results, we employed a model-averaging approach [@grueber2011multimodel] in which we 'dredged' possible subsets of the global model.
+This was done using the dredge function in the MuMIn package [@barton2016mumin], and we allowed a total of 11 candidate predictors to be fit at any one time in order to avoid over-fitting.
+This left a total of 784,626 models which were assessed, and the model-averaged parameter estimates were calculated from the top models of those with ???AIC < 4.
+As with the global model, we present the standardized estimates which represent relative effect sizes
+
+### Phylogenetically constrained models
+Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package [@ho2014phylolm].
+Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.
+
+## Taxonomic considerations
 Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed.
 The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/].
-There were minor differences which existed in common names, English names, and recognized species among the three datasets.
+There were minor differences which existed in common names, scientific names, and recognized species among the three datasets.
 A taxonomic key was made which combined the three different taxonomies (Appendix 2).
@@
 
-## Models
-Before modelling, the explanatory variables (Appendix 1) were investigated for collinearity, and brain size and body size were highly correlated.
-Due to the affect that collinearity can have on modelling [@cade2015model], we were inclined to drop one of the two correlated explanatory variables.
-Given that others using a continuous response measure of urbanization have found brain size to have no evidence of species' response to urban environments [@evans2011makes], we eliminated this trait from consideration.
-In so doing, it also increased the number of species to be included in our models.
-
-We had two linear modelling approaches; one with and one without phylogenetic constraints.
-Our urbanization index was log-transformed in order to meet model assumptions.
-The explanatory variables used in the models were those listed in Appendix 1, with the exception of brain size.
-
-For both approaches, we employed a model-averaging technique [@grueber2011multimodel] in which we 'dredged' all possible subsets of the model [@barton2016mumin].
-The global model was weighted by a measure of error associated with the response variable.
-This was assessed as the proportion of the number of total observations of a species divided by the number of unique localities a species was observed from.
-This helps to give more weight in the model-fitting procedure to those species that had more observations in more unique locations (i.e., an assessment of reliability of the observations).
-The global model fit was assessed by averaging the R2 values [@nakagawa2013general; @nakagawa2017coefficient] for each model in the top-selected model-set.
-Additionally, we calculated the relative contribution of each explanatory variable [*sensu* @evans2011makes] by **XXXXXXXXXXXXXXXX**.
-The phylogenetically constrained model was fit using the 'phylolm' function from the phylolm package [@ho2014phylolm].
-In order to assess the model fit and relative effect sizes of explanatory variables in the phylogenetically constrained models we **XXXXXXXXXXXX**.
-
-All models and analyses were conducted in the R computing environment [@r2017coreteam] and relied heavily on the tidyverse environment [@wickham2017tidyverse].
-
 # RESULTS
@@
 The least recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
-The mean ± SD number of observations for a species was 10,250 ± 19,269.19.
+The mean ï¿½ SD number of observations for a species was 10,250 ï¿½ 19,269.19.
+Density plots of all 580 species' distributions of average radiance values are presented in Appendix 3.
 
-We ended with 569 candidate species (Appendix 3) from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data.
+We ended with 477 candidate species (Appendix 4) from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data.
 
 ## Urbanization index
-For the 569 species included in the models the mean ± SD urbanization index was 1.11 ± 2.17, with a range from 0.001 (Black Noddy and Black-naped Tern) to 20.15 (Laughing Dove; Fig. 2).
+For the 477 species included in the models the mean ± SD urbanization index was 0.91 ± 1.76, with a range from 0.008 (Black-backed Butcherbird) to 11.85 (Powerful Owl; Fig. 2).
 
 ### Phylogenetic signal
-There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; XXXXX SIGNIFICANCE TEST).
+There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; Table 1).
 
-**MAYBE DO SOME MEAN index by FAMILY PRESENTATIONS HERE**.
+**MAYBE DO SOME MEAN index by FAMILY PRESENTATIONS HERE????**
 
@@
 ### Non phylogenetic models
+The most important predictors of urbanization in the non-phylogenetic global model (Fig.
+4) were feeding habitat generalism (p < 0.001), breeding habitat generalism (p < 0.001), clutch size (p < 0.001), and diet generalism (p < 0.001).
+Urban birds were also positively significantly associated with agricultural habitats (p < 0.001) and migratory status (p < 0.001).
+Urban birds were significantly negatively associated with solitary nesting (p < 0.001), feeding in pairs and flocks (p < 0.001), being granivorous (p < 0.001), using forested habitats (p < 0.001), using grassland/shrubland habitats (p < 0.001), being insectivorous (p < 0.001), and feeding solitary (p < 0.001).
+
+When this global model was dredged and model-averaged, patterns remained similar (Fig.
+4).
+Clutch size (p < 0.001), diet generalism (p < 0.001), breeding habitat generalism (p < 0.001), and feeding habitat generalism (p < 0.001) all remained significant predictors.
+Additionally, agricultural birds were significantly associated with urbanization (p < 0.001).
+Urban birds remained significantly negatively associated with solitary nesting (p < 0.001), feeding in pairs and flocks (p < 0.001), being granivorous (p < 0.001), using forested habitats (p < 0.001), using grassland/shrubland habitats (p < 0.001), being insectivorous (p < 0.001), and feeding solitary (p < 0.001).
+
 ### Phylogenetically constrained models
+The most significant predictors of urbanization in the phylogenetically controlled global model (Fig.
+4) were feeding habitat generalism (p < 0.001), diet generalism (p < 0.001), clutch size (p < 0.001), brain residual size (p < 0.001), and breeding habitat generalism (p < 0.001).
+Granivores (p < 0.001), birds without any known nest aggregation (p < 0.001), insectivores (p < 0.001) and birds who are associated with grassland/shrubland habitat (p < 0.001) were significantly negatively associated with the urbanization index.
+
+When the phylogenetic global model was model-averaged, the patterns were similar (Fig.
+4).
+Feeding habitat generalism (p < 0.001), brain residual size (p < 0.001), clutch size (p < 0.001), and breeding habitat generalism (p < 0.001) remained significant predictors of urbanization index.
+Further, cooperative breeders (p < 0.001), range size (p < 0.001), and species which nest aggregation was unknown (p < 0.001) remained significantly negative predictors.
+
+### Differences in non-phylogenetic model and phylogenetic model
+The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the significance of individual predictors (Fig.
+4).
+However, brain residual size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig.
+5).
+Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger parameter estimates) in the phylogenetic model compared with the non-phylogenetic model (Fig.
+5).
+Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled model (Fig.
+4 & Fig.
+5).
+
+#### Phylogenetic uncertainty
+**WE ASSESSED PHYLOGENETIC UNCERTAINY AND FOUND THAT....**
 
 # DISCUSSION
+We found a significant effect of phylogenetic signal in our study, which contrasts with previous studies [@evans2011makes].
+Niche breadth, as measured by generalism, was significantly associated with our urbanization index across all analyses.
+Conversely, granivores and insectivores were consistently selected against.
+Residual brain mass was significantly important within our phylogenetic models, suggesting that within a clade, it is the species with the largest residual brain mass that are most readily observed in urban areas.
+A similar pattern was found for body size, although not significant.
+Our results suggests that urbanization selects against specialist species.
+
+Generalists are ultimately better prepared to thrive, survive, and adapt within urban areas.
+
+We provide the largest study to date using a continuous urbanization metric in order to assess biological and ecological traits which predict a species' adaptability to urban environments.
+Understanding specific species' traits which are associated with urban birds allows us to fully understand how to best preserve and maintain avian diversity in cities.
+We recommend that urban areas should focus on providing habitats which allow granivores, insectivores, and others to be found more readily in cities.
+
+## Acknowledgments
+We thank the > 6,000 eBirders who have contributed data in Australia.
+We also thank the dedicated eBird team at the Cornell Lab of Ornithology and the global network of reviewers.
+Thanks to Stephen Garnett and Glenn Ehmke for providing us with extent of occurrence estimates for each of the species in our analysis.
 
@@
 
-```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance values"}
+```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance values."}
 knitr::include_graphics("final figures for paper/Figure1.pdf")
@@
 
-```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a log-scale, of the response variable for the 569 species included in the analysis, with five example species highlighted."}
+```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a log-scale, of the response variable for the 477 species included in the analysis, with five example species highlighted."}
 knitr::include_graphics("final figures for paper/Figure2.pdf")
@@
 
-```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree of the birds yata yata yata...."}
+```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree of the birds with the tips colored based on the urbanization index.
+The urbanization index is also shown based on the size of the bar associated with a species."}
 knitr::include_graphics("final figures for paper/Figure3.pdf")
 ```
+
+```{r Figure 4, echo=FALSE, fig.cap="Parameter estimates for each of the four modelling approaches employed in the analysis.
+A star represents a significant effect.
+Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set."}
+knitr::include_graphics("final figures for paper/Figure4.pdf")
+```
+
+```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for parameter estimates from the non-phylogenetically constrained and phylogenetically constrained models."}
+knitr::include_graphics("final figures for paper/Figure5.pdf")
+```
+
+# Supplementary figures
+```{r Figure S1, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the modelling procedures."}
+knitr::include_graphics("figures/corrplot_of_continuous_variables.pdf")
+```
+
+```{r Figure S2, echo=FALSE, fig.cap="Phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
+knitr::include_graphics("figures/accounting_for_phylo_uncertainty.pdf")
+```
+
+# Supplementary tables
+```{r Table S1, echo=FALSE, fig.cap="Variance inflation factors for the explanatory variables used in the analysis."}
+knitr::include_graphics("tables/collinearity_investigation.pdf")
+```

### 2018-05-04 ms.Rmd
more formatting stuff

@@
 # Supplementary figures
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\thetable}{S\arabic{table}}%
+        \setcounter{figure}{0}
+        \renewcommand{\thefigure}{S\arabic{figure}}%
+     }
+     \beginsupplement
+
 ```{r Figure S1, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the modelling procedures."}

### 2018-05-07 ms.Rmd
adding ms to remake workflow

@@
 Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting one another.
-Indeed, residual brain size was found to be important by Mï¿½ller and Erritzï¿½e [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains] but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on species' response to urbanization.
+Indeed, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains] but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on species' response to urbanization.
 Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011makes with @moller2009successful & @croci2008does] and niche breadth [cf., @kark2007living with @evans2011makes].
@@
 A total of 637,482 checklists provided a total of 5,944,819 observations of 580 different species which had > 100 observations.
-The least recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
-The mean ï¿½ SD number of observations for a species was 10,250 ï¿½ 19,269.19.
+The least often recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
+The mean number of observations for a species was 10,250.
 Density plots of all 580 species' distributions of average radiance values are presented in Appendix 3.
@@
 ## Urbanization index
-For the 477 species included in the models the mean ± SD urbanization index was 0.91 ± 1.76, with a range from 0.008 (Black-backed Butcherbird) to 11.85 (Powerful Owl; Fig. 2).
+For the 477 species included in the models the mean ? SD urbanization index was 0.91 ? 1.76, with a range from 0.008 (Black-backed Butcherbird) to 11.85 (Powerful Owl; Fig. 2).
 
@@
 ```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance values."}
-knitr::include_graphics("final figures for paper/Figure1.pdf")
+knitr::include_graphics("finalFigs/Figure1.pdf")
 ```
@@
 ```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a log-scale, of the response variable for the 477 species included in the analysis, with five example species highlighted."}
-knitr::include_graphics("final figures for paper/Figure2.pdf")
+knitr::include_graphics("finalFigs/Figure2.pdf")
 ```
@@
 The urbanization index is also shown based on the size of the bar associated with a species."}
-knitr::include_graphics("final figures for paper/Figure3.pdf")
+knitr::include_graphics("finalFigs/Figure3.pdf")
 ```
@@
 Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set."}
-knitr::include_graphics("final figures for paper/Figure4.pdf")
+knitr::include_graphics("finalFigs/Figure4.pdf")
 ```
@@
 ```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for parameter estimates from the non-phylogenetically constrained and phylogenetically constrained models."}
-knitr::include_graphics("final figures for paper/Figure5.pdf")
+knitr::include_graphics("finalFigs/Figure5.pdf")
 ```

### 2018-05-07 ms.Rmd
working on manuscript creation workflow

--- 2018-05-04 ms.Rmd
@@
 The least recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
-The mean ï¿½ SD number of observations for a species was 10,250 ï¿½ 19,269.19.
+The mean ± SD number of observations for a species was 10,250 ± 19,269.19.
 Density plots of all 580 species' distributions of average radiance values are presented in Appendix 3.
@@
 ```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance values."}
-knitr::include_graphics("final figures for paper/Figure1.pdf")
+knitr::include_graphics("final figures & tables for ms/Figure1.pdf")
 ```
@@
 ```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a log-scale, of the response variable for the 477 species included in the analysis, with five example species highlighted."}
-knitr::include_graphics("final figures for paper/Figure2.pdf")
+knitr::include_graphics("final figures & tables for ms/Figure2.pdf")
 ```
@@
 The urbanization index is also shown based on the size of the bar associated with a species."}
-knitr::include_graphics("final figures for paper/Figure3.pdf")
+knitr::include_graphics("final figures & tables for ms/Figure3.pdf")
 ```
@@
 Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set."}
-knitr::include_graphics("final figures for paper/Figure4.pdf")
+knitr::include_graphics("final figures & tables for ms/Figure4.pdf")
 ```
@@
 ```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for parameter estimates from the non-phylogenetically constrained and phylogenetically constrained models."}
-knitr::include_graphics("final figures for paper/Figure5.pdf")
+knitr::include_graphics("final figures & tables for ms/Figure5.pdf")
+```
+
+# Tables
+```{r Table 1, echo=FALSE, warning=FALSE, message=FALSE}
+remake::dump_environment()
+
+library(phylosignal)
+   dplyr::select(analysis_data,body_size_logged,clutch_size_logged,response,range_size, #select appropriate rows
+          nest_generalism,diet_generalism,breeding_habitat_generalism,feeding_habitat_generalism,brain_residual) ->dd
+
+row.names(dd) <- analysis_data$binom #name rows so that it matches the tree
+dd$rand <- rnorm(dim(dd)[2]) #random numbers to test package
+p4d <- phylo4d(aus_bird_tree, dd) #create phylobase object
+
+ps <- phyloSignal(p4d,reps = 9999)
+
+Table1 <- ps$pvalue
+
+row.names(Table1) <- c("log(Body size)", "log(Clutch size)", "Urbanization index", "Range size (1000s km2)",
+                       "Nest generalism", "Diet generalism", "Breeding habitat generalism",
+                       "Feeding habitat generalism", "Brain residual", "Rand")
+
+library(kableExtra)
+
+kable(Table1, format = "latex", caption = "Results showing the significance of the phylogenetic signal analyses.", booktabs = T) %>%
+kable_styling(latex_options = c("striped", "hold_position"))
+
 ```
@@
 ```{r Figure S1, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the modelling procedures."}
-knitr::include_graphics("figures/corrplot_of_continuous_variables.pdf")
+knitr::include_graphics("final figures & tables for ms/FigureS1.pdf")
 ```
@@
 ```{r Figure S2, echo=FALSE, fig.cap="Phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
-knitr::include_graphics("figures/accounting_for_phylo_uncertainty.pdf")
+knitr::include_graphics("final figures & tables for ms/FigureS2.pdf")
 ```
@@
 # Supplementary tables
-```{r Table S1, echo=FALSE, fig.cap="Variance inflation factors for the explanatory variables used in the analysis."}
-knitr::include_graphics("tables/collinearity_investigation.pdf")
+```{r Table S1, echo=FALSE, warning=FALSE, message=FALSE}
+library(kableExtra)
+
+read_csv("Data/TableS1.csv")
+
+kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the explanatory variables used in the analysis.", booktabs = T) %>%
+kable_styling(latex_options = c("striped", "hold_position"))
+
 ```
+
+```{r Table S2, echo=FALSE, warning=FALSE, message=FALSE}
+remake::dump_environment()
+
+library(phylosignal)
+   dplyr::select(analysis_data,body_size_logged,clutch_size_logged,response,range_size, #select appropriate rows
+          nest_generalism,diet_generalism,breeding_habitat_generalism,feeding_habitat_generalism,brain_residual) ->dd
+
+row.names(dd) <- analysis_data$binom #name rows so that it matches the tree
+dd$rand <- rnorm(dim(dd)[2]) #random numbers to test package
+p4d <- phylo4d(aus_bird_tree, dd) #create phylobase object
+
+ps <- phyloSignal(p4d,reps = 9999)
+
+TableS2 <- ps$stat
+
+row.names(TableS2) <- c("log(Body size)", "log(Clutch size)", "Urbanization index", "Range size (1000s km2)",
+                       "Nest generalism", "Diet generalism", "Breeding habitat generalism",
+                       "Feeding habitat generalism", "Brain residual", "Rand")
+
+kable(TableS2, format = "latex", caption = "Summary statistics of the phylogenetic signal analysis.", booktabs = T) %>%
+kable_styling(latex_options = c("striped", "hold_position"))
+
+```

### 2018-05-07 R/non_phylo_analysis.R

@@ -101,6 +101,8 @@ collinearity_investigation_function <- function(global_model) {
   is.num <- sapply(collinearity_investigation, is.numeric)
   collinearity_investigation[is.num] <- lapply(collinearity_investigation[is.num], round, 2)
   
+  write_csv(collinearity_investigation, "Data/TableS1.csv")
+  
   pdf("tables/collinearity_investigation.pdf")
   grid.table(collinearity_investigation)
   dev.off()

### 2018-05-07 R/phylo_fxns.R

@@ -354,3 +354,52 @@ phy_v_non_phy<-function(global_model,phy_mod_rescaled){
   print(p)
   dev.off()
 }
+
+
+
+## phylosignal analysis
+phylosignal_analysis <- function(analysis_data, aus_bird_tree){
+  
+  require(phylosignal)
+  dplyr::select(analysis_data,body_size_logged,clutch_size_logged,response,range_size, #select appropriate rows
+                nest_generalism,diet_generalism,breeding_habitat_generalism,feeding_habitat_generalism,brain_residual) ->dd
+  
+  row.names(dd)<-analysis_data$binom #name rows so that it matches the tree
+  dd$rand<-rnorm(dim(dd)[2]) #random numbers to test package
+  p4d <- phylo4d(aus_bird_tree, dd) #create phylobase object
+  
+  ps<-phyloSignal(p4d,reps = 9999) #run calculation, p values a bit unstable at 999 reps
+  
+  
+  Table1 <- ps$pvalue
+  
+  row.names(Table1) <- c("log(Body size)", "log(Clutch size)", "Urbanization index", "Range size (1000s km2)",
+                         "Nest generalism", "Diet generalism", "Breeding habitat generalism",
+                         "Feeding habitat generalism", "Brain residual", "Rand")
+  
+  
+  TableS2 <- ps$stat
+  
+  row.names(TableS2) <- c("log(Body size)", "log(Clutch size)", "Urbanization index", "Range size (1000s km2)",
+                          "Nest generalism", "Diet generalism", "Breeding habitat generalism",
+                          "Feeding habitat generalism", "Brain residual", "Rand")
+  
+  write_csv(Table1, "Data/Table1.csv")
+  write_csv(TableS2, "Data/TableS2.csv")
+  
+  is.num <- sapply(Table1, is.numeric)
+  Table1[is.num] <- lapply(Table1[is.num], round, 2)
+  pdf("tables/phylosignal_pvalues.pdf")
+  grid.table(Table1)
+  dev.off()
+  
+  is.num <- sapply(TableS2, is.numeric)
+  TableS2[is.num] <- lapply(TableS2[is.num], round, 2)
+  pdf("tables/phylosignal_summary_stats.pdf")
+  grid.table(TableS2)
+  dev.off()
+  
+}
+
+
+

### 2018-05-07 scripts/results_script.R

@@ -1,4 +1,4 @@
-### A script for summary of the results and to create final figures using the final set of species
+### A script for summary of the results and to create final figures/tables using the final set of species
 
 ## first dump the remake environment
 remake::dump_environment()
@@ -39,7 +39,7 @@ Figure1 <- species_urban %>%
   theme(legend.position = c(0.26, 0.8))+
   theme(axis.text.x=element_text(hjust=0.7))
 
-pdf("final figures for paper/Figure1.pdf", width=5.5, height=4.5)
+pdf("final figures & tables for ms/Figure1.pdf", width=5.5, height=4.5)
 print(Figure1)
 dev.off()
 
@@ -71,7 +71,7 @@ Figure2 <- ggplot(data=analysis_data, aes(log(urban_median))) +
   ylab("Count")+
   geom_vline(data=species_values, aes(xintercept=urban_median), color='red', size=1.62, alpha=0.9)
 
-pdf("final figures for paper/Figure2.pdf", width=5.5, height=4.5)
+pdf("final figures & tables for ms/Figure2.pdf", width=5.5, height=4.5)
 print(Figure2)
 dev.off()
 
@@ -87,7 +87,7 @@ my.file.copy <- function(from, to) {
 }
 
 my.file.copy(from = "figures/bird_urbanness_phylo.pdf",
-               to = "final figures for paper/Figure3.pdf")
+               to = "final figures & tables for ms/Figure3.pdf")
 
 
 
@@ -314,7 +314,7 @@ Figure4 <- ggplot(p, aes(x=fct_inorder(variable), y=estimate, group=model, color
     theme(axis.title.y=element_text(size=16))+
     theme(axis.text.y=element_text(size=7.5))
     
-  pdf("final figures for paper/Figure4.pdf", width=9.45, height=6.4)
+  pdf("final figures & tables for ms/Figure4.pdf", width=9.45, height=6.4)
   print(Figure4)
   dev.off()
 
@@ -329,9 +329,30 @@ my.file.copy <- function(from, to) {
 }
 
 my.file.copy(from = "figures/phy_v_non_phy.pdf",
-             to = "final figures for paper/Figure5.pdf")
+             to = "final figures & tables for ms/Figure5.pdf")
   
   
-  
+##########################################
+###### Supplementary Figures #############
+##########################################
+
+## Figure S1
+my.file.copy(from = "figures/corrplot_of_continuous_variables.pdf",
+             to = "final figures & tables for ms/FigureS1.pdf")
+
+
+## Figure S2
[…]

### 2018-05-07 ms.Rmd
cleaning TableS1

--- 2018-05-07 ms.Rmd
@@
 
+# Tables
+```{r Table 1, echo=FALSE, warning=FALSE, message=FALSE}
+library(readr)
+Table1 <- read_csv("Data/Table1.csv")
+
+library(kableExtra)
+
+kable(Table1, format = "latex", caption = "Results showing the significance of the phylogenetic signal analyses.", booktabs = T) %>%
+kable_styling(latex_options = c("striped", "hold_position"))
+
+```
+
 # Supplementary figures
@@
 ```{r Figure S1, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the modelling procedures."}
-knitr::include_graphics("figures/corrplot_of_continuous_variables.pdf")
+knitr::include_graphics("finalFigs/FigureS1.pdf")
 ```
@@
 ```{r Figure S2, echo=FALSE, fig.cap="Phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
-knitr::include_graphics("figures/accounting_for_phylo_uncertainty.pdf")
+knitr::include_graphics("finalFigs/FigureS2.pdf")
 ```
@@
 # Supplementary tables
-```{r Table S1, echo=FALSE, fig.cap="Variance inflation factors for the explanatory variables used in the analysis."}
-knitr::include_graphics("tables/collinearity_investigation.pdf")
+```{r Table S1, echo=FALSE, warning=FALSE, message=FALSE}
+library(kableExtra)
+
+collinearity_investigation <- read_csv("Data/TableS1.csv")
+collinearity_investigation <- collinearity_investigation[-1,]
+
+kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the explanatory variables used in the analysis.", booktabs = T) %>%
+kable_styling(latex_options = c("striped", "hold_position"))
+
 ```
+
+```{r Table S2, echo=FALSE, warning=FALSE, message=FALSE}
+library(kableExtra)
+
+TableS2 <- read_csv("Data/TableS2.csv")
+
+kable(TableS2, format = "latex", caption = "Summary statistics of the phylogenetic signal analysis.", booktabs = T) %>%
+kable_styling(latex_options = c("striped", "hold_position"))
+```

### 2018-05-07 ms.Rmd
some latex formatting

@@
 output:
-  pdf_document: null
-  bookdown::word_document2: null
-  bookdown::tufte_html2:
-    number_sections: no
-    toc: no
-  html_document:
-    df_print: paged
-    toc: no
+  rmarkdown::pdf_document:
+    fig_caption: yes    
+    includes:  
+      in_header: preamble-latex.tex
 fontsize: 12pt
@@
 bibliography: references.bib
-header-includes: 
-  - \usepackage{setspace}\doublespacing
-  - \usepackage{lineno}
-  - \linenumbers
+
 ---
@@
 
+\newpage
+
 # Figures
@@
 
+\newpage
+
 # Tables
+
 ```{r Table 1, echo=FALSE, warning=FALSE, message=FALSE}
@@
 kable(Table1, format = "latex", caption = "Results showing the significance of the phylogenetic signal analyses.", booktabs = T) %>%
-kable_styling(latex_options = c("striped", "hold_position"))
+kable_styling(latex_options = c("hold_position"))
 
 ```
+
+\newpage
 
@@
 
+\newpage
+
 # Supplementary tables
@@
 kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the explanatory variables used in the analysis.", booktabs = T) %>%
-kable_styling(latex_options = c("striped", "hold_position"))
+kable_styling(latex_options = c("hold_position"))
 
@@
 kable(TableS2, format = "latex", caption = "Summary statistics of the phylogenetic signal analysis.", booktabs = T) %>%
-kable_styling(latex_options = c("striped", "hold_position"))
+kable_styling(latex_options = c("hold_position"))
 ```

### 2018-05-07 R/phylo_fxns.R
minor change

@@ -368,34 +368,34 @@ phylosignal_analysis <- function(analysis_data, aus_bird_tree){
   dd$rand<-rnorm(dim(dd)[2]) #random numbers to test package
   p4d <- phylo4d(aus_bird_tree, dd) #create phylobase object
   
-  ps<-phyloSignal(p4d,reps = 9999) #run calculation, p values a bit unstable at 999 reps
+  ps <- phyloSignal(p4d,reps = 9999) #run calculation, p values a bit unstable at 999 reps
   
   
-  Table1 <- ps$pvalue
+  Table1 <- ps$stat
   
   row.names(Table1) <- c("log(Body size)", "log(Clutch size)", "Urbanization index", "Range size (1000s km2)",
                          "Nest generalism", "Diet generalism", "Breeding habitat generalism",
                          "Feeding habitat generalism", "Brain residual", "Rand")
   
   
-  TableS2 <- ps$stat
+  TableS2 <- ps$pvalue
   
   row.names(TableS2) <- c("log(Body size)", "log(Clutch size)", "Urbanization index", "Range size (1000s km2)",
                           "Nest generalism", "Diet generalism", "Breeding habitat generalism",
                           "Feeding habitat generalism", "Brain residual", "Rand")
   
-  write_csv(Table1, "Data/Table1.csv")
-  write_csv(TableS2, "Data/TableS2.csv")
+  write.csv(Table1, "Data/Table1.csv")
+  write.csv(TableS2, "Data/TableS2.csv")
   
   is.num <- sapply(Table1, is.numeric)
   Table1[is.num] <- lapply(Table1[is.num], round, 2)
-  pdf("tables/phylosignal_pvalues.pdf")
+  pdf("tables/phylosignal_summary_stats.pdf")
   grid.table(Table1)
   dev.off()
   
   is.num <- sapply(TableS2, is.numeric)
   TableS2[is.num] <- lapply(TableS2[is.num], round, 2)
-  pdf("tables/phylosignal_summary_stats.pdf")
+  pdf("tables/phylosignal_pvalues.pdf")
   grid.table(TableS2)
   dev.off()
   

### 2018-05-07 ms.Rmd
really close to having the table formatted correctly... but not quite!

@@
 library(readr)
-Table1 <- read_csv("Data/Table1.csv")
+Table1a <- read_csv("Data/Table1.csv")
+colnames(Table1a) <- gsub("X1", "Variable", colnames(Table1a))
+is.num <- sapply(Table1a, is.numeric)
+Table1a[is.num] <- lapply(Table1a[is.num], round, 3)
+
+Table1b <- read_csv("Data/TableS2.csv")
+colnames(Table1b) <- gsub("X1", "Variable", colnames(Table1b))
+is.num <- sapply(Table1b, is.numeric)
+Table1b[is.num] <- lapply(Table1b[is.num], round, 3)
+
+grouped <- rbind(cbind(Value="Statistic", Table1a), cbind(Value="p-values", Table1b))
+tgrp <- xtabs(cbind(Cmean, I, K, K.star, Lambda) ~ Value+Variable, grouped)
+
+table <- ftable(aperm(tgrp, c(3, 1, 2)), row.vars=c("Variable", "Value"))
 
@@
 
-kable(Table1, format = "latex", caption = "Results showing the significance of the phylogenetic signal analyses.", booktabs = T) %>%
-kable_styling(latex_options = c("hold_position"))
+kable(table, format = "latex", caption = "Results showing the significance of the phylogenetic signal analyses.")
 
@@
 ```
-
-```{r Table S2, echo=FALSE, warning=FALSE, message=FALSE}
-library(kableExtra)
-
-TableS2 <- read_csv("Data/TableS2.csv")
-
-kable(TableS2, format = "latex", caption = "Summary statistics of the phylogenetic signal analysis.", booktabs = T) %>%
-kable_styling(latex_options = c("hold_position"))
-```

### 2018-05-07 scripts/results_script.R
updated results script and some figures

@@ -39,7 +39,7 @@ Figure1 <- species_urban %>%
   theme(legend.position = c(0.26, 0.8))+
   theme(axis.text.x=element_text(hjust=0.7))
 
-pdf("final figures & tables for ms/Figure1.pdf", width=5.5, height=4.5)
+pdf("finalFigs/Figure1.pdf", width=5.5, height=4.5)
 print(Figure1)
 dev.off()
 
@@ -71,7 +71,7 @@ Figure2 <- ggplot(data=analysis_data, aes(log(urban_median))) +
   ylab("Count")+
   geom_vline(data=species_values, aes(xintercept=urban_median), color='red', size=1.62, alpha=0.9)
 
-pdf("final figures & tables for ms/Figure2.pdf", width=5.5, height=4.5)
+pdf("finalFigs/Figure2.pdf", width=5.5, height=4.5)
 print(Figure2)
 dev.off()
 
@@ -87,7 +87,7 @@ my.file.copy <- function(from, to) {
 }
 
 my.file.copy(from = "figures/bird_urbanness_phylo.pdf",
-               to = "final figures & tables for ms/Figure3.pdf")
+               to = "finalFigs/Figure3.pdf")
 
 
 
@@ -138,7 +138,7 @@ estimates_non_phylo <- tidy(std.mod) %>%
                    "Feeding aggregation \n (solitary)")) %>%
   arrange(estimate) %>%
   mutate(trend=ifelse(.$estimate >0, "positive", "negative")) %>%
-  dplyr::select(term2, estimate, lwr, upr, p.value, significance, trend) %>%
+  dplyr::select(term2, estimate, lwr, upr, statistic, p.value, significance, trend) %>%
   rename(variable=term2) %>%
   rename(p_value=p.value) %>%
   mutate(model = "Non-phylo global model")
@@ -150,8 +150,12 @@ summary <- readRDS("Data/dredged_model_summary_results.rds")
 p_values <- data.frame(p_value=summary$coefmat.full[,5]) %>%
   rownames_to_column("variable")
 
+statistic <- data.frame(statistic=summary$coefmat.full[,4]) %>%
+  rownames_to_column("variable")
+
 estimates_non_phylo_averaged <- model_results %>%
     inner_join(., p_values, by="variable") %>%
+    inner_join(., statistic, by="variable") %>%
     filter(variable != "(Intercept)") %>%
     droplevels() %>%
     arrange(desc(estimate)) %>%
[…]

### 2018-05-07 scripts/results_script.R
tables1 is called with kable now in ms

@@ -354,14 +354,5 @@ my.file.copy(from = "figures/corrplot_of_continuous_variables.pdf",
 my.file.copy(from = "figures/accounting_for_phylo_uncertainty.pdf",
              to = "finalFigs/FigureS2.pdf")
 
-
-
-##########################################
-###### Supplementary Tables ##############
-##########################################
-
-## Table S1
-my.file.copy(from = "tables/collinearity_investigation.pdf",
-             to = "finalFigs/TableS1.pdf")
   
   

### 2018-05-07 ms.Rmd
use pander instead of kable
adding statistics and proper p-values to model results section.

@@
 Our first objective was to investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
-Second, we developed models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using continuous and categorical explanatory variables.
+Second, we developed models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using continuous and categorical predictor variables.
 These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth.
@@
 ## Biological traits
-Explanatory variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological].
+Predictor variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological].
 Potential traits (Appendix 1a) were chosen on the basis of their existing support and hypotheses in the current literature (e.g., [@mcclure1989characterizes; @kark2007living; @evans2011makes]).
@@
 This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global].
-In order to assess whether there was a phylogenetic signal of the urbanization index used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
-This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the explanatory variables.
+In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic, @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
+This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.
 
 ## Statistical analyses
-All models and analyses were conducted in the R computing environment [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse].
+All models and analyses were conducted in R [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse].
 
-Before modelling, the explanatory variables were investigated for multi-collinearity.
+Before modelling, the predictor variables were investigated for multi-collinearity.
 We used the corrplot package [@corrplot2017] to show correlation among continuous variables (Supplementary Figure 1), and used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other (Supplementary Table 1).
@@
 Our urbanization index was log-transformed in order to meet model assumptions.
-The explanatory variables used in the models were those listed in Appendix 1, with the exception of IUCN status.
+The predictor variables used in the models were those listed in Appendix 1, with the exception of IUCN status.
 
@@
 ### Non-phylogenetically constrained models
-Given that we hypothesized any of the potential candidate traits could significantly drive the response in our urbanization index, we first ran one global model including all 22 explanatory variables.
+Given that we hypothesized any of the potential candidate traits could significantly drive the response in our urbanization index, we first ran one global model including all 22 predictor variables.
 We then used the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes.
@@
 
-We ended with 477 candidate species (Appendix 4) from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data.
+We ended with 477 candidate species from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data (Appendix 4).
 
@@
 The most important predictors of urbanization in the non-phylogenetic global model (Fig.
-4) were feeding habitat generalism (p < 0.001), breeding habitat generalism (p < 0.001), clutch size (p < 0.001), and diet generalism (p < 0.001).
-Urban birds were also positively significantly associated with agricultural habitats (p < 0.001) and migratory status (p < 0.001).
-Urban birds were significantly negatively associated with solitary nesting (p < 0.001), feeding in pairs and flocks (p < 0.001), being granivorous (p < 0.001), using forested habitats (p < 0.001), using grassland/shrubland habitats (p < 0.001), being insectivorous (p < 0.001), and feeding solitary (p < 0.001).
+4) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006).
+Urban birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and migratory status (t = 2.15, p = 0.032).
+Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitary (t = -2.90, p = 0.004).
 
@@
 4).
-Clutch size (p < 0.001), diet generalism (p < 0.001), breeding habitat generalism (p < 0.001), and feeding habitat generalism (p < 0.001) all remained significant predictors.
-Additionally, agricultural birds were significantly associated with urbanization (p < 0.001).
-Urban birds remained significantly negatively associated with solitary nesting (p < 0.001), feeding in pairs and flocks (p < 0.001), being granivorous (p < 0.001), using forested habitats (p < 0.001), using grassland/shrubland habitats (p < 0.001), being insectivorous (p < 0.001), and feeding solitary (p < 0.001).
+Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors.
+Additionally, agricultural birds were significantly associated with urbanization (z = 3.40, p = 0.001).
+Urban birds remained significantly negatively associated with feeding solitary (z = 3.39, p = 0.001), being insectivorous (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), feeding in pairs and flocks (z = 3.16, p = 0.002), being granivorous (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).
 
@@
 The most significant predictors of urbanization in the phylogenetically controlled global model (Fig.
-4) were feeding habitat generalism (p < 0.001), diet generalism (p < 0.001), clutch size (p < 0.001), brain residual size (p < 0.001), and breeding habitat generalism (p < 0.001).
-Granivores (p < 0.001), birds without any known nest aggregation (p < 0.001), insectivores (p < 0.001) and birds who are associated with grassland/shrubland habitat (p < 0.001) were significantly negatively associated with the urbanization index.
+4) were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism ( t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005).
+Nest generlaism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding species (t = -3.25, p = 0.001) and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.
 
@@
 4).
-Feeding habitat generalism (p < 0.001), brain residual size (p < 0.001), clutch size (p < 0.001), and breeding habitat generalism (p < 0.001) remained significant predictors of urbanization index.
-Further, cooperative breeders (p < 0.001), range size (p < 0.001), and species which nest aggregation was unknown (p < 0.001) remained significantly negative predictors.
+Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significant predictors of urbanization index.
+Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization.
 
@@
 library(kableExtra)
-
-kable(table, format = "latex", caption = "Results showing the significance of the phylogenetic signal analyses.")
+library(pander)
+pander(table, caption = "Results showing the significance of the phylogenetic signal analyses.")
 
@@
 
-kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the explanatory variables used in the analysis.", booktabs = T) %>%
+kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the predictor variables used in the analysis.", booktabs = T) %>%
 kable_styling(latex_options = c("hold_position"))

### 2018-05-07 ms.Rmd
some rubbish discussion coming into shape...

@@
 # DISCUSSION
-We found a significant effect of phylogenetic signal in our study, which contrasts with previous studies [@evans2011makes].
-Niche breadth, as measured by generalism, was significantly associated with our urbanization index across all analyses.
+Our continuous urbanization index showed a significant phylogenetic signal [@revell2008phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence, @ibanez2017global, @sol2017urbanisation, @la2018phylogenetic].
+This contrasts with previous studies [@evans2011makes], but is likely a result of sample size considerations.
+Niche breadth, as measured by generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban].
 Conversely, granivores and insectivores were consistently selected against.
-Residual brain mass was significantly important within our phylogenetic models, suggesting that within a clade, it is the species with the largest residual brain mass that are most readily observed in urban areas.
+Residual brain mass was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade, the species with the largest residual brain mass are most readily observed in urban areas.
 A similar pattern was found for body size, although not significant.
-Our results suggests that urbanization selects against specialist species.
+
+Traditionally, the literature which aims to associate avian species' traits with urbanization has been limited, and as such has lead to less generalizable patterns.
+The majority of studies in this space have focused on discrete classifications of species into adapters, avoiders, and exploiters [e.g., ], or even binary classification based on presence/absence in cities or towns [e.g., ].
+We highlight that our analysis used a total of 477 species (Appendix 4), which is greater than many previous studies.
+
+ Our results suggest that urbanization selects against specialist species.
 
@@
 We provide the largest study to date using a continuous urbanization metric in order to assess biological and ecological traits which predict a species' adaptability to urban environments.
-Understanding specific species' traits which are associated with urban birds allows us to fully understand how to best preserve and maintain avian diversity in cities.
+
+Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with rapidly expanding remote sensing technology [@gorelick2017google].
+
+ is rapidly expanding our understanding of urbanization impacts on birds [e.g., ] Understanding specific species' traits which are associated with urban birds allows us to fully understand how to best preserve and maintain avian diversity in cities.
 We recommend that urban areas should focus on providing habitats which allow granivores, insectivores, and others to be found more readily in cities.
+
+Our study looked over a broad spatial and temporal scale to identify a novel approach to quantifying a continuous urbanization index for birds.
+We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global], but migratory status was considered as a predictor.
+We highlight that future studies, especially within migrant-dominated landscapes (i.e., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [CITATIONS].
+Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2017non]).
+Future studies should aim to (1) broaden the spatial scale, (2) investigate the temporal changes in a species' urbanization indices [@la2014role, @lepczyk2017global], and (3) investigate the role that city size plays on selected traits and phylogenetic diversity [@sol2017urbanisation, @batary2017non].
+
+Ultimately, we highlight the applicability of broad-scale empirical data to assess species' traits which predict presence and adaptability in urban environments.
+Our results suggest that generalist species are better prepared to forage and survive in the ecologically novel aspects of urban environments [CITATIONS].
+Australia is a relatively young country and fully understanding how birds respond to urban development is critical to inform future management.
+Urban avian biodiversity could be increased by improving the suitability of foraging for granivores and insectivores, especially important given the recent noted collapse of insects in other parts of the world [CITATION].
+Furthermore, future studies should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved modelling performance.
 

### 2018-05-07 ms.Rmd
slowly adding structure to discussion (very slowly)

@@
 Cornwell $^1$$^,$$^4$
-  | $^1$Center for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
+  | $^1$Centre for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
 date: ""
@@
 
-Traditionally, the literature which aims to associate avian species' traits with urbanization has been limited, and as such has lead to less generalizable patterns.
-The majority of studies in this space have focused on discrete classifications of species into adapters, avoiders, and exploiters [e.g., ], or even binary classification based on presence/absence in cities or towns [e.g., ].
-We highlight that our analysis used a total of 477 species (Appendix 4), which is greater than many previous studies.
+**A PARAGRAPH ON SPECIFIC TRAITS AND CITATIONS WHICH CONFIRM OR DENY THEM**
 
- Our results suggest that urbanization selects against specialist species.
+ Fully understanding specific species' traits which are associated with urban birds allows us to fully understand how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].
+However, previous studies which have assessed this have generally focused on discrete classifications of species based on their responses to urbanization, limiting the applicability of results found.
+We further highlight that our analysis used a total of 477 species (Appendix 4), greater than many previous studies, ultimately allowing for more generalizable results.
+The aforementioned differences in our study compared with others is likely a result of our increased sample size, as well as our use of a continuous measure of urbanization.
 
-Generalists are ultimately better prepared to thrive, survive, and adapt within urban areas.
+Our results suggest that urbanization selects against specialist species, granivores, and insectivores.
+As such, future urban development should focus on maintaining a high diversity of habitat (CITATIONS) both for foraging (CITATION) and nesting (CITATIONS).
+Further, a focus should be on providing habitat for insects (CITATION), which will in turn increase the presence of insectivores in urban environments.
 
-We provide the largest study to date using a continuous urbanization metric in order to assess biological and ecological traits which predict a species' adaptability to urban environments.
+Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advancing remote sensing technology [@gorelick2017google].
+We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds.
+We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global], but migratory status was considered as a predictor.
+We highlight that future studies, especially within migrant-dominated landscapes (i.e., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call].
+Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2017non]).
 
-Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with rapidly expanding remote sensing technology [@gorelick2017google].
-
- is rapidly expanding our understanding of urbanization impacts on birds [e.g., ] Understanding specific species' traits which are associated with urban birds allows us to fully understand how to best preserve and maintain avian diversity in cities.
-We recommend that urban areas should focus on providing habitats which allow granivores, insectivores, and others to be found more readily in cities.
-
-Our study looked over a broad spatial and temporal scale to identify a novel approach to quantifying a continuous urbanization index for birds.
-We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global], but migratory status was considered as a predictor.
-We highlight that future studies, especially within migrant-dominated landscapes (i.e., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [CITATIONS].
-Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2017non]).
-Future studies should aim to (1) broaden the spatial scale, (2) investigate the temporal changes in a species' urbanization indices [@la2014role, @lepczyk2017global], and (3) investigate the role that city size plays on selected traits and phylogenetic diversity [@sol2017urbanisation, @batary2017non].
-
-Ultimately, we highlight the applicability of broad-scale empirical data to assess species' traits which predict presence and adaptability in urban environments.
-Our results suggest that generalist species are better prepared to forage and survive in the ecologically novel aspects of urban environments [CITATIONS].
+Ultimately, we provide the largest study to date using a continuous urbanization metric in order to assess biological and ecological traits which predict a species' adaptability to urban environments.
+In doing so, we highlight the applicability of broad-scale empirical data to assess species' traits which predict presence and adaptability in urban environments.
+Our results suggest that generalist species are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
 Australia is a relatively young country and fully understanding how birds respond to urban development is critical to inform future management.

### 2018-05-07 ms.Rmd
tightening up the discussion a little bit

@@
 Our continuous urbanization index showed a significant phylogenetic signal [@revell2008phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence, @ibanez2017global, @sol2017urbanisation, @la2018phylogenetic].
-This contrasts with previous studies [@evans2011makes], but is likely a result of sample size considerations.
-Niche breadth, as measured by generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban].
-Conversely, granivores and insectivores were consistently selected against.
-Residual brain mass was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade, the species with the largest residual brain mass are most readily observed in urban areas.
-A similar pattern was found for body size, although not significant.
+This contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of sample size considerations.
+Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban].
+Conversely, granivores and insectivores were consistently selected against in urban areas.
+Residual brain mass, and body size to an extent, was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade, the species with the largest residual brain mass and large body size are most suited to surviving in urban areas.
 
-**A PARAGRAPH ON SPECIFIC TRAITS AND CITATIONS WHICH CONFIRM OR DENY THEM**
+Our finding that ecological generalists are more readily observed in urban areas confirms that of other work [@mckinney2006urbanization, @bonier2007urban, @moller2009successful].
+We also confirm work which has found a relationship between annual fecundity and urbanization in birds [@croci2008does, @moller2009successful].
+While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living, @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful, @maklakov2011brains, @moller2015brain].
+Furthermore, we do confirm that urbanization appears to select against insectivorous species [@chace2006urban, @evans2011makes].
+While others have found that urbanization supports granivorous species [@chace2006urban, @kark2007living, @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [@fuller2008garden].
+We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@kark2007living].
 
- Fully understanding specific species' traits which are associated with urban birds allows us to fully understand how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].
-However, previous studies which have assessed this have generally focused on discrete classifications of species based on their responses to urbanization, limiting the applicability of results found.
-We further highlight that our analysis used a total of 477 species (Appendix 4), greater than many previous studies, ultimately allowing for more generalizable results.
-The aforementioned differences in our study compared with others is likely a result of our increased sample size, as well as our use of a continuous measure of urbanization.
-
-Our results suggest that urbanization selects against specialist species, granivores, and insectivores.
-As such, future urban development should focus on maintaining a high diversity of habitat (CITATIONS) both for foraging (CITATION) and nesting (CITATIONS).
-Further, a focus should be on providing habitat for insects (CITATION), which will in turn increase the presence of insectivores in urban environments.
+Clearly, many traits have been used to identify a species' adaptability with urban-living (e.g., Appendix 1), but often times results differ among studies.
+Fully understanding the suite of specific species' traits which are associated with urban birds allows for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].
+As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization.
+Our analysis used a total of 477 species (Appendix 4), more than any similar study, to the best of our knowledge.
+Furthermore, our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done).
+The aforementioned differences in our study's results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
+Our results suggest that urbanization selects against specialist species [@sorace2009distribution], granivores, and insectivores.
+As such, future urban development aimed at increasing avian diversity should focus on maintaining a high diversity of habitat [@melles2003urban] both for foraging [@shochat2004linking], and nesting [@fernandez2001habitat].
+A species-specific approach to assess particular resources necessary for specialist species may be necessary.
+Lastly, a focal point should be on increasing insect abundance and diversity in urban environments (CITATION), which will in turn increase the presence of insectivorours bird species in urban environments.
 
@@
 We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds.
-We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global], but migratory status was considered as a predictor.
+We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global, @katewaleffect], but migratory status was considered as a predictor.
 We highlight that future studies, especially within migrant-dominated landscapes (i.e., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call].
-Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2017non]).
+Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2018non]).
 
@@
 In doing so, we highlight the applicability of broad-scale empirical data to assess species' traits which predict presence and adaptability in urban environments.
-Our results suggest that generalist species are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
-Australia is a relatively young country and fully understanding how birds respond to urban development is critical to inform future management.
-Urban avian biodiversity could be increased by improving the suitability of foraging for granivores and insectivores, especially important given the recent noted collapse of insects in other parts of the world [CITATION].
-Furthermore, future studies should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved modelling performance.
+Our results suggest that generalist species, with large relative brain size, are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
+Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management.
+Urban avian biodiversity could be increased by improving the suitability of foraging for granivores and insectivores, especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more].
+Furthermore, future studies should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance.
 

### 2018-05-08 scripts/results_script.R
added export of 'model results' to csv for appendix for paper

@@ -285,11 +285,11 @@ estimates_phylo_averaged <- model_results %>%
     dplyr::select(variable2, estimate, lwr, upr, statistic, p_value, significance, trend, model) %>%
     rename(variable = variable2)
 
-
-
 ## Now need to merge the dfs together
 param_estimates <- bind_rows(estimates_non_phylo, estimates_phylo, estimates_non_phylo_averaged, estimates_phylo_averaged)
 
+## write model results for an appendix
+write_csv(param_estimates, "H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/Appendices/Appendix 5/Appendix5.csv")
 
 ## Make figure which combines all model results into one
 p <- param_estimates %>%

### 2018-05-14 scripts/Appendix_5-model_output_results.Rmd
R markdown doc to make Appendix 5 - raw model outputs

---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
```

## Model: Non-phylogenetically controlled global model results

```{r Non-phylogenetically controlled global model results, echo=FALSE, message=FALSE}
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
summary(global_model)
```

## Model: Non-phylogenetically controlled model-averaged results

```{r Non-phylogenetically controlled model-averaged results, echo=FALSE, message=FALSE}
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
summary <- readRDS("Data/dredged_model_summary_results.rds")
summary(summary)
```

## Model: Phylogenetically controlled global model results

```{r Phylogenetically controlled global model results, echo=FALSE, message=FALSE}
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
summary(phy_mod_rescaled)
```

## Model: Phylogenetically controlled model-averaged results

```{r Phylogenetically controlled model-averaged results, echo=FALSE, message=FALSE}
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
summary2 <- readRDS("Data/PHYLO_dredged_model_summary_results.rds")
summary(summary2)
```

### 2018-05-14 ms.Rmd
A little work on results and discussion

@@
 
-Here, we improve upon current methods of assessing biological traits which predicts urbanization in birds by using a series of publicly available datasets available for Australian birds.
-Importantly, we developed and used a continuous urbanization index as a response variable in the analyses.
-We had two objectives.
-Our first objective was to investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
-Second, we developed models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using continuous and categorical predictor variables.
+Here, we improve upon current methods of assessing biological traits which predicts urbanization in birds by using a series of publicly available datasets.
+We developed and used a continuous urbanization index as a response variable in the analyses.
+We had two objectives: (1) investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes], and (2) develop models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using a suite of continuous and categorical predictor variables.
 These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth.
@@
 ### Phylogenetic signal
-There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; Table 1).
-
-**MAYBE DO SOME MEAN index by FAMILY PRESENTATIONS HERE????**
+There was a distinct phylogenetic signal in our urbanization index measure (Fig.
+3; Table 1), as it was significantly different than random for 4 out of the 5 indices.
+Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1).
+The result for body size-- $k=3.3$ --is among the strongest ever recorded measures of phylogenetic signal in the literature (see REFS) and for $k$ the statistic suggests much stronger signal than expected from Brownian motion.
+We found weaker signal for behavioral traits compared to morphological traits.
+For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 1).
 
 ## Model results
-### Non phylogenetic models
+### Non-phylogenetic models
 The most important predictors of urbanization in the non-phylogenetic global model (Fig.
@@
 
-### Differences in non-phylogenetic model and phylogenetic model
+### Differences in non-phylogenetic and phylogenetic models
 The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the significance of individual predictors (Fig.
@@
 5).
-Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled model (Fig.
+Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled models (Fig.
 4 & Fig.
@@
 
-#### Phylogenetic uncertainty
-**WE ASSESSED PHYLOGENETIC UNCERTAINY AND FOUND THAT....**
+# DISCUSSION
+Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban].
+Conversely, granivores and insectivores were consistently negatively associated with urbanization.
+Furthermore, we found a significant phylogenetic signal [@revell2008phylogenetic] of the urbanization index in our analysis, suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence, @ibanez2017global, @sol2017urbanisation, @la2018phylogenetic].
+This contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our increased sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
+Our results demonstrate that considering phylogeny in regards to this question is important.
 
-# DISCUSSION
-Our continuous urbanization index showed a significant phylogenetic signal [@revell2008phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence, @ibanez2017global, @sol2017urbanisation, @la2018phylogenetic].
-This contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of sample size considerations.
-Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban].
-Conversely, granivores and insectivores were consistently selected against in urban areas.
-Residual brain mass, and body size to an extent, was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade, the species with the largest residual brain mass and large body size are most suited to surviving in urban areas.
+The effect of treating species as non-independent by considering phylogeny in the context of the linear model was apparent.
+For example, there was an increase in parameter estimates of the most important parameters (e.g., clutch size and feeding habitat generalism; Figures 4 and 5), and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
+S2).
+Further, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass and largest body size are most suited to surviving in urban areas [@major2010museum].
+Conversely, other predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
+For those variables which became less important, it is likely that there may be particular clades that include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].
 
@@
 Furthermore, we do confirm that urbanization appears to select against insectivorous species [@chace2006urban, @evans2011makes].
-While others have found that urbanization supports granivorous species [@chace2006urban, @kark2007living, @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [@fuller2008garden].
+While others have found that urbanization supports granivorous species [@chace2006urban, @kark2007living, @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
 We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@kark2007living].
@@
 Furthermore, our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done).
-The aforementioned differences in our study's results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
+The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
 Our results suggest that urbanization selects against specialist species [@sorace2009distribution], granivores, and insectivores.
 As such, future urban development aimed at increasing avian diversity should focus on maintaining a high diversity of habitat [@melles2003urban] both for foraging [@shochat2004linking], and nesting [@fernandez2001habitat].
-A species-specific approach to assess particular resources necessary for specialist species may be necessary.
-Lastly, a focal point should be on increasing insect abundance and diversity in urban environments (CITATION), which will in turn increase the presence of insectivorours bird species in urban environments.
+A species-specific approach to assess particular resources necessary for specialist species may be necessary [@fernandez2004spatial].
+Lastly, a focal point should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], which will in turn increase the presence of insectivorous bird species in urban environments.
 
-Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advancing remote sensing technology [@gorelick2017google].
+Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
 We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds.
 We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global, @katewaleffect], but migratory status was considered as a predictor.
-We highlight that future studies, especially within migrant-dominated landscapes (i.e., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call].
-Further, future studies should look to expand this analysis outside of the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2018non]).
+We highlight that future studies, especially within migrant-dominated landscapes (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call].
+Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2018non]).
 

### 2018-05-14 ms.Rmd
Fixed one reference and added a reference on spatial scale of ecological traits

@@
 As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization.
-Our analysis used a total of 477 species (Appendix 4), more than any similar study, to the best of our knowledge.
-Furthermore, our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done).
+Our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done).
 The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
@@
 We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds.
-We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global, @katewaleffect], but migratory status was considered as a predictor.
+We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global, @katuwaleffect], but migratory status was considered as a predictor.
 We highlight that future studies, especially within migrant-dominated landscapes (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call].
-Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2018non]).
+Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2018non]).
 

### 2018-05-14 ms.Rmd
fixing #48

@@
 # Tables
-
+\small
 ```{r Table 1, echo=FALSE, warning=FALSE, message=FALSE}
@@
 library(pander)
-pander(table, caption = "Results showing the significance of the phylogenetic signal analyses.")
+pander(table, caption = "Results showing the significance of the phylogenetic signal analyses.", split.table=Inf)
 

### 2018-05-15 ms.Rmd
rough go at an Abstract...

@@
 # ABSTRACT
-
-TBD
+Understanding species' responses to urbanization is critical to the conservation of urban biodiversity.
+Previous work assessing ecological and life history traits that are associated with urban birds has elucidated differing and contrasting patterns for a wide variety of traits.
+Much of this work has relied on a priori grouping of species into discrete groups (e.g., urban avoiders, exploiters, and adapters or even binary presence/absence in urban areas), and has been spatially and/or temporally limited.
+Here, we provide a novel and easily adapted methodology which uses a continuous urbanization index to assess which ecological and life history traits most influence a species' ability to be found in urban environments.
+A continuous measure of urbanization is critical as it fully accounts for the inherent differences in species' responses to urban environments.
+We further expand on previous studies by incorporating a total of 477 species in our analyses, a number far greater than similar studies, to the best of our knowledge.
+We find a significant phylogenetic effect of our urbanization index and thus account for the non-linear relationships among species in our modelling.
+Niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors of urbanization.
+We also find that residual brain mass is a critical predictor within our phylogenetic models, suggesting that those species within a clade which have the largest relative brain size are most readily adapting to urban development.
+Urbanization is selecting against specialists, granivores, and insectivores.
+These results have important implications for aiding the preservation and persistence of urban avian biodiversity.
 

### 2018-05-15 ms.Rmd
some minor odds and ends (keywords, formatting, refs, etc.)

@@
 bibliography: references.bib
-
 ---
@@
 
+*keywords*: birds; brain size; generalists; specialists; life history traits; ecological traits; urbanization index
+
+\newpage
 # INTRODUCTION
@@
 This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global].
-In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic, @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
+In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
 This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.
@@
 This was done using the dredge function in the MuMIn package [@barton2016mumin], and we allowed a total of 11 candidate predictors to be fit at any one time in order to avoid over-fitting.
-This left a total of 784,626 models which were assessed, and the model-averaged parameter estimates were calculated from the top models of those with ???AIC < 4.
-As with the global model, we present the standardized estimates which represent relative effect sizes
+This left a total of 784,626 models which were assessed, and the model-averaged parameter estimates were calculated from the top models of those with $\Delta$AIC < 4.
+As with the global model, we present the standardized estimates which represent relative effect sizes.
 
@@
 ## Urbanization index
-For the 477 species included in the models the mean ? SD urbanization index was 0.91 ? 1.76, with a range from 0.008 (Black-backed Butcherbird) to 11.85 (Powerful Owl; Fig. 2).
+For the 477 species included in the models the mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, with a range from 0.008 (Black-backed Butcherbird) to 11.85 (Powerful Owl; Fig. 2).
 
@@
 5).
+Full model results for each of the four analyses can be found in Appendix 5.
 
@@
 Conversely, granivores and insectivores were consistently negatively associated with urbanization.
-Furthermore, we found a significant phylogenetic signal [@revell2008phylogenetic] of the urbanization index in our analysis, suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence, @ibanez2017global, @sol2017urbanisation, @la2018phylogenetic].
+Furthermore, we found a significant phylogenetic signal [@revell2008phylogenetic] of the urbanization index in our analysis, suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @la2018phylogenetic].
 This contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our increased sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
-Our results demonstrate that considering phylogeny in regards to this question is important.
+Our results demonstrate that considering phylogeny in regards to the association between ecological and life history traits and urbanization in birds is important.
 
@@
 Conversely, other predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
-For those variables which became less important, it is likely that there may be particular clades that include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].
+For the variables which became less important, it is likely that there may be particular clades that include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].
 
-Our finding that ecological generalists are more readily observed in urban areas confirms that of other work [@mckinney2006urbanization, @bonier2007urban, @moller2009successful].
-We also confirm work which has found a relationship between annual fecundity and urbanization in birds [@croci2008does, @moller2009successful].
-While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living, @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful, @maklakov2011brains, @moller2015brain].
-Furthermore, we do confirm that urbanization appears to select against insectivorous species [@chace2006urban, @evans2011makes].
-While others have found that urbanization supports granivorous species [@chace2006urban, @kark2007living, @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
-We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@kark2007living].
+Our finding that ecological generalists are more readily observed in urban areas confirms that of other work [@mckinney2006urbanization; @bonier2007urban; @moller2009successful].
+We also confirm work which has found a relationship between annual fecundity and urbanization in birds [@croci2008does; @moller2009successful].
+While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain].
+Furthermore, we do confirm that urbanization appears to select against insectivorous species [@chace2006urban; @evans2011makes].
+While others have found that urbanization supports granivorous species [@chace2006urban; @kark2007living; @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
+We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
 
@@
 
-Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global, @aronson2016hierarchical, @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
+Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
 We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds.
-We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role, @lepczyk2017global, @katuwaleffect], but migratory status was considered as a predictor.
-We highlight that future studies, especially within migrant-dominated landscapes (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full, @marra2015call].
-Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation, @batary2018non]).
+We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor.
+We highlight that future studies, especially within migrant-dominated landscapes (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full; @marra2015call].
+Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation; @batary2018non]).
 
@@
 
+\newpage
 # REFERENCES
@@
 A star represents a significant effect.
-Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set."}
+Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set.
+Full model results can be found in Appendix 5."}
 knitr::include_graphics("finalFigs/Figure4.pdf")

### 2018-05-15 ms.Rmd
wording changes

@@
 # METHODS
-## Study area
-We looked at bird observations throughout the whole of continental Australia. Bird observations from outlying islands were eliminated from potential analyses.
 
 ## Bird observation data
-We collated observations from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists.
+We collated bird observations throughout the whole of continental Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists.
 Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird].
@@
 This time-frame was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see below) over the given time frame in which bird observations were collected.
+Bird observations from outlying islands were eliminated from potential analyses.
 
@@
 
-## Phylogenetic analysis
-For the phylogenetic portion of our analyses, we used the phylogenetic tree provided by Jetz et al.
-[@jetz2012global].
-This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global].
-In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
-This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.
-
 ## Statistical analyses
@@
 
-### Phylogenetically constrained models
+### Phylogenetic analysis
+
+For the phylogenetic portion of our analyses, we used the phylogenetic trees provided by Jetz et al.
+[@jetz2012global].
+This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global].
+In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
+This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.
+
 Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package [@ho2014phylolm].
 Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.
+We also  fit the same model to a range of the equally likely trees from [@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.
 
@@
 There were minor differences which existed in common names, scientific names, and recognized species among the three datasets.
-A taxonomic key was made which combined the three different taxonomies (Appendix 2).
+A taxonomic look-up table was made which combined the three different taxonomies (Appendix 2).
 Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset.
@@
 # DISCUSSION
-Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban].
-Conversely, granivores and insectivores were consistently negatively associated with urbanization.
-Furthermore, we found a significant phylogenetic signal [@revell2008phylogenetic] of the urbanization index in our analysis, suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @la2018phylogenetic].
-This contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our increased sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
-Our results demonstrate that considering phylogeny in regards to the association between ecological and life history traits and urbanization in birds is important.
+
+Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index.
+This result was robust across non-phylogenetic and phylogenetic models, model averaging approaches, and across phylogenetic uncertaintly.
+Conversely, specilized feeding strategies were negatively associated with the urbanization index: granivore and insectivore specialists were much less likely to be found in urban environments.
+These results suggest that urbanization is, in general, excluding most specialist species.
+SOMETHING ABOUT IMPLICATIONS FOR CONSERVATION / FRAGMENTATION.
+
+The non-random distribution of the urbanization index across the phylogeny (Table 1) is interesting.
+This result contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
+Interestingly many of the predictor traits---generalism, clutch size---have even stronger phylogenetic signal, so it's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny.
+It is worth noting that this results is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @major2010museum].
 
@@
 
-Our finding that ecological generalists are more readily observed in urban areas confirms that of other work [@mckinney2006urbanization; @bonier2007urban; @moller2009successful].
-We also confirm work which has found a relationship between annual fecundity and urbanization in birds [@croci2008does; @moller2009successful].
+Brain size at a given body size is an especially interesting trait--it's becomes more important in the phylogenetic model compared to the non-phylogenetic model.
 While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain].
-Furthermore, we do confirm that urbanization appears to select against insectivorous species [@chace2006urban; @evans2011makes].
+Because of the phylogenetic signal in both brain size and urbanization, it is crutial to investigate this relationship in a phylogenetically informed manner.
+
+Certian aspects of the trait-urbanization relationship appear to consistent across the globe.
+CHANGE PHRASING BELOW HERE: The relationship with annual fecundity and urbanization in birds is consistent with work in Europe  [@croci2008does; @moller2009successful].
+The that urbanization appears to select against insectivorous species [@chace2006urban; @evans2011makes].
 While others have found that urbanization supports granivorous species [@chace2006urban; @kark2007living; @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].

### 2018-05-15 ms.Rmd
combining the 'phylo' para and the 'brain size' para into one

@@
 # ABSTRACT
-Understanding species' responses to urbanization is critical to the conservation of urban biodiversity.
-Previous work assessing ecological and life history traits that are associated with urban birds has elucidated differing and contrasting patterns for a wide variety of traits.
-Much of this work has relied on a priori grouping of species into discrete groups (e.g., urban avoiders, exploiters, and adapters or even binary presence/absence in urban areas), and has been spatially and/or temporally limited.
-Here, we provide a novel and easily adapted methodology which uses a continuous urbanization index to assess which ecological and life history traits most influence a species' ability to be found in urban environments.
-A continuous measure of urbanization is critical as it fully accounts for the inherent differences in species' responses to urban environments.
-We further expand on previous studies by incorporating a total of 477 species in our analyses, a number far greater than similar studies, to the best of our knowledge.
-We find a significant phylogenetic effect of our urbanization index and thus account for the non-linear relationships among species in our modelling.
-Niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors of urbanization.
-We also find that residual brain mass is a critical predictor within our phylogenetic models, suggesting that those species within a clade which have the largest relative brain size are most readily adapting to urban development.
-Urbanization is selecting against specialists, granivores, and insectivores.
-These results have important implications for aiding the preservation and persistence of urban avian biodiversity.
+Understanding the suite of ecological and life history traits which are most associated with urbanization enables a predictive ability to understand the species which are readily suited to adapt with our increasingly urbanizing world.
+Previous work investigating this relationship has found contrasting patterns for a specific trait's association with urbanization.
+Here, we provide a novel and scalable methodology relying on a continuous urbanization index to assess ecological and life history traits which most influence a species' ability to be found in urban environments.
+In a continental-scale analysis, incorporating 477 species, we find that niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors of urbanization.
+Conversely, specialized feeding strategies (i.e., insectivores and granivores) were negatively associated with urban environments.
+Our results suggest that the enhancement and persistence of avian biodiversity in urban areas relies on developing suitable habitat to serve a wide range of life history strategies.
 
@@
 Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index.
-This result was robust across non-phylogenetic and phylogenetic models, model averaging approaches, and across phylogenetic uncertaintly.
-Conversely, specilized feeding strategies were negatively associated with the urbanization index: granivore and insectivore specialists were much less likely to be found in urban environments.
+This result was robust across non-phylogenetic and phylogenetic models, model averaging approaches, and across phylogenetic uncertainty.
+Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments.
 These results suggest that urbanization is, in general, excluding most specialist species.
@@
 
-The non-random distribution of the urbanization index across the phylogeny (Table 1) is interesting.
-This result contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
-Interestingly many of the predictor traits---generalism, clutch size---have even stronger phylogenetic signal, so it's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny.
-It is worth noting that this results is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @major2010museum].
+The non-random distribution of the urbanization index across the phylogeny (Table 1) contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
+It's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny and we demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear model.
+For example, some predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
+For these variables, it is likely that there may be particular clades which include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].
+Conversely, some of the most important predictor variables (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization when phylogeny was accounted for, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
+S2).
+Furthermore, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
+While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], our results suggest that brain size is indeed heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain].
+This finding highlights the necessity to investigate these relationships in a phylogenetically informed manner.
+Lastly, it is worth noting that these results are consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @major2010museum].
 
-The effect of treating species as non-independent by considering phylogeny in the context of the linear model was apparent.
-For example, there was an increase in parameter estimates of the most important parameters (e.g., clutch size and feeding habitat generalism; Figures 4 and 5), and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
-S2).
-Further, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass and largest body size are most suited to surviving in urban areas [@major2010museum].
-Conversely, other predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
-For the variables which became less important, it is likely that there may be particular clades that include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].
-
-Brain size at a given body size is an especially interesting trait--it's becomes more important in the phylogenetic model compared to the non-phylogenetic model.
-While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain].
-Because of the phylogenetic signal in both brain size and urbanization, it is crutial to investigate this relationship in a phylogenetically informed manner.
-
-Certian aspects of the trait-urbanization relationship appear to consistent across the globe.
+Certian aspects of the trait-urbanization relationship appear to be consistent across the globe.
 CHANGE PHRASING BELOW HERE: The relationship with annual fecundity and urbanization in birds is consistent with work in Europe  [@croci2008does; @moller2009successful].

### 2018-05-15 ms.Rmd
a lot of reworking the discussion!

@@
 # DISCUSSION
-
-Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index.
+Niche breadth, as measured by different types of generalism, was significantly associated with urbanization in our analysis.
 This result was robust across non-phylogenetic and phylogenetic models, model averaging approaches, and across phylogenetic uncertainty.
 Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments.
-These results suggest that urbanization is, in general, excluding most specialist species.
-SOMETHING ABOUT IMPLICATIONS FOR CONSERVATION / FRAGMENTATION.
+These results suggest that urbanization is, in general, excluding specialist species [@sorace2009distribution].
+As such, future urban development aimed at increasing avian diversity in urban areas should focus on maintaining a high diversity of habitat [@melles2003urban] both for foraging [@shochat2004linking], and nesting [@fernandez2001habitat], with a focus on intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role].
+A species-specific approach to assess particular resources necessary for specialist species may be necessary [@fernandez2004spatial].
+Lastly, a focal point should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], which will in turn increase the presence of insectivorous bird species in urban environments.
 
@@
 
-Certian aspects of the trait-urbanization relationship appear to be consistent across the globe.
-CHANGE PHRASING BELOW HERE: The relationship with annual fecundity and urbanization in birds is consistent with work in Europe  [@croci2008does; @moller2009successful].
-The that urbanization appears to select against insectivorous species [@chace2006urban; @evans2011makes].
-While others have found that urbanization supports granivorous species [@chace2006urban; @kark2007living; @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
-We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
+Certian aspects of the trait-urbanization relationship appear to be consistent, and thus generalizable, across the globe.
+For instance, our finding of the relative importance of annual fecundity is consistent with work in Europe [@croci2008does; @moller2009successful].
+Urbanization appears to generally select against insectivorous species [@chace2006urban; @evans2011makes].
+Although we did not find any strong support for our hypotheses on different nesting types, we did find that in general species which tended to be more social were closely associated with urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
+While these patterns may be generalizable, we also found patterns in our analysis which seem to contrast the global norm.
+For example, we found that urbanization selects against granivorous species, contrasting the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
 
-Clearly, many traits have been used to identify a species' adaptability with urban-living (e.g., Appendix 1), but often times results differ among studies.
-Fully understanding the suite of specific species' traits which are associated with urban birds allows for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].
+As illustrated (above and in Appendix 1), many traits have been used to identify a species' adaptability to urban environments, but results have frequently differed among studies.
 As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization.
@@
 The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
-Our results suggest that urbanization selects against specialist species [@sorace2009distribution], granivores, and insectivores.
-As such, future urban development aimed at increasing avian diversity should focus on maintaining a high diversity of habitat [@melles2003urban] both for foraging [@shochat2004linking], and nesting [@fernandez2001habitat].
-A species-specific approach to assess particular resources necessary for specialist species may be necessary [@fernandez2004spatial].
-Lastly, a focal point should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], which will in turn increase the presence of insectivorous bird species in urban environments.
-
-Global broad-scale empirical datasets are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen], in large part due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
+This analysis was made possible through the use of global broad-scale empirical datasets, which are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
+This is largely due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
 We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds.
@@
 Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation; @batary2018non]).
+Fully understanding the suite of specific species' traits which are associated with urban birds, and placing them in a theoretical landscape ecology context, will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].
 
-Ultimately, we provide the largest study to date using a continuous urbanization metric in order to assess biological and ecological traits which predict a species' adaptability to urban environments.
-In doing so, we highlight the applicability of broad-scale empirical data to assess species' traits which predict presence and adaptability in urban environments.
+We provide the largest study to date using a continuous urbanization metric in order to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
 Our results suggest that generalist species, with large relative brain size, are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
+This work should be extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species.
 Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management.

### 2018-05-15 scripts/Appendix_5-model_output_results.Rmd
making the appendix look moderately more aesthetically pleasing.

@@
   html_document: default
+classoption: landscape
+geometry: margin=2cm
 ---
@@
 ```
+\newpage
 
 ## Model: Non-phylogenetically controlled global model results
-
+\footnotesize
 ```{r Non-phylogenetically controlled global model results, echo=FALSE, message=FALSE}
@@
 remake::dump_environment()
+options(digits=4)
 summary(global_model)
 ```
+\newpage
 
 ## Model: Non-phylogenetically controlled model-averaged results
-
+\footnotesize
 ```{r Non-phylogenetically controlled model-averaged results, echo=FALSE, message=FALSE}
@@
 summary <- readRDS("Data/dredged_model_summary_results.rds")
+options(digits=4)
 summary(summary)
 ```
+\newpage
 
 ## Model: Phylogenetically controlled global model results
-
+\footnotesize
 ```{r Phylogenetically controlled global model results, echo=FALSE, message=FALSE}
@@
 remake::dump_environment()
+options(digits=4)
 summary(phy_mod_rescaled)
 ```
+\newpage
 
 ## Model: Phylogenetically controlled model-averaged results
-
+\footnotesize
 ```{r Phylogenetically controlled model-averaged results, echo=FALSE, message=FALSE}
@@
 summary2 <- readRDS("Data/PHYLO_dredged_model_summary_results.rds")
+options(digits=4)
 summary(summary2)

### 2018-05-15 ms.Rmd
Polishing the introduction a little bit .
Added a linking sentence between para 2 and para 3. Revamped para 4 to make it less like we are 'extending' other's work, and hanging the hat on the novelty a little bit more. And a couple other odd word changes, etc.

@@
 # ABSTRACT
-Understanding the suite of ecological and life history traits which are most associated with urbanization enables a predictive ability to understand the species which are readily suited to adapt with our increasingly urbanizing world.
-Previous work investigating this relationship has found contrasting patterns for a specific trait's association with urbanization.
-Here, we provide a novel and scalable methodology relying on a continuous urbanization index to assess ecological and life history traits which most influence a species' ability to be found in urban environments.
-In a continental-scale analysis, incorporating 477 species, we find that niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors of urbanization.
-Conversely, specialized feeding strategies (i.e., insectivores and granivores) were negatively associated with urban environments.
-Our results suggest that the enhancement and persistence of avian biodiversity in urban areas relies on developing suitable habitat to serve a wide range of life history strategies.
+Understanding the suite of ecological and life history traits which are associated with urbanization enables a predictive ability to understand the species which are readily suited to adapt with our increasingly urbanizing world.
+Research investigating this relationship has found contrasting patterns for a specific trait's association with urbanization.
+Here, we provide a novel and scalable methodology relying on a continuous urbanization index to assess ecological and life history traits which most influence a species' adaptability to persist in urban environments.
+In a continental-scale analysis, incorporating 477 species, we find that niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors.
+Conversely, specialized feeding strategies (i.e., insectivores and granivores) were negatively associated with urbanization.
+Our results suggest that the enhancement and persistence of avian biodiversity in urban envioronments relies on developing suitable habitat to serve a wide range of life history strategies.
 
@@
 Thus, urbanization globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects] and is a major threat to conservation [@czech2000economic; @dearborn2010motivations].
-Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @lowry2013behavioural; @mcdonnell2015adaptation].
-For instance, some species which rarely occur within urban areas can be relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
+Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation].
+However, some species rarely occur within urban areas, despite being relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
 
-In order to best predict species' responses to future urbanization changes, it is essential to understand what biological traits are associated with urban birds [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
+In order to best predict species' responses to future urbanization changes, it is essential to understand what ecological and life history traits are most associated with urban birds [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
 To date, a wide array of traits have been used to predict presence of birds in urban environments.
@@
 Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011makes with @moller2009successful & @croci2008does] and niche breadth [cf., @kark2007living with @evans2011makes].
+Forming generalizable patterns, at multiple spatial scales [@suarez2018scale], which associate ecological and life history traits with urbanization in birds is critical to help curb the impact urbanization has on avian biodiversity [@sol2017urbanisation].
 
-The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even birds which do or do not occur within urban environments [e.g., @bonier2007urban; @moller2009successful].
-This type of discrete response variable severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011makes].
+The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even binary grouping based on presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
+These types of discrete response variables severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011makes].
 Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
@@
 
-Despite the number of studies which have investigated the relationships between species' traits and urban adaptedness, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two month survey period [@kark2007living]].
-Fortunately, citizen science [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen] can provide broad-scale empirical datasets, allowing us to test hypotheses at spatial and temporal scales previously difficult to attain.
-This data provides a cost-effective method to research the relationship between species' traits and urban observations.
+Furthermore, despite the number of studies which have investigated the relationships between traits and urbanization, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two month survey period [@kark2007living].
+Fortunately, citizen science [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen] provides broad-scale empirical data, allowing us to test hypotheses at spatial and temporal scales previously difficult to attain.
+These data provide a cost-effective method to research the relationship between species' traits and urbanization in birds.
 Indeed, various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
-However, we are unaware of any projects which use citizen science observations to assess species' traits in relation to urban environments over a large spatial and temporal scale.
+Yet, we are unaware of any projects which use citizen science observations to assess species' traits in relation to urban environments over a large spatial and temporal scale.
 
-Here, we improve upon current methods of assessing biological traits which predicts urbanization in birds by using a series of publicly available datasets.
-We developed and used a continuous urbanization index as a response variable in the analyses.
-We had two objectives: (1) investigate whether there was a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes], and (2) develop models which were and were not phylogenetically constrained to assess species' traits that predict pre-adaptation to urban living, using a suite of continuous and categorical predictor variables.
-These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth.
+We use a series of publicly available datasets to implement a novel continuous urbanization index in order to assess ecological and life history traits which most influence a bird's response to urbanization.
+We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against).
+Our first objective was to investigate the presence of a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
+Our second objective was to develop models which were and were not phylogenetically constrained to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables.
+These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth [@coffee2016visualising].
 

### 2018-05-15 ms.Rmd
Some polishing/minor changes of results

@@
 
-## Taxonomic considerations
+### Taxonomic considerations
 Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed.
@@
 # RESULTS
-A total of 637,482 checklists provided a total of 5,944,819 observations of 580 different species which had > 100 observations.
+A total of 637,482 checklists resulting in 5,944,819 observations of 580 different species which had > 100 observations were initially investigated for analyses.
 The least often recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
 The mean number of observations for a species was 10,250.
-Density plots of all 580 species' distributions of average radiance values are presented in Appendix 3.
-
-We ended with 477 candidate species from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data (Appendix 4).
+Plots of all 580 species' distributions of average radiance values are presented in Appendix 3.
+After the 580 candidate species from the eBird dataset were considered based on complete data following both the trait taxonomy and phylogenetic tree taxonomy (e.g., Appendix 2), we were left with 477 species for analysis.
+A complete list of the 477 species, in addition to their traits, number of observations, and their urbanization index can be found in Appendix 4.
 
@@
 There was a distinct phylogenetic signal in our urbanization index measure (Fig.
-3; Table 1), as it was significantly different than random for 4 out of the 5 indices.
+3; Table 1), as it was significantly different than random for 4 out of the 5 indices considered.
 Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1).
-The result for body size-- $k=3.3$ --is among the strongest ever recorded measures of phylogenetic signal in the literature (see REFS) and for $k$ the statistic suggests much stronger signal than expected from Brownian motion.
-We found weaker signal for behavioral traits compared to morphological traits.
+The result for body size ($k=3.3$) is among the strongest ever recorded measures of phylogenetic signal in the literature, and for $k$ the statistic suggests much stronger signal than expected from Brownian motion.
+We generally found weaker signal for behavioral traits compared with morphological traits.
 For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 1).
@@
 The most important predictors of urbanization in the non-phylogenetic global model (Fig.
-4) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006).
+4; Appendix 5) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006).
 Urban birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and migratory status (t = 2.15, p = 0.032).
-Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitary (t = -2.90, p = 0.004).
+Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitarily (t = -2.90, p = 0.004).
 
 When this global model was dredged and model-averaged, patterns remained similar (Fig.
-4).
+4; Appendix 5).
 Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors.
 Additionally, agricultural birds were significantly associated with urbanization (z = 3.40, p = 0.001).
-Urban birds remained significantly negatively associated with feeding solitary (z = 3.39, p = 0.001), being insectivorous (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), feeding in pairs and flocks (z = 3.16, p = 0.002), being granivorous (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).
+Urban birds remained significantly negatively associated with feeding solitarily (z = 3.39, p = 0.001), being insectivorous (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), feeding in pairs and flocks (z = 3.16, p = 0.002), being granivorous (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).
 
@@
 The most significant predictors of urbanization in the phylogenetically controlled global model (Fig.
-4) were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism ( t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005).
+4; Appendix 5) were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism ( t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005).
 Nest generlaism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding species (t = -3.25, p = 0.001) and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.
@@
 When the phylogenetic global model was model-averaged, the patterns were similar (Fig.
-4).
+4; Appendix 5).
 Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significant predictors of urbanization index.
@@
 ### Differences in non-phylogenetic and phylogenetic models
-The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the significance of individual predictors (Fig.
+The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the statistical significance of individual predictors (Fig.
 4).
-However, brain residual size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig.
+However, residual brain size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig.
 5).
-Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger parameter estimates) in the phylogenetic model compared with the non-phylogenetic model (Fig.
+Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger standardized parameter estimates) in the phylogenetic models compared with the non-phylogenetic models (Fig.
 5).

### 2018-05-16 ms.Rmd
some minor work on discussion

@@
 # DISCUSSION
-Niche breadth, as measured by different types of generalism, was significantly associated with urbanization in our analysis.
-This result was robust across non-phylogenetic and phylogenetic models, model averaging approaches, and across phylogenetic uncertainty.
+Niche breadth, as measured by different types of generalism, was significantly associated with urbanization in our analyses, evidenced by the robustness of this result across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
+S2).
 Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments.
@@
 A species-specific approach to assess particular resources necessary for specialist species may be necessary [@fernandez2004spatial].
-Lastly, a focal point should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], which will in turn increase the presence of insectivorous bird species in urban environments.
+Furthermore, a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], which will in turn benefit avian biodiversity in urban environments.
 
-The non-random distribution of the urbanization index across the phylogeny (Table 1) contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
-It's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny and we demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear model.
+The non-random distribution of the urbanization index across the phylogeny (Table 1) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
+This finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
+It's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny and we demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models.
 For example, some predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
@@
 Furthermore, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
-While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], our results suggest that brain size is indeed heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain].
-This finding highlights the necessity to investigate these relationships in a phylogenetically informed manner.
-Lastly, it is worth noting that these results are consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @major2010museum].
+While some studies have found a lack of evidence that residual brain size is an important predictor in urban birds [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed heavily correlated with urbanization in birds [@moller2009successful; @maklakov2011brains; @moller2015brain].
+This finding highlights the necessity to investigate the trait-urbanization relationships in a phylogenetically informed manner.
 
@@
 
-We provide the largest study to date using a continuous urbanization metric in order to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
-Our results suggest that generalist species, with large relative brain size, are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
-This work should be extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species.
-Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management.
+We provide the largest study to date using a continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
+Our results suggest that generalist species, with large residual brain size, are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
+We highlight that future studies investigating the trait-urbanization relationship should focus on broad-scale data  to form generalizable patterns which rely on continuous measures for improved predictive performance.
+This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species.
+In the context of our specific findings, Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management.
 Urban avian biodiversity could be increased by improving the suitability of foraging for granivores and insectivores, especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more].
-Furthermore, future studies should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance.
 

### 2018-05-17 ms.Rmd
some changes based on RM comments/thoughts

@@
 Urbanization fragments and isolates natural habitat [@marzluff2001restoration; @melles2003urban; @loram2007urban], drastically altering species' composition and ecosystem functions.
-Thus, urbanization globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects] and is a major threat to conservation [@czech2000economic; @dearborn2010motivations].
+Thus, urbanization globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects] and is a major concern for conservation [@czech2000economic; @dearborn2010motivations].
 Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation].
-However, some species rarely occur within urban areas, despite being relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
+Conversely, some species rarely occur within urban areas, despite being relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
 
-In order to best predict species' responses to future urbanization changes, it is essential to understand what ecological and life history traits are most associated with urban birds [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
+In order to best predict species' responses to future urbanization changes, it is essential to understand what ecological and life history traits are most associated with birds that are either successful or unsuccessful in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
 To date, a wide array of traits have been used to predict presence of birds in urban environments.
-For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does; @conole2011functional], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have been hypothesized to influence a birds' presence in urban environments.
+For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does; @conole2011functional], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have been hypothesized to influence a species' presence in urban environments.
 Rather than a specific trait that predicts a species' ability to be found in urban environments, it is most likely a combination of traits.
-For instance, urban birds' presence in Jerusalem was predicted by traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-nesters, and widely distributed [@croci2008does].
-Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting one another.
+For instance, the presence of urban birds in Jerusalem was predicted by traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
+Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting with one another.
 Indeed, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains] but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on species' response to urbanization.
@@
 The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even binary grouping based on presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
-These types of discrete response variables severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011makes].
+These types of discrete response variables severely limits the capacity to generalize results as it assumes species within a group respond to urbanization equally [@evans2011makes].
 Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
-In a notable exception, [@evans2011makes] advocates for a continuous measure of urbanization in which a species' is assessed on their entire distribution in response to urban environments.
+In a notable exception, Evans *et al.* [-@evans2011makes] advocates for a continuous measure of urbanization in which a species is assessed on its entire distribution in response to urban environments.
 
-Furthermore, despite the number of studies which have investigated the relationships between traits and urbanization, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two month survey period [@kark2007living].
-Fortunately, citizen science [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen] provides broad-scale empirical data, allowing us to test hypotheses at spatial and temporal scales previously difficult to attain.
-These data provide a cost-effective method to research the relationship between species' traits and urbanization in birds.
+Furthermore, despite the number of studies which have investigated the relationships between traits and urbanization, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living].
+There is a need for broad-scale empirical datasets, which allow us to test hypotheses at large spatiotemporal scales, and citizen science is beginning to provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen].
 Indeed, various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
-Yet, we are unaware of any projects which use citizen science observations to assess species' traits in relation to urban environments over a large spatial and temporal scale.
 
-We use a series of publicly available datasets to implement a novel continuous urbanization index in order to assess ecological and life history traits which most influence a bird's response to urbanization.
+The aim of this study is to determine whether there is a suite of life history traits that predicts avian tolerance to urbanization at a continental scale.
+For such an analysis to have generality it must (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3), include a broad-range of traits, and (4) employ a continuous measure of a specie's response to urbanization.
+We conduct an analysis that meets these criteria, using a series of publicly available datasets to implement a novel continuous urbanization index in order to assess ecological and life history traits which most influence a bird's response to urbanization.
 We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against).
@@
 Our second objective was to develop models which were and were not phylogenetically constrained to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables.
-These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth [@coffee2016visualising].
 
@@
 We filtered the bird observations [@callaghan2017assessing; @la2014role] by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barrett2003new].
-For further information on eBird protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types].
+For further information on eBird protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types.
 Further, we filtered the checklists to those with a distance travelled of < 5 km or an area covered which was < 500 Ha.
@@
 Only species which included > 100 observations in the eBird database over the specified time-frame, and that met our criteria were considered for further analyses.
-This left us with 580 potential species from the eBird dataset which had > 100 observations on eligible checklists in the specified time frame.
+This left us with 580 potential species from the eBird dataset which had > 100 observations on eligible checklists in the specified time frame, which was a substantial proportion of the ~730 bird species which are resident in Australia.
+
+Given the diverse datasets (see below) used in this analysis, there were differing taxonomic authorities followed (see Appendix 2 for elaboration).
 
@@
 Therefore, we excluded IUCN status as a trait given that the majority of our candidate species were of Least Concern.
-
 Only species which had complete data for the list of candidate traits were considered in the analyses, and we thereby ended with 477 species which met this criteria.
@@
 
-Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package [@ho2014phylolm].
+Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package which fits a linear regression model where the likelihood is linear in the number of tips in the tree [@ho2014phylolm].
 Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.
@@
 
-### Taxonomic considerations
-Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed.
-The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/].
-There were minor differences which existed in common names, scientific names, and recognized species among the three datasets.
-A taxonomic look-up table was made which combined the three different taxonomies (Appendix 2).
-Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset.
-If it did not match both, then it was not considered for analysis (e.g., species currently only recognized by eBird/clements such as Western Whistler, Silver-backed Butcherbird, or Paperbark Flycatcher).
-Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia and thus not included in the analysis.
-
 # RESULTS
-A total of 637,482 checklists resulting in 5,944,819 observations of 580 different species which had > 100 observations were initially investigated for analyses.
+A total of 316,306 checklists resulting in 5,944,819 observations of 477 different species were included in analyses.
 The least often recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
 The mean number of observations for a species was 10,250.
-Plots of all 580 species' distributions of average radiance values are presented in Appendix 3.
-After the 580 candidate species from the eBird dataset were considered based on complete data following both the trait taxonomy and phylogenetic tree taxonomy (e.g., Appendix 2), we were left with 477 species for analysis.
-A complete list of the 477 species, in addition to their traits, number of observations, and their urbanization index can be found in Appendix 4.
+Plots of all 477 species' distributions of average radiance values are presented in Appendix 3.
+A complete list of the 477 species, in addition to their traits, number of observations, and their urbanization indices can be found in Appendix 4.
 
@@
 4; Appendix 5) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006).
-Urban birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and migratory status (t = 2.15, p = 0.032).
-Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitarily (t = -2.90, p = 0.004).
+Urban-tolerant birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032).
+Conversely, urban-tolerant birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), group foraging (t = -2.24, p =0.025), granivorous diet (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), insectivorous diet (t = -3.96, p < 0.001), and solitary feeding (t = -2.90, p = 0.004).
 
-When this global model was dredged and model-averaged, patterns remained similar (Fig.
+When this global model was dredged and model-averaged to investigate for specific suites of traits, patterns remained similar (Fig.
 4; Appendix 5).
 Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors.
-Additionally, agricultural birds were significantly associated with urbanization (z = 3.40, p = 0.001).
-Urban birds remained significantly negatively associated with feeding solitarily (z = 3.39, p = 0.001), being insectivorous (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), feeding in pairs and flocks (z = 3.16, p = 0.002), being granivorous (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).
+Additionally, birds which were classified in the traits database as having strong association with agricultural habitats were significantly associated with urbanization (z = 3.40, p = 0.001).
+Urban-tolerant birds remained significantly negatively associated with solitary feeding (z = 3.39, p = 0.001),  insectivorous diet (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), group foraging (z = 3.16, p = 0.002), granivorous diet (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).
 
@@
 4; Appendix 5) were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism ( t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005).
-Nest generlaism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding species (t = -3.25, p = 0.001) and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.
+Nest generalism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding (t = -3.25, p = 0.001), and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.
 
@@
 Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significant predictors of urbanization index.
-Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization.
+Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization.
 

### 2018-05-17 scripts/results_script.R
overhaul of preparation of results for various aspects.

@@ -15,6 +15,24 @@ min(analysis_data$urban_median)
 
 max(analysis_data$urban_median)
 
+## get total number of checklists and observations used for the analysis, based on the 477 species only
+## need to subset the eBird data to only those species in the analysis data
+
+## get list of species
+species <- analysis_data %>%
+  dplyr::select(COMMON_NAME_ebird) %>%
+  rename(COMMON_NAME = COMMON_NAME_ebird) %>%
+  .$COMMON_NAME
+
+## subset eBird data based on this list
+eBird_final_data <- ebird_data %>%
+  filter(COMMON_NAME %in% species)
+
+## number of checklists
+length(unique(eBird_final_data$SAMPLING_EVENT_IDENTIFIER))
+
+## number of observations
+nrow(eBird_final_data)
 
 ## Figure 1
 ## A figure showing the distribution of 5 different species included in the analysis
@@ -26,18 +44,18 @@ species_to_plot <- c("Dusky Moorhen",
                      "Hooded Robin",
                      "Noisy Friarbird",
                      "Squatter Pigeon")
-
+library(scales)
 Figure1 <- species_urban %>%
   filter(COMMON_NAME %in% species_to_plot) %>%
   ggplot(., aes(avg_rad, fill=COMMON_NAME)) +
   geom_density(alpha = 0.7) + 
-  scale_x_log10() + 
+  scale_x_continuous(trans='log10', breaks=c(0.0001, 0.0100, 1.0000, 100.0000), labels=c("0.0001", "0.01", "1.0", "100.0"))+ 
   theme_classic() +
-  xlab("log(Average radiance)") +
+  xlab(bquote('Average radiance ('* 'nW' ~cm^-2~sr^-1*')'))+
   ylab("Density")+
   guides(fill=guide_legend(title="  Common Name")) +
   theme(legend.position = c(0.26, 0.8))+
-  theme(axis.text.x=element_text(hjust=0.7))
+  theme(axis.text.x=element_text(hjust=0.6))
 
 pdf("finalFigs/Figure1.pdf", width=5.5, height=4.5)
 print(Figure1)
@@ -58,18 +76,23 @@ species_to_plot <- c("Dusky Moorhen",
 
 species_values <- analysis_data %>%
   filter(COMMON_NAME_ebird %in% species_to_plot) %>%
-  dplyr::select(COMMON_NAME_ebird, urban_median) %>%
-  mutate(urban_median = log(urban_median))
[…]

### 2018-05-17 scripts/results_script.R
removing log (forgot to save before)

@@ -86,7 +86,7 @@ Figure2 <- ggplot(data=analysis_data, aes(urban_median)) +
                  alpha = 0.3)+
   scale_x_continuous(trans='log10', breaks=c(0.005, 0.1, 1.0, 12.0), labels=c("0.005", "0.1", "1.0", "12.0"))+
   theme_classic()+
-  xlab("log(Urbanization index)")+
+  xlab("Urbanization index")+
   ylab("Count")+
   geom_segment(aes(x=0.025, xend=0.025, y=23, yend=16), color='firebrick4', size=1.2, arrow=arrow(length=unit(0.5, "cm")))+
   geom_segment(aes(x=0.034, xend=0.034, y=25, yend=18), color='firebrick4', size=1.2, arrow=arrow(length=unit(0.5, "cm")))+

### 2018-05-17 ms.Rmd
looking better... try remaking now...

@@
 There was a distinct phylogenetic signal in our urbanization index measure (Fig.
-3; Table 1), as it was significantly different than random for 4 out of the 5 indices considered.
+3; Fig.
+S2; Table 1), as it was significantly different than random for 4 out of the 5 indices considered.
 Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1).
@@
 Niche breadth, as measured by different types of generalism, was significantly associated with urbanization in our analyses, evidenced by the robustness of this result across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
-S2).
+S3).
 Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments.
@@
 Conversely, some of the most important predictor variables (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization when phylogeny was accounted for, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
-S2).
+S3).
 Furthermore, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
@@
 
-Certian aspects of the trait-urbanization relationship appear to be consistent, and thus generalizable, across the globe.
+Certain aspects of the trait-urbanization relationship appear to be consistent, and thus generalizable, across the globe.
 For instance, our finding of the relative importance of annual fecundity is consistent with work in Europe [@croci2008does; @moller2009successful].
@@
 
-```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance values."}
+```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance value distributions, on a logarithmic scale."}
 knitr::include_graphics("finalFigs/Figure1.pdf")
@@
 
-```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a log-scale, of the response variable for the 477 species included in the analysis, with five example species highlighted."}
+```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a logarithmic scale, of the response variable for the 477 species included in the analysis, with five example species highlighted."}
 knitr::include_graphics("finalFigs/Figure2.pdf")
@@
 
-```{r Figure S2, echo=FALSE, fig.cap="Phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
+```{r Figure S2, echo=FALSE, fig.cap="A reference phylogenetic tree, with labeled tips, for each of the 477 species included in the analysis. Appendix 4 lists the 477 species."}
 knitr::include_graphics("finalFigs/FigureS2.pdf")
+```
+
+```{r Figure S3, echo=FALSE, fig.cap="Phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
+knitr::include_graphics("finalFigs/FigureS3.pdf")
 ```

### 2018-05-17 ms.Rmd

@@
 # ABSTRACT
-Understanding the suite of ecological and life history traits which are associated with urbanization enables a predictive ability to understand the species which are readily suited to adapt with our increasingly urbanizing world.
-Research investigating this relationship has found contrasting patterns for a specific trait's association with urbanization.
-Here, we provide a novel and scalable methodology relying on a continuous urbanization index to assess ecological and life history traits which most influence a species' adaptability to persist in urban environments.
-In a continental-scale analysis, incorporating 477 species, we find that niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors.
-Conversely, specialized feeding strategies (i.e., insectivores and granivores) were negatively associated with urbanization.
-Our results suggest that the enhancement and persistence of avian biodiversity in urban envioronments relies on developing suitable habitat to serve a wide range of life history strategies.
+Understanding the suite of ecological and life history traits which influence a species' tolerance to urbanization enables a predictive ability to elucidate those species which are most likely to persist in our increasingly urbanizing world.
+Previous work investigating these relationships has found contrasting patterns for a trait's association with urbanization.
+In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we provide a novel and scalable methodology which assesses the ecological and life history traits which most influence a species' adaptability to persist in urban environments.
+We elucidate generalized, phylogenetically controlled, patterns: bird species which are specialists, with high clutch size, and large residual brain size are among the most urban-tolerant bird species.
+Conversely, specialized feeding strategies (i.e., insectivores and granivores) are negatively associated with urbanization.
+Our results suggest that the enhancement and persistence of avian biodiversity in urban environments relies on maintaining and developing a suite of habitats to serve a range of life history strategies.
 
@@
 # DISCUSSION
-Niche breadth, as measured by different types of generalism, was significantly associated with urbanization in our analyses, evidenced by the robustness of this result across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
+Niche breadth, as measured by different types of generalism, was a significantly important life-history characteristic associated with the most urban-tolerant species in our analyses.
+This result was robust across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
 S3).
@@
 
-The non-random distribution of the urbanization index across the phylogeny (Table 1) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
+The non-random distribution of the urbanization index across the phylogeny (Table 1) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that particular phylogenetic constraints allow certain species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
 This finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
 It's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny and we demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models.
-For example, some predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
+For example, some predictor variables became less crucial in the phylogenetic models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
 For these variables, it is likely that there may be particular clades which include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].
@@
 Furthermore, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
-While some studies have found a lack of evidence that residual brain size is an important predictor in urban birds [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed heavily correlated with urbanization in birds [@moller2009successful; @maklakov2011brains; @moller2015brain].
-This finding highlights the necessity to investigate the trait-urbanization relationships in a phylogenetically informed manner.
+While some studies have found a lack of evidence that residual brain size is an important predictor in urban-tolerant species [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed heavily correlated with urban-tolerance [@moller2009successful; @maklakov2011brains; @moller2015brain].
+This finding highlights the importance of investigating trait-urbanization relationships in a phylogenetically informed manner.
 
@@
 Urbanization appears to generally select against insectivorous species [@chace2006urban; @evans2011makes].
-Although we did not find any strong support for our hypotheses on different nesting types, we did find that in general species which tended to be more social were closely associated with urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
-While these patterns may be generalizable, we also found patterns in our analysis which seem to contrast the global norm.
+Although we did not find any strong support for our hypotheses on different nesting types, we found that species which tended to be more social were closely associated with urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
+While the aformentioned patterns may be generalizable, we also found patterns in our analysis which seem to contrast the global norm.
 For example, we found that urbanization selects against granivorous species, contrasting the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
 
-As illustrated (above and in Appendix 1), many traits have been used to identify a species' adaptability to urban environments, but results have frequently differed among studies.
+As illustrated above and in Appendix 1, many traits have been used to identify a species' adaptability to urban environments, but results have frequently differed among studies.
 As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization.
-Our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done).
+Our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done; e.g., [@kark2007living]).
 The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
-This analysis was made possible through the use of global broad-scale empirical datasets, which are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
+Our generalizable analysis was made possible through the use of global broad-scale empirical datasets, which are rapidly expanding our knowledge of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
 This is largely due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
-We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds.
-We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor.
+We did not explicitly account for changes throughout the annual cycle in a species' association with urban environments [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor.
 We highlight that future studies, especially within migrant-dominated landscapes (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full; @marra2015call].
 Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation; @batary2018non]).
-Fully understanding the suite of specific species' traits which are associated with urban birds, and placing them in a theoretical landscape ecology context, will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].
+Fully understanding the suite of traits which are associated with urban birds, and placing them in a theoretical landscape ecology context, will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].
+We highlight that future studies investigating the trait-urbanization relationship should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance.
+This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species.
 
 We provide the largest study to date using a continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
-Our results suggest that generalist species, with large residual brain size, are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
-We highlight that future studies investigating the trait-urbanization relationship should focus on broad-scale data  to form generalizable patterns which rely on continuous measures for improved predictive performance.
-This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species.
+Our results suggest that generalist species, with high clutch sizes, and large residual brain sizes, are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
 In the context of our specific findings, Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management.
@@
 
-```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance value distributions, on a logarithmic scale."}
+```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance value distributions, on a logarithmic scale.
+These species were chosen because they represent species with differnt distributional responses."}
 knitr::include_graphics("finalFigs/Figure1.pdf")
@@
 
-```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a logarithmic scale, of the response variable for the 477 species included in the analysis, with five example species highlighted."}
+```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a logarithmic scale, of the urbanization index for the 477 species included in the analysis, with five example species highlighted.
+The urbanization index was taken as the median of the distribution for a species' average radiance values."}
 knitr::include_graphics("finalFigs/Figure2.pdf")
@@
 
-```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree of the birds with the tips colored based on the urbanization index.
-The urbanization index is also shown based on the size of the bar associated with a species."}
+```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree, based on Jetz et al.
+(2012) of the 477 bird species in the analysis.
+The tree demonstrates the urbanization index for each species in two ways: (1) coloring of the tips where dark red equates to a low urbanization index and dark blue equates to a high urbanization index, and (2) the size of the barplot portrayed for a species at its tip show the relative urbanization index.
+The urbanization index was taken as the median of the distribution for a species' average radiance values.
+Figure S2 shows a reference phylogenetic tree with the names of each tip displayed, and this link () provides an interactive visualization in order to easily identify each of the 477 species in the tree.
+For a full list of the 477 species included and their urbanization index see Appendix 4."}
 knitr::include_graphics("finalFigs/Figure3.pdf")
@@
 A star represents a significant effect.
-Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set.
+Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set which was model-averaged.
 Full model results can be found in Appendix 5."}
@@
 
-```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for parameter estimates from the non-phylogenetically constrained and phylogenetically constrained models."}
+```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for standardized parameter estimates from the non-phylogenetically controlled and phylogenetically controlled global models."}
 knitr::include_graphics("finalFigs/Figure5.pdf")
@@
 library(pander)
-pander(table, caption = "Results showing the significance of the phylogenetic signal analyses.", split.table=Inf)
+pander(table, caption = "Results showing the significance of the phylogenetic signal analyses, for a suite of indices.", split.table=Inf)
 
@@
 
-```{r Figure S1, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the modelling procedures."}
+```{r Figure S1, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the analysis."}
 knitr::include_graphics("finalFigs/FigureS1.pdf")
@@
 
-```{r Figure S3, echo=FALSE, fig.cap="Phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
+```{r Figure S3, echo=FALSE, fig.cap="An assessment of phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
 knitr::include_graphics("finalFigs/FigureS3.pdf")

### 2018-05-18 ms.Rmd
waxing poetically by adding some results of standardized parameter estimates in (sensu Richard Major)

@@
 Our first objective was to investigate the presence of a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
-Our second objective was to develop models which were and were not phylogenetically constrained to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables.
+Our second objective was to develop models which were and were not phylogenetically controlled to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables.
 
@@
 
-### Non-phylogenetically constrained models
+### Non-phylogenetically controlled models
 Given that we hypothesized any of the potential candidate traits could significantly drive the response in our urbanization index, we first ran one global model including all 22 predictor variables.
@@
 ### Non-phylogenetic models
-The most important predictors of urbanization in the non-phylogenetic global model (Fig.
-4; Appendix 5) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006).
-Urban-tolerant birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032).
+The most important predictors of urbanization (i.e., the largest parameter estimates - shown in parantheses) in the non-phylogenetic global model (Fig.
+4; Appendix 5) were feeding habitat generalism (0.79), breeding habitat generalism (0.76), and clutch size (0.73).
+Further, significance was found for feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006).
+Urban-tolerant birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032), but less so than the previous traits.
 Conversely, urban-tolerant birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), group foraging (t = -2.24, p =0.025), granivorous diet (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), insectivorous diet (t = -3.96, p < 0.001), and solitary feeding (t = -2.90, p = 0.004).
+The strongest negative responses, in rank of their parameter estimates, were solitary feeding (-0.92), insectivorous diet (-0.85), grassland/shrubland habitat association (-0.733), and tree/forest habitat association (-0.63).
 
@@
 4; Appendix 5).
+However, the most important variable, based on standardized parameter estimates was clutch size (0.75), followed by birds associated with agricultural habitats (0.58), and then diet generalism (0.28).
+Breeding habitat (0.15) and feeding habitat (0.12) generalism parameter estimates were substantially smaller than in the non model-averaged approach.
 Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors.
@@
 Urban-tolerant birds remained significantly negatively associated with solitary feeding (z = 3.39, p = 0.001),  insectivorous diet (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), group foraging (z = 3.16, p = 0.002), granivorous diet (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).
+Solitary feeding (-1.00) remained the strongest negatively associated trait with urbanization, followed by insectivorous diet (-0.98).
 
-### Phylogenetically constrained models
-The most significant predictors of urbanization in the phylogenetically controlled global model (Fig.
-4; Appendix 5) were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism ( t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005).
+### Phylogenetically controlled models
+Brain residual size (1.25) had the largest standardized parameter estimate for the phylogenetically controlled global model (Fig.
+4; Appendix 5), followed by feeding habitat generalism (1.18), clutch size (0.91), and body size (0.82).The most significant predictors of urbanization in the phylogenetically controlled global model  were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism (t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005).
 Nest generalism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding (t = -3.25, p = 0.001), and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.
 
-When the phylogenetic global model was model-averaged, the patterns were similar (Fig.
-4; Appendix 5).
-Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significant predictors of urbanization index.
-Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization.
+Similarly, upon model-averaging (Fig.
+4; Appendix 5), feeding habitat generalism (1.30), clutch size (1.15), body size (0.66), and residual brain size (0.32) showed the strongest association with urbanization.
+Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were all statistically significant.
+Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization, with no known nest aggregation (-1.22), and species associated with grasslands/shrublands (-1.16) showing the strongest negative response.
 
 ### Differences in non-phylogenetic and phylogenetic models
-The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the statistical significance of individual predictors (Fig.
+The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates (Appendix 5), and the statistical significance of individual predictors (Appendix 5; Fig.
 4).
 However, residual brain size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig.
-5).
-Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger standardized parameter estimates) in the phylogenetic models compared with the non-phylogenetic models (Fig.
+5), showing the strongest relative response in the phylogenetic global model.
+Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger standardized parameter estimates) in the phylogenetic models compared with the non-phylogenetic models (Appendix 5; Fig.
 5).

### 2018-05-20 ms.Rmd
some more tweaks and changes based on RM comments

@@
 ---
-title: "A continuous urbanization index to assess phylogenetic adaptations in urban bird living"
+title: "A continental-scale analysis finds generalists, with large clutch size and residual brain mass, are the most urban-tolerant of birds"
 
@@
 # ABSTRACT
-Understanding the suite of ecological and life history traits which influence a species' tolerance to urbanization enables a predictive ability to elucidate those species which are most likely to persist in our increasingly urbanizing world.
-Previous work investigating these relationships has found contrasting patterns for a trait's association with urbanization.
+Understanding the suite of ecological and life history traits which influence a species' tolerance to urbanization is critical to understand which species are most likely to persist in our increasingly urbanizing world.
+Previous work investigating these relationships has found a wide array of contrasting patterns for a given trait's association with urbanization.
 In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we provide a novel and scalable methodology which assesses the ecological and life history traits which most influence a species' adaptability to persist in urban environments.
@@
 Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting with one another.
-Indeed, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains] but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on species' response to urbanization.
-Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011makes with @moller2009successful & @croci2008does] and niche breadth [cf., @kark2007living with @evans2011makes].
-Forming generalizable patterns, at multiple spatial scales [@suarez2018scale], which associate ecological and life history traits with urbanization in birds is critical to help curb the impact urbanization has on avian biodiversity [@sol2017urbanisation].
+Indeed, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains], but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on a species' response to urbanization.
+Similarly, contrasting results have been reported for annual fecundity [cf., @croci2008does; @moller2009successful; @evans2011makes] and niche breadth [cf., @kark2007living; @evans2011makes].
+Forming generalizable patterns which associate ecological and life history traits with response to urbanization in birds, at multiple spatial scales [@suarez2018scale], is critical to help curb the impact urbanization has on avian biodiversity [@sol2017urbanisation].
 
-The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even binary grouping based on presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
+The majority of previous studies assign species to discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even binary grouping based on presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
 These types of discrete response variables severely limits the capacity to generalize results as it assumes species within a group respond to urbanization equally [@evans2011makes].
-Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
+Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., @kark2007living; @croci2008does], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
 In a notable exception, Evans *et al.* [-@evans2011makes] advocates for a continuous measure of urbanization in which a species is assessed on its entire distribution in response to urban environments.
@@
 The aim of this study is to determine whether there is a suite of life history traits that predicts avian tolerance to urbanization at a continental scale.
-For such an analysis to have generality it must (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3), include a broad-range of traits, and (4) employ a continuous measure of a specie's response to urbanization.
-We conduct an analysis that meets these criteria, using a series of publicly available datasets to implement a novel continuous urbanization index in order to assess ecological and life history traits which most influence a bird's response to urbanization.
-We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against).
+For such an analysis to have generality it must (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3), include a broad-range of traits, and (4) employ a continuous measure of response to urbanization.
+We conduct an analysis that meets these criteria, integrating a series of publicly available datasets to implement a novel continuous urbanization index in order to assess ecological and life history traits which most influence a bird's response to urbanization.
 Our first objective was to investigate the presence of a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
 Our second objective was to develop models which were and were not phylogenetically controlled to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables.
+We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against).
 
@@
 This was done based on the assumption that those species which had the most observations would have the most reliable urbanization indices.
-This approach also addresses biases in sampling of eBird data in which some birds can have a high number of observations, but from known roosting sites [e.g., @callaghanavian].
+This approach also addresses biases in sampling of eBird data in which some birds can have a high number of observations, but from known roosting sites [e.g., @callaghan2018avian].
 
@@
 Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments.
-These results suggest that urbanization is, in general, excluding specialist species [@sorace2009distribution].
-As such, future urban development aimed at increasing avian diversity in urban areas should focus on maintaining a high diversity of habitat [@melles2003urban] both for foraging [@shochat2004linking], and nesting [@fernandez2001habitat], with a focus on intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role].
-A species-specific approach to assess particular resources necessary for specialist species may be necessary [@fernandez2004spatial].
-Furthermore, a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], which will in turn benefit avian biodiversity in urban environments.
+These results suggest that the process of urbanization is, in general, operating to exclude specialist species [@sorace2009distribution].
+Given specialist species are known to have strong negative responses to landscape fragmentation and disturbance [@devictor2008distribution], sustainable urban development with a goal of maintaining biological diversity in urban areas should focus on providing intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role].
+A species-specific approach, which assesses particular resources necessary for specialist species, for example focusing on foraging [@shochat2004linking] and nesting [@fernandez2001habitat] requirements, is likely necessary.
+We recommend that a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more].
 
@@
 This finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
-It's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny and we demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models.
-For example, some predictor variables became less crucial in the phylogenetic models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
-For these variables, it is likely that there may be particular clades which include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].
+Unsurprisingly, our larger sample size found a non-random pattern in urbanization across the bird phylogeny and we further demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models.
+For example, some predictor variables became less important in the phylogenetic models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
+For these variables, it is likely that there may be particular clades which include a number of species that are associated with both that particular trait and degree of urbanization [see @felsenstein1985confidence].
 Conversely, some of the most important predictor variables (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization when phylogeny was accounted for, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
@@
 Although we did not find any strong support for our hypotheses on different nesting types, we found that species which tended to be more social were closely associated with urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
-While the aformentioned patterns may be generalizable, we also found patterns in our analysis which seem to contrast the global norm.
-For example, we found that urbanization selects against granivorous species, contrasting the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
+While the aforementioned patterns may be generalizable, we also found patterns in our analysis which seem to contrast the global norm.
+For example, we found that urbanization selects against granivorous species, contrasting with the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
 
@@
 The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
-Our generalizable analysis was made possible through the use of global broad-scale empirical datasets, which are rapidly expanding our knowledge of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
+Our broad-ranging analysis was made possible through the integration of broad-scale empirical datasets, which are rapidly expanding our knowledge of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
 This is largely due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
 We did not explicitly account for changes throughout the annual cycle in a species' association with urban environments [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor.
-We highlight that future studies, especially within migrant-dominated landscapes (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full; @marra2015call].
-Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation; @batary2018non]).
-Fully understanding the suite of traits which are associated with urban birds, and placing them in a theoretical landscape ecology context, will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].
+We emphasize that future studies, especially within migrant-dominated systems (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full; @marra2015call].
+Further, future studies should look to expand this analysis to other continents, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation; @batary2018non]).
+Fully understanding the suite of traits which are associated with urban-tolerant species, and placing them in a theoretical landscape ecology context will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes].
 We highlight that future studies investigating the trait-urbanization relationship should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance.
@@
 We provide the largest study to date using a continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
-Our results suggest that generalist species, with high clutch sizes, and large residual brain sizes, are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
+Our results suggest that generalist species, with high clutch sizes and large residual brain sizes, are best prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
 In the context of our specific findings, Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management.
-Urban avian biodiversity could be increased by improving the suitability of foraging for granivores and insectivores, especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more].
+Urban avian biodiversity could be increased by emphasizing the critical importance of intact vegetation [@sushinsky2013should] for specialist species [@devictor2008distribution], and concomitantly improving the suitability of foraging for granivores and insectivores.
 

### 2018-05-21 scripts/results_script.R
updating supp figures for paper

@@ -368,18 +368,21 @@ my.file.copy(from = "figures/phy_v_non_phy.pdf",
 ##########################################
 
 ## Figure S1
-my.file.copy(from = "figures/corrplot_of_continuous_variables.pdf",
-             to = "finalFigs/FigureS1.pdf")
-
+## made using google earth engine, outside of R
 
 ## Figure S2
-my.file.copy(from = "figures/ref_tree.pdf",
+my.file.copy(from = "figures/corrplot_of_continuous_variables.pdf",
              to = "finalFigs/FigureS2.pdf")
 
 
 ## Figure S3
-my.file.copy(from = "figures/accounting_for_phylo_uncertainty.pdf",
+my.file.copy(from = "figures/ref_tree.pdf",
              to = "finalFigs/FigureS3.pdf")
+
+
+## Figure S4
+my.file.copy(from = "figures/accounting_for_phylo_uncertainty.pdf",
+             to = "finalFigs/FigureS4.pdf")
 
 
 

### 2018-05-21 ms.Rmd
more minor tweaks to manuscript

@@
 After this sampling, we were left with 637,482 eligible checklists throughout Australia.
-
 Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we did not expect any correlation with seabirds using urban areas under normal circumstances.
@@
 This left us with 580 potential species from the eBird dataset which had > 100 observations on eligible checklists in the specified time frame, which was a substantial proportion of the ~730 bird species which are resident in Australia.
-
 Given the diverse datasets (see below) used in this analysis, there were differing taxonomic authorities followed (see Appendix 2 for elaboration).
@@
 We used VIIRS night-time lights [@elvidge2017viirs] as a proxy of a bird's association with urban habitat.
-The VIIRS night-time lights measures the electric lighting present on Earth's surface which is generally associated with human settlement.
+The VIIRS night-time lights measures the electric lighting present on Earth's surface which is generally associated with human settlement (Fig.
+S1).
 There are also a significant number of steps (e.g., filtering out background noise, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@elvidge2017viirs].
@@
 Any radiance values which were negative were forced to 0.00001 [*sensu* @ou2015evaluation].
-
 Each checklist's average radiance value was then passed to every bird on that checklist, which provided every species with a distribution of average radiance values (Fig.
 1).
-The median for each species' given distribution of average radiance values served as our response variable in our analyses.
+The median for each species' given distribution of average radiance values served as the response variable in our analyses.
 
@@
 Before modelling, the predictor variables were investigated for multi-collinearity.
-We used the corrplot package [@corrplot2017] to show correlation among continuous variables (Supplementary Figure 1), and used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other (Supplementary Table 1).
+We used the corrplot package [@corrplot2017] to show correlation among continuous variables (Fig.
+S2), and used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other (Supplementary Table 1).
 
@@
 ### Phylogenetic analysis
-
 For the phylogenetic portion of our analyses, we used the phylogenetic trees provided by Jetz et al.
@@
 A total of 316,306 checklists resulting in 5,944,819 observations of 477 different species were included in analyses.
-The least often recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615).
+The least often recorded species was *Turnix melanogaster* (Black-breasted Buttonquail; N=104) while the most recorded species was *Cracticus tibicen* (Australian Magpie; N=158,615).
 The mean number of observations for a species was 10,250.
@@
 ## Urbanization index
-For the 477 species included in the models the mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, with a range from 0.008 (Black-backed Butcherbird) to 11.85 (Powerful Owl; Fig. 2).
+For the 477 species included in the models the mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, with a range from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig.
+2).
 
@@
 3; Fig.
-S2; Table 1), as it was significantly different than random for 4 out of the 5 indices considered.
+S3; Table 1), as it was significantly different than random for 4 out of the 5 indices considered.
 Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1).
@@
 This result was robust across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
-S3).
+S4).
 Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments.
@@
 Conversely, some of the most important predictor variables (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization when phylogeny was accounted for, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
-S3).
+S4).
 Furthermore, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
@@
 As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization.
-Our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done; e.g., [@kark2007living]).
+Our models assessed the applicability of traits in relation to one another [i.e., we did not run separate models or statistical tests for each specific trait as previously done; e.g., @kark2007living].
 The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
@@
 We emphasize that future studies, especially within migrant-dominated systems (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full; @marra2015call].
-Further, future studies should look to expand this analysis to other continents, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation; @batary2018non]).
+Further, future studies should look to expand this analysis to other continents, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale], by investigating the effect of city size [@sol2017urbanisation; @batary2018non] on these patterns.
 Fully understanding the suite of traits which are associated with urban-tolerant species, and placing them in a theoretical landscape ecology context will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes].
@@
 ```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance value distributions, on a logarithmic scale.
-These species were chosen because they represent species with differnt distributional responses."}
+These species were chosen because they represent species with different distributional responses."}
 knitr::include_graphics("finalFigs/Figure1.pdf")
@@
 ```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree, based on Jetz et al.
-(2012) of the 477 bird species in the analysis.
+(2012) of the 477 bird species in the analysis, demonstrating the clustering of the urbanization index in regards to phylogeny.
 The tree demonstrates the urbanization index for each species in two ways: (1) coloring of the tips where dark red equates to a low urbanization index and dark blue equates to a high urbanization index, and (2) the size of the barplot portrayed for a species at its tip show the relative urbanization index.
 The urbanization index was taken as the median of the distribution for a species' average radiance values.
-Figure S2 shows a reference phylogenetic tree with the names of each tip displayed, and this link () provides an interactive visualization in order to easily identify each of the 477 species in the tree.
+Figure S3 shows a reference phylogenetic tree with the names of each tip displayed, and this link () provides an interactive visualization in order to easily identify each of the 477 species in the tree and their respective urbanization indices.
 For a full list of the 477 species included and their urbanization index see Appendix 4."}
@@
 
-```{r Figure S1, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the analysis."}
+```{r Figure S1, echo=FALSE, fig.cap="The VIIRS nightime-lights shown for Australia, Brisbane, Melbourne, and Sydney. Values are measured as average radiance shown in red."}
 knitr::include_graphics("finalFigs/FigureS1.pdf")
@@
 
-```{r Figure S2, echo=FALSE, fig.cap="A reference phylogenetic tree, with labeled tips, for each of the 477 species included in the analysis. Appendix 4 lists the 477 species."}
+```{r Figure S2, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the analysis."}
 knitr::include_graphics("finalFigs/FigureS2.pdf")
@@
 
-```{r Figure S3, echo=FALSE, fig.cap="An assessment of phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
+```{r Figure S3, echo=FALSE, fig.cap="A reference phylogenetic tree, with labeled tips, for each of the 477 species included in the analysis. Appendix 4 lists the 477 species."}
 knitr::include_graphics("finalFigs/FigureS3.pdf")
+```
+
+```{r Figure S4, echo=FALSE, fig.cap="An assessment of phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
+knitr::include_graphics("finalFigs/FigureS4.pdf")
 ```

### 2018-05-21 ms.Rmd
added link for web-app

@@
 The urbanization index was taken as the median of the distribution for a species' average radiance values.
-Figure S3 shows a reference phylogenetic tree with the names of each tip displayed, and this link () provides an interactive visualization in order to easily identify each of the 477 species in the tree and their respective urbanization indices.
+Figure S3 shows a reference phylogenetic tree with the names of each tip displayed, and this link (https://johnwilshire.github.io/bird_urbanness_web_app/) provides an interactive visualization in order to easily identify each of the 477 species in the tree and their respective urbanization indices.
 For a full list of the 477 species included and their urbanization index see Appendix 4."}

### 2018-05-29 ms.Rmd
some more phylo justification

@@
 ### Phylogenetic analysis
-For the phylogenetic portion of our analyses, we used the phylogenetic trees provided by Jetz et al.
+As the correlation between traits and urbanization may be influenced by the shared evolutionary history of the species [felsenstein1985phylogenies], we also a used phylogenetically informed model to test the hypothesis.
+Phylogenetically informed models are descendents of phylogenetic indepedent contrasts [felsenstein1985phylogenies], which in simple terms, tests for repeated evolutionary association between the traits of interest.
+Or in this particular case, whether the traits of interest (e.g.
+large brain size at a given body size) repeatedly allow for the use of urban areas wherever it evolved across the bird phylogeny.
+We used the phylogenetic trees provided by Jetz et al.
 [@jetz2012global].
@@
 Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.
-We also  fit the same model to a range of the equally likely trees from [@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.
+To account for aspects of phylogenetic uncertainty, we also fit the same model to a range of the equally likely trees from [@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.
 

### 2018-05-29 ms.Rmd
rewrite of first para. of introduction

@@
 # INTRODUCTION
-Urbanization fragments and isolates natural habitat [@marzluff2001restoration; @melles2003urban; @loram2007urban], drastically altering species' composition and ecosystem functions.
-Thus, urbanization globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects] and is a major concern for conservation [@czech2000economic; @dearborn2010motivations].
-Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation].
-Conversely, some species rarely occur within urban areas, despite being relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
+Species specialization is a fundamentally important biological trait which can dictate population-level responses to changing environments.
+Generally, species with limited niche breadth experience higher extinction risk [@boulangeat2012niche].
+Further, certain organisms can more readily adapt their niches in response to changing environments leading to potential ecological inheritance within specific species [@odling2013niche].
+The process of urbanization drastically fragments and isolates natural habitats [@marzluff2001restoration; @melles2003urban; @loram2007urban], providing ecologically novel landscapes for the species within.
+Indeed, some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others remain relatively common in surrounding habitats but rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
+Urbanization, by selecting for a specific set of species, globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], which is a major concern for conservation [@czech2000economic; @dearborn2010motivations].
+Hence, it is important to fully understand the suite of ecological and life history traits most associated with birds that are successful or unsuccessful in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
+Only then will the species most at-risk of extinction, and thus in need of conservation action, be best understood.
 
-In order to best predict species' responses to future urbanization changes, it is essential to understand what ecological and life history traits are most associated with birds that are either successful or unsuccessful in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
 To date, a wide array of traits have been used to predict presence of birds in urban environments.

### 2018-05-29 ms.Rmd
adding a bit more polish to the intro

@@
 Hence, it is important to fully understand the suite of ecological and life history traits most associated with birds that are successful or unsuccessful in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
-Only then will the species most at-risk of extinction, and thus in need of conservation action, be best understood.
+Only then will the species most at-risk of extinction, and thus in need of conservation attention, be best understood and appropriate action subsequently taken (e.g., necessary habitat improvements).
 
-To date, a wide array of traits have been used to predict presence of birds in urban environments.
+To date, a wide array of ecological and life history traits have been used to predict presence of birds in urban environments.
 For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does; @conole2011functional], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have been hypothesized to influence a species' presence in urban environments.
-Rather than a specific trait that predicts a species' ability to be found in urban environments, it is most likely a combination of traits.
-For instance, the presence of urban birds in Jerusalem was predicted by traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
-Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting with one another.
-Indeed, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains], but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on a species' response to urbanization.
+For instance, the presence of urban birds in Jerusalem was best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
+Notably, a wide range of often contrasting results using aforementioned traits have been reported in the literature.
+For instance, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains], but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on a species' response to urbanization.
 Similarly, contrasting results have been reported for annual fecundity [cf., @croci2008does; @moller2009successful; @evans2011makes] and niche breadth [cf., @kark2007living; @evans2011makes].
-Forming generalizable patterns which associate ecological and life history traits with response to urbanization in birds, at multiple spatial scales [@suarez2018scale], is critical to help curb the impact urbanization has on avian biodiversity [@sol2017urbanisation].
 
@@
 These types of discrete response variables severely limits the capacity to generalize results as it assumes species within a group respond to urbanization equally [@evans2011makes].
-Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., @kark2007living; @croci2008does], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
+Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., @kark2007living; @croci2008does; @geschkecompact], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
 In a notable exception, Evans *et al.* [-@evans2011makes] advocates for a continuous measure of urbanization in which a species is assessed on its entire distribution in response to urban environments.
@@
 Furthermore, despite the number of studies which have investigated the relationships between traits and urbanization, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living].
-There is a need for broad-scale empirical datasets, which allow us to test hypotheses at large spatiotemporal scales, and citizen science is beginning to provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen].
+There is a need for broad-scale empirical datasets, which allow us to test hypotheses at large spatiotemporal scales to make generalizable patterns, and citizen science can provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen].
 Indeed, various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
 
-The aim of this study is to determine whether there is a suite of life history traits that predicts avian tolerance to urbanization at a continental scale.
+The aim of this study was to determine whether there is a suite of life history traits that predicts avian tolerance to urbanization at a continental scale.
+We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against).
 For such an analysis to have generality it must (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3), include a broad-range of traits, and (4) employ a continuous measure of response to urbanization.
-We conduct an analysis that meets these criteria, integrating a series of publicly available datasets to implement a novel continuous urbanization index in order to assess ecological and life history traits which most influence a bird's response to urbanization.
+We conducted an analysis that meets these criteria, integrating a series of publicly available datasets to implement a novel continuous urbanization index to assess ecological and life history traits which most influence a bird's response to urbanization.
 Our first objective was to investigate the presence of a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
 Our second objective was to develop models which were and were not phylogenetically controlled to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables.
-We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against).
 

### 2018-05-29 ms.Rmd
some polish and addressing RK and JM comments and some editing of figure captions

@@
 S1).
-There are also a significant number of steps (e.g., filtering out background noise, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@elvidge2017viirs].
+There are also a significant number of steps (e.g., filtering out background noise such as fire, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@elvidge2017viirs].
 For the 637,482 checklists which met our criteria above, we used the checklists' associated spatial coordinates and Google Earth Engine [@gorelick2017google] to assign the average radiance value calculated from the VIIRS layer.
@@
 The median for each species' given distribution of average radiance values served as the response variable in our analyses.
+Inherently, there is no definitive range for this index, as it solely depends on a species' distribution of observations.
 
 ## Biological traits
-Predictor variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological].
-Potential traits (Appendix 1a) were chosen on the basis of their existing support and hypotheses in the current literature (e.g., [@mcclure1989characterizes; @kark2007living; @evans2011makes]).
+Many biological and life history traits have been used in previous studies, and potential traits in our study were chosen on the basis of their existing support and hypotheses in the current literature [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes]; we sought to encompass a broad array of traits previously tested.
+Hence, predictor variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological].
+Table 1 provides a brief overview of the traits that we investigated, while Appendix 1 Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated.
 Traits were parsed in different manners (Appendix 1b), and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat).
-Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated.
 Preliminary analyses including IUCN status (exotic/native) showed that exotic birds were strongly associated with urbanization, but given the preferential selection by acclimatization societies for introduction of species with proven synanthropy [@blackburn2009avian], exotic species were omitted from the final analysis to avoid bias.
-Therefore, we excluded IUCN status as a trait given that the majority of our candidate species were of Least Concern.
-Only species which had complete data for the list of candidate traits were considered in the analyses, and we thereby ended with 477 species which met this criteria.
+Therefore, we excluded IUCN status as a trait because the majority of our candidate species were of Least Concern.
+Only species which had complete data for the list of candidate traits were considered in the analyses, and 477 species met this criteria.
 
@@
 Our urbanization index was log-transformed in order to meet model assumptions.
-The predictor variables used in the models were those listed in Appendix 1, with the exception of IUCN status.
+The predictor variables used in the models were those listed in Table 1.
 
@@
 This was done based on the assumption that those species which had the most observations would have the most reliable urbanization indices.
-This approach also addresses biases in sampling of eBird data in which some birds can have a high number of observations, but from known roosting sites [e.g., @callaghan2018avian].
+This approach also addresses biases in sampling of eBird data in which some birds can have a high number of observations, but from a small number of locations such as known roosting sites [e.g., @callaghan2018avian].
 
@@
 Given that we hypothesized any of the potential candidate traits could significantly drive the response in our urbanization index, we first ran one global model including all 22 predictor variables.
-We then used the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes.
+We used the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes.
 In order to confirm our linear model results, we employed a model-averaging approach [@grueber2011multimodel] in which we 'dredged' possible subsets of the global model.
@@
 ### Phylogenetic analysis
-As the correlation between traits and urbanization may be influenced by the shared evolutionary history of the species [felsenstein1985phylogenies], we also a used phylogenetically informed model to test the hypothesis.
-Phylogenetically informed models are descendents of phylogenetic indepedent contrasts [felsenstein1985phylogenies], which in simple terms, tests for repeated evolutionary association between the traits of interest.
-Or in this particular case, whether the traits of interest (e.g.
-large brain size at a given body size) repeatedly allow for the use of urban areas wherever it evolved across the bird phylogeny.
+As the correlation between traits and urbanization may be influenced by the shared evolutionary history of the species [@felsenstein1985phylogenies], we also used phylogenetically informed models to test our hypotheses.
+Phylogenetically informed models are descendants of phylogenetic independent contrasts [@felsenstein1985phylogenies], which in simple terms, tests for repeated evolutionary association between the traits of interest.
+Or in this particular case, whether the traits of interest (e.g., large brain size at a given body size) repeatedly allow for the use of urban environments wherever it evolved across the bird phylogeny.
 We used the phylogenetic trees provided by Jetz et al.
@@
 Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.
-To account for aspects of phylogenetic uncertainty, we also fit the same model to a range of the equally likely trees from [@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.
+To account for aspects of phylogenetic uncertainty, we also fit the same model to a range of equally likely trees from Jetz *et al.* [-@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.
 
@@
 ## Urbanization index
-For the 477 species included in the models the mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, with a range from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig.
+For the 477 species included in the analysis the mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, with a range from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig.
 2).
@@
 3; Fig.
-S3; Table 1), as it was significantly different than random for 4 out of the 5 indices considered.
-Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1).
+S3; Table 2), as it was significantly different than random for 4 out of the 5 indices considered.
+For example, note the distinct signal located throughout the waterbirds' clades in Figure 3 or the significantly large index for *Manorina melanocephala*, Noisy Miner, compared to the rest of the honeyeaters.
+Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 2).
 The result for body size ($k=3.3$) is among the strongest ever recorded measures of phylogenetic signal in the literature, and for $k$ the statistic suggests much stronger signal than expected from Brownian motion.
 We generally found weaker signal for behavioral traits compared with morphological traits.
-For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 1).
+For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 2).
 
@@
 ```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a logarithmic scale, of the urbanization index for the 477 species included in the analysis, with five example species highlighted.
-The urbanization index was taken as the median of the distribution for a species' average radiance values."}
+The urbanization index was taken as the median of the distribution for a species' average radiance values, as demonstrated in Figure 1."}
 knitr::include_graphics("finalFigs/Figure2.pdf")
@@
 (2012) of the 477 bird species in the analysis, demonstrating the clustering of the urbanization index in regards to phylogeny.
-The tree demonstrates the urbanization index for each species in two ways: (1) coloring of the tips where dark red equates to a low urbanization index and dark blue equates to a high urbanization index, and (2) the size of the barplot portrayed for a species at its tip show the relative urbanization index.
+The tree demonstrates two representations of the urbanization index for each species: (1) coloring of the tips where dark red equates to a low urbanization index and dark blue equates to a high urbanization index, and (2) the size of the barplot at a tip shows the relative magnitude of the urbanization index.
 The urbanization index was taken as the median of the distribution for a species' average radiance values.
@@
 
-```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for standardized parameter estimates from the non-phylogenetically controlled and phylogenetically controlled global models."}
+```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for 22 standardized parameter estimates from the non-phylogenetically controlled and phylogenetically controlled global models, as modelled based on 477 species."}
 knitr::include_graphics("finalFigs/Figure5.pdf")
@@
 \small
-```{r Table 1, echo=FALSE, warning=FALSE, message=FALSE}
+```{r Table 2, echo=FALSE, warning=FALSE, message=FALSE}
 library(readr)
@@
 
-```{r Figure S1, echo=FALSE, fig.cap="The VIIRS nightime-lights shown for Australia, Brisbane, Melbourne, and Sydney. Values are measured as average radiance shown in red."}
+```{r Figure S1, echo=FALSE, fig.cap="The VIIRS nightime-lights shown for Australia, Brisbane (top right), Melbourne (bottom left), and Sydney (bottom right).
+Values are measured as average radiance, shown in red.
+For each checklist in our analysis, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included, and this led to the distributions shown in Figure 1."}
 knitr::include_graphics("finalFigs/FigureS1.pdf")
@@
 
-```{r Figure S3, echo=FALSE, fig.cap="A reference phylogenetic tree, with labeled tips, for each of the 477 species included in the analysis. Appendix 4 lists the 477 species."}
+```{r Figure S3, echo=FALSE, fig.cap="A reference phylogenetic tree, with labeled tips, based on Jetz et al. (2012) of the 477 bird species in the analysis. Appendix 4 lists the 477 species."}
 knitr::include_graphics("finalFigs/FigureS3.pdf")

### 2018-05-29 ms.Rmd
couple minor things

@@
 
-The non-random distribution of the urbanization index across the phylogeny (Table 1) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that particular phylogenetic constraints allow certain species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
+The non-random distribution of the urbanization index across the phylogeny (Table 2) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that particular phylogenetic constraints allow certain species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
 This finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
@@
 
-We provide the largest study to date using a continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
+We provide a novel continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
 Our results suggest that generalist species, with high clutch sizes and large residual brain sizes, are best prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
@@
 Thanks to Stephen Garnett and Glenn Ehmke for providing us with extent of occurrence estimates for each of the species in our analysis.
+Thanks to Corrine Edwards for providing the illustrations used in Fig.
+3.
 

### 2018-05-30 ms.Rmd
hacking at the discussion

@@
 # DISCUSSION
-Niche breadth, as measured by different types of generalism, was a significantly important life-history characteristic associated with the most urban-tolerant species in our analyses.
-This result was robust across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
+The most urban-tolerant species in our analyses (i.e., the ones with a higher urbanization index) had a larger niche breadth, as measured by different forms of generalism.
+Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) were less likely to be found in urban environments.
+These results were robust across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
 S4).
-Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments.
 These results suggest that the process of urbanization is, in general, operating to exclude specialist species [@sorace2009distribution].
-Given specialist species are known to have strong negative responses to landscape fragmentation and disturbance [@devictor2008distribution], sustainable urban development with a goal of maintaining biological diversity in urban areas should focus on providing intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role].
-A species-specific approach, which assesses particular resources necessary for specialist species, for example focusing on foraging [@shochat2004linking] and nesting [@fernandez2001habitat] requirements, is likely necessary.
-We recommend that a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more].
+Further, the non-random distribution of the urbanization index across the phylogeny (Fig.
+3 & Table 2) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that particular phylogenetic constraints allow certain species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
 
-The non-random distribution of the urbanization index across the phylogeny (Table 2) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that particular phylogenetic constraints allow certain species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
-This finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
-Unsurprisingly, our larger sample size found a non-random pattern in urbanization across the bird phylogeny and we further demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models.
-For example, some predictor variables became less important in the phylogenetic models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural).
-For these variables, it is likely that there may be particular clades which include a number of species that are associated with both that particular trait and degree of urbanization [see @felsenstein1985confidence].
-Conversely, some of the most important predictor variables (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization when phylogeny was accounted for, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
-S4).
-Furthermore, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
+Our finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is unsurprising given our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
+We further demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models, and thus highlight the importance of investigating trait-urbanization relationships in a phylogenetically informed manner.
+For example, some of the most important predictor variables in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
+S4); and other traits showed the opposite trend.
+Most notably, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
 While some studies have found a lack of evidence that residual brain size is an important predictor in urban-tolerant species [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed heavily correlated with urban-tolerance [@moller2009successful; @maklakov2011brains; @moller2015brain].
-This finding highlights the importance of investigating trait-urbanization relationships in a phylogenetically informed manner.
 
-Certain aspects of the trait-urbanization relationship appear to be consistent, and thus generalizable, across the globe.
+Many traits have been used to identify a species' adaptability to urban environments, but results have frequently differed among studies.
+As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization.
+Our models assessed the applicability of traits in relation to one another [i.e., we did not run separate models or statistical tests for each specific trait as previously done; e.g., @kark2007living] thus allowing comparable effect sizes.
+Despite this, there appear to be some consistent, and thus generalizable, trends within the trait-urbanization complex across the globe.
 For instance, our finding of the relative importance of annual fecundity is consistent with work in Europe [@croci2008does; @moller2009successful].
-Urbanization appears to generally select against insectivorous species [@chace2006urban; @evans2011makes].
-Although we did not find any strong support for our hypotheses on different nesting types, we found that species which tended to be more social were closely associated with urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
-While the aforementioned patterns may be generalizable, we also found patterns in our analysis which seem to contrast the global norm.
-For example, we found that urbanization selects against granivorous species, contrasting with the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden].
+Additionally, urbanization appears to generally select against insectivorous species [@chace2006urban; @evans2011makes].
+Although we did not find any strong support for our hypotheses on different nesting types, we found a positive association between the degree of sociality and urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
+While the aforementioned patterns may be generalizable, we also found patterns in our analyses which seem to contrast the global norm.
+For example, we found that urbanization selects against granivorous species, contrasting with the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [compared to European cities for instance; @fuller2008garden].
 
-As illustrated above and in Appendix 1, many traits have been used to identify a species' adaptability to urban environments, but results have frequently differed among studies.
-As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization.
-Our models assessed the applicability of traits in relation to one another [i.e., we did not run separate models or statistical tests for each specific trait as previously done; e.g., @kark2007living].
-The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization.
-Our broad-ranging analysis was made possible through the integration of broad-scale empirical datasets, which are rapidly expanding our knowledge of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
+Our analysis was made possible through integration of broad-scale empirical datasets, which are rapidly expanding our knowledge of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
 This is largely due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
 We did not explicitly account for changes throughout the annual cycle in a species' association with urban environments [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor.
-We emphasize that future studies, especially within migrant-dominated systems (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full; @marra2015call].
-Further, future studies should look to expand this analysis to other continents, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale], by investigating the effect of city size [@sol2017urbanisation; @batary2018non] on these patterns.
-Fully understanding the suite of traits which are associated with urban-tolerant species, and placing them in a theoretical landscape ecology context will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes].
-We highlight that future studies investigating the trait-urbanization relationship should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance.
+We emphasize that future studies, especially within migrant-dominated systems (e.g., Europe, North America), should account for changes in a species' urbanization index throughout the full-annual cycle [@hostetler2015full; @marra2015call].
 This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species.
 
+Fully understanding the suite of traits which are associated with urban-tolerant species, and placing them in a theoretical landscape ecology context will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes].
+Given specialist species are known to have strong negative responses to landscape fragmentation and disturbance [@devictor2008distribution], sustainable urban development with a goal of maintaining biological diversity in urban areas should focus on providing intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role].
+We recommend that a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more].
+A species-specific approach, which assesses particular resources necessary for specialist species, for example focusing on foraging [@shochat2004linking] and nesting [@fernandez2001habitat] requirements, is likely necessary.
+To this end, our urbanization index can be used as a potential means to provide quantifiable targets for local restoration projects.
+For example, a particular restoration project should aim to add species which have lower urbanization indices into their local avifauna, but we note that these indices may differ at different spatial scales [@suarez2018scale].
+
 We provide a novel continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
+Future studies investigating the trait-urbanization relationship should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance.
 Our results suggest that generalist species, with high clutch sizes and large residual brain sizes, are best prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].

### 2018-05-30 ms.Rmd
minor changes

@@
 # DISCUSSION
-The most urban-tolerant species in our analyses (i.e., the ones with a higher urbanization index) had a larger niche breadth, as measured by different forms of generalism.
+The most urban-tolerant species in our analyses (i.e., the ones with a higher urbanization index) had larger niche breadths, as measured by different forms of generalism.
 Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) were less likely to be found in urban environments.
@@
 
-Fully understanding the suite of traits which are associated with urban-tolerant species, and placing them in a theoretical landscape ecology context will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes].
+Fully understanding the suite of traits which are associated with urban-tolerant species will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes].
 Given specialist species are known to have strong negative responses to landscape fragmentation and disturbance [@devictor2008distribution], sustainable urban development with a goal of maintaining biological diversity in urban areas should focus on providing intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role].

### 2018-06-03 ms.Rmd
some work on phylo sig table

@@
 library(pander)
-pander(table, caption = "Results showing the significance of the phylogenetic signal analyses, for a suite of indices.", split.table=Inf)
+pander(table, caption = "Results showing the  phylogenetic signal for both predictor and response traits.
+Because there are a number of indices in the literature, with specific strengths and weaknesses, we present here all available indices (Keck et al.
+2016).", split.table=Inf)
 

### 2018-06-03 R/phylo_fxns.R

@@ -365,7 +365,7 @@ phylosignal_analysis <- function(analysis_data, aus_bird_tree){
                 nest_generalism,diet_generalism,breeding_habitat_generalism,feeding_habitat_generalism,brain_residual) ->dd
   
   row.names(dd)<-analysis_data$binom #name rows so that it matches the tree
-  dd$rand<-rnorm(dim(dd)[2]) #random numbers to test package
+  #dd$rand<-rnorm(dim(dd)[2]) #random numbers to test package
   p4d <- phylo4d(aus_bird_tree, dd) #create phylobase object
   
   ps <- phyloSignal(p4d,reps = 9999) #run calculation, p values a bit unstable at 999 reps
@@ -375,14 +375,14 @@ phylosignal_analysis <- function(analysis_data, aus_bird_tree){
   
   row.names(Table1) <- c("log(Body size)", "log(Clutch size)", "Urbanization index", "Range size (1000s km2)",
                          "Nest generalism", "Diet generalism", "Breeding habitat generalism",
-                         "Feeding habitat generalism", "Brain residual", "Rand")
+                         "Feeding habitat generalism", "Brain residual")
   
   
   TableS2 <- ps$pvalue
   
   row.names(TableS2) <- c("log(Body size)", "log(Clutch size)", "Urbanization index", "Range size (1000s km2)",
                           "Nest generalism", "Diet generalism", "Breeding habitat generalism",
-                          "Feeding habitat generalism", "Brain residual", "Rand")
+                          "Feeding habitat generalism", "Brain residual")
   
   write.csv(Table1, "Data/Table1.csv")
   write.csv(TableS2, "Data/TableS2.csv")

### 2018-06-03 ms.Rmd
more caption work

@@
 
-```{r Figure S4, echo=FALSE, fig.cap="An assessment of phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
+```{r Figure S4, echo=FALSE, fig.cap="To account for one aspect of phylogenetic uncertainty, we fits the same phylolm model to 1000 of the family of trees presented by Jetz et al.
+(2012).
+The density plots are the estimated parameters for most important variables across models fit to the collection trees from Jetz et al.
+(2012)"}
 knitr::include_graphics("finalFigs/FigureS4.pdf")

### 2018-06-04 ms.Rmd
caption work

@@
 In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we provide a novel and scalable methodology which assesses the ecological and life history traits which most influence a species' adaptability to persist in urban environments.
-We elucidate generalized, phylogenetically controlled, patterns: bird species which are specialists, with high clutch size, and large residual brain size are among the most urban-tolerant bird species.
+We elucidate generalized, phylogenetically controlled, patterns: bird species which are generalists, with high clutch size, and large residual brain size are among the most urban-tolerant bird species.
 Conversely, specialized feeding strategies (i.e., insectivores and granivores) are negatively associated with urbanization.
@@
 ```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a logarithmic scale, of the urbanization index for the 477 species included in the analysis, with five example species highlighted.
-The urbanization index was taken as the median of the distribution for a species' average radiance values, as demonstrated in Figure 1."}
+The urbanization index was taken as the median of the distribution for a species' average radiance values, as demonstrated in Figure 1.
+Figure S1 shows randomly chosen indices for a given point."}
 knitr::include_graphics("finalFigs/Figure2.pdf")
@@
 Values are measured as average radiance, shown in red.
-For each checklist in our analysis, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included, and this led to the distributions shown in Figure 1."}
+For each checklist in our analysis, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included, and this led to the distributions shown in Figure 1.
+We show random points and their average associated average radiance values calculated as the mean within a 5 km buffer of that point."}
 knitr::include_graphics("finalFigs/FigureS1.pdf")
@@
 
-```{r Figure S4, echo=FALSE, fig.cap="To account for one aspect of phylogenetic uncertainty, we fits the same phylolm model to 1000 of the family of trees presented by Jetz et al.
+```{r Figure S4, echo=FALSE, fig.cap="To account for one aspect of phylogenetic uncertainty, we fit the same phylolm model to 1000 of the family of trees presented by Jetz et al.
 (2012).
 The density plots are the estimated parameters for most important variables across models fit to the collection trees from Jetz et al.
-(2012)"}
+(2012)."}
 knitr::include_graphics("finalFigs/FigureS4.pdf")

### 2018-06-04 ms.Rmd
changed the title.

@@
 ---
-title: "A continental-scale analysis finds generalists, with large clutch size and residual brain mass, are the most urban-tolerant of birds"
+title: "Generalists are the most urban-tolerant of birds: a phylogenetically controlled analysis of ecological and life-history traits using a novel continuous urbanization index"
 

### 2018-06-04 ms.Rmd
Couple of changes and adding a sentence for #63

@@
 For such an analysis to have generality it must (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3), include a broad-range of traits, and (4) employ a continuous measure of response to urbanization.
-We conducted an analysis that meets these criteria, integrating a series of publicly available datasets to implement a novel continuous urbanization index to assess ecological and life history traits which most influence a bird's response to urbanization.
+We conducted an analysis that meets these criteria, integrating a series of datasets to implement a novel continuous urbanization index to assess ecological and life history traits which most influence a bird's response to urbanization.
 Our first objective was to investigate the presence of a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
@@
 
-We employed two linear modelling approaches; one with and one without phylogenetic constraints.
+We employed two linear modelling approaches to provide robustness of our results; one with and one without phylogenetic constraints.
+Phylogenetic models were used because of the importance of considering evolutionary distinctedness in ecological research [@losos2008phylogenetic; @graham2018phylogenetic], but given that phylogenetically designed analyses can have a number of shortcomings [@cadotte2017phylogenies], we also included non-phylogenetic models in our analysis.
 Our urbanization index was log-transformed in order to meet model assumptions.
-The predictor variables used in the models were those listed in Table 1.
+The predictor variables used in the models are listed in Table 1.
 
@@
 S4).
-These results suggest that the process of urbanization is, in general, operating to exclude specialist species [@sorace2009distribution].
+These results suggest that the process of urbanization is, in general, operating to exclude specialist species [@sorace2009distribution; concepcion2015impacts].
 Further, the non-random distribution of the urbanization index across the phylogeny (Fig.

### 2018-06-07 ms.Rmd
some RK wordsmithing for introduction and start of methods

@@
 # ABSTRACT
-Understanding the suite of ecological and life history traits which influence a species' tolerance to urbanization is critical to understand which species are most likely to persist in our increasingly urbanizing world.
-Previous work investigating these relationships has found a wide array of contrasting patterns for a given trait's association with urbanization.
-In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we provide a novel and scalable methodology which assesses the ecological and life history traits which most influence a species' adaptability to persist in urban environments.
-We elucidate generalized, phylogenetically controlled, patterns: bird species which are generalists, with high clutch size, and large residual brain size are among the most urban-tolerant bird species.
-Conversely, specialized feeding strategies (i.e., insectivores and granivores) are negatively associated with urbanization.
-Our results suggest that the enhancement and persistence of avian biodiversity in urban environments relies on maintaining and developing a suite of habitats to serve a range of life history strategies.
-
-*keywords*: birds; brain size; generalists; specialists; life history traits; ecological traits; urbanization index
+Identifying which ecological and life history traits influence a species' tolerance to urbanization is critical to understanding species trajectories of biodiversity in an increasingly urbanizing world.
+There is evidence for a wide array of contrasting patterns for single trait associations with urbanization.
+In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we provided a novel and scalable methodology that evaluated the ecological and life history traits which most influence a species' adaptability to persist in urban environments.
+We identified generalized, phylogenetically controlled patterns: bird species which are generalists (i.e., large niche breadth), with high clutch size, and large residual brain size are among the most urban-tolerant bird species.
+Conversely, specialized feeding strategies (i.e., insectivores and granivores) were negatively associated with urbanization.
+Enhancement and persistence of avian biodiversity in urban environments probably relies on protecting, maintaining, and restoring diverse habitats serving a range of life history strategies.
 
@@
 # INTRODUCTION
-Species specialization is a fundamentally important biological trait which can dictate population-level responses to changing environments.
-Generally, species with limited niche breadth experience higher extinction risk [@boulangeat2012niche].
-Further, certain organisms can more readily adapt their niches in response to changing environments leading to potential ecological inheritance within specific species [@odling2013niche].
-The process of urbanization drastically fragments and isolates natural habitats [@marzluff2001restoration; @melles2003urban; @loram2007urban], providing ecologically novel landscapes for the species within.
-Indeed, some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others remain relatively common in surrounding habitats but rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
-Urbanization, by selecting for a specific set of species, globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], which is a major concern for conservation [@czech2000economic; @dearborn2010motivations].
-Hence, it is important to fully understand the suite of ecological and life history traits most associated with birds that are successful or unsuccessful in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation].
-Only then will the species most at-risk of extinction, and thus in need of conservation attention, be best understood and appropriate action subsequently taken (e.g., necessary habitat improvements).
+Species' specialization, based on a suite of biological traits, can dictate population-level responses to changing environments.
+Generally, species with limited niche breadth experience high extinction risk [@boulangeat2012niche].
+Further, some organisms have traits that allow for adaptation of their niches with changing environments, leading to potential ecological inheritance [@odling2013niche].
+Urbanization drastically changes natural habitats through destruction, fragmentation, and isolation [@marzluff2001restoration; @melles2003urban; @loram2007urban], providing ecologically novel landscapes for species.
+Some species successfully adapt, colonize, and thrive in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others remain relatively common in surrounding natural habitats but rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
+Urbanization selects for a specific set of species, globally threatening biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], a major concern for conservation [@czech2000economic; @dearborn2010motivations].
+It is important to identify which ecological and life history traits most associate with successful or unsuccessful species in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful; @mcdonnell2015adaptation].
+This can provide generalized understanding of relationships between biodiversity and urbanization, guiding protection of the species most at-risk of extinction, through habitat protection, management, and restoration.
 
-To date, a wide array of ecological and life history traits have been used to predict presence of birds in urban environments.
-For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does; @conole2011functional], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have been hypothesized to influence a species' presence in urban environments.
-For instance, the presence of urban birds in Jerusalem was best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
-Notably, a wide range of often contrasting results using aforementioned traits have been reported in the literature.
-For instance, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains], but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on a species' response to urbanization.
+A wide array of ecological and life history traits are associated with bird species in urban environments, including degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution]; nesting substrate [@kark2007living; @croci2008does; @conole2011functional]; diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes]; distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does]; migratory status [@friesen1995effects]; niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes]; and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific] (see Appendix 1).
+In Jerusalem, urban bird communities were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while throughout France and Switzerland they found that urban bird species were positively associated with forested habitats, were sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
+There are also contrasting results: residual brain size in birds is sometimes positively associated with urbanization [@maklakov2011brains; @moller2015brain] and sometimes not [@kark2007living; @evans2011makes].
 Similarly, contrasting results have been reported for annual fecundity [cf., @croci2008does; @moller2009successful; @evans2011makes] and niche breadth [cf., @kark2007living; @evans2011makes].
 
-The majority of previous studies assign species to discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even binary grouping based on presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
-These types of discrete response variables severely limits the capacity to generalize results as it assumes species within a group respond to urbanization equally [@evans2011makes].
-Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., @kark2007living; @croci2008does; @geschkecompact], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].
-In a notable exception, Evans *et al.* [-@evans2011makes] advocates for a continuous measure of urbanization in which a species is assessed on its entire distribution in response to urban environments.
+Most analyses assign species to discrete a priori groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
+These approaches severely limit the capacity to generalize, as it assumes grouped species respond equally to urbanization [@evans2011makes].
+The terminology also varies among studies [cf., @kark2007living; @croci2008does; @geschkecompact], affecting interpretation [@mcdonnell2015adaptation; @fischer2015categorizing].
+There is clearly a need for continuous measures of urbanization [@evans2011makes] where a species is assessed on its entire distribution in response to urban environments.
 
-Furthermore, despite the number of studies which have investigated the relationships between traits and urbanization, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living].
-There is a need for broad-scale empirical datasets, which allow us to test hypotheses at large spatiotemporal scales to make generalizable patterns, and citizen science can provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen].
-Indeed, various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].
+Most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does] or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living].
+Broad-scale empirical data allow testing of hypotheses at large spatiotemporal scales.
+Citizen science can provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen], and can be used to describe patterns of abundance, distribution, and functional composition of birds in urban ecosystems [@mccaffrey2005using; @la2014role; @morelli2016evidence; @lepczyk2017global @callaghan2017assessing].
 
-The aim of this study was to determine whether there is a suite of life history traits that predicts avian tolerance to urbanization at a continental scale.
-We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against).
-For such an analysis to have generality it must (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3), include a broad-range of traits, and (4) employ a continuous measure of response to urbanization.
-We conducted an analysis that meets these criteria, integrating a series of datasets to implement a novel continuous urbanization index to assess ecological and life history traits which most influence a bird's response to urbanization.
-Our first objective was to investigate the presence of a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes].
-Our second objective was to develop models which were and were not phylogenetically controlled to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables.
+We aimed to determine what suite of life history traits predicts avian tolerance to urbanization at a continental scale.
+We predicted that different traits would positively and negatively associate with urbanization.
+For generality, our analyses had to (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3) include a broad-range of traits, and (4) employ a continuous measure of response to urbanization.
+We integrated a series of large datasets to establish a novel continuous urbanization index to test associations of ecological and life history traits which influenced a species' responses to urbanization.
+We first investigated evidence of a phylogenetic signal in our urbanization index [*sensu* @evans2011makes], given our comprehensive sample size.
+We then developed models, both including and excluding phylogenetic constraints, in order to test for association between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.
 
@@
 ## Bird observation data
-We collated bird observations throughout the whole of continental Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists.
+We collated bird observations throughout mainland Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists.
 Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird].
 We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018.
-This time-frame was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see below) over the given time frame in which bird observations were collected.
-Bird observations from outlying islands were eliminated from potential analyses.
+This corresponded to the richest period of eBird data, minimizing undue leverage of mismatch between changes in response variables over which bird observations were collected.
 
-We filtered the bird observations [@callaghan2017assessing; @la2014role] by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barrett2003new].
-For further information on eBird protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types.
-Further, we filtered the checklists to those with a distance travelled of < 5 km or an area covered which was < 500 Ha.
-Any checklists which were shared among multiple individuals were subsampled to include only one checklist in order to avoid duplicate results.
-After this sampling, we were left with 637,482 eligible checklists throughout Australia.
-Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we did not expect any correlation with seabirds using urban areas under normal circumstances.
-Only species which included > 100 observations in the eBird database over the specified time-frame, and that met our criteria were considered for further analyses.
-This left us with 580 potential species from the eBird dataset which had > 100 observations on eligible checklists in the specified time frame, which was a substantial proportion of the ~730 bird species which are resident in Australia.
+We filtered the bird observations [@callaghan2017assessing; @la2014role] only including observations on complete checklists and which were either travelling, random, stationary, area, or followed BirdLife Australia protocols [@barrett2003new] (see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types).
+We then filtered the checklists to those which travelled < 5 km or covered < 500 Ha.
+Any checklists shared among multiple individuals were subsampled, including only one checklist in order to avoid duplicate results.
+This provided 637,482 eligible checklists throughout Australia.
+Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we did not expect any correlation with seabirds using urban areas.
+Further, species with > 100 observations were considered for further analyses.
+Totally, 580 species, a substantial proportion of the ~730 bird species which are resident in Australia, were considered for analyses.
 Given the diverse datasets (see below) used in this analysis, there were differing taxonomic authorities followed (see Appendix 2 for elaboration).

### 2018-06-11 ms.Rmd
some edits from RK

@@
 # ABSTRACT
-Identifying which ecological and life history traits influence a species' tolerance to urbanization is critical to understanding species trajectories of biodiversity in an increasingly urbanizing world.
+Identifying which ecological and life history traits influence a species' tolerance to urbanization is critical to understanding the trajectory of biodiversity in an increasingly urbanizing world.
 There is evidence for a wide array of contrasting patterns for single trait associations with urbanization.
@@
 Totally, 580 species, a substantial proportion of the ~730 bird species which are resident in Australia, were considered for analyses.
-Given the diverse datasets (see below) used in this analysis, there were differing taxonomic authorities followed (see Appendix 2 for elaboration).
+We used differing taxonomic authorities (see Appendix 2 for elaboration).
 
 ## Urbanization index
-We used VIIRS night-time lights [@elvidge2017viirs] as a proxy of a bird's association with urban habitat.
-The VIIRS night-time lights measures the electric lighting present on Earth's surface which is generally associated with human settlement (Fig.
+We used VIIRS night-time lights [@elvidge2017viirs], for urban habitat, measuring the electric lighting present on Earth's surface, generally associated with human settlement (Fig.
 S1).
-There are also a significant number of steps (e.g., filtering out background noise such as fire, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@elvidge2017viirs].
-For the 637,482 checklists which met our criteria above, we used the checklists' associated spatial coordinates and Google Earth Engine [@gorelick2017google] to assign the average radiance value calculated from the VIIRS layer.
+The night-time lights dataset filters out background noise (e.g., fires, degraded data, and other light source contamination) to ensure validity of the data and its association with human settlement [@elvidge2017viirs].
+For our 637,482 checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], assigning the average radiance value calculated from the VIIRS layer.
 For each checklist, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
-Any radiance values which were negative were forced to 0.00001 [*sensu* @ou2015evaluation].
-Each checklist's average radiance value was then passed to every bird on that checklist, which provided every species with a distribution of average radiance values (Fig.
-1).
-The median for each species' given distribution of average radiance values served as the response variable in our analyses.
-Inherently, there is no definitive range for this index, as it solely depends on a species' distribution of observations.
+Any negative radiance values were forced to 0.00001 [*sensu* @ou2015evaluation].
+Every bird species subsequently had a distribution of average radiance values (Fig.
+1), with each median used as the response variable in our analysis.
+There was no definitive range for this index, depending on a species' distribution of observations.
 
-## Biological traits
-Many biological and life history traits have been used in previous studies, and potential traits in our study were chosen on the basis of their existing support and hypotheses in the current literature [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes]; we sought to encompass a broad array of traits previously tested.
-Hence, predictor variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological].
-Table 1 provides a brief overview of the traits that we investigated, while Appendix 1 Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated.
+## Ecological and life history traits
+Ecological and life history traits were chosen on the basis of existing support [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes; Table 1].
+This included a wide array of potential traits for all of Australia's birds [@garnett2015biological].
+Traits included body size, residual brain size, clutch size, breeding habitat generalism, feeding habitat generalism, diet generalism, nesting generalism, range size, nomadic/irruptiveness, whether a species was migratory, Table 1 provides a brief overview of the traits that we investigated, while Appendix 1 Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated.
 Traits were parsed in different manners (Appendix 1b), and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat).
@@
 ## Statistical analyses
+Before modelling of predictor variables (Table 1) we investigated multi-collinearity, using the corrplot package [@corrplot2017], among continuous variables (Fig.
+S2).
+We used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other, and all traits were < 2 suggesting negligible correlation (Supplementary Table 1).
+We used two independent linear modelling approaches: one without and one with phylogenetic constraints.
+Both approaches were used because of the associated shortcomings [@cadotte2017phylogenies] and importance [@losos2008phylogenetic; @graham2018phylogenetic] of incorporating evolutionary distinctness in ecological research.
+
+We weighted the linear regression models [@solon2015weighting] using a source of error associated with our urbanization indices, which were log-transformed to meet model assumptions.
+This error was the proportion of the number of total observations of a species, divided by the number of its unique localities.
+This weighted the model-fitting to those species with more observations in more unique locations, likely to be the most reliable urbanization indices.
+This approach also addressed biases in sampling of eBird data where some birds have a high number of observations from a small number of locations such as known roosting sites [e.g., @callaghan2018avian].
+
 All models and analyses were conducted in R [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse].
 
-Before modelling, the predictor variables were investigated for multi-collinearity.
-We used the corrplot package [@corrplot2017] to show correlation among continuous variables (Fig.
-S2), and used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other (Supplementary Table 1).
-
-We employed two linear modelling approaches to provide robustness of our results; one with and one without phylogenetic constraints.
-Phylogenetic models were used because of the importance of considering evolutionary distinctedness in ecological research [@losos2008phylogenetic; @graham2018phylogenetic], but given that phylogenetically designed analyses can have a number of shortcomings [@cadotte2017phylogenies], we also included non-phylogenetic models in our analysis.
-Our urbanization index was log-transformed in order to meet model assumptions.
-The predictor variables used in the models are listed in Table 1.
-
-We weighted the linear regression models [@solon2015weighting] using a source of error associated with our urbanization indices.
-This error was the proportion of the number of total observations of a species divided by the number of unique localities a species was observed from.
-This gave more weight in the model-fitting procedure to those species that had more observations in more unique locations.
-This was done based on the assumption that those species which had the most observations would have the most reliable urbanization indices.
-This approach also addresses biases in sampling of eBird data in which some birds can have a high number of observations, but from a small number of locations such as known roosting sites [e.g., @callaghan2018avian].
-
 ### Non-phylogenetically controlled models
-Given that we hypothesized any of the potential candidate traits could significantly drive the response in our urbanization index, we first ran one global model including all 22 predictor variables.
-We used the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes.
-In order to confirm our linear model results, we employed a model-averaging approach [@grueber2011multimodel] in which we 'dredged' possible subsets of the global model.
-This was done using the dredge function in the MuMIn package [@barton2016mumin], and we allowed a total of 11 candidate predictors to be fit at any one time in order to avoid over-fitting.
-This left a total of 784,626 models which were assessed, and the model-averaged parameter estimates were calculated from the top models of those with $\Delta$AIC < 4.
-As with the global model, we present the standardized estimates which represent relative effect sizes.
+We first ran one global model, including all 22 predictor variables, using the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes.
+To corroborate our linear model results, we employed a model-averaging approach [@grueber2011multimodel] where we 'dredged' possible subsets of the global model, using the dredge function in the MuMIn package [@barton2016mumin].
+We fitted a total of 11 candidate predictors at any one time to avoid over-fitting, totalling 784,626 models assessed.
+Model-averaged parameter estimates were calculated from the subsequent top models, with $\Delta$AIC < 4, with standardized estimates representing relative effect sizes.
 
 ### Phylogenetic analysis
-As the correlation between traits and urbanization may be influenced by the shared evolutionary history of the species [@felsenstein1985phylogenies], we also used phylogenetically informed models to test our hypotheses.
-Phylogenetically informed models are descendants of phylogenetic independent contrasts [@felsenstein1985phylogenies], which in simple terms, tests for repeated evolutionary association between the traits of interest.
-Or in this particular case, whether the traits of interest (e.g., large brain size at a given body size) repeatedly allow for the use of urban environments wherever it evolved across the bird phylogeny.
-We used the phylogenetic trees provided by Jetz et al.
-[@jetz2012global].
-This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global].
-In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam].
-This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.
+As associations between traits and urbanization may be influenced by shared evolutionary history of the species [@felsenstein1985phylogenies], we also used phylogenetically informed models.
+These test for repeated evolutionary association between traits of interest [@felsenstein1985phylogenies].
+We used published phylogenetic trees for 9,993 species [@jetz2012global] to assess existence of a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] associated with our urbanization index and for each of the other continuous predictor variables (Table 1).
+This was to assess the relative contribution of phylogeny on the predictor variables, and we used the phylosignal package [@keck2016phylosignal] in R.
 
-Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package which fits a linear regression model where the likelihood is linear in the number of tips in the tree [@ho2014phylolm].
-Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.
-To account for aspects of phylogenetic uncertainty, we also fit the same model to a range of equally likely trees from Jetz *et al.* [-@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.
+Phylogenetic models were fit based on the same global model as with the non-phylogenetic models, using the 'phylolm' function from the phylolm package which fits a linear regression model where the likelihood is linear in the number of tips in the tree [@ho2014phylolm].
+We also model-averaged the global phylogenetic model using the aforementioned methods.
+To account for phylogenetic uncertainty, we also fitted the same model to a range of equally likely trees from Jetz *et al.* [-@jetz2012global], examining the distribution of putatively important traits across the distribution of trees.
 
 # RESULTS
-A total of 316,306 checklists resulting in 5,944,819 observations of 477 different species were included in analyses.
-The least often recorded species was *Turnix melanogaster* (Black-breasted Buttonquail; N=104) while the most recorded species was *Cracticus tibicen* (Australian Magpie; N=158,615).
-The mean number of observations for a species was 10,250.
-Plots of all 477 species' distributions of average radiance values are presented in Appendix 3.
-A complete list of the 477 species, in addition to their traits, number of observations, and their urbanization indices can be found in Appendix 4.
-
-## Urbanization index
-For the 477 species included in the analysis the mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, with a range from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig.
+We analyzed 316,306 checklists including 5,944,819 observations of 477 different species.
+The least recorded species was *Turnix melanogaster* (Black-breasted Buttonquail; N=104) while the most recorded species was *Cracticus tibicen* (Australian Magpie; N=158,615).
+The mean ($\pm$ SD) number of observations for a species was 10,250 $\pm$ xxxx.
+Distributions of measured VIIRS average radiance values (Appendix 3), their traits, number of observations, and urbanization indices for the 477 species (Appendix 4) varied considerably.
+The mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, ranging from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig.
 2).

### 2018-06-11 R/non_phylo_analysis.R
fixing a spelling error

@@ -48,7 +48,7 @@ corrplot_figure <- function(analysis_data){
     rename(`Diet generalism` = diet_generalism) %>%
     rename(`Nest generalism` = nest_generalism) %>%
     rename(`Number of records` = N) %>%
-    rename(`Unique localitieis observed` = unique_localities) %>%
+    rename(`Unique localities observed` = unique_localities) %>%
     rename(`log(Body size)` = body_size_logged) %>%
     rename(`log(Clutch size)` = clutch_size_logged) %>%
     rename(`Range size (1000s km2)` = range_size) %>%

### 2018-06-12 R/non_phylo_analysis.R
adding column names that got dropped

@@ -101,7 +101,7 @@ collinearity_investigation_function <- function(global_model) {
   is.num <- sapply(collinearity_investigation, is.numeric)
   collinearity_investigation[is.num] <- lapply(collinearity_investigation[is.num], round, 2)
   
-  write_csv(collinearity_investigation, "Data/TableS1.csv")
+  write.csv(collinearity_investigation, "Data/TableS1.csv")
   
   pdf("tables/collinearity_investigation.pdf")
   grid.table(collinearity_investigation)

### 2018-06-12 ms.Rmd
More changes and edits

@@
 Species' specialization, based on a suite of biological traits, can dictate population-level responses to changing environments.
-Generally, species with limited niche breadth experience high extinction risk [@boulangeat2012niche].
+Generally, species with limited niche breadth experience higher extinction risk than those with wider niche breadth [@boulangeat2012niche].
 Further, some organisms have traits that allow for adaptation of their niches with changing environments, leading to potential ecological inheritance [@odling2013niche].
@@
 
-A wide array of ecological and life history traits are associated with bird species in urban environments, including degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution]; nesting substrate [@kark2007living; @croci2008does; @conole2011functional]; diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes]; distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does]; migratory status [@friesen1995effects]; niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes]; and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific] (see Appendix 1).
-In Jerusalem, urban bird communities were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while throughout France and Switzerland they found that urban bird species were positively associated with forested habitats, were sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
+A wide array of ecological and life history traits are associated with bird species in urban environments, including degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution]; nesting substrate [@kark2007living; @croci2008does; @conole2011functional]; diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes]; range size [@jokimaki2000artificial; @chace2006urban; @croci2008does]; migratory status [@friesen1995effects]; niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes]; and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific] (see Appendix 1).
+In Jerusalem, urban bird communities were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while throughout France and Switzerland urban bird species were positively associated with forested habitats, sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
 There are also contrasting results: residual brain size in birds is sometimes positively associated with urbanization [@maklakov2011brains; @moller2015brain] and sometimes not [@kark2007living; @evans2011makes].
@@
 We first investigated evidence of a phylogenetic signal in our urbanization index [*sensu* @evans2011makes], given our comprehensive sample size.
-We then developed models, both including and excluding phylogenetic constraints, in order to test for association between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.
+We then developed models, including and excluding phylogenetic constraints, in order to test for association between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.
 
 # METHODS
-
 ## Bird observation data
@@
 
-We filtered the bird observations [@callaghan2017assessing; @la2014role] only including observations on complete checklists and which were either travelling, random, stationary, area, or followed BirdLife Australia protocols [@barrett2003new] (see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types).
+We filtered bird observations [@callaghan2017assessing; @la2014role], only including observations on complete checklists and which were either travelling, random, stationary, area, or followed BirdLife Australia protocols [@barrett2003new] (see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types).
 We then filtered the checklists to those which travelled < 5 km or covered < 500 Ha.
@@
 Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we did not expect any correlation with seabirds using urban areas.
-Further, species with > 100 observations were considered for further analyses.
-Totally, 580 species, a substantial proportion of the ~730 bird species which are resident in Australia, were considered for analyses.
+Further, only species with > 100 observations were considered for further analyses.
+Totally, 580 species, a substantial proportion of the ~730 bird species which are resident in Australia, were initially considered for analyses.
 We used differing taxonomic authorities (see Appendix 2 for elaboration).
@@
 The night-time lights dataset filters out background noise (e.g., fires, degraded data, and other light source contamination) to ensure validity of the data and its association with human settlement [@elvidge2017viirs].
-For our 637,482 checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], assigning the average radiance value calculated from the VIIRS layer.
+For our 637,482 checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], to assign the average radiance value calculated from the VIIRS layer.
 For each checklist, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
@@
 Every bird species subsequently had a distribution of average radiance values (Fig.
-1), with each median used as the response variable in our analysis.
+1), with the median of each species' distribution used as the response variable in our analysis.
 There was no definitive range for this index, depending on a species' distribution of observations.
@@
 ## Ecological and life history traits
-Ecological and life history traits were chosen on the basis of existing support [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes; Table 1].
-This included a wide array of potential traits for all of Australia's birds [@garnett2015biological].
-Traits included body size, residual brain size, clutch size, breeding habitat generalism, feeding habitat generalism, diet generalism, nesting generalism, range size, nomadic/irruptiveness, whether a species was migratory, Table 1 provides a brief overview of the traits that we investigated, while Appendix 1 Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated.
-Traits were parsed in different manners (Appendix 1b), and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat).
-Preliminary analyses including IUCN status (exotic/native) showed that exotic birds were strongly associated with urbanization, but given the preferential selection by acclimatization societies for introduction of species with proven synanthropy [@blackburn2009avian], exotic species were omitted from the final analysis to avoid bias.
-Therefore, we excluded IUCN status as a trait because the majority of our candidate species were of Least Concern.
-Only species which had complete data for the list of candidate traits were considered in the analyses, and 477 species met this criteria.
+Ecological and life history traits were chosen on the basis of existing support [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes; Table 1; Appendix 1a].
+These included a wide array of potential traits for all of Australia's birds [@garnett2015biological].
+A total of 22 traits were included in analyses: clutch size, body size, brain residual, nest generalism, feeding habitat generalism, breeding habitat generalism, diet generalism, habitat association with agricultural landscapes, habitat association with treed/forested habitats, habitat association with grass/shrubland habitats, whether a species nests in hollows, whether a species nests on the ground, whether a species eats plants, whether a species is a granivore, whether a species is an insectivore, whether a species eats carrion, whether a species is nomadic/irruptive, whether a species is migratory, the range size of a species, nesting aggregation of a species, feeding aggregation of a species, and whether a species is a cooperative breeder (Table 1; Appendix 1a).
+For any missing trait in the dataset [@garnett2015biological] for a species we assumed it did not exist.
+For example, if a species had no data for a specific habitat, it was deemed not to occur in the habitat.
+Our preliminary analyses included IUCN status and showed that exotic birds were strongly associated with urbanization, but given the preferential selection by acclimatization societies for introduction of species with proven synanthropy [@blackburn2009avian], exotic species were omitted from the final analyses to avoid bias.
+We also excluded IUCN status because most (96%) of our candidate species were of Least Concern.
+Totally, 477 species with complete data for all candidate traits were included in analyses.
 
 ## Statistical analyses
-Before modelling of predictor variables (Table 1) we investigated multi-collinearity, using the corrplot package [@corrplot2017], among continuous variables (Fig.
-S2).
+Before modelling of predictor variables (Table 1) we investigated multi-collinearity among continuous variables (Fig.
+S2), using the corrplot package [@corrplot2017].
 We used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other, and all traits were < 2 suggesting negligible correlation (Supplementary Table 1).
 We used two independent linear modelling approaches: one without and one with phylogenetic constraints.
-Both approaches were used because of the associated shortcomings [@cadotte2017phylogenies] and importance [@losos2008phylogenetic; @graham2018phylogenetic] of incorporating evolutionary distinctness in ecological research.
+Both approaches were used because of shortcomings [@cadotte2017phylogenies] and importance [@losos2008phylogenetic; @graham2018phylogenetic] of incorporating evolutionary distinctness in ecological research.
 
-We weighted the linear regression models [@solon2015weighting] using a source of error associated with our urbanization indices, which were log-transformed to meet model assumptions.
-This error was the proportion of the number of total observations of a species, divided by the number of its unique localities.
+Linear regression models were weighted [@solon2015weighting] using a source of error associated with our urbanization indices, which were log-transformed to meet model assumptions.
+This error was the proportion of the number of total observations of a species, divided by the number of its unique localities (Fig.
+S2 shows the correlation between the two).
 This weighted the model-fitting to those species with more observations in more unique locations, likely to be the most reliable urbanization indices.
@@
 
-All models and analyses were conducted in R [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse].
+All models and analyses were conducted in R [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse]. Statistical significance was concluded as $\alpha$ < 0.05.
 
-### Non-phylogenetically controlled models
+## Non-phylogenetically controlled models
 We first ran one global model, including all 22 predictor variables, using the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes.
-To corroborate our linear model results, we employed a model-averaging approach [@grueber2011multimodel] where we 'dredged' possible subsets of the global model, using the dredge function in the MuMIn package [@barton2016mumin].
-We fitted a total of 11 candidate predictors at any one time to avoid over-fitting, totalling 784,626 models assessed.
-Model-averaged parameter estimates were calculated from the subsequent top models, with $\Delta$AIC < 4, with standardized estimates representing relative effect sizes.
+To corroborate our global model results, we employed a model-averaging approach [@grueber2011multimodel] where we 'dredged' possible subsets of the global model, using the dredge function in the MuMIn package [@barton2016mumin].
+We fitted a total of 11 candidate predictors at any one time to avoid over-fitting, totaling 784,626 models assessed.
+Model-averaged parameter estimates were calculated from the subsequent top models, with $\Delta$AIC < 4, with standardized estimates representing relative effect sizes of respective traits.
 
-### Phylogenetic analysis
-As associations between traits and urbanization may be influenced by shared evolutionary history of the species [@felsenstein1985phylogenies], we also used phylogenetically informed models.
-These test for repeated evolutionary association between traits of interest [@felsenstein1985phylogenies].
-We used published phylogenetic trees for 9,993 species [@jetz2012global] to assess existence of a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] associated with our urbanization index and for each of the other continuous predictor variables (Table 1).
-This was to assess the relative contribution of phylogeny on the predictor variables, and we used the phylosignal package [@keck2016phylosignal] in R.
+## Phylogenetic analysis
+### Phylogenetic signal
+As associations between traits and urbanization may be influenced by shared evolutionary history of a species [@felsenstein1985phylogenies], we also used a phylogenetically informed analytical approach.
+First, we assessed the existence of a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] associated with our urbanization index and for each of the continuous predictor variables (Table 1), using published phylogenetic trees for 9,993 species [@jetz2012global].
+A phylogenetic signal is defined as the statistical nonindependence among species trait values due to their phylogenetic relatedness [@revell2008phylogenetic], and it assesses the relative contribution of phylogeny on a given trait.
+There are a number of indices used to assess phylogenetic signal, each with specific strengths and weaknesses [@keck2016phylosignal], and we assessed all available indices (Blomberg's K and K*, Abouheif's Cmean, Moran's I, and Pagel's Lambda), using the phylosignal package [@keck2016phylosignal] in R.
 
-Phylogenetic models were fit based on the same global model as with the non-phylogenetic models, using the 'phylolm' function from the phylolm package which fits a linear regression model where the likelihood is linear in the number of tips in the tree [@ho2014phylolm].
+### Phylogenetically controlled models
+Second, we used phylogenetically informed models to test for repeated evolutionary association between traits of interest [@felsenstein1985phylogenies].
+Phylogenetic models were fit based on the same global model as the non-phylogenetic models, but using the 'phylolm' function from the phylolm package which fits a linear regression model where the likelihood is linear in the number of tips in the tree [@ho2014phylolm].
 We also model-averaged the global phylogenetic model using the aforementioned methods.
-To account for phylogenetic uncertainty, we also fitted the same model to a range of equally likely trees from Jetz *et al.* [-@jetz2012global], examining the distribution of putatively important traits across the distribution of trees.
+To account for phylogenetic uncertainty, we fitted the same model to 1000 highly probable trees from Jetz *et al.* [-@jetz2012global], examining the distribution of putatively important traits across the distribution of trees.
 
@@
 The least recorded species was *Turnix melanogaster* (Black-breasted Buttonquail; N=104) while the most recorded species was *Cracticus tibicen* (Australian Magpie; N=158,615).
-The mean ($\pm$ SD) number of observations for a species was 10,250 $\pm$ xxxx.
-Distributions of measured VIIRS average radiance values (Appendix 3), their traits, number of observations, and urbanization indices for the 477 species (Appendix 4) varied considerably.
+The mean ($\pm$ SD) number of observations for a species was 10,930 $\pm$ 19,498.
+Distributions of measured VIIRS average radiance values (shown for each species in Appendix 3), traits, number of observations, and urbanization indices for the 477 species (Appendix 4) varied considerably.
 The mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, ranging from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig.
@@
 
-### Phylogenetic signal
-There was a distinct phylogenetic signal in our urbanization index measure (Fig.
-3; Fig.
-S3; Table 2), as it was significantly different than random for 4 out of the 5 indices considered.
-For example, note the distinct signal located throughout the waterbirds' clades in Figure 3 or the significantly large index for *Manorina melanocephala*, Noisy Miner, compared to the rest of the honeyeaters.
-Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 2).
-The result for body size ($k=3.3$) is among the strongest ever recorded measures of phylogenetic signal in the literature, and for $k$ the statistic suggests much stronger signal than expected from Brownian motion.
-We generally found weaker signal for behavioral traits compared with morphological traits.
-For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 2).
+## Non-phylogenetic models
+The most important predictors of urbanization (i.e., the largest parameter estimates - shown in parentheses) in the non-phylogenetic global model (Fig.
+4; Appendix 5) were feeding habitat generalism (0.79), breeding habitat generalism (0.76), and clutch size (0.73).
+Further, feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006) were significant.
+Urban-tolerant birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032), but less so than the previous traits.
+Conversely, urban-tolerant birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), group foraging (t = -2.24, p =0.025), granivorous diet (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), and grass/shrubland habitats (t = -4.26, p < 0.001), with insectivorous diet (t = -3.96, p < 0.001) and solitary feeding (t = -2.90, p = 0.004).
+The strongest negative associations with the urbanization index were solitary feeding (-0.92), insectivorous diet (-0.85), grass/shrubland habitat association (-0.733), and tree/forest habitat association (-0.63).
 
-## Model results
-### Non-phylogenetic models
-The most important predictors of urbanization (i.e., the largest parameter estimates - shown in parantheses) in the non-phylogenetic global model (Fig.
-4; Appendix 5) were feeding habitat generalism (0.79), breeding habitat generalism (0.76), and clutch size (0.73).
-Further, significance was found for feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006).
-Urban-tolerant birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032), but less so than the previous traits.
-Conversely, urban-tolerant birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), group foraging (t = -2.24, p =0.025), granivorous diet (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), insectivorous diet (t = -3.96, p < 0.001), and solitary feeding (t = -2.90, p = 0.004).
-The strongest negative responses, in rank of their parameter estimates, were solitary feeding (-0.92), insectivorous diet (-0.85), grassland/shrubland habitat association (-0.733), and tree/forest habitat association (-0.63).
-
-When this global model was dredged and model-averaged to investigate for specific suites of traits, patterns remained similar (Fig.
+With model-averaging, patterns remained similar (Fig.
 4; Appendix 5).
-However, the most important variable, based on standardized parameter estimates was clutch size (0.75), followed by birds associated with agricultural habitats (0.58), and then diet generalism (0.28).
+However, the most strongly associated trait, based on standardized parameter estimates was clutch size (0.75), followed by birds associated with agricultural habitats (0.58), and then diet generalism (0.28).
 Breeding habitat (0.15) and feeding habitat (0.12) generalism parameter estimates were substantially smaller than in the non model-averaged approach.
-Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors.
-Additionally, birds which were classified in the traits database as having strong association with agricultural habitats were significantly associated with urbanization (z = 3.40, p = 0.001).
+Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) were all significantly associated with our urbanization index.
+Additionally, species associated with agricultural habitats were significantly associated with urbanization (z = 3.40, p = 0.001).
 Urban-tolerant birds remained significantly negatively associated with solitary feeding (z = 3.39, p = 0.001),  insectivorous diet (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), group foraging (z = 3.16, p = 0.002), granivorous diet (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).
 Solitary feeding (-1.00) remained the strongest negatively associated trait with urbanization, followed by insectivorous diet (-0.98).
+
+## Phylogenetic analyses
+### Phylogenetic signal
+There was a distinct relationship of phylogenetic signal to our urbanization index (Fig.
+3; Fig.
+S3; Table 2), as it was significantly different than random for 4 out of the 5 indices considered (Cmean, I, K.star, and Lambda).
+For example, the urbanization index is disproportionate in respect to the phylogenetic tree (Fig.
+3), evidenced by the highest urbanization indices being clustered throughout the waterbird clade.
+Further, within a specific clade there were some species which had much larger urbanization indices (e.g., note the high index for *Manorina melanocephala*, Noisy Miner, compared to the rest of the honeyeaters).
+We generally found weaker signal for behavioral traits, compared with morphological traits.
+For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 2).
+Additionally, there was strong phylogenetic signal for body size and brain residual size (Table 2).
 
@@
 
-Similarly, upon model-averaging (Fig.
-4; Appendix 5), feeding habitat generalism (1.30), clutch size (1.15), body size (0.66), and residual brain size (0.32) showed the strongest association with urbanization.
-Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were all statistically significant.
-Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization, with no known nest aggregation (-1.22), and species associated with grasslands/shrublands (-1.16) showing the strongest negative response.
+With model-averaging (Fig.
+4; Appendix 5), feeding habitat generalism (1.30), clutch size (1.15), body size (0.66), and residual brain size (0.32) showed the strongest association with our urbanization index, while species with no known nest aggregation (-1.22) and species associated with grass/shrubland (-1.16) showed the strongest negative response.
+Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were positively statistically significant, while species which were cooperative breeders (z = 2.53, p = 0.011), used grass/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were negatively statistically significant.
 
-### Differences in non-phylogenetic and phylogenetic models
-The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates (Appendix 5), and the statistical significance of individual predictors (Appendix 5; Fig.
-4).
-However, residual brain size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig.
-5), showing the strongest relative response in the phylogenetic global model.
-Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger standardized parameter estimates) in the phylogenetic models compared with the non-phylogenetic models (Appendix 5; Fig.
+## Differences in non-phylogenetic and phylogenetic models
+Non-phylogenetic and phylogenetic models largely agreed in the relative rank of standardized predictor estimates (Appendix 5) and the statistical significance of individual predictors (Fig.
+4; Appendix 5).
+However, residual brain size was not significant in either non-phylogenetic models (Fig.
+4), but was significantly associated with urbanization in our phylogenetic models, showing the strongest relative response in the phylogenetic global model (Fig.
+4; Fig.
 5).
+Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (large standardized parameter estimates) in the phylogenetic models, compared with the non-phylogenetic models (Fig.
+5; Appendix 5).
 Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled models (Fig.
@@
 5).
-Full model results for each of the four analyses can be found in Appendix 5.
 
@@
 S4).
-These results suggest that the process of urbanization is, in general, operating to exclude specialist species [@sorace2009distribution; concepcion2015impacts].
+These results suggest that the process of urbanization is, in general, operating to exclude specialist species [@sorace2009distribution; @concepcion2015impacts].
 Further, the non-random distribution of the urbanization index across the phylogeny (Fig.
@@
 Most notably, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
-While some studies have found a lack of evidence that residual brain size is an important predictor in urban-tolerant species [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed heavily correlated with urban-tolerance [@moller2009successful; @maklakov2011brains; @moller2015brain].
+While some studies have found a lack of evidence that residual brain size is an important predictor in urban-tolerant species [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed highly correlated with urban-tolerance [@moller2009successful; @maklakov2011brains; @moller2015brain].
 
@@
 While the aforementioned patterns may be generalizable, we also found patterns in our analyses which seem to contrast the global norm.
-For example, we found that urbanization selects against granivorous species, contrasting with the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [compared to European cities for instance; @fuller2008garden].
+For example, we found that urbanization selects against granivorous species, contrasting with the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [@jones2018birdfeeding] generally involving different species and foods  [compared to European cities for instance; @fuller2008garden].
 
@@
 We emphasize that future studies, especially within migrant-dominated systems (e.g., Europe, North America), should account for changes in a species' urbanization index throughout the full-annual cycle [@hostetler2015full; @marra2015call].
-This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species.
+This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species, thus allowing prediction of which species face the greatest risk of impending urbanization, specifically in developing countries.
 
@@
 
-```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance value distributions, on a logarithmic scale.
-These species were chosen because they represent species with different distributional responses."}
+```{r Figure 1, echo=FALSE, fig.cap="Average radiance density distributions, on a logarithmic scale, across the range of five species with different distributional responses.
+The urbanization index, used as the response variable in our analyses, was taken as the median for a given species' distribution."}
 knitr::include_graphics("finalFigs/Figure1.pdf")
@@
 
-```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a logarithmic scale, of the urbanization index for the 477 species included in the analysis, with five example species highlighted.
-The urbanization index was taken as the median of the distribution for a species' average radiance values, as demonstrated in Figure 1.
-Figure S1 shows randomly chosen indices for a given point."}
+```{r Figure 2, echo=FALSE, fig.cap="Histogram showing the count of the 477 species, relative to the urbanization index, on a logarithmic scale, highlighting the five species in Fig.
+1.
+(See Figure S1 for randomly chosen indices for a given point to help orient the urbanization index.)"}
 knitr::include_graphics("finalFigs/Figure2.pdf")
@@
 
-```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree, based on Jetz et al.
-(2012) of the 477 bird species in the analysis, demonstrating the clustering of the urbanization index in regards to phylogeny.
-The tree demonstrates two representations of the urbanization index for each species: (1) coloring of the tips where dark red equates to a low urbanization index and dark blue equates to a high urbanization index, and (2) the size of the barplot at a tip shows the relative magnitude of the urbanization index.
-The urbanization index was taken as the median of the distribution for a species' average radiance values.
-Figure S3 shows a reference phylogenetic tree with the names of each tip displayed, and this link (https://johnwilshire.github.io/bird_urbanness_web_app/) provides an interactive visualization in order to easily identify each of the 477 species in the tree and their respective urbanization indices.
-For a full list of the 477 species included and their urbanization index see Appendix 4."}
+```{r Figure 3, echo=FALSE, fig.cap="Reproduced phylogenetic tree (Jetz et al.
+2012) of our 477 species (see Appendix 4 for a full list of species with their urbanization indices), demonstrating clustering of species' associations with the urbanization index relative to phylogeny.
+There are two representations of the urbanization index for each species: (1) coloring ranges from a low urbanization index (dark red) to a high urbanization index (dark blue), with (2) magnitude of the urbanization index (barplot).
+The urbanization index was the median of the distribution for a species' average radiance values, as shown in Fig.
+1.
+For reference, Figure S3 is a phylogenetic tree with the names of each tip displayed, and we also provide an interactive visualization (https://johnwilshire.github.io/bird_urbanness_web_app/)."}
 knitr::include_graphics("finalFigs/Figure3.pdf")
@@
 
-```{r Figure 4, echo=FALSE, fig.cap="Parameter estimates for each of the four modelling approaches employed in the analysis.
-A star represents a significant effect.
-Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set which was model-averaged.
-Full model results can be found in Appendix 5."}
+```{r Figure 4, echo=FALSE, fig.cap="Parameter estimates for each of the four modelling approaches employed (See Appendix 5 for full model results), where a star represent significant effects.
+Model-averaged results did not contain every possible trait variable."}
 knitr::include_graphics("finalFigs/Figure4.pdf")
@@
 
-```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for 22 standardized parameter estimates from the non-phylogenetically controlled and phylogenetically controlled global models, as modelled based on 477 species."}
+```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for each of the 22 traits' standardized parameter estimates, for the 477 species.
+Proportional differences represent the difference between a given trait's parameter estimate for the non-phylogenetically controlled and phylogenetically controlled global models."}
 knitr::include_graphics("finalFigs/Figure5.pdf")
@@
 library(pander)
-pander(table, caption = "Results showing the  phylogenetic signal for both predictor and response traits.
-Because there are a number of indices in the literature, with specific strengths and weaknesses, we present here all available indices (Keck et al.
-2016).", split.table=Inf)
+pander(table, caption = "Summary of results showing an assessment of a phylogenetic signal for our response variable (urbanization index) and the continuous predictor traits.
+We show all available indices (Blomberg's K and K*, Abouheif's Cmean, Moran's I, and Pagel's Lambda; Keck et al.
+2016), each of which represents the statistical nonindependence of a trait's value due to its phylogenetic relatedness (Revell et al.
+2008).", split.table=Inf)
 
@@
 ```{r Figure S1, echo=FALSE, fig.cap="The VIIRS nightime-lights shown for Australia, Brisbane (top right), Melbourne (bottom left), and Sydney (bottom right).
-Values are measured as average radiance, shown in red.
-For each checklist in our analysis, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included, and this led to the distributions shown in Figure 1.
-We show random points and their average associated average radiance values calculated as the mean within a 5 km buffer of that point."}
+Average radiance of VIIRS night-time lights, in red, was calculated for a given checklist within a 5 km buffer, to match the spatial scale of checklists of Australia's bird species.
+Values shown demonstrate the average radiance value for seven different points in Australia."}
 knitr::include_graphics("finalFigs/FigureS1.pdf")
@@
 
-```{r Figure S2, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the analysis."}
+```{r Figure S2, echo=FALSE, fig.cap="Correlations between continuous traits, number of records, and number of unique localities used in the analysis."}
 knitr::include_graphics("finalFigs/FigureS2.pdf")
@@
 
-```{r Figure S3, echo=FALSE, fig.cap="A reference phylogenetic tree, with labeled tips, based on Jetz et al. (2012) of the 477 bird species in the analysis. Appendix 4 lists the 477 species."}
+```{r Figure S3, echo=FALSE, fig.cap="Reference phylogenetic tree of the 477 bird species in the analysis, with labeled tips, based on Jetz et al.
+(2012).
+We also provide an interactive visualization for better interpreting the phylogenetic tree (https://johnwilshire.github.io/bird_urbanness_web_app/)"}
 knitr::include_graphics("finalFigs/FigureS3.pdf")
@@
 
-```{r Figure S4, echo=FALSE, fig.cap="To account for one aspect of phylogenetic uncertainty, we fit the same phylolm model to 1000 of the family of trees presented by Jetz et al.
-(2012).
-The density plots are the estimated parameters for most important variables across models fit to the collection trees from Jetz et al.
-(2012)."}
+```{r Figure S4, echo=FALSE, fig.cap="The same phylolm model was fitted to 1000 of the family of trees presented by Jetz et al.
+(2012) to account for one aspect of phylogenetic uncertainty.
+Density plots show estimated parameters for the most important traits across models."}
 knitr::include_graphics("finalFigs/FigureS4.pdf")
@@
 library(kableExtra)
+library(dplyr)
 
@@
 collinearity_investigation <- collinearity_investigation[-1,]
+collinearity_investigation <- collinearity_investigation %>%
+  rename(Trait = X1)
 
-kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the predictor variables used in the analysis.", booktabs = T) %>%
+kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the predictor traits used in the analysis.", booktabs = T) %>%
 kable_styling(latex_options = c("hold_position"))

### 2018-06-12 ms.Rmd
More minor edits

@@
 4; Appendix 5).
-However, residual brain size was not significant in either non-phylogenetic models (Fig.
-4), but was significantly associated with urbanization in our phylogenetic models, showing the strongest relative response in the phylogenetic global model (Fig.
+However, residual brain size was not significant in either non-phylogenetic model approaches (Fig.
+4), but was significantly associated with urbanization in our phylogenetic modelling, showing the strongest relative response in the phylogenetic global model (Fig.
 4; Fig.
@@
 5; Appendix 5).
-Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled models (Fig.
+Lastly, although not significant, body size was minimally important for the non-phylogenetic models, but substantially more important for the phylogenetically controlled models (Fig.
 4 & Fig.

### 2018-06-18 ms.Rmd
more RK edits

@@
 There is evidence for a wide array of contrasting patterns for single trait associations with urbanization.
-In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we provided a novel and scalable methodology that evaluated the ecological and life history traits which most influence a species' adaptability to persist in urban environments.
+In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we developed a novel and scalable methodology that evaluated the ecological and life history traits which most influence a species' adaptability to persist in urban environments.
 We identified generalized, phylogenetically controlled patterns: bird species which are generalists (i.e., large niche breadth), with high clutch size, and large residual brain size are among the most urban-tolerant bird species.
@@
 Further, some organisms have traits that allow for adaptation of their niches with changing environments, leading to potential ecological inheritance [@odling2013niche].
-Urbanization drastically changes natural habitats through destruction, fragmentation, and isolation [@marzluff2001restoration; @melles2003urban; @loram2007urban], providing ecologically novel landscapes for species.
-Some species successfully adapt, colonize, and thrive in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others remain relatively common in surrounding natural habitats but rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
-Urbanization selects for a specific set of species, globally threatening biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], a major concern for conservation [@czech2000economic; @dearborn2010motivations].
+Urbanization drastically changes natural habitats through destruction, fragmentation, and isolation [@marzluff2001restoration; @melles2003urban; @loram2007urban], often providing ecologically novel landscapes for species.
+Some species successfully adapt, colonize, and thrive in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others, common in surrounding natural habitats, rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
+Urbanization selects for specific sets of species, globally threatening biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], a major concern for conservation [@czech2000economic; @dearborn2010motivations].
 It is important to identify which ecological and life history traits most associate with successful or unsuccessful species in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful; @mcdonnell2015adaptation].
-This can provide generalized understanding of relationships between biodiversity and urbanization, guiding protection of the species most at-risk of extinction, through habitat protection, management, and restoration.
+This can improve understanding of relationships between biodiversity and urbanization, and guide protection of the species most at-risk of extinction, through habitat protection, management, and restoration.
 
-A wide array of ecological and life history traits are associated with bird species in urban environments, including degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution]; nesting substrate [@kark2007living; @croci2008does; @conole2011functional]; diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes]; range size [@jokimaki2000artificial; @chace2006urban; @croci2008does]; migratory status [@friesen1995effects]; niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes]; and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific] (see Appendix 1).
-In Jerusalem, urban bird communities were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while throughout France and Switzerland urban bird species were positively associated with forested habitats, sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
+Many ecological and life history traits are associated with bird species in urban environments, including degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution]; nesting substrate [@kark2007living; @croci2008does; @conole2011functional]; diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes]; range size [@jokimaki2000artificial; @chace2006urban; @croci2008does]; migratory status [@friesen1995effects]; niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes]; and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific] (see Appendix 1).
+In Jerusalem, urban bird species were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while, throughout France and Switzerland, urban bird species were positively associated with forested habitats, sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
 There are also contrasting results: residual brain size in birds is sometimes positively associated with urbanization [@maklakov2011brains; @moller2015brain] and sometimes not [@kark2007living; @evans2011makes].
@@
 
-Most analyses assign species to discrete a priori groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
+Most analyses assign species to discrete a priori groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or strictly presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
 These approaches severely limit the capacity to generalize, as it assumes grouped species respond equally to urbanization [@evans2011makes].
-The terminology also varies among studies [cf., @kark2007living; @croci2008does; @geschkecompact], affecting interpretation [@mcdonnell2015adaptation; @fischer2015categorizing].
-There is clearly a need for continuous measures of urbanization [@evans2011makes] where a species is assessed on its entire distribution in response to urban environments.
+Terminology also varies among studies [cf., @kark2007living; @croci2008does; @geschkecompact], affecting interpretation [@mcdonnell2015adaptation; @fischer2015categorizing].
+There is clearly a need for continuous measures of urbanization [@evans2011makes], where a species is assessed over its entire distribution in response to urban environments.
 
 Most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does] or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living].
-Broad-scale empirical data allow testing of hypotheses at large spatiotemporal scales.
+Broad-scale empirical data allow hypotheses to be tested at large spatiotemporal scales.
 Citizen science can provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen], and can be used to describe patterns of abundance, distribution, and functional composition of birds in urban ecosystems [@mccaffrey2005using; @la2014role; @morelli2016evidence; @lepczyk2017global @callaghan2017assessing].
 
-We aimed to determine what suite of life history traits predicts avian tolerance to urbanization at a continental scale.
+We aimed to determine what suite of life history traits predicted avian tolerance to urbanization at a continental scale.
 We predicted that different traits would positively and negatively associate with urbanization.
-For generality, our analyses had to (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3) include a broad-range of traits, and (4) employ a continuous measure of response to urbanization.
-We integrated a series of large datasets to establish a novel continuous urbanization index to test associations of ecological and life history traits which influenced a species' responses to urbanization.
-We first investigated evidence of a phylogenetic signal in our urbanization index [*sensu* @evans2011makes], given our comprehensive sample size.
-We then developed models, including and excluding phylogenetic constraints, in order to test for association between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.
+For generality, our analyses had to (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3) include a broad-range of traits, and (4) employ a continuous measure of urbanization.
+We integrated a series of large datasets to establish a novel continuous urbanization index to test associations with ecological and life history traits influencing species' responses to urbanization.
+We first investigated whether our urbanization index was statistically independent in relation to phylogeny (i.e., phylogenetic signal - Revell et al.
+2008; Evans et al.
+2011), given our comprehensive sample size.
+We then developed models, including and excluding phylogenetic constraints, to test for associations between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.
 
@@
 We collated bird observations throughout mainland Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists.
-Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird].
+Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird], often comparable to trained observers [@callaghan2015efficacy].
 We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018.
@@
 We filtered bird observations [@callaghan2017assessing; @la2014role], only including observations on complete checklists and which were either travelling, random, stationary, area, or followed BirdLife Australia protocols [@barrett2003new] (see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types).
-We then filtered the checklists to those which travelled < 5 km or covered < 500 Ha.
-Any checklists shared among multiple individuals were subsampled, including only one checklist in order to avoid duplicate results.
+We then filtered the checklists to those which travelled < 5 km or covered < 500 Ha, to ensure the highest quality data were used.
+Any checklists shared among multiple observers were subsampled, including only one checklist to avoid duplication.
 This provided 637,482 eligible checklists throughout Australia.
-Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we did not expect any correlation with seabirds using urban areas.
-Further, only species with > 100 observations were considered for further analyses.
-Totally, 580 species, a substantial proportion of the ~730 bird species which are resident in Australia, were initially considered for analyses.
-We used differing taxonomic authorities (see Appendix 2 for elaboration).
+Seabirds (e.g., gannets, petrels, shearwaters) were omitted from analyses as we did not expect any correlation with seabirds using urban areas.
+Further, only species with > 100 observations were analyzed.
+In total, 580 species, most (~80%) of Australia's bird species were candidates for analyses.
+We used differing taxonomic authorities (see Appendix 2).
 
 ## Urbanization index
-We used VIIRS night-time lights [@elvidge2017viirs], for urban habitat, measuring the electric lighting present on Earth's surface, generally associated with human settlement (Fig.
+We used VIIRS night-time lights [@elvidge2017viirs] as a continuous measure for urban habitat, measuring the electric lighting associated with human settlement (Fig.
 S1).
-The night-time lights dataset filters out background noise (e.g., fires, degraded data, and other light source contamination) to ensure validity of the data and its association with human settlement [@elvidge2017viirs].
-For our 637,482 checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], to assign the average radiance value calculated from the VIIRS layer.
-For each checklist, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
+This has filtered out background noise (e.g., fires, degraded data, and other light source contamination) to ensure validity of the data and its association with human settlement [@elvidge2017viirs].
+For our checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], to assign the average radiance value, calculated from the VIIRS layer.
+For each checklist, the average radiance was calculated within a 5 km buffer, to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols.
 Any negative radiance values were forced to 0.00001 [*sensu* @ou2015evaluation].
 Every bird species subsequently had a distribution of average radiance values (Fig.
-1), with the median of each species' distribution used as the response variable in our analysis.
-There was no definitive range for this index, depending on a species' distribution of observations.
+1), with the median of each species' distribution becoming the response variable.
+There was no definitive range for this index, dependent on a species' distribution of observations.
 
 ## Ecological and life history traits
-Ecological and life history traits were chosen on the basis of existing support [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes; Table 1; Appendix 1a].
+Ecological and life history traits were chosen on the basis of existing support (e.g., McClure 1989; Kark *et al.* 2007; Evans *et al.* 2011; Table 1; Appendix 1).
 These included a wide array of potential traits for all of Australia's birds [@garnett2015biological].
 A total of 22 traits were included in analyses: clutch size, body size, brain residual, nest generalism, feeding habitat generalism, breeding habitat generalism, diet generalism, habitat association with agricultural landscapes, habitat association with treed/forested habitats, habitat association with grass/shrubland habitats, whether a species nests in hollows, whether a species nests on the ground, whether a species eats plants, whether a species is a granivore, whether a species is an insectivore, whether a species eats carrion, whether a species is nomadic/irruptive, whether a species is migratory, the range size of a species, nesting aggregation of a species, feeding aggregation of a species, and whether a species is a cooperative breeder (Table 1; Appendix 1a).
-For any missing trait in the dataset [@garnett2015biological] for a species we assumed it did not exist.
-For example, if a species had no data for a specific habitat, it was deemed not to occur in the habitat.
-Our preliminary analyses included IUCN status and showed that exotic birds were strongly associated with urbanization, but given the preferential selection by acclimatization societies for introduction of species with proven synanthropy [@blackburn2009avian], exotic species were omitted from the final analyses to avoid bias.
+For any missing trait in the dataset [@garnett2015biological] for a species, we assumed it did not exist.
+For example, if a species had no data for a specific habitat, it was deemed not to occur in that habitat.
+Our preliminary analyses, which included IUCN status, showed that exotic birds strongly associated with urbanization; these were omitted from the final analyses to avoid bias, given the preferential selection by acclimatization societies for their proven synanthropy [@blackburn2009avian].
 We also excluded IUCN status because most (96%) of our candidate species were of Least Concern.
-Totally, 477 species with complete data for all candidate traits were included in analyses.
+In total, 477 species had complete data for all candidate traits and were included in analyses.
 
 ## Statistical analyses
-Before modelling of predictor variables (Table 1) we investigated multi-collinearity among continuous variables (Fig.
+Before modelling of predictor variables (Table 1), we investigated multi-collinearity among continuous variables (Fig.
 S2), using the corrplot package [@corrplot2017].
-We used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other, and all traits were < 2 suggesting negligible correlation (Supplementary Table 1).
-We used two independent linear modelling approaches: one without and one with phylogenetic constraints.
+We used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation among each other; all traits < 2 suggested negligible correlation (Supplementary Table 1).
+We used two independent linear modelling approaches: one without and one with phylogenetic effects.
+Phylogenetic effects are defined as the tendency of closely related organisms to have similar life history components due to their shared evolutionary history [@derrickson1988taxon; @mckitrick1993phylogenetic].
 Both approaches were used because of shortcomings [@cadotte2017phylogenies] and importance [@losos2008phylogenetic; @graham2018phylogenetic] of incorporating evolutionary distinctness in ecological research.
 
-Linear regression models were weighted [@solon2015weighting] using a source of error associated with our urbanization indices, which were log-transformed to meet model assumptions.
-This error was the proportion of the number of total observations of a species, divided by the number of its unique localities (Fig.
+Linear regression models were weighted [@solon2015weighting], using a source of error associated with our urbanization indices, which were log-transformed to meet model assumptions.
+This error was the proportion of the number of total observations for a species, divided by the number of its unique localities (Fig.
 S2 shows the correlation between the two).
 This weighted the model-fitting to those species with more observations in more unique locations, likely to be the most reliable urbanization indices.
-This approach also addressed biases in sampling of eBird data where some birds have a high number of observations from a small number of locations such as known roosting sites [e.g., @callaghan2018avian].
-
-All models and analyses were conducted in R [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse]. Statistical significance was concluded as $\alpha$ < 0.05.
+This approach also addressed biases in sampling of eBird data, where some birds have a high number of observations from a small number of locations such as known roosting sites [e.g., @callaghan2018avian].
+All models and analyses were conducted in R [@r2017coreteam], relying on the tidyverse workflow [@wickham2017tidyverse].
+Statistical significance was concluded as $\alpha$ < 0.05.
+Variability of means is reported as standard deviations.
 
 ## Non-phylogenetically controlled models
-We first ran one global model, including all 22 predictor variables, using the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes.
-To corroborate our global model results, we employed a model-averaging approach [@grueber2011multimodel] where we 'dredged' possible subsets of the global model, using the dredge function in the MuMIn package [@barton2016mumin].
-We fitted a total of 11 candidate predictors at any one time to avoid over-fitting, totaling 784,626 models assessed.
-Model-averaged parameter estimates were calculated from the subsequent top models, with $\Delta$AIC < 4, with standardized estimates representing relative effect sizes of respective traits.
+We first ran a global model, including all 22 predictor traits, standardizing model estimates to represent relative effect sizes, using the standardize function from the arm package [@gelman2008scaling; @gelman2016arm].
+To corroborate our global model results, we employed a model-averaging approach [@grueber2011multimodel], 'dredging' all possible subsets of the global model with the dredge function in the MuMIn package [@barton2016mumin].
+We fitted 11 candidate traits at any one time to avoid over-fitting, assessing 784,626 models.
+Model-averaged parameter estimates were calculated from the subsequent top models, with $\Delta$AIC < 4, where standardized estimates represented relative effect sizes of respective traits.
 
@@
 ### Phylogenetic signal
-As associations between traits and urbanization may be influenced by shared evolutionary history of a species [@felsenstein1985phylogenies], we also used a phylogenetically informed analytical approach.
-First, we assessed the existence of a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] associated with our urbanization index and for each of the continuous predictor variables (Table 1), using published phylogenetic trees for 9,993 species [@jetz2012global].
-A phylogenetic signal is defined as the statistical nonindependence among species trait values due to their phylogenetic relatedness [@revell2008phylogenetic], and it assesses the relative contribution of phylogeny on a given trait.
-There are a number of indices used to assess phylogenetic signal, each with specific strengths and weaknesses [@keck2016phylosignal], and we assessed all available indices (Blomberg's K and K*, Abouheif's Cmean, Moran's I, and Pagel's Lambda), using the phylosignal package [@keck2016phylosignal] in R.
+As associations between traits and urbanization may be influenced by shared evolutionary history of a species [@felsenstein1985phylogenies], we also modelled this phylogenetically.
+First, we tested for a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] associated with our urbanization index and for each of the continuous predictor traits (Table 1), using published phylogenetic trees [@jetz2012global].
+This was defined as the statistical nonindependence among species' traits, due to their phylogenetic relatedness [@revell2008phylogenetic], assessing relative contribution of phylogeny on a given trait.
+There are different indices used to assess phylogenetic signal, each with strengths and weaknesses [@keck2016phylosignal]; we assessed all available indices (Blomberg's K and K*, Abouheif's Cmean, Moran's I, and Pagel's Lambda), using the phylosignal package [@keck2016phylosignal] in R.
 
 ### Phylogenetically controlled models
-Second, we used phylogenetically informed models to test for repeated evolutionary association between traits of interest [@felsenstein1985phylogenies].
-Phylogenetic models were fit based on the same global model as the non-phylogenetic models, but using the 'phylolm' function from the phylolm package which fits a linear regression model where the likelihood is linear in the number of tips in the tree [@ho2014phylolm].
-We also model-averaged the global phylogenetic model using the aforementioned methods.
+Second, we used phylogenetically informed models to test for repeated evolutionary associations between traits of interest [@felsenstein1985phylogenies].
+Phylogenetic models were fitted based on the same global model as the non-phylogenetic models, but using the 'phylolm' function [@ho2014phylolm], which fits a linear regression model where the likelihood is linear in the number of tips in the tree.
+We similarly model-averaged the global phylogenetic model.
 To account for phylogenetic uncertainty, we fitted the same model to 1000 highly probable trees from Jetz *et al.* [-@jetz2012global], examining the distribution of putatively important traits across the distribution of trees.
@@
 # RESULTS
-We analyzed 316,306 checklists including 5,944,819 observations of 477 different species.
-The least recorded species was *Turnix melanogaster* (Black-breasted Buttonquail; N=104) while the most recorded species was *Cracticus tibicen* (Australian Magpie; N=158,615).
-The mean ($\pm$ SD) number of observations for a species was 10,930 $\pm$ 19,498.
-Distributions of measured VIIRS average radiance values (shown for each species in Appendix 3), traits, number of observations, and urbanization indices for the 477 species (Appendix 4) varied considerably.
-The mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, ranging from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig.
+We analyzed 316,306 checklists, including 5,944,819 observations of our 477 different species.
+The least recorded species was *Turnix melanogaster* (Black-breasted Buttonquail; N=104), while the most recorded species was *Cracticus tibicen* (Australian Magpie; N=158,615).
+The mean number of observations for a species was 10,930 $\pm$ 19,498.
+Distributions of measured VIIRS average radiance values (shown for each species in Appendix 3), traits, number of observations, and urbanization indices for the 477 species varied considerably (Appendix 4).
+The mean urbanization index was 0.91 $\pm$ 1.76, ranging from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig.
 2).
@@
 Further, feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006) were significant.
-Urban-tolerant birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032), but less so than the previous traits.
+Urban-tolerant birds were also significantly positively associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032), but had smaller effect sizes than previous traits.
 Conversely, urban-tolerant birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), group foraging (t = -2.24, p =0.025), granivorous diet (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), and grass/shrubland habitats (t = -4.26, p < 0.001), with insectivorous diet (t = -3.96, p < 0.001) and solitary feeding (t = -2.90, p = 0.004).
@@
 4; Appendix 5).
-However, the most strongly associated trait, based on standardized parameter estimates was clutch size (0.75), followed by birds associated with agricultural habitats (0.58), and then diet generalism (0.28).
+However, the most strongly associated trait was clutch size (0.75), followed by birds associated with agricultural habitats (0.58), and then diet generalism (0.28).
 Breeding habitat (0.15) and feeding habitat (0.12) generalism parameter estimates were substantially smaller than in the non model-averaged approach.
@@
 Urban-tolerant birds remained significantly negatively associated with solitary feeding (z = 3.39, p = 0.001),  insectivorous diet (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), group foraging (z = 3.16, p = 0.002), granivorous diet (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).
-Solitary feeding (-1.00) remained the strongest negatively associated trait with urbanization, followed by insectivorous diet (-0.98).
+Solitary feeding (-1.00) was the strongest negatively associated trait with urbanization, followed by insectivorous diet (-0.98).
 
@@
 ### Phylogenetic signal
-There was a distinct relationship of phylogenetic signal to our urbanization index (Fig.
+There was a distinct relationship between the phylogenetic signal and our urbanization index (Fig.
 3; Fig.
-S3; Table 2), as it was significantly different than random for 4 out of the 5 indices considered (Cmean, I, K.star, and Lambda).
-For example, the urbanization index is disproportionate in respect to the phylogenetic tree (Fig.
-3), evidenced by the highest urbanization indices being clustered throughout the waterbird clade.
-Further, within a specific clade there were some species which had much larger urbanization indices (e.g., note the high index for *Manorina melanocephala*, Noisy Miner, compared to the rest of the honeyeaters).
-We generally found weaker signal for behavioral traits, compared with morphological traits.
-For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 2).
-Additionally, there was strong phylogenetic signal for body size and brain residual size (Table 2).
+S3; Table 2), significantly different from random for 4 out of 5 indices (Cmean, I, K.star, and Lambda).
+For example, the urbanization index was disproportionately important with respect to the phylogenetic tree (Fig.
+3), evidenced by the highest urbanization indices clustered throughout the waterbird clade.
+Further, within a specific clade, there were some species which substantially higher urbanization indices (e.g., *Manorina melanocephala*, Noisy Miner, cf.
+honeyeaters).
+We generally found weaker signals for behavioral, compared with morphological traits.
+For behavioural traits, the estimates of phylogenetic signal were weaker than Brownian motion, but usually significant (Table 2).
+Additionally, there was a strong phylogenetic signal for body size and brain residual size (Table 2).
 
@@
 Brain residual size (1.25) had the largest standardized parameter estimate for the phylogenetically controlled global model (Fig.
-4; Appendix 5), followed by feeding habitat generalism (1.18), clutch size (0.91), and body size (0.82).The most significant predictors of urbanization in the phylogenetically controlled global model  were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism (t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005).
-Nest generalism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding (t = -3.25, p = 0.001), and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.
+4; Appendix 5), followed by feeding habitat generalism (1.18), clutch size (0.91), and body size (0.82).The most significantly positive traits associated with urbanization were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism (t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005).
+Conversely, nest generalism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding (t = -3.25, p = 0.001), and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with urbanization.
 
 With model-averaging (Fig.
-4; Appendix 5), feeding habitat generalism (1.30), clutch size (1.15), body size (0.66), and residual brain size (0.32) showed the strongest association with our urbanization index, while species with no known nest aggregation (-1.22) and species associated with grass/shrubland (-1.16) showed the strongest negative response.
-Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were positively statistically significant, while species which were cooperative breeders (z = 2.53, p = 0.011), used grass/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were negatively statistically significant.
+4; Appendix 5), feeding habitat generalism (1.30), clutch size (1.15), body size (0.66), and residual brain size (0.32) were most strongly associated with our urbanization index, while species with no known nest aggregation (-1.22) and species associated with grass/shrubland (-1.16) showed the strongest negative response.
+Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significantly positively asspcoated, while species which were cooperative breeders (z = 2.53, p = 0.011), used grass/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were statistically negatively associated.
 
 ## Differences in non-phylogenetic and phylogenetic models
-Non-phylogenetic and phylogenetic models largely agreed in the relative rank of standardized predictor estimates (Appendix 5) and the statistical significance of individual predictors (Fig.
+Non-phylogenetic and phylogenetic models had similar relative ranks of standardized trait estimates and statistical significance (Fig.
 4; Appendix 5).
 However, residual brain size was not significant in either non-phylogenetic model approaches (Fig.
-4), but was significantly associated with urbanization in our phylogenetic modelling, showing the strongest relative response in the phylogenetic global model (Fig.
+4), but significantly associated with urbanization in our phylogenetic modelling, showing the strongest relative response in the phylogenetic global model (Fig.
 4; Fig.
 5).
-Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (large standardized parameter estimates) in the phylogenetic models, compared with the non-phylogenetic models (Fig.
+Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important in the phylogenetic models, compared with the non-phylogenetic models (Fig.
 5; Appendix 5).
-Lastly, although not significant, body size was minimally important for the non-phylogenetic models, but substantially more important for the phylogenetically controlled models (Fig.
+Lastly, body size was minimally important for the non-phylogenetic models, but substantially more important for the phylogenetically controlled models (Fig.
 4 & Fig.
@@
 # DISCUSSION
-The most urban-tolerant species in our analyses (i.e., the ones with a higher urbanization index) had larger niche breadths, as measured by different forms of generalism.
-Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) were less likely to be found in urban environments.
-These results were robust across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
+The most urban-tolerant species in our analyses, defined based on their urbanization indices, had large niche breadths, reflected in generalism.
+Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) seldom occurred in urban environments.
+These results were robust across separate non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
+S4), indicating that urbanization is generally excluding specialist species [@sorace2009distribution; @concepcion2015impacts].
+Further, the non-random distribution of the urbanization index across the phylogeny (Fig.
+3 & Table 2) is consistent of reduced phylogenetic diversity in urban areas [@la2018phylogenetic].
+Particular phylogenetic effects [@derrickson1988taxon; @mckitrick1993phylogenetic] probably allow a subset of species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
+
+This strong phylogenetic signal contradicts other studies [@evans2011makes], but was unsurprising given our large sample size of 477 species compared to the next highest of 55 species [@evans2011makes].
+We also demonstrated the importance of treating species as non-independent, by considering phylogeny in the context of the linear models, highlighting the importance of investigating trait-urbanization relationships in a phylogenetically informed manner.
+For example, some of the more important traits in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) had a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models, compared to non-phylogenetic models.
+This was robust relative to phylogenetic uncertainty (Fig.
 S4).
-These results suggest that the process of urbanization is, in general, operating to exclude specialist species [@sorace2009distribution; @concepcion2015impacts].
-Further, the non-random distribution of the urbanization index across the phylogeny (Fig.
-3 & Table 2) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that particular phylogenetic constraints allow certain species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
+Other traits showed the opposite trend.
+Most notably, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models.
+This indicates that within clades, species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] were most suited to urban-tolerance.
+Some have indicated poor evidence that urban-tolerant species have large residual brain size [@kark2007living; @evans2011makes], but we found it was highly correlated with urban-tolerance [Fig.
+4; Fig.
+5; @moller2009successful; @maklakov2011brains; @moller2015brain].
 
-Our finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is unsurprising given our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].
-We further demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models, and thus highlight the importance of investigating trait-urbanization relationships in a phylogenetically informed manner.
-For example, some of the most important predictor variables in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig.
-S4); and other traits showed the opposite trend.
-Most notably, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas.
-While some studies have found a lack of evidence that residual brain size is an important predictor in urban-tolerant species [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed highly correlated with urban-tolerance [@moller2009successful; @maklakov2011brains; @moller2015brain].
-
-Many traits have been used to identify a species' adaptability to urban environments, but results have frequently differed among studies.
-As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization.
-Our models assessed the applicability of traits in relation to one another [i.e., we did not run separate models or statistical tests for each specific trait as previously done; e.g., @kark2007living] thus allowing comparable effect sizes.
-Despite this, there appear to be some consistent, and thus generalizable, trends within the trait-urbanization complex across the globe.
+Although many traits can identify a species' adaptability to urban environments, the published evidence has remained equivocal.
+Previously, much of the current focus was on discrete classifications of species, based on their responses to urbanization.
+Contrastingly, we assessed the applicability of traits relative to one another; we did not run separate models or statistical tests for each specific trait [e.g., @kark2007living], allowing comparable effect sizes.
+Despite this, there appear to be some consistent and generalizable trends within the trait-urbanization complex across the globe.
 For instance, our finding of the relative importance of annual fecundity is consistent with work in Europe [@croci2008does; @moller2009successful].
@@
 
-```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for each of the 22 traits' standardized parameter estimates, for the 477 species.
-Proportional differences represent the difference between a given trait's parameter estimate for the non-phylogenetically controlled and phylogenetically controlled global models."}
+```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for each of the 22 traits' (Table 1) standardized parameter estimates, for the 477 species.
+Proportional differences represented the difference between a given trait's parameter estimate for the non-phylogenetically controlled and phylogenetically controlled global models."}
 knitr::include_graphics("finalFigs/Figure5.pdf")

### 2018-06-18 ms.Rmd
more changes/edits/polish

@@
 Further, the non-random distribution of the urbanization index across the phylogeny (Fig.
-3 & Table 2) is consistent of reduced phylogenetic diversity in urban areas [@la2018phylogenetic].
-Particular phylogenetic effects [@derrickson1988taxon; @mckitrick1993phylogenetic] probably allow a subset of species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
+3 & Table 2) is consistent with reduced phylogenetic diversity in urban areas [@la2018phylogenetic].
+Particular phylogenetic effects [@derrickson1988taxon; @mckitrick1993phylogenetic] probably allow specific subsets of species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].
 
@@
 We also demonstrated the importance of treating species as non-independent, by considering phylogeny in the context of the linear models, highlighting the importance of investigating trait-urbanization relationships in a phylogenetically informed manner.
-For example, some of the more important traits in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) had a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models, compared to non-phylogenetic models.
-This was robust relative to phylogenetic uncertainty (Fig.
-S4).
+For example, some of the more important traits in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) had a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models than non-phylogenetic models.
 Other traits showed the opposite trend.
@@
 Although many traits can identify a species' adaptability to urban environments, the published evidence has remained equivocal.
-Previously, much of the current focus was on discrete classifications of species, based on their responses to urbanization.
+Previously, much of the current focus was on discrete classifications of species, based on their responses to urbanization [@geschkecompact].
 Contrastingly, we assessed the applicability of traits relative to one another; we did not run separate models or statistical tests for each specific trait [e.g., @kark2007living], allowing comparable effect sizes.
-Despite this, there appear to be some consistent and generalizable trends within the trait-urbanization complex across the globe.
-For instance, our finding of the relative importance of annual fecundity is consistent with work in Europe [@croci2008does; @moller2009successful].
-Additionally, urbanization appears to generally select against insectivorous species [@chace2006urban; @evans2011makes].
-Although we did not find any strong support for our hypotheses on different nesting types, we found a positive association between the degree of sociality and urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
-While the aforementioned patterns may be generalizable, we also found patterns in our analyses which seem to contrast the global norm.
-For example, we found that urbanization selects against granivorous species, contrasting with the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [@jones2018birdfeeding] generally involving different species and foods  [compared to European cities for instance; @fuller2008garden].
+There appear to be some consistent and generalizable trends within the trait-urbanization complex across the globe.
+For instance, our identification of the relative importance of annual fecundity was consistent with work from Europe [@croci2008does; @moller2009successful].
+Urbanization also appears to select against insectivorous species [@chace2006urban; @evans2011makes].
+The degree of sociality was also generally positively associated with urbanization, as found elsewhere [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
+However, our study also highlighted differences with global norms.
+For our data, urbanization selected against granivorous species, not previously identified [cf.
+@chace2006urban; @kark2007living; @evans2011makes]; this could be an artefact of relatively little supplementary feeding in Australian cities [@jones2018birdfeeding] compared to European cities [@fuller2008garden].
 
-Our analysis was made possible through integration of broad-scale empirical datasets, which are rapidly expanding our knowledge of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
-This is largely due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google].
-We did not explicitly account for changes throughout the annual cycle in a species' association with urban environments [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor.
-We emphasize that future studies, especially within migrant-dominated systems (e.g., Europe, North America), should account for changes in a species' urbanization index throughout the full-annual cycle [@hostetler2015full; @marra2015call].
-This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species, thus allowing prediction of which species face the greatest risk of impending urbanization, specifically in developing countries.
+The large scale of our analysis was only possible following successful integration of broad-scale empirical datasets (i.e., citizen science data [@silvertown2009new] and remote sensing technology [@gorelick2017google]), which aid our understanding of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen].
+We did not explicitly incorporate temporal changes in relation to the life cycle for different species' [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered.
+In future, especially within migrant-dominated systems (e.g., Europe, North America), temporal changes in association with our urbanization index would add considerable value [@hostetler2015full; @marra2015call].
+Australia has only a recent history of rapid urbanization [@coffee2016visualising] and our results may differ for parts of the world with long histories of development, and at different spatial scales [@suarez2018scale].
+Yet, there is an opportunity to test our model predictions in urban environments in other parts of the world, potentially allowing prediction of which species face the greatest risk of impending urbanization.
 
-Fully understanding the suite of traits which are associated with urban-tolerant species will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes].
-Given specialist species are known to have strong negative responses to landscape fragmentation and disturbance [@devictor2008distribution], sustainable urban development with a goal of maintaining biological diversity in urban areas should focus on providing intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role].
-We recommend that a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more].
-A species-specific approach, which assesses particular resources necessary for specialist species, for example focusing on foraging [@shochat2004linking] and nesting [@fernandez2001habitat] requirements, is likely necessary.
-To this end, our urbanization index can be used as a potential means to provide quantifiable targets for local restoration projects.
-For example, a particular restoration project should aim to add species which have lower urbanization indices into their local avifauna, but we note that these indices may differ at different spatial scales [@suarez2018scale].
+Understanding which traits promote urban-tolerance or intolerance can inform actions to maintain avian biodiversity in urban environments [@evans2011makes].
+In particular, future urban development could maintain biodiversity through improved protection of intact native vegetation [@sushinsky2013should] connected by corridors [@litteral2017role], advantaging specialist species detrimentally affected by landscape fragmentation and disturbance [@devictor2008distribution].
+Further, increasing insect abundance and diversity in urban environments is potentially important [@baldock2015uk], given collapsing insect populations around the world [@hallmann2017more].
+A species-specific approach could help to enhance urban biodiversity, which assesses foraging [@shochat2004linking] or nesting [@fernandez2001habitat] requirements of each species, for example, used to guide protection, mitigation, and restoration of necessary habitats in cities.
+Our urbanization index could provide quantifiable targets for local restoration projects, where a project aims to reduce the mean urbanization index for their local avifauna.
 
-We provide a novel continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments.
-Future studies investigating the trait-urbanization relationship should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance.
-Our results suggest that generalist species, with high clutch sizes and large residual brain sizes, are best prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns].
-In the context of our specific findings, Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management.
-Urban avian biodiversity could be increased by emphasizing the critical importance of intact vegetation [@sushinsky2013should] for specialist species [@devictor2008distribution], and concomitantly improving the suitability of foraging for granivores and insectivores.
+In conclusion, our novel continuous urbanization index showed relationships between ecological and life history traits with avian tolerance to urban environments.
+Improving understanding of the temporal nature of these relationships will be important, but we highlight the importance of considering these questions using a continuous measure of urbanization [@evans2011makes] as well as using a phylogenetically informed approach.
+It is the generalist species [@shochat2006patterns], with high clutch sizes and large residual brain sizes, which are best prepared to forage and survive in the ecologically novel aspects of urban environments.
 

### 2018-06-25 ms.Rmd
changes for Proc B.

@@
 ---
-title: "Generalists are the most urban-tolerant of birds: a phylogenetically controlled analysis of ecological and life-history traits using a novel continuous urbanization index"
+title: "Generalists are the most urban-tolerant of birds: an analysis of ecological and life history traits using a novel continuous urbanization index"
 
@@
 link-citations: no
-csl: https://raw.githubusercontent.com/citation-style-language/styles/master/ecology-letters.csl
+csl: https://raw.githubusercontent.com/citation-style-language/styles/master/proceedings-of-the-royal-society-b.csl
 bibliography: references.bib
@@
 Enhancement and persistence of avian biodiversity in urban environments probably relies on protecting, maintaining, and restoring diverse habitats serving a range of life history strategies.
+
+*Keywords*: eBird, citizen science, phylogenetics, life history traits, avian, brain size
 
@@
 We integrated a series of large datasets to establish a novel continuous urbanization index to test associations with ecological and life history traits influencing species' responses to urbanization.
-We first investigated whether our urbanization index was statistically independent in relation to phylogeny (i.e., phylogenetic signal - Revell et al.
-2008; Evans et al.
-2011), given our comprehensive sample size.
-We then developed models, including and excluding phylogenetic constraints, to test for associations between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.
+We first investigated whether our urbanization index was statistically independent in relation to phylogeny (i.e., phylogenetic signal - [@revell2008phylogenetic; @evans2011makes]), given our comprehensive sample size.
+We then developed models, including and excluding phylogenetic effects, to test for associations between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.
 
@@
 Further, only species with > 100 observations were analyzed.
-In total, 580 species, most (~80%) of Australia's bird species were candidates for analyses.
+In total 580 species, most (~80%) of Australia's bird species, were candidates for analyses.
 We used differing taxonomic authorities (see Appendix 2).
@@
 ## Ecological and life history traits
-Ecological and life history traits were chosen on the basis of existing support (e.g., McClure 1989; Kark *et al.* 2007; Evans *et al.* 2011; Table 1; Appendix 1).
-These included a wide array of potential traits for all of Australia's birds [@garnett2015biological].
-A total of 22 traits were included in analyses: clutch size, body size, brain residual, nest generalism, feeding habitat generalism, breeding habitat generalism, diet generalism, habitat association with agricultural landscapes, habitat association with treed/forested habitats, habitat association with grass/shrubland habitats, whether a species nests in hollows, whether a species nests on the ground, whether a species eats plants, whether a species is a granivore, whether a species is an insectivore, whether a species eats carrion, whether a species is nomadic/irruptive, whether a species is migratory, the range size of a species, nesting aggregation of a species, feeding aggregation of a species, and whether a species is a cooperative breeder (Table 1; Appendix 1a).
+Ecological and life history traits were chosen on the basis of existing support [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes] (Table 1; Appendix 1), including a wide array of potential traits for all of Australia's birds [@garnett2015biological].
+A total of 22 traits, grouped into 7 categories, were included (Table 1): life history traits (n=3), measures of niche breadth (n=4), specific habitat preferences (n=3), specific nesting preferences (n=2), specific diet preferences (n=4), movement and range size (n=3), and degree of sociality (n=3).
 For any missing trait in the dataset [@garnett2015biological] for a species, we assumed it did not exist.
@@
 3), evidenced by the highest urbanization indices clustered throughout the waterbird clade.
-Further, within a specific clade, there were some species which substantially higher urbanization indices (e.g., *Manorina melanocephala*, Noisy Miner, cf.
+Further, within a specific clade, there were some species with substantially higher urbanization indices (e.g., *Manorina melanocephala*, Noisy Miner, cf.
 honeyeaters).
@@
 
-## Acknowledgments
+# Data, code and materials
+This work relied heavily on open-source, fully accessible datasets.
+The necessary portions of those datasets, along with code to reproduce the analyses are available at
+this github repository (https://github.com/cornwell-lab-unsw/bird_urbanness), and will be
+permanently archived upon acceptance of the manuscript. The supporting data has been included in the article's supplementary material (Appendix 4).
+
+# Competing interests
+We have no competing interests.
+
+# Author's Contributions
+CTC, WKC, and JHW conceived and designed the methodology of the
+study. CTC, WKC performed the statistical analyses. JMM, REM, and RTK contributed in drafting
+the manuscript. All authors provided editorial advice, approved the final version of the manuscript,
+and agree to be accountable for all aspects of the work.
+
+# Acknowledgments
 We thank the > 6,000 eBirders who have contributed data in Australia.
@@
 3.
+
+# Funding
+The authors declare no sources of funding specific to this work.
 
@@
 library(readr)
+library(dplyr)
 Table1a <- read_csv("Data/Table1.csv")
@@
 Table1a[is.num] <- lapply(Table1a[is.num], round, 3)
+Table1a <- Table1a %>%
+  rename(`Abouheif's Cmean` = Cmean) %>%
+  rename(`Moran's I` = I) %>%
+  rename(`Blomberg's K` = K) %>%
+  rename(`K*` = K.star) %>%
+  rename(`Pagel's Lambda` = Lambda)
 
@@
 Table1b[is.num] <- lapply(Table1b[is.num], round, 3)
+Table1b <- Table1b %>%
+  rename(`Abouheif's Cmean` = Cmean) %>%
+  rename(`Moran's I` = I) %>%
+  rename(`Blomberg's K` = K) %>%
+  rename(`K*` = K.star) %>%
+  rename(`Pagel's Lambda` = Lambda)
 
 grouped <- rbind(cbind(Value="Statistic", Table1a), cbind(Value="p-values", Table1b))
-tgrp <- xtabs(cbind(Cmean, I, K, K.star, Lambda) ~ Value+Variable, grouped)
+tgrp <- xtabs(cbind(`Abouheif's Cmean`, `Moran's I`, `Blomberg's K`, `K*`, `Pagel's Lambda`) ~ Value+Variable, grouped)
 
@@
 pander(table, caption = "Summary of results showing an assessment of a phylogenetic signal for our response variable (urbanization index) and the continuous predictor traits.
-We show all available indices (Blomberg's K and K*, Abouheif's Cmean, Moran's I, and Pagel's Lambda; Keck et al.
+We show all available indices (Keck et al.
 2016), each of which represents the statistical nonindependence of a trait's value due to its phylogenetic relatedness (Revell et al.
@@
   rename(Trait = X1)
+collinearity_investigation <- collinearity_investigation[c(2, 1, 4, 16, 3, 6, 11, 5, 22, 20, 15, 14, 10, 7, 8, 9, 13, 12, 21, 18, 19, 17), ]
 

### 2018-09-03 ms.Rmd
revisions for next version of paper

@@
 link-citations: no
-csl: https://raw.githubusercontent.com/citation-style-language/styles/master/proceedings-of-the-royal-society-b.csl
+csl: https://raw.githubusercontent.com/citation-style-language/styles/master/oikos.csl
 bibliography: references.bib
@@
 Some species successfully adapt, colonize, and thrive in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others, common in surrounding natural habitats, rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird].
-Urbanization selects for specific sets of species, globally threatening biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], a major concern for conservation [@czech2000economic; @dearborn2010motivations].
+Urbanization selects for specific sets of species, based on the regional pool [@aronson2016hierarchical], globally threatening biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], a major concern for conservation [@czech2000economic; @dearborn2010motivations].
 It is important to identify which ecological and life history traits most associate with successful or unsuccessful species in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful; @mcdonnell2015adaptation].
-This can improve understanding of relationships between biodiversity and urbanization, and guide protection of the species most at-risk of extinction, through habitat protection, management, and restoration.
+Identifying these traits can improve understanding of relationships between biodiversity and urbanization, and guide protection of the species most at-risk of extinction, through habitat protection, management, and restoration.
 
@@
 In Jerusalem, urban bird species were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while, throughout France and Switzerland, urban bird species were positively associated with forested habitats, sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does].
-There are also contrasting results: residual brain size in birds is sometimes positively associated with urbanization [@maklakov2011brains; @moller2015brain] and sometimes not [@kark2007living; @evans2011makes].
+Despite the breadth of research which has focused on ecological and life history traits associated with urbanization in birds, the results are severely mixed and frequently contradictory.
+For example, residual brain size in birds is sometimes positively associated with urbanization [@maklakov2011brains; @moller2015brain] and sometimes not [@kark2007living; @evans2011makes].
 Similarly, contrasting results have been reported for annual fecundity [cf., @croci2008does; @moller2009successful; @evans2011makes] and niche breadth [cf., @kark2007living; @evans2011makes].
+Ultimately, a more comprehensive and generalized understanding of ecological and life history traits associated with urban-tolerant birds is necessary.
 
+Owing to the contrasting and mixed results are methodological shortcomings.
 Most analyses assign species to discrete a priori groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or strictly presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful].
-These approaches severely limit the capacity to generalize, as it assumes grouped species respond equally to urbanization [@evans2011makes].
+These approaches severely limit the capacity to generalize, as it assumes grouped species respond equally to urbanization [@lepczyk2008human; @evans2011makes].
 Terminology also varies among studies [cf., @kark2007living; @croci2008does; @geschkecompact], affecting interpretation [@mcdonnell2015adaptation; @fischer2015categorizing].
 There is clearly a need for continuous measures of urbanization [@evans2011makes], where a species is assessed over its entire distribution in response to urban environments.
-
-Most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does] or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living].
+Further, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does] or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living], restricting the generalizability of the results.
 Broad-scale empirical data allow hypotheses to be tested at large spatiotemporal scales.
@@
 We aimed to determine what suite of life history traits predicted avian tolerance to urbanization at a continental scale.
+Our study represented the largest sample size and largest spatial scale ever investigated, with regards to ecological and life history traits and urban-tolerance in birds.
 We predicted that different traits would positively and negatively associate with urbanization.
@@
 We then developed models, including and excluding phylogenetic effects, to test for associations between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.
+Importantly, given the mixed results of previous research assessing this question and thus competing a priori hypotheses, we found it important to include all predictor variables in the models.
 
@@
 We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018.
-This corresponded to the richest period of eBird data, minimizing undue leverage of mismatch between changes in response variables over which bird observations were collected.
+This temporal scale corresponded to the richest period of eBird data, minimizing undue leverage of mismatch between changes in response variables over which bird observations were collected.
 
@@
 Any checklists shared among multiple observers were subsampled, including only one checklist to avoid duplication.
-This provided 637,482 eligible checklists throughout Australia.
+We were left with 637,482 eligible checklists throughout Australia.
 Seabirds (e.g., gannets, petrels, shearwaters) were omitted from analyses as we did not expect any correlation with seabirds using urban areas.
@@
 ## Urbanization index
-We used VIIRS night-time lights [@elvidge2017viirs] as a continuous measure for urban habitat, measuring the electric lighting associated with human settlement (Fig.
+A number of measures of urbanization have previously been used [@sanderson2002human; @hahs2006selecting], but we focused on a continuous measure that is globally applicable (i.e., applicable and easily downloaded for anywhere in the world).
+As such, we used VIIRS night-time lights [@elvidge2017viirs] as a continuous measure for urban habitat, measuring the electric lighting associated with human settlement (Fig.
 S1).
 This has filtered out background noise (e.g., fires, degraded data, and other light source contamination) to ensure validity of the data and its association with human settlement [@elvidge2017viirs].
+There is a strong positive relationship between VIIRS night-time lights and human population density (Fig.
+S2), another commonly used proxy for level of urbanization.
 For our checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], to assign the average radiance value, calculated from the VIIRS layer.
@@
 Before modelling of predictor variables (Table 1), we investigated multi-collinearity among continuous variables (Fig.
-S2), using the corrplot package [@corrplot2017].
+S3), using the corrplot package [@corrplot2017].
 We used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation among each other; all traits < 2 suggested negligible correlation (Supplementary Table 1).
@@
 
-Linear regression models were weighted [@solon2015weighting], using a source of error associated with our urbanization indices, which were log-transformed to meet model assumptions.
+We performed linear regression models with the response variable being the species-specific urbanization index and the predictor variables the 22 ecological and life history traits.
+This was done without and with controlling for phylogenetic effects (see below).
+All linear regression models were weighted [@solon2015weighting], using a source of error associated with our urbanization indices.
 This error was the proportion of the number of total observations for a species, divided by the number of its unique localities (Fig.
-S2 shows the correlation between the two).
+S3 shows the correlation between the two).
 This weighted the model-fitting to those species with more observations in more unique locations, likely to be the most reliable urbanization indices.
@@
 3; Fig.
-S3; Table 2), significantly different from random for 4 out of 5 indices (Cmean, I, K.star, and Lambda).
+S4; Table 2), significantly different from random for 4 out of 5 indices (Cmean, I, K.star, and Lambda).
 For example, the urbanization index was disproportionately important with respect to the phylogenetic tree (Fig.
@@
 These results were robust across separate non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig.
-S4), indicating that urbanization is generally excluding specialist species [@sorace2009distribution; @concepcion2015impacts].
+S5), indicating that urbanization is generally excluding specialist species [@sorace2009distribution; @concepcion2015impacts].
 Further, the non-random distribution of the urbanization index across the phylogeny (Fig.
@@
 In conclusion, our novel continuous urbanization index showed relationships between ecological and life history traits with avian tolerance to urban environments.
+Our results are also generalizable and comprehensive, given the largest sample size and spatial scale to be used to address this question.
+Further, the methodology used is globally applicable, dependent only on observational data and trait data, combined with remotely-sensed products.
 Improving understanding of the temporal nature of these relationships will be important, but we highlight the importance of considering these questions using a continuous measure of urbanization [@evans2011makes] as well as using a phylogenetically informed approach.
@@
 
-# Data, code and materials
+# Data archiving
 This work relied heavily on open-source, fully accessible datasets.
 The necessary portions of those datasets, along with code to reproduce the analyses are available at
-this github repository (https://github.com/cornwell-lab-unsw/bird_urbanness), and will be
+a github repository, and will be
 permanently archived upon acceptance of the manuscript. The supporting data has been included in the article's supplementary material (Appendix 4).
-
-# Competing interests
-We have no competing interests.
-
-# Author's Contributions
-CTC, WKC, and JHW conceived and designed the methodology of the
-study. CTC, WKC performed the statistical analyses. JMM, REM, and RTK contributed in drafting
-the manuscript. All authors provided editorial advice, approved the final version of the manuscript,
-and agree to be accountable for all aspects of the work.
 
@@
 1.
-For reference, Figure S3 is a phylogenetic tree with the names of each tip displayed, and we also provide an interactive visualization (https://johnwilshire.github.io/bird_urbanness_web_app/)."}
+For reference, Figure S3 is a phylogenetic tree with the names of each tip displayed."}
 knitr::include_graphics("finalFigs/Figure3.pdf")
@@
 
-```{r Figure S2, echo=FALSE, fig.cap="Correlations between continuous traits, number of records, and number of unique localities used in the analysis."}
+```{r Figure S2, echo=FALSE, fig.cap="The relationship between VIIRS night-time lights and human population density, fitted with a linear model (blue line)."}
 knitr::include_graphics("finalFigs/FigureS2.pdf")
@@
 
-```{r Figure S3, echo=FALSE, fig.cap="Reference phylogenetic tree of the 477 bird species in the analysis, with labeled tips, based on Jetz et al.
-(2012).
-We also provide an interactive visualization for better interpreting the phylogenetic tree (https://johnwilshire.github.io/bird_urbanness_web_app/)"}
+```{r Figure S3, echo=FALSE, fig.cap="Correlations between continuous traits, number of records, and number of unique localities used in the analysis."}
 knitr::include_graphics("finalFigs/FigureS3.pdf")
@@
 
-```{r Figure S4, echo=FALSE, fig.cap="The same phylolm model was fitted to 1000 of the family of trees presented by Jetz et al.
+```{r Figure S4, echo=FALSE, fig.cap="Reference phylogenetic tree of the 477 bird species in the analysis, with labeled tips, based on Jetz et al.
+(2012).
+We also provide an interactive visualization for better interpreting the phylogenetic tree (https://johnwilshire.github.io/bird_urbanness_web_app/)"}
+knitr::include_graphics("finalFigs/FigureS4.pdf")
+```
+
+```{r Figure S5, echo=FALSE, fig.cap="The same phylolm model was fitted to 1000 of the family of trees presented by Jetz et al.
 (2012) to account for one aspect of phylogenetic uncertainty.
 Density plots show estimated parameters for the most important traits across models."}
-knitr::include_graphics("finalFigs/FigureS4.pdf")
+knitr::include_graphics("finalFigs/FigureS5.pdf")
 ```

### 2018-09-03 scripts/results_script.R

@@ -371,18 +371,31 @@ my.file.copy(from = "figures/phy_v_non_phy.pdf",
 ## made using google earth engine, outside of R
 
 ## Figure S2
-my.file.copy(from = "figures/corrplot_of_continuous_variables.pdf",
-             to = "finalFigs/FigureS2.pdf")
+FigureS2 <- ggplot(urbanness_table, aes(x=avg_rad, y=`population-density_mean`))+
+  geom_point()+
+  geom_smooth(method="lm")+
+  theme_classic()+
+  xlab(bquote('Average radiance ('* 'nW' ~cm^-2~sr^-1*')'))+
+  ylab("Human population density")
 
+
+pdf("finalFigs/FigureS2.pdf", width=5.5, height=4.5)
+print(FigureS2)
+dev.off()
 
 ## Figure S3
-my.file.copy(from = "figures/ref_tree.pdf",
+my.file.copy(from = "figures/corrplot_of_continuous_variables.pdf",
              to = "finalFigs/FigureS3.pdf")
 
 
 ## Figure S4
-my.file.copy(from = "figures/accounting_for_phylo_uncertainty.pdf",
+my.file.copy(from = "figures/ref_tree.pdf",
              to = "finalFigs/FigureS4.pdf")
+
+
+## Figure S5
+my.file.copy(from = "figures/accounting_for_phylo_uncertainty.pdf",
+             to = "finalFigs/FigureS5.pdf")
 
 
 
